### 2022-03-28 article.Rnw
begin jss paper

\documentclass[article]{jss}

%% -- LaTeX packages and custom commands ---------------------------------------

%% recommended packages
\usepackage{orcidlink,thumbpdf,lmodern}

%% another package (only for this demo article)
\usepackage{framed}

%% new custom commands
\newcommand{\class}[1]{`\code{#1}'}
\newcommand{\fct}[1]{\code{#1()}}

%% For Sweave-based articles about R packages:
%% need no \usepackage{Sweave}
\SweaveOpts{engine=R, eps=FALSE, keep.source = TRUE}
<<preliminaries, echo=FALSE, results=hide>>=
options(prompt = "R> ", continue = "+  ", width = 70, useFancyQuotes = FALSE)
library("MASS")
@

%% -- Article metainformation (author, title, ...) -----------------------------

%% - \author{} with primary affiliation (and optionally ORCID link)
%% - \Plainauthor{} without affiliations
%% - Separate authors by \And or \AND (in \author) or by comma (in \Plainauthor).
%% - \AND starts a new line, \And does not.
\author{Toby Dylan Hocking~\orcidlink{0000-0002-3146-0865}\\Northern Arizona University}
\Plainauthor{Toby Dylan Hocking}

%% - \title{} in title case
%% - \Plaintitle{} without LaTeX markup (if any)
%% - \Shorttitle{} with LaTeX markup (if any), used as running title
\title{Comparing binsegRcpp with other implementations of binary segmentation}
\Plaintitle{A Short Demo Article: Regression Models for Count Data in R}
\Shorttitle{A Short Demo Article in \proglang{R}}

%% - \Abstract{} almost as usual
\Abstract{ Binary segmentation is a classic algorithm for detecting
 changepoints in sequential data. In theory it should be extremely fast for $N$ data and $K$ segments, $O(N K)$ in the worst case, and $O(N \log K)$ in the best case. In practice existing implementations such as ruptures (\proglang{Python} module) and changepoint (\proglang{R} package) are much slower, and in fact sometimes do not return a correct result. We present the \proglang{R} package \pkg{binsegRcpp}, which provides an efficient C++ implementation of binary segmentation, and always returns correct results. We discuss some advanced C++ coding techniques which were used to avoid repetition, for efficiency, and for readability.  We additionally include detailed theoretical and empirical comparisons with other implementations of binary segmentation in \proglang{R} packages \pkg{fpop} and \pkg{wbs}. }

%% - \Keywords{} with LaTeX markup, at least one required
%% - \Plainkeywords{} without LaTeX markup (if necessary)
%% - Should be comma-separated and in sentence case.
\Keywords{JSS, style guide, comma-separated, not capitalized, \proglang{R}}
\Plainkeywords{JSS, style guide, comma-separated, not capitalized, R}

%% - \Address{} of at least one author
%% - May contain multiple affiliations for each author
%%   (in extra lines, separated by \emph{and}\\).
%% - May contain multiple authors for the same affiliation
%%   (in the same first line, separated by comma).
\Address{
  Achim Zeileis\\
  Journal of Statistical Software\\
  \emph{and}\\
  Department of Statistics\\
  Faculty of Economics and Statistics\\
  Universit\"at Innsbruck\\
  Universit\"atsstr.~15\\
  6020 Innsbruck, Austria\\
  E-mail: \email{Achim.Zeileis@R-project.org}\\
  URL: \url{https://www.zeileis.org/}
}

\begin{document}

%% -- Introduction -------------------------------------------------------------

%% - In principle "as usual".
%% - But should typically have some discussion of both _software_ and _methods_.
%% - Use \proglang{}, \pkg{}, and \code{} markup throughout the manuscript.
%% - If such markup is in (sub)section titles, a plain text version has to be
%%   added as well.
%% - All software mentioned should be properly \cite-d.
%% - All abbreviations should be introduced.
%% - Unless the expansions of abbreviations are proper names (like "Journal
%%   of Statistical Software" above) they should be in sentence case (like
%%   "generalized linear models" below).

\section[Introduction: Count data regression in R]{Introduction: Count data regression in \proglang{R}} \label{sec:intro}

\begin{leftbar}

The introduction is in principle ``as usual''. However, it should usually embed
both the implemented \emph{methods} and the \emph{software} into the respective
relevant literature. For the latter both competing and complementary software
should be discussed (within the same software environment and beyond), bringing
out relative (dis)advantages. All software mentioned should be properly
\verb|\cite{}|d. (See also Appendix~\ref{app:bibtex} for more details on
\textsc{Bib}{\TeX}.)

For writing about software JSS requires authors to use the markup
\verb|\proglang{}| (programming languages and large programmable systems),
\verb|\pkg{}| (software packages), \verb|\code{}| (functions, commands,
arguments, etc.). If there is such markup in (sub)section titles (as above), a
plain text version has to be provided in the {\LaTeX} command as well. Below we
also illustrate how abbrevations should be introduced and citation commands can
be employed. See the {\LaTeX} code for more details.
\end{leftbar}

TODO (Standard Template Library containers, virtual classes, static
variables, function pointers, templates, macros).

TODO \code{fpop::multiBinSeg} \citep{Maidstone2017}.

TODO \code{wbs::sbs} \citep{Baranowski2019}.

TODO \code{changepoint::cpt.mean(method="BinSeg")} \citep{Killick2014,Killick2022}.

TODO \code{ruptures.Binseg} \citep{Truong2020}.

Modeling count variables is a common task in economics and the social sciences.
The classical Poisson regression model for count data is often of limited use in
these disciplines because empirical count data sets typically exhibit
overdispersion and/or an excess number of zeros. The former issue can be
addressed by extending  the plain Poisson regression model in various
directions: e.g., using sandwich covariances or estimating an additional
dispersion parameter (in a so-called quasi-Poisson model). Another more formal
way is to use a negative binomial (NB) regression. All of these models belong to
the family of generalized linear models (GLMs). However, although these models
typically can capture overdispersion rather well, they are in many applications
not sufficient for  modeling excess zeros. Since \cite{Mullahy:1986} there is
increased interest in zero-augmented models that address this issue by a second
model component capturing zero counts. An overview of count data models in
econometrics, including  hurdle and zero-inflated models, is provided in
\cite{Cameron+Trivedi:2013}.

In \proglang{R} \citep{R}, GLMs are provided by the model fitting functions
\fct{glm} in the \pkg{stats} package and \fct{glm.nb} in the \pkg{MASS} package
\citep[][Chapter~7.4]{Venables+Ripley:2002} along with associated methods for
diagnostics and inference. The manuscript that this document is based on
\citep{Zeileis+Kleiber+Jackman:2008} then introduced hurdle and zero-inflated
count models in the functions \fct{hurdle} and \fct{zeroinfl} in the \pkg{pscl}
package \citep{Jackman:2015}. Of course, much more software could be discussed
here, including (but not limited to) generalized additive models for count data
as available in the \proglang{R} packages \pkg{mgcv} \cite{Wood:2006}, 
\pkg{gamlss} \citep{Stasinopoulos+Rigby:2007}, or \pkg{VGAM} \citep{Yee:2009}.

%% -- Manuscript ---------------------------------------------------------------

%% - In principle "as usual" again.
%% - When using equations (e.g., {equation}, {eqnarray}, {align}, etc.
%%   avoid empty lines before and after the equation (which would signal a new
%%   paragraph.
%% - When describing longer chunks of code that are _not_ meant for execution
%%   (e.g., a function synopsis or list of arguments), the environment {Code}
%%   is recommended. Alternatively, a plain {verbatim} can also be used.
%%   (For executed code see the next section.)

\section{Models and software} \label{sec:models}

The basic Poisson regression model for count data is a special case of the GLM
framework \cite{McCullagh+Nelder:1989}. It describes the dependence of a count
response variable $y_i$ ($i = 1, \dots, n$) by assuming a Poisson distribution
$y_i \sim \mathrm{Pois}(\mu_i)$. The dependence of the conditional mean
$\E[y_i \, | \, x_i] = \mu_i$ on the regressors $x_i$ is then specified via a
log link and a linear predictor
%
\begin{equation} \label{eq:mean}
\log(\mu_i) \quad = \quad x_i^\top \beta,
\end{equation}
%
where the regression coefficients $\beta$ are estimated by maximum likelihood
(ML) using the iterative weighted least squares (IWLS) algorithm.

\begin{leftbar}
Note that around the \verb|{equation}| above there should be no spaces (avoided
in the {\LaTeX} code by \verb|%| lines) so that ``normal'' spacing is used and
not a new paragraph started.
\end{leftbar}

\proglang{R} provides a very flexible implementation of the general GLM
framework in the function \fct{glm} \citep{Chambers+Hastie:1992} in the
\pkg{stats} package. Its most important arguments are
\begin{Code}
glm(formula, data, subset, na.action, weights, offset,
  family = gaussian, start = NULL, control = glm.control(...),
  model = TRUE, y = TRUE, x = FALSE, ...)
\end{Code}
where \code{formula} plus \code{data} is the now standard way of specifying
regression relationships in \proglang{R}/\proglang{S} introduced in
\cite{Chambers+Hastie:1992}. The remaining arguments in the first line
(\code{subset}, \code{na.action}, \code{weights}, and \code{offset}) are also
standard  for setting up formula-based regression models in
\proglang{R}/\proglang{S}. The arguments in the second line control aspects
specific to GLMs while the arguments in the last line specify which components
are returned in the fitted model object (of class \class{glm} which inherits
from \class{lm}). For further arguments to \fct{glm} (including alternative
specifications of starting values) see \code{?glm}. For estimating a Poisson
model \code{family = poisson} has to be specified.

\begin{leftbar}
As the synopsis above is a code listing that is not meant to be executed,
one can use either the dedicated \verb|{Code}| environment or a simple
\verb|{verbatim}| environment for this. Again, spaces before and after should be
avoided.

Finally, there might be a reference to a \verb|{table}| such as
Table~\ref{tab:overview}. Usually, these are placed at the top of the page
(\verb|[t!]|), centered (\verb|\centering|), with a caption below the table,
column headers and captions in sentence style, and if possible avoiding vertical
lines.
\end{leftbar}

\begin{table}[t!]
\centering
\begin{tabular}{lllp{7.4cm}}
\hline
Type           & Distribution & Method   & Description \\ \hline
GLM            & Poisson      & ML       & Poisson regression: classical GLM,
                                           estimated by maximum likelihood (ML) \\
               &              & Quasi    & ``Quasi-Poisson regression'':
                                           same mean function, estimated by
                                           quasi-ML (QML) or equivalently
                                           generalized estimating equations (GEE),
                                           inference adjustment via estimated
                                           dispersion parameter \\
               &              & Adjusted & ``Adjusted Poisson regression'':
                                           same mean function, estimated by
                                           QML/GEE, inference adjustment via
                                           sandwich covariances\\
               & NB           & ML       & NB regression: extended GLM,
                                           estimated by ML including additional
                                           shape parameter \\ \hline
Zero-augmented & Poisson      & ML       & Zero-inflated Poisson (ZIP),
                                           hurdle Poisson \\
               & NB           & ML       & Zero-inflated NB (ZINB),
                                           hurdle NB \\ \hline
\end{tabular}
\caption{\label{tab:overview} Overview of various count regression models. The
table is usually placed at the top of the page (\texttt{[t!]}), centered
(\texttt{centering}), has a caption below the table, column headers and captions
are in sentence style, and if possible vertical lines should be avoided.}
\end{table}

%% -- Illustrations ------------------------------------------------------------

%% - Virtually all JSS manuscripts list source code along with the generated
%%   output. The style files provide dedicated environments for this.
%% - In R, the environments {Sinput} and {Soutput} - as produced by Sweave() or
%%   or knitr using the render_sweave() hook - are used (without the need to
%%   load Sweave.sty).
%% - Equivalently, {CodeInput} and {CodeOutput} can be used.
%% - The code input should use "the usual" command prompt in the respective
%%   software system.
%% - For R code, the prompt "R> " should be used with "+  " as the
%%   continuation prompt.
%% - Comments within the code chunks should be avoided - these should be made
%%   within the regular LaTeX text.

\section{Illustrations} \label{sec:illustrations}

For a simple illustration of basic Poisson and NB count regression the
\code{quine} data from the \pkg{MASS} package is used. This provides the number
of \code{Days} that children were absent from school in Australia in a
particular year, along with several covariates that can be employed as regressors.
The data can be loaded by
%
<<data>>=
data("quine", package = "MASS")
@
%
and a basic frequency distribution of the response variable is displayed in
Figure~\ref{fig:quine}.

\begin{leftbar}
For code input and output, the style files provide dedicated environments.
Either the ``agnostic'' \verb|{CodeInput}| and \verb|{CodeOutput}| can be used
or, equivalently, the environments \verb|{Sinput}| and \verb|{Soutput}| as
produced by \fct{Sweave} or \pkg{knitr} when using the \code{render_sweave()}
hook. Please make sure that all code is properly spaced, e.g., using
\code{y = a + b * x} and \emph{not} \code{y=a+b*x}. Moreover, code input should
use ``the usual'' command prompt in the respective software system. For
\proglang{R} code, the prompt \code{"R> "} should be used with \code{"+  "} as
the continuation prompt. Generally, comments within the code chunks should be
avoided -- and made in the regular {\LaTeX} text instead. Finally, empty lines
before and after code input/output should be avoided (see above).
\end{leftbar}

\begin{figure}[t!]
\centering
<<visualization, echo=FALSE, fig=TRUE, height=5.2, width=7>>=
par(mar = c(4, 4, 1, 1))
plot(table(quine$Days), xlab = "Days", ylab = "Frequency", axes = FALSE)
axis(2)
axis(1, at = 0:16 * 5, labels = FALSE)
axis(1, at = 0:8 * 10)
box()
@
\caption{\label{fig:quine} Frequency distribution for number of days absent
from school.}
\end{figure}

As a first model for the \code{quine} data, we fit the basic Poisson regression
model. (Note that JSS prefers when the second line of code is indented by two
spaces.)
%
<<poisson>>=
m_pois <- glm(Days ~ (Eth + Sex + Age + Lrn)^2, data = quine,
  family = poisson)
@
%
To account for potential overdispersion we also consider a negative binomial
GLM.
%
<<negbin>>=
library("MASS")
m_nbin <- glm.nb(Days ~ (Eth + Sex + Age + Lrn)^2, data = quine)
@
%
In a comparison with the BIC the latter model is clearly preferred.
%
<<comparison>>=
BIC(m_pois, m_nbin)
@
%
Hence, the full summary of that model is shown below.
%
<<summary>>=
summary(m_nbin)
@

%% -- Summary/conclusions/discussion -------------------------------------------

\section{Summary and discussion} \label{sec:summary}

\begin{leftbar}
As usual \dots
\end{leftbar}

%% -- Optional special unnumbered sections -------------------------------------

\section*{Computational details}

\begin{leftbar}
If necessary or useful, information about certain computational details
such as version numbers, operating systems, or compilers could be included
in an unnumbered section. Also, auxiliary packages (say, for visualizations,
maps, tables, \dots) that are not cited in the main text can be credited here.
\end{leftbar}

The results in this paper were obtained using
\proglang{R}~\Sexpr{paste(R.Version()[6:7], collapse = ".")} with the
\pkg{MASS}~\Sexpr{packageVersion("MASS")} package. \proglang{R} itself
and all packages used are available from the Comprehensive
\proglang{R} Archive Network (CRAN) at
\url{https://CRAN.R-project.org/}.

\section*{Acknowledgments}

\begin{leftbar}
All acknowledgments (note the AE spelling) should be collected in this
unnumbered section before the references. It may contain the usual information
about funding and feedback from colleagues/reviewers/etc. Furthermore,
information such as relative contributions of the authors may be added here
(if any).
\end{leftbar}

%% -- Bibliography -------------------------------------------------------------
%% - References need to be provided in a .bib BibTeX database.
%% - All references should be made with \cite, \citet, \citep, \citealp etc.
%%   (and never hard-coded). See the FAQ for details.
%% - JSS-specific markup (\proglang, \pkg, \code) should be used in the .bib.
%% - Titles in the .bib should be in title case.
%% - DOIs should be included where available.

\bibliography{refs}

%% -- Appendix (if any) --------------------------------------------------------
%% - After the bibliography with page break.
%% - With proper section titles and _not_ just "Appendix".

\newpage

\begin{appendix}

\section{More technical details} \label{app:technical}

\begin{leftbar}
Appendices can be included after the bibliography (with a page break). Each
section within the appendix should have a proper section title (rather than
just \emph{Appendix}).

For more technical style details, please check out JSS's style FAQ at
\url{https://www.jstatsoft.org/pages/view/style#frequently-asked-questions}
which includes the following topics:
\begin{itemize}
  \item Title vs.\ sentence case.
  \item Graphics formatting.
  \item Naming conventions.
  \item Turning JSS manuscripts into \proglang{R} package vignettes.
  \item Trouble shooting.
  \item Many other potentially helpful details\dots
\end{itemize}
\end{leftbar}

\section[Using BibTeX]{Using \textsc{Bib}{\TeX}} \label{app:bibtex}

\begin{leftbar}
References need to be provided in a \textsc{Bib}{\TeX} file (\code{.bib}). All
references should be made with \verb|\cite|, \verb|\citet|, \verb|\citep|,
\verb|\citealp| etc.\ (and never hard-coded). This commands yield different
formats of author-year citations and allow to include additional details (e.g.,
pages, chapters, \dots) in brackets. In case you are not familiar with these
commands see the JSS style FAQ for details.

Cleaning up \textsc{Bib}{\TeX} files is a somewhat tedious task -- especially
when acquiring the entries automatically from mixed online sources. However,
it is important that informations are complete and presented in a consistent
style to avoid confusions. JSS requires the following format.
\begin{itemize}
  \item JSS-specific markup (\verb|\proglang|, \verb|\pkg|, \verb|\code|) should
    be used in the references.
  \item Titles should be in title case.
  \item Journal titles should not be abbreviated and in title case.
  \item DOIs should be included where available.
  \item Software should be properly cited as well. For \proglang{R} packages
    \code{citation("pkgname")} typically provides a good starting point.
\end{itemize}
\end{leftbar}

\end{appendix}

%% -----------------------------------------------------------------------------

\end{document}

### 2022-03-31 article.Rnw
comparison table

@@
 %% recommended packages
-\usepackage{orcidlink,thumbpdf,lmodern}
+\usepackage{orcidlink,thumbpdf,lmodern,amsmath,amssymb}
 
@@
 
-Modeling count variables is a common task in economics and the social sciences.
-The classical Poisson regression model for count data is often of limited use in
-these disciplines because empirical count data sets typically exhibit
-overdispersion and/or an excess number of zeros. The former issue can be
-addressed by extending  the plain Poisson regression model in various
-directions: e.g., using sandwich covariances or estimating an additional
-dispersion parameter (in a so-called quasi-Poisson model). Another more formal
-way is to use a negative binomial (NB) regression. All of these models belong to
-the family of generalized linear models (GLMs). However, although these models
-typically can capture overdispersion rather well, they are in many applications
-not sufficient for  modeling excess zeros. Since \cite{Mullahy:1986} there is
-increased interest in zero-augmented models that address this issue by a second
-model component capturing zero counts. An overview of count data models in
-econometrics, including  hurdle and zero-inflated models, is provided in
-\cite{Cameron+Trivedi:2013}.
-
-In \proglang{R} \citep{R}, GLMs are provided by the model fitting functions
-\fct{glm} in the \pkg{stats} package and \fct{glm.nb} in the \pkg{MASS} package
-\citep[][Chapter~7.4]{Venables+Ripley:2002} along with associated methods for
-diagnostics and inference. The manuscript that this document is based on
-\citep{Zeileis+Kleiber+Jackman:2008} then introduced hurdle and zero-inflated
-count models in the functions \fct{hurdle} and \fct{zeroinfl} in the \pkg{pscl}
-package \citep{Jackman:2015}. Of course, much more software could be discussed
-here, including (but not limited to) generalized additive models for count data
-as available in the \proglang{R} packages \pkg{mgcv} \cite{Wood:2006}, 
-\pkg{gamlss} \citep{Stasinopoulos+Rigby:2007}, or \pkg{VGAM} \citep{Yee:2009}.
-
 %% -- Manuscript ---------------------------------------------------------------
@@
 \centering
-\begin{tabular}{lllp{7.4cm}}
+\begin{tabular}{l|ccccc}
+  package & binsegRcpp & changepoint & wbs & fpop & ruptures \\
+  function & binseg & cpt.mean & sbs & multiBinSeg & Binseg \\
+%   > ruptures$version$version
+% [1] "1.1.6"
+% > sapply(c("binsegRcpp", "changepoint", "wbs", "fpop"), function(p)paste(packageVersion(p)))
+%  binsegRcpp changepoint         wbs        fpop 
+% "2022.3.29"     "2.2.3"       "1.4" "2019.8.26" 
+  version & 2022.3.29 & 2.2.3 & 1.4 &2019.8.26 & 1.1.6\\
 \hline
-Type           & Distribution & Method   & Description \\ \hline
-GLM            & Poisson      & ML       & Poisson regression: classical GLM,
-                                           estimated by maximum likelihood (ML) \\
-               &              & Quasi    & ``Quasi-Poisson regression'':
-                                           same mean function, estimated by
-                                           quasi-ML (QML) or equivalently
-                                           generalized estimating equations (GEE),
-                                           inference adjustment via estimated
-                                           dispersion parameter \\
-               &              & Adjusted & ``Adjusted Poisson regression'':
-                                           same mean function, estimated by
-                                           QML/GEE, inference adjustment via
-                                           sandwich covariances\\
-               & NB           & ML       & NB regression: extended GLM,
-                                           estimated by ML including additional
-                                           shape parameter \\ \hline
-Zero-augmented & Poisson      & ML       & Zero-inflated Poisson (ZIP),
-                                           hurdle Poisson \\
-               & NB           & ML       & Zero-inflated NB (ZINB),
-                                           hurdle NB \\ \hline
+  weights & yes & no & no & no & no \\
+  max segs & yes & yes & no & yes & yes \\
+  dim & one & one & one & multi & multi \\
+  correct & yes & no & yes & yes & no \\
+  losses & C++ & C & L2 & L2 & Python \\
+  % all propose some loss computation.
+  storage & STL multiset & arrays & recursion & heap & LRU cache \\
+  space & $O(S)$ & $O(S^2)$ & $O(S)$ & $O(S)$ & $O(S)$ \\
+  cumsum & yes & yes & yes & yes & no \\
+  best & $O(N\log N)$ & $O(N^3)$ & $O(N\log N)$ & $O(N\log N)$ & $>O(N\log N)$ \\
+  worst & $O(N^2)$ & $O(N^3)$ & $O(N^2)$ & $O(N^2)$ & $O(N^2)$ \\
+  CV & yes & no & no & no & no\\
+  params & yes & largest & no & no & no \\
 \end{tabular}
-\caption{\label{tab:overview} Overview of various count regression models. The
-table is usually placed at the top of the page (\texttt{[t!]}), centered
-(\texttt{centering}), has a caption below the table, column headers and captions
-are in sentence style, and if possible vertical lines should be avoided.}
+\caption{\label{tab:overview} TODO}
 \end{table}
+
+\paragraph{Loss computation.} We run each algorithm on \code{N.data}
+points up to \code{max.segments =
+  max.changes+1}. \code{binsegRcpp::binseg} result is a list which
+contains a data table with \code{max.segments} rows and column
+\code{loss} that is the square loss.  \code{changepoint::cpt.mean}
+result is a list of class \code{cpt.range} with method \code{logLik}
+which returns the square loss of one of the models. \code{wbs::sbs}
+result is a list which contains a data frame with \code{N.data-1} rows
+and \code{CUSUM} and \code{min.th} columns
+TODO.\code{fpop::multiBinSeg} result is a list with element
+\code{J.est}, which is a vector of \code{max.changes} square loss
+decrease values.  \code{ruptures.Binseg.predict} result is a vector of
+segment ends for one model size, which can be passed to
+\code{sum_of_costs} method to compute the square loss.
+
+Three packages have implemented the normal change in mean and variance
+model. \code{binsegRcpp::binseg} loss values are the normal negative
+log likelihood (NLL). The \code{changepoint::logLik} function returns two
+times the NLL. The \code{ruptures.Binseg}
+loss is related via this equation,
+
+\begin{equation}
+  \text{NLL} = (\text{rupturesLoss} + N[1+\log(2\pi)])/2
+\end{equation}
 

### 2022-03-31 figure-comparison-mean.R

@@ -1,0 +1,38 @@
+source("packages.R")
+Sys.setenv(RETICULATE_PYTHON=if(.Platform$OS.type=="unix")
+  "/home/tdhock/.local/share/r-miniconda/envs/cs570s22/bin/python"
+  else "~/Miniconda3/envs/cs570s22/python.exe")
+reticulate::use_condaenv("cs570s22", required=TRUE)
+ruptures <- reticulate::import("ruptures")
+x <- c(0,0.3,0.2,0.1, 10,11,12,13)
+N.data <- length(x)
+x.mat <- matrix(x)
+max_bkps <- 3
+changepoint::cpt.mean(x, method="BinSeg", Q=max_bkps)
+mbs.fit <- fpop::multiBinSeg(x,max_bkps)
+bs.fit <- binsegRcpp::binseg("mean_norm", x, max_bkps+1)
+
+binseg_instance <- ruptures$Binseg(model="l2", jump=1L)$fit(x.mat)
+(b0 <- binseg_instance$predict(n_bkps=0))
+binseg_instance$cost$sum_of_costs(as.list(b0))
+binseg_instance$predict(n_bkps=1)
+binseg_instance$predict(n_bkps=2)
+ruptures.loss <- sapply(seq(0,n_bkps), function(n_bkps){
+  end.vec <- binseg_instance$predict(n_bkps=n_bkps)
+  tryCatch({
+    binseg_instance$cost$sum_of_costs(as.list(end.vec))
+  }, error=function(e){
+    NA_real_
+  })
+})
+sbs.fit <- wbs::sbs(x)
+cpt.fit <- changepoint::cpt.mean(x, method="BinSeg", Q=max_bkps)
+changepoint::logLik(cpt.fit)
+changepoint::cpts(cpt.fit)
+cpt.fit@cpts.full
+sapply(c("binsegRcpp", "changepoint", "wbs", "fpop"), function(p)paste(packageVersion(p)))
+ruptures$version$version
+rbind(
+  ruptures=ruptures.loss,
+  fpop=cumsum(c(sum((x-mean(x))^2), mbs.fit$J.est)),
+  binsegRcpp=bs.fit$splits$loss)

### 2022-03-31 figure-comparison-meanvar.R

@@ -1,0 +1,32 @@
+source("packages.R")
+Sys.setenv(RETICULATE_PYTHON=if(.Platform$OS.type=="unix")
+  "/home/tdhock/.local/share/r-miniconda/envs/cs570s22/bin/python"
+  else "~/Miniconda3/envs/cs570s22/python.exe")
+reticulate::use_condaenv("cs570s22", required=TRUE)
+ruptures <- reticulate::import("ruptures")
+x <- c(0,0.3,0.2,0.1, 10,11,12,13)
+N.data <- length(x)
+x.mat <- matrix(x)
+n_bkps <- 3
+changepoint::cpt.meanvar(x, method="BinSeg", Q=n_bkps)
+bs.fit <- binsegRcpp::binseg("meanvar_norm", x, n_bkps+1)
+seg.dt <- coef(bs.fit, 4L)
+seg.dt[, n.data := end-start+1]
+seg.dt[, data.table(
+  data=x,
+  mean=dput(rep(mean,n.data)),
+  var=dput(rep(var,n.data)))]
+binseg_instance <- ruptures$Binseg(model="normal", jump=1L)$fit(x.mat)
+bs.fit$splits$loss
+## Converting ruptures loss to normal negative log likelihood.
+convert <- function(rcost)(rcost+N.data*(1+log(2*pi)))/2
+(b0 <- binseg_instance$predict(n_bkps=0))
+myvar <- function(y)mean((y-mean(y))^2)
+v <- myvar(x)
+eps <- 1e-6
+log(v+eps)*N.data
+binseg_instance$cost$sum_of_costs(list(b0))
+(b1 <- binseg_instance$predict(n_bkps=1))
+convert(binseg_instance$cost$sum_of_costs(b1))
+binseg_instance$predict(n_bkps=2)
+binseg_instance$predict(n_bkps=3)

### 2022-04-06 figure-timings-meanvar_norm-data.R
initial timings

@@ -1,4 +1,5 @@
 source("packages.R")
+
 seconds.limit <- 100
 do.sub <- function(...){
   mcall <- match.call()
@@ -18,7 +19,8 @@ Sys.setenv(RETICULATE_PYTHON=if(.Platform$OS.type=="unix")
   else "~/Miniconda3/envs/cs570s22/python.exe")
 reticulate::use_condaenv("cs570s22", required=TRUE)
 ruptures <- reticulate::import("ruptures")
-binseg_instance <- ruptures$Binseg(min_size=1L, jump=1L, model="l1")
+binseg_instance <- ruptures$Binseg(min_size=1L, jump=1L, model="normal")
+
 timing.dt.list <- list()
 done.list <- list()
 for(N.data.exp in 2:20){#2^20 = 1,048,576
@@ -34,17 +36,27 @@ for(N.data.exp in 2:20){#2^20 = 1,048,576
   for(case in names(data.list)){
     data.vec <- data.list[[case]]
     cum.data.vec <- cumsum(c(0,data.vec))
+    cum.square.vec <- cumsum(c(0,data.vec^2))
     data.mat <- matrix(data.vec)
     result.list <- list()
     m.args <- do.sub("ruptures"={
       binseg_instance$fit(data.mat)$predict(n_bkps=max.changes)
+    }, "changepoint"={
+      cpt.fit <- changepoint::cpt.meanvar(
+        data.vec, method="BinSeg", Q=max.changes)
+      sort(c(N.data,cpt.fit@cpts.full[max.changes,]))
+    }, blockcpd={
+      bcpd.fit <- blockcpd::fit_blockcpd(
+        t(data.mat), method="hierseg", family="normal", lambda=0,
+        min_block_size = 2L)
+      c(bcpd.fit$changepoints, N.data)
     }, "binsegRcpp.multiset"={
       binseg.fit <- binsegRcpp::binseg(
-        "l1",data.vec, max.segs, container.str="multiset")
+        "meanvar_norm",data.vec, max.segs, container.str="multiset")
       sort(binseg.fit$splits$end)
     }, "binsegRcpp.list"={
       binseg.fit <- binsegRcpp::binseg(
-        "l1",data.vec, max.segs, container.str="list")
+        "meanvar_norm",data.vec, max.segs, container.str="list")
       sort(binseg.fit$splits$end)
     }, times=5)
     m.args[names(done.list[[case]])] <- NULL
[…]

### 2022-04-06 figure-timings-meanvar_norm.R

@@ -1,5 +1,5 @@
 source("packages.R")
-timing.stats <- data.table::fread("figure-timings-laplace-data.csv")
+timing.stats <- data.table::fread("figure-timings-meanvar_norm-data.csv")
 disp.pkg <- c(
   changepoint="changepoint.array",
   ruptures="ruptures.LRU cache",
@@ -31,12 +31,12 @@ gg <- ggplot()+
   scale_y_log10(
     "Mean squared error (log scale)")
 (dl <- directlabels::direct.label(gg, list(cex=0.6, "last.polygons")))
-png("figure-timings-laplace-loss.png", width=9, height=3.5, units="in", res=200)
+png("figure-timings-meanvar_norm-loss.png", width=9, height=3.5, units="in", res=200)
 print(dl)
 dev.off()
 
 gg <- ggplot()+
-  ggtitle("L1 loss")+
+  ggtitle("Normal change in mean and variance")+
   theme_bw()+
   theme(
     legend.position="none",
@@ -74,6 +74,6 @@ gg <- ggplot()+
   scale_y_log10(
     "Computation time (seconds, log scale)\nMedian line and min/max band over 5 timings",
     breaks=10^seq(-10,10))
-png("figure-timings-laplace.png", width=9, height=3.5, units="in", res=200)
+png("figure-timings-meanvar_norm.png", width=9, height=3.5, units="in", res=200)
 print(gg)
 dev.off()

### 2022-04-06 figure-timings-poisson-data.R

@@ -1,4 +1,5 @@
 source("packages.R")
+
 seconds.limit <- 100
 do.sub <- function(...){
   mcall <- match.call()
@@ -13,12 +14,7 @@ do.sub <- function(...){
   }
   L
 }
-Sys.setenv(RETICULATE_PYTHON=if(.Platform$OS.type=="unix")
-  "/home/tdhock/.local/share/r-miniconda/envs/cs570s22/bin/python"
-  else "~/Miniconda3/envs/cs570s22/python.exe")
-reticulate::use_condaenv("cs570s22", required=TRUE)
-ruptures <- reticulate::import("ruptures")
-binseg_instance <- ruptures$Binseg(min_size=1L, jump=1L, model="l1")
+
 timing.dt.list <- list()
 done.list <- list()
 for(N.data.exp in 2:20){#2^20 = 1,048,576
@@ -28,7 +24,7 @@ for(N.data.exp in 2:20){#2^20 = 1,048,576
   print(N.data)
   set.seed(1)
   data.list <- list(
-    flat=rnorm(N.data),
+    flat=rpois(N.data,100),
     best=1:N.data,
     worst=rep(c(0,1),l=N.data))
   for(case in names(data.list)){
@@ -36,15 +32,21 @@ for(N.data.exp in 2:20){#2^20 = 1,048,576
     cum.data.vec <- cumsum(c(0,data.vec))
     data.mat <- matrix(data.vec)
     result.list <- list()
-    m.args <- do.sub("ruptures"={
-      binseg_instance$fit(data.mat)$predict(n_bkps=max.changes)
+    m.args <- do.sub("changepoint"={
+      cpt.fit <- changepoint::cpt.meanvar(
+        data.vec, method="BinSeg", Q=max.changes, test.stat="Poisson")
+      sort(c(N.data,cpt.fit@cpts.full[max.changes,]))
+    }, blockcpd={
+      bcpd.fit <- blockcpd::fit_blockcpd(
+        t(data.mat), method="hierseg", family="poisson", lambda=0)
+      c(bcpd.fit$changepoints, N.data)
     }, "binsegRcpp.multiset"={
       binseg.fit <- binsegRcpp::binseg(
-        "l1",data.vec, max.segs, container.str="multiset")
+        "poisson",data.vec, max.segs, container.str="multiset")
       sort(binseg.fit$splits$end)
     }, "binsegRcpp.list"={
       binseg.fit <- binsegRcpp::binseg(
[…]

### 2022-04-06 figure-timings-poisson.R

@@ -1,5 +1,5 @@
 source("packages.R")
-timing.stats <- data.table::fread("figure-timings-laplace-data.csv")
+timing.stats <- data.table::fread("figure-timings-poisson-data.csv")
 disp.pkg <- c(
   changepoint="changepoint.array",
   ruptures="ruptures.LRU cache",
@@ -31,12 +31,12 @@ gg <- ggplot()+
   scale_y_log10(
     "Mean squared error (log scale)")
 (dl <- directlabels::direct.label(gg, list(cex=0.6, "last.polygons")))
-png("figure-timings-laplace-loss.png", width=9, height=3.5, units="in", res=200)
+png("figure-timings-poisson-loss.png", width=9, height=3.5, units="in", res=200)
 print(dl)
 dev.off()
 
 gg <- ggplot()+
-  ggtitle("L1 loss")+
+  ggtitle("Poisson loss")+
   theme_bw()+
   theme(
     legend.position="none",
@@ -74,6 +74,6 @@ gg <- ggplot()+
   scale_y_log10(
     "Computation time (seconds, log scale)\nMedian line and min/max band over 5 timings",
     breaks=10^seq(-10,10))
-png("figure-timings-laplace.png", width=9, height=3.5, units="in", res=200)
+png("figure-timings-poisson.png", width=9, height=3.5, units="in", res=200)
 print(gg)
 dev.off()

### 2022-04-06 packages.R

@@ -51,6 +51,7 @@ works_with_R <- function(Rvers,...){
 }
 .libPaths("library")
 library(data.table)
+library(blockcpd)
 ##remotes::install_github("tdhock/binsegRcpp@61ad58075ae0399d69a8565949ff7aa944f0a461")
 ##install.packages("~/R/binsegRcpp", repo=NULL,type="source")
 works_with_R(

### 2022-04-06 figure-splits-loss.R
splits loss fig

@@ -1,0 +1,49 @@
+source("packages.R")
+
+N.data.exp <- 4
+N.data <- 2^N.data.exp
+print(N.data)
+set.seed(1)
+data.list <- list(
+  flat=rnorm(N.data),
+  linear=1:N.data,
+  zero.one=rep(c(0,1),l=N.data))
+out.dt.list <- list()
+for(loss in c("l1", "poisson", "meanvar_norm", "mean_norm")){
+  for(data.name in names(data.list)){
+    data.vec <- data.list[[data.name]]
+    binseg.fit <- binsegRcpp::binseg(loss, data.vec)
+    clist <- binsegRcpp::get_complexity(binseg.fit)
+    for(out.name in names(clist)){
+      out.dt.list[[out.name]][[paste(loss, data.name)]] <- data.table(
+        loss, data.name, clist[[out.name]])
+    }
+  }
+}
+out <- lapply(out.dt.list, function(L)do.call(rbind, L))
+library(ggplot2)
+gg <- ggplot()+
+  facet_grid(loss ~ data.name, labeller=label_both)+
+  geom_line(aes(
+    segments, splits, color=case, size=case),
+    data=out$iterations[case!="empirical"])+
+  geom_point(aes(
+    segments, splits, color=case),
+    data=out$iterations[case=="empirical"])+
+  geom_text(aes(
+    x, y,
+    label=label,
+    color=case),
+    hjust=1,
+    data=out$totals)+
+  scale_color_manual(
+    values=binsegRcpp::case.colors,
+    guide="none")+
+  scale_size_manual(
+    values=binsegRcpp::case.sizes,
+    guide="none")+
+  theme_bw()+
+  theme(panel.spacing=grid::unit(0,"lines"))
+png("figure-splits-loss.png", width=10, height=6, units="in", res=200)
+print(gg)
+dev.off()

### 2022-04-06 figure-splits-loss.R
splits loss update

@@ -1,19 +1,22 @@
 source("packages.R")
-
-N.data.exp <- 4
+N.data.exp <- 6
 N.data <- 2^N.data.exp
 print(N.data)
 set.seed(1)
 data.list <- list(
-  flat=rnorm(N.data),
+  rpois1000=rpois(N.data,1000),
   linear=1:N.data,
+  zero.ten=rep(c(0,1,10,11),l=N.data),
   zero.one=rep(c(0,1),l=N.data))
 out.dt.list <- list()
+data.name <- "zero.one"
+loss <- "meanvar_norm"
 for(loss in c("l1", "poisson", "meanvar_norm", "mean_norm")){
   for(data.name in names(data.list)){
     data.vec <- data.list[[data.name]]
     binseg.fit <- binsegRcpp::binseg(loss, data.vec)
     clist <- binsegRcpp::get_complexity(binseg.fit)
+    clist$iterations[case=="worst"]
     for(out.name in names(clist)){
       out.dt.list[[out.name]][[paste(loss, data.name)]] <- data.table(
         loss, data.name, clist[[out.name]])
@@ -29,6 +32,7 @@ gg <- ggplot()+
     data=out$iterations[case!="empirical"])+
   geom_point(aes(
     segments, splits, color=case),
+    shape=1,
     data=out$iterations[case=="empirical"])+
   geom_text(aes(
     x, y,
@@ -44,6 +48,7 @@ gg <- ggplot()+
     guide="none")+
   theme_bw()+
   theme(panel.spacing=grid::unit(0,"lines"))
-png("figure-splits-loss.png", width=10, height=6, units="in", res=200)
+print(gg)
+png("figure-splits-loss.png", width=14, height=6, units="in", res=200)
 print(gg)
 dev.off()

### 2022-04-07 article.Rnw
container tab
meanvar_norm timings redo

@@
 We discuss some advanced C++ coding techniques which were used to avoid repetition, for efficiency, and for readability.
-We additionally include detailed theoretical and empirical comparisons with other implementations of binary segmentation in \proglang{R} packages \pkg{fpop} and \pkg{wbs}.
+We additionally include detailed theoretical and empirical comparisons with other implementations of binary segmentation in \proglang{R} packages \pkg{fpop}, \pkg{blockcpd} and \pkg{wbs}.
 }
@@
 
+\begin{table}
+  \centering
+  \begin{tabular}{c|cc|cc|cc}
+    case/splits: & \multicolumn{2}{c}{one } & \multicolumn{2}{c}{best/equal} & \multicolumn{2}{c}{worst/unequal } \\
+    operation: & insert &
+                             %remove &
+                                      argmin & insert & argmin & insert & argmin\\
+    \hline
+    list & $O(1)$ &
+    % $O(1)$ &
+               $O(n)$ &
+    $O(S\log S)$ & $O(S)$ & $O(S)$ & $O(S)$\\
+    heap/multiset & $O(\log n)$ &
+    % $O(\log n)$ &
+                    $O(1)$ &
+                             $O(S)$ & $O(S^2)$ & $O(S)$ & $O(S)$\\
+    \hline
+    Find new split & \multicolumn{2}{c}{} & \multicolumn{2}{c}{$O(N\log S)$} & \multicolumn{2}{c}{$O(NS)$}
+  \end{tabular}
+  \caption{container properties}
+  \label{tab:containers}
+\end{table}
+
 %% -- Illustrations ------------------------------------------------------------

### 2022-04-07 figure-timings-meanvar_norm-data.R

@@ -23,6 +23,8 @@ binseg_instance <- ruptures$Binseg(min_size=1L, jump=1L, model="normal")
 
 timing.dt.list <- list()
 done.list <- list()
+N.data.exp <- 20
+case <- "best"
 for(N.data.exp in 2:20){#2^20 = 1,048,576
   N.data <- 2^N.data.exp
   max.segs <- as.integer(N.data/2)
@@ -32,7 +34,7 @@ for(N.data.exp in 2:20){#2^20 = 1,048,576
   data.list <- list(
     flat=rnorm(N.data),
     best=1:N.data,
-    worst=rep(c(0,1),l=N.data))
+    worst=rep(c(0,1,10,11),l=N.data))
   for(case in names(data.list)){
     data.vec <- data.list[[case]]
     cum.data.vec <- cumsum(c(0,data.vec))

### 2022-04-08 figure-timings-laplace.R
uniform colors across figures, stop at 2^18 for meanvar_norm

@@ -1,10 +1,5 @@
 source("packages.R")
 timing.stats <- data.table::fread("figure-timings-laplace-data.csv")
-disp.pkg <- c(
-  changepoint="changepoint.array",
-  ruptures="ruptures.LRU cache",
-  "wbs::sbs"="wbs::sbs.recursion",
-  "fpop::multiBinSeg"="fpop::multiBinSeg.heap")
 timing.stats[, new.pkg := ifelse(
   package %in% names(disp.pkg), disp.pkg[package], package)]
 timing.stats[, Package := sub("[.]", "\n", new.pkg, perl=TRUE)]
@@ -17,6 +12,7 @@ ref.dt <- rbind(
   data.table(seconds=1, unit="1 second"),
   data.table(seconds=60, unit="1 minute"))
 gg <- ggplot()+
+  scale_color_manual(values=pkg.colors)+
   theme_bw()+
   theme(panel.spacing=grid::unit(0, "lines"))+
   facet_grid(. ~ case, labeller=label_both)+
@@ -36,6 +32,8 @@ print(dl)
 dev.off()
 
 gg <- ggplot()+
+  scale_color_manual(values=pkg.colors)+
+  scale_fill_manual(values=pkg.colors)+
   ggtitle("L1 loss")+
   theme_bw()+
   theme(

### 2022-04-08 figure-timings-meanvar_norm-data.R

@@ -23,7 +23,7 @@ binseg_instance <- ruptures$Binseg(min_size=1L, jump=1L, model="normal")
 
 timing.dt.list <- list()
 done.list <- list()
-N.data.exp <- 20
+N.data.exp <- 17
 case <- "best"
 for(N.data.exp in 2:20){#2^20 = 1,048,576
   N.data <- 2^N.data.exp
@@ -53,13 +53,13 @@ for(N.data.exp in 2:20){#2^20 = 1,048,576
         min_block_size = 2L)
       c(bcpd.fit$changepoints, N.data)
     }, "binsegRcpp.multiset"={
-      binseg.fit <- binsegRcpp::binseg(
+      binseg.fit.multiset <- binsegRcpp::binseg(
         "meanvar_norm",data.vec, max.segs, container.str="multiset")
-      sort(binseg.fit$splits$end)
+      sort(binseg.fit.multiset$splits$end)
     }, "binsegRcpp.list"={
-      binseg.fit <- binsegRcpp::binseg(
+      binseg.fit.list <- binsegRcpp::binseg(
         "meanvar_norm",data.vec, max.segs, container.str="list")
-      sort(binseg.fit$splits$end)
+      sort(binseg.fit.list$splits$end)
     }, times=5)
     m.args[names(done.list[[case]])] <- NULL
     if(length(m.args) > 1){
@@ -96,6 +96,14 @@ for(N.data.exp in 2:20){#2^20 = 1,048,576
   }
 }
 
+if(FALSE){
+  binseg1819[["17"]] <- list(
+      list=binseg.fit.list,
+      multiset=binseg.fit.multiset)
+  save(binseg1819,file="binseg1819.RData")
+}
+
 timing.dt <- do.call(rbind, timing.dt.list)
+timing.dt[, computed.N := sapply(computed.ends, length)]
 timing.dt[N.data > 100, `:=`(computed.ends=NA, valid.ends=NA)]
 data.table::fwrite(timing.dt, "figure-timings-meanvar_norm-data.csv")

### 2022-04-08 figure-timings-meanvar_norm.R

@@ -1,10 +1,7 @@
 source("packages.R")
-timing.stats <- data.table::fread("figure-timings-meanvar_norm-data.csv")
-disp.pkg <- c(
-  changepoint="changepoint.array",
-  ruptures="ruptures.LRU cache",
-  "wbs::sbs"="wbs::sbs.recursion",
-  "fpop::multiBinSeg"="fpop::multiBinSeg.heap")
+timing.stats <- data.table::fread(
+"figure-timings-meanvar_norm-data.csv"
+)[N.data<2^19]
 timing.stats[, new.pkg := ifelse(
   package %in% names(disp.pkg), disp.pkg[package], package)]
 timing.stats[, Package := sub("[.]", "\n", new.pkg, perl=TRUE)]
@@ -17,6 +14,7 @@ ref.dt <- rbind(
   data.table(seconds=1, unit="1 second"),
   data.table(seconds=60, unit="1 minute"))
 gg <- ggplot()+
+  scale_color_manual(values=pkg.colors)+
   theme_bw()+
   theme(panel.spacing=grid::unit(0, "lines"))+
   facet_grid(. ~ case, labeller=label_both)+
@@ -36,6 +34,8 @@ print(dl)
 dev.off()
 
 gg <- ggplot()+
+  scale_color_manual(values=pkg.colors)+
+  scale_fill_manual(values=pkg.colors)+
   ggtitle("Normal change in mean and variance")+
   theme_bw()+
   theme(
@@ -70,7 +70,7 @@ gg <- ggplot()+
   coord_cartesian(
     expand=FALSE,
     ylim=c(1e-4, 1e3),
-    xlim=c(8, 3e7))+
+    xlim=c(8, 5e6))+
   scale_y_log10(
     "Computation time (seconds, log scale)\nMedian line and min/max band over 5 timings",
     breaks=10^seq(-10,10))

### 2022-04-08 figure-timings-poisson.R

@@ -1,10 +1,5 @@
 source("packages.R")
 timing.stats <- data.table::fread("figure-timings-poisson-data.csv")
-disp.pkg <- c(
-  changepoint="changepoint.array",
-  ruptures="ruptures.LRU cache",
-  "wbs::sbs"="wbs::sbs.recursion",
-  "fpop::multiBinSeg"="fpop::multiBinSeg.heap")
 timing.stats[, new.pkg := ifelse(
   package %in% names(disp.pkg), disp.pkg[package], package)]
 timing.stats[, Package := sub("[.]", "\n", new.pkg, perl=TRUE)]
@@ -17,6 +12,7 @@ ref.dt <- rbind(
   data.table(seconds=1, unit="1 second"),
   data.table(seconds=60, unit="1 minute"))
 gg <- ggplot()+
+  scale_color_manual(values=pkg.colors)+
   theme_bw()+
   theme(panel.spacing=grid::unit(0, "lines"))+
   facet_grid(. ~ case, labeller=label_both)+
@@ -36,6 +32,8 @@ print(dl)
 dev.off()
 
 gg <- ggplot()+
+  scale_color_manual(values=pkg.colors)+
+  scale_fill_manual(values=pkg.colors)+
   ggtitle("Poisson loss")+
   theme_bw()+
   theme(

### 2022-04-08 figure-timings.R

@@ -1,10 +1,5 @@
 source("packages.R")
 timing.stats <- data.table::fread("figure-timings-data.csv")
-disp.pkg <- c(
-  changepoint="changepoint.array",
-  ruptures="ruptures.LRU cache",
-  "wbs::sbs"="wbs::sbs.recursion",
-  "fpop::multiBinSeg"="fpop::multiBinSeg.heap")
 timing.stats[, new.pkg := ifelse(
   package %in% names(disp.pkg), disp.pkg[package], package)]
 timing.stats[, Package := sub("[.]", "\n", new.pkg, perl=TRUE)]
@@ -17,6 +12,7 @@ ref.dt <- rbind(
   data.table(seconds=1, unit="1 second"),
   data.table(seconds=60, unit="1 minute"))
 gg <- ggplot()+
+  scale_color_manual(values=pkg.colors)+
   theme_bw()+
   theme(panel.spacing=grid::unit(0, "lines"))+
   facet_grid(. ~ case, labeller=label_both)+
@@ -36,6 +32,8 @@ print(dl)
 dev.off()
 
 gg <- ggplot()+
+  scale_color_manual(values=pkg.colors)+
+  scale_fill_manual(values=pkg.colors)+
   theme_bw()+
   theme(
     legend.position="none",

### 2022-04-08 packages.R

@@ -67,6 +67,21 @@ works_with_R(
   xtable="1.8.4",
   fpop="2019.8.26",
   changepoint="2.2.3")
-
-
+dput(RColorBrewer::brewer.pal(Inf,"Set1"))
+c("#E41A1C", "#377EB8", "#984EA3", "#FF7F00", 
+"#A65628", "#F781BF", "#999999")
+pkg.colors <- c(
+  "changepoint\narray"="#E41A1C",
+  "ruptures\nLRU cache"="#377EB8",
+  "blockcpd\nheap"="#984EA3", 
+  "fpop::multiBinSeg\nheap"="#FF7F00", 
+  "wbs::sbs\nrecursion"="#A65628",
+  "binsegRcpp\nmultiset"="black",
+  "binsegRcpp\nlist"="grey50")
+disp.pkg <- c(
+  blockcpd="blockcpd.heap",
+  changepoint="changepoint.array",
+  ruptures="ruptures.LRU cache",
+  "wbs::sbs"="wbs::sbs.recursion",
+  "fpop::multiBinSeg"="fpop::multiBinSeg.heap")
 

### 2022-07-08 article.Rnw
figures

@@
   \centering
-<<visualization, echo=FALSE, fig=TRUE, height=5.2, width=7>>=
-par(mar = c(4, 4, 1, 1))
-plot(table(quine$Days), xlab = "Days", ylab = "Frequency", axes = FALSE)
-axis(2)
-axis(1, at = 0:16 * 5, labels = FALSE)
-axis(1, at = 0:8 * 10)
-box()
-@
-\caption{\label{fig:quine} Frequency distribution for number of days absent
-from school.}
+  \includegraphics[width=\textwidth]{figure-neuroblastoma.png}
+  \caption{\label{fig:neuroblastoma} For a real cancer DNA copy number
+    data set with 273 observations, we show the first four
+    changepoints detected by several different implementations of
+    binary segmentation.}
+\end{figure}
+
+\begin{figure}[t!]
+  \centering
+  \includegraphics[width=\textwidth]{figure-timings.png}
+\caption{\label{fig:timings} Timings using square loss.}
+\end{figure}
+
+\begin{figure}[t!]
+  \centering
+  \includegraphics[width=\textwidth]{figure-timings-laplace.png}
+\caption{\label{fig:timings-laplace} Timings using L1 loss.}
+\end{figure}
+
+\begin{figure}[t!]
+  \centering
+  \includegraphics[width=\textwidth]{figure-timings-meanvar_norm.png}
+\caption{\label{fig:timings-meanvar_norm} Timings using Gaussian change in mean and variance model.}
+\end{figure}
+
+\begin{figure}[t!]
+  \centering
+  \includegraphics[width=\textwidth]{figure-timings-poisson.png}
+\caption{\label{fig:timings-poisson} Timings using Poisson loss.}
+\end{figure}
+
+\begin{figure}[t!]
+  \centering
+  \includegraphics[width=\textwidth]{figure-compare-distributions.png}
+  \caption{\label{fig:compare-distributions} Timings using binsegRcpp multiset with
+    several different loss functions.}
 \end{figure}

### 2022-07-11 article.Rnw
minor

@@
 %% - \Shorttitle{} with LaTeX markup (if any), used as running title
-\title{Comparing binsegRcpp with other implementations of binary segmentation}
-\Plaintitle{A Short Demo Article: Regression Models for Count Data in R}
-\Shorttitle{A Short Demo Article in \proglang{R}}
+\title{Comparing binsegRcpp with other implementations of binary segmentation for changepoint detection}
+\Plaintitle{Comparing binsegRcpp with other implementations of binary segmentation for changepoint detection}
+\Shorttitle{binsegRcpp for changepoint detection}
 
@@
 \Address{
-  Achim Zeileis\\
-  Journal of Statistical Software\\
-  \emph{and}\\
-  Department of Statistics\\
-  Faculty of Economics and Statistics\\
-  Universit\"at Innsbruck\\
-  Universit\"atsstr.~15\\
-  6020 Innsbruck, Austria\\
-  E-mail: \email{Achim.Zeileis@R-project.org}\\
+  Toby Dylan Hocking\\
+  School of Informatics, Computing and Cyber Systems\\
+  Northern Arizona University\\
+  1295 S. Knoles Dr.\\
+  Flagstaff, AZ 86011\\
+  E-mail: \email{Toby.Hocking@R-project.org}\\
   URL: \url{https://www.zeileis.org/}

### 2022-07-12 figure-timings-versions-data.R
timings versions code

@@ -1,0 +1,93 @@
+source("packages.R")
+sha.vec <- c(
+  "before l1loss"="5a150dfbf4f8934e802b6cead6add3c09c111ace",
+  "unordered_map"="bcef04479edcd69377b5aade566f26765da5f6da",
+  "mvl_construct"="5942af606641428315b0e63c7da331c4cd44c091",
+  "var_param"="2a8dbd5346543bffa2faca293e14735c9ee4e9ee",
+  "all l1loss"="71de83db39d14c75d6d85f9d2e133a951d0b0122")
+pkg.path <- "~/R/binsegRcpp"
+pkg.DESC <- file.path(pkg.path, "DESCRIPTION")
+DESC.mat <- read.dcf(pkg.DESC)
+Package <- DESC.mat[,"Package"]
+new.path <- file.path(tempdir(), Package)
+new.DESC <- file.path(new.path, "DESCRIPTION")
+unlink(new.path, recursive=TRUE, force=TRUE)
+system(paste("cp -r", pkg.path, new.path))
+owd <- setwd(new.path)
+lib.path <- file.path(owd, "library")
+for(commit.name in names(sha.vec)){
+  sha <- sha.vec[[commit.name]]
+  new.Package <- sprintf("%s.%s", Package, sha)
+  system("git checkout -- NAMESPACE DESCRIPTION R/*.R src/*.cpp")
+  system(paste("git checkout", sha))
+  DESC.mat[,"Package"] <- new.Package
+  write.dcf(DESC.mat, new.DESC)
+  unlink(file.path(lib.path, new.Package), recursive=TRUE, force=TRUE)
+  system(sprintf("sed -i 's/%s/%s/g' R/*.R NAMESPACE", Package, new.Package))
+  Rcpp::compileAttributes(new.path)
+  system(paste("R CMD INSTALL -l",lib.path,new.path))
+}
+setwd(owd)
+.libPaths("library")
+seconds.limit <- 100
+do.sub <- function(...){
+  mcall <- match.call()
+  L <- as.list(mcall[-1])
+  for(arg.name in names(L)){
+    maybe.lang <- L[[arg.name]]
+    if(is.language(maybe.lang)){
+      L[[arg.name]] <- substitute(
+        result.list[[NAME]] <- EXPR,
+        list(NAME=arg.name, EXPR=maybe.lang))
+    }
+  }
+  L
+}
+
+timing.dt.list <- list()
[…]

### 2022-07-12 figure-timings-versions.R

@@ -1,0 +1,53 @@
+source("packages.R")
+timing.stats <- data.table::fread("figure-timings-versions-data.csv")
+timing.stats[, new.pkg := ifelse(
+  package %in% names(disp.pkg), disp.pkg[package], package)]
+timing.stats[, Package := sub("[.]", "\n", new.pkg, perl=TRUE)]
+(total.minutes <- timing.stats[, sum(seconds_median*seconds_timings) / 60])
+ref.dt <- rbind(
+  data.table(seconds=1, unit="1 second"),
+  data.table(seconds=60, unit="1 minute"))
+gg <- ggplot()+
+  ggtitle("Normal change in mean with constant variance (square loss)")+
+  scale_color_manual(values=pkg.colors)+
+  scale_fill_manual(values=pkg.colors)+
+  theme_bw()+
+  theme(
+    legend.position="none",
+    panel.spacing=grid::unit(0, "lines"))+
+  facet_grid(. ~ case, labeller=label_both)+
+  geom_hline(aes(
+    yintercept=seconds),
+    data=ref.dt,
+    color="grey")+
+  geom_text(aes(
+    10, seconds, label=unit),
+    data=ref.dt,
+    size=3,
+    hjust=0,
+    vjust=1.1,
+    color="grey50")+
+  directlabels::geom_dl(aes(
+    N.data, seconds_median, color=Package, label=Package),
+    data=timing.stats,
+    method=list(cex=0.6, "last.polygons"))+
+  geom_ribbon(aes(
+    N.data, ymin=seconds_min, ymax=seconds_max, fill=Package),
+    alpha=0.5,
+    data=timing.stats)+
+  geom_line(aes(
+    N.data, seconds_median, color=Package),
+    data=timing.stats)+
+  scale_x_log10(
+    "Number of data points to segment (log scale)",
+    breaks=10^seq(1, 6, by=1))+
+  coord_cartesian(
+    expand=FALSE,
+    ylim=c(1e-4, 1e3),
+    xlim=c(8, 9e7))+
+  scale_y_log10(
+    "Computation time (seconds, log scale)\nMedian line and min/max band over 5 timings",
+    breaks=10^seq(-10,10))
+png("figure-timings-versions.png", width=9, height=3.5, units="in", res=200)
+print(gg)
+dev.off()

### 2022-07-13 figure-timings-versions-data.R
more faster new versions

@@ -1,10 +1,22 @@
 source("packages.R")
 sha.vec <- c(
-  "before l1loss"="5a150dfbf4f8934e802b6cead6add3c09c111ace",
-  "unordered_map"="bcef04479edcd69377b5aade566f26765da5f6da",
-  "mvl_construct"="5942af606641428315b0e63c7da331c4cd44c091",
-  "var_param"="2a8dbd5346543bffa2faca293e14735c9ee4e9ee",
-  "all l1loss"="71de83db39d14c75d6d85f9d2e133a951d0b0122")
+  cv="908b77c411bc7f4fcbcf53759245e738ae724c3e",
+  ##"unordered_map"="bcef04479edcd69377b5aade566f26765da5f6da",
+  ##"mvl_construct"="5942af606641428315b0e63c7da331c4cd44c091",
+  ##"new dist meth"="544a4df46ba53ccae8a82a3c67911007cd008000",
+  ##"laplace tests"="61ad58075ae0399d69a8565949ff7aa944f0a461",
+  ## "vloss tests"="53ab142f5b866e80a2d8581f694499f3531c47c1",
+  ## "std::vector"="4732a3af89776461f0398892221c1d09236cda1f",
+  ## "consolidate"="03af3dcec278118808bad5f1bf4dc803f28030e8",
+  ## "simplify con"="2352f6efc0c11fac94e802c960b9ea38a30e0539",
+  ## "var_param"="2a8dbd5346543bffa2faca293e14735c9ee4e9ee",
+  ## "all l1loss"="71de83db39d14c75d6d85f9d2e133a951d0b0122",
+  ## "candidate_ptr arg"="94277b39558fc8c51da2689a738e1b990623bb68",
+  ## "Split pointer arg"="0200e7c55603707b2393bcaf29f1ca64250a3986",
+  ## "ParamsLoss ptr"="b77a6b8f3b74da4e0f3eb1c0b99cf9dfa310ab11",
+  ##"mean var ptr"="26279384340413007c0839b8ba8652b0f6c59f1e",
+  "rm unord map"="dcd0808f52b0b9858352106cc7852e36d7f5b15d",
+  "no var est"="28f2e5aa56a5c7e4147fc65a86a4f4936cecaa03")
 pkg.path <- "~/R/binsegRcpp"
 pkg.DESC <- file.path(pkg.path, "DESCRIPTION")
 DESC.mat <- read.dcf(pkg.DESC)
[…]

### 2022-07-13 figure-timings-versions.R

@@ -9,24 +9,11 @@ ref.dt <- rbind(
   data.table(seconds=60, unit="1 minute"))
 gg <- ggplot()+
   ggtitle("Normal change in mean with constant variance (square loss)")+
-  scale_color_manual(values=pkg.colors)+
-  scale_fill_manual(values=pkg.colors)+
   theme_bw()+
   theme(
     legend.position="none",
     panel.spacing=grid::unit(0, "lines"))+
   facet_grid(. ~ case, labeller=label_both)+
-  geom_hline(aes(
-    yintercept=seconds),
-    data=ref.dt,
-    color="grey")+
-  geom_text(aes(
-    10, seconds, label=unit),
-    data=ref.dt,
-    size=3,
-    hjust=0,
-    vjust=1.1,
-    color="grey50")+
   directlabels::geom_dl(aes(
     N.data, seconds_median, color=Package, label=Package),
     data=timing.stats,
@@ -43,8 +30,7 @@ gg <- ggplot()+
     breaks=10^seq(1, 6, by=1))+
   coord_cartesian(
     expand=FALSE,
-    ylim=c(1e-4, 1e3),
-    xlim=c(8, 9e7))+
+    xlim=c(8, 5e5))+
   scale_y_log10(
     "Computation time (seconds, log scale)\nMedian line and min/max band over 5 timings",
     breaks=10^seq(-10,10))

### 2022-07-14 figure-compare-distributions.R
recompute timings with constant factor faster binsegRcpp

@@ -10,7 +10,12 @@ for(distribution in names(timings.csv.vec)){
   timings.csv <- timings.csv.vec[[distribution]]
   timings.dt.list[[distribution]] <- data.table(
     distribution, 
-    data.table::fread(timings.csv)[grepl("binsegRcpp", package)])
+    data.table::fread(timings.csv)[
+      grepl("binsegRcpp", package),
+      data.table(
+        case, 
+        seconds_min, seconds_max, seconds_median,
+        package, N.data)])
 }
 (timings.dt <- do.call(rbind, timings.dt.list))
 timings.dt[, container := sub("binsegRcpp.", "", package)]

### 2022-07-14 figure-timings-laplace-data.R

@@ -31,7 +31,7 @@ for(N.data.exp in 2:20){#2^20 = 1,048,576
   data.list <- list(
     flat=rnorm(N.data),
     best=1:N.data,
-    worst=rep(c(0,1),l=N.data))
+    worst=2^seq(-N.data, -1))
   for(case in names(data.list)){
     data.vec <- data.list[[case]]
     cum.data.vec <- cumsum(c(0,data.vec))

### 2022-08-08 article.Rnw
orcid
ref line fig

@@
 %% - \AND starts a new line, \And does not.
-\author{Toby Dylan Hocking~\orcidlink{0000-0002-3146-0865}\\Northern Arizona University}
+\author{Toby Dylan Hocking~\orcid{0000-0002-3146-0865}\\Northern Arizona University}
 \Plainauthor{Toby Dylan Hocking}
@@
 \Abstract{ Binary segmentation is a classic algorithm for detecting
- changepoints in sequential data.
-In theory it should be extremely fast for $N$ data and $K$ segments, $O(N K)$ in the worst case, and $O(N \log K)$ in the best case.
-In practice existing implementations such as ruptures (\proglang{Python} module) and changepoint (\proglang{R} package) are much slower, and in fact sometimes do not return a correct result.
-We present the \proglang{R} package \pkg{binsegRcpp}, which provides an efficient C++ implementation of binary segmentation, and always returns correct results.
-We discuss some advanced C++ coding techniques which were used to avoid repetition, for efficiency, and for readability.
-We additionally include detailed theoretical and empirical comparisons with other implementations of binary segmentation in \proglang{R} packages \pkg{fpop}, \pkg{blockcpd} and \pkg{wbs}.
-}
+  changepoints in sequential data. In theory, using a simple loss
+  function like Gaussian or Poisson, binary segmentation should be
+  extremely fast for $N$ data and $K$ segments, $O(N K)$ in the worst
+  case, and $O(N \log K)$ in the best case. In practice existing
+  implementations such as ruptures (\proglang{Python} module) and
+  changepoint (\proglang{R} package) are much slower, and in fact
+  sometimes do not return a correct result. We present the
+  \proglang{R} package \pkg{binsegRcpp}, which provides an efficient
+  C++ implementation of binary segmentation, and always returns
+  correct results. We discuss some advanced C++ coding techniques
+  which were used to avoid repetition, for efficiency, and for
+  readability.  We additionally include detailed theoretical and
+  empirical comparisons with other implementations of binary
+  segmentation in \proglang{R} packages \pkg{fpop}, \pkg{blockcpd} and
+  \pkg{wbs}. }
 

### 2022-08-08 figure-compare-ruptures.R

@@ -1,0 +1,131 @@
+source("packages.R")
+references_funs <- list(
+  "N log N"=function(N)log10(N*log(N)),
+  "N^2"=function(N)2*log10(N),
+  "N^3"=function(N)3*log10(N))
+references <- function
+(N, empirical, lower.limit=min(empirical),
+  fun.list=NULL
+){
+  fun.latex <- reference <- . <- NULL
+  if(is.null(fun.list))fun.list <- references_funs
+  data.table(fun.latex=names(fun.list))[, {
+    fun <- fun.list[[fun.latex]]
+    log10.vec <- fun(N)
+    one.fun <- data.table(
+      N, empirical,
+      reference=10^(log10.vec-max(log10.vec)+max(log10(empirical)))
+    )
+    above <- one.fun[lower.limit < reference]
+    if(1 < nrow(above)){
+      above
+    }else{
+      one.fun[(.N-1):.N]
+    }
+  }, by=.(fun.latex, fun.name=gsub("\\", "", fun.latex, fixed=TRUE))]
+}
+references_best <- function(DT, fun.list=NULL){
+  unit.col.vec <- c(seconds="seconds_median")
+  ref.dt.list <- list()
+  metric.dt.list <- list()
+  for(unit in names(unit.col.vec)){
+    col.name <- unit.col.vec[[unit]]
+    values <- DT[[col.name]]
+    only.positive <- values[0 < values]
+    if(length(only.positive)){
+      lower.limit <- min(only.positive)
+      all.refs <- DT[
+      , references(N.data, .SD[[col.name]], lower.limit, fun.list)
+      , by=.(Package, case)]
+      all.refs[, rank := rank(N), by=.(Package, case, fun.name)]
+      second <- all.refs[rank==2]
+      second[, dist := log10(empirical/reference) ]
+      second[, sign := sign(dist)]
+      l.cols <- list(
+        overall=c("Package","case"), 
+        each.sign=c("Package","case","sign"))
+      for(best.type in names(l.cols)){
+        by <- l.cols[[best.type]]
+        second[
+        , paste0(best.type,".rank") := rank(abs(dist))
+        , by=by]
+      }
+      ref.dt.list[[unit]] <- data.table(unit, all.refs[
+        second,
+        on=.(Package, case, fun.name, fun.latex)])
+      best <- second[overall.rank==1, .(Package, case, fun.name, fun.latex)]
[…]

### 2022-08-08 figure-compare-ruptures.R
show binsegRcpp

@@ -76,7 +76,9 @@ ref.dt <- rbind(
   data.table(seconds=1, unit="1 second"),
   data.table(seconds=60, unit="1 minute"))
 ruptures <- timing.stats[
-  package %in% c("ruptures.cumsum","ruptures.l2","changepoint")]
+  package %in% c(
+    "ruptures.cumsum","ruptures.l2","changepoint",
+    "binsegRcpp.list","binsegRcpp.multiset")]
 rbest <- references_best(ruptures)
 rank1 <- rbest$references[reference > 1e-3]
 gg <- ggplot()+
@@ -126,6 +128,6 @@ gg <- ggplot()+
     method="bottom.polygons",
     data=rank1)+
   coord_cartesian(ylim=c(1e-4, 1e3))
-png("figure-compare-ruptures.png", width=9, height=9, units="in", res=200)
+png("figure-compare-ruptures.png", width=12, height=8, units="in", res=200)
 print(gg)
 dev.off()

### 2025-04-28 article.Rnw
cite fastcpd
C++ coding

@@
 \documentclass[article]{jss}
-
+\usepackage[utf8]{inputenc}
 %% -- LaTeX packages and custom commands ---------------------------------------
@@
   C++ implementation of binary segmentation, and always returns
-  correct results. We discuss some advanced C++ coding techniques
+  correct results. We discuss some C++ coding techniques
   which were used to avoid repetition, for efficiency, and for
@@
   Toby Dylan Hocking\\
-  School of Informatics, Computing and Cyber Systems\\
-  Northern Arizona University\\
-  1295 S. Knoles Dr.\\
-  Flagstaff, AZ 86011\\
-  E-mail: \email{Toby.Hocking@R-project.org}\\
-  URL: \url{https://www.zeileis.org/}
+  Université de Sherbrooke\\
+  E-mail: \email{Toby.Hocking@R-project.org}
 }
@@
 
-%% -- Introduction -------------------------------------------------------------
-
-%% - In principle "as usual".
-%% - But should typically have some discussion of both _software_ and _methods_.
-%% - Use \proglang{}, \pkg{}, and \code{} markup throughout the manuscript.
-%% - If such markup is in (sub)section titles, a plain text version has to be
-%%   added as well.
-%% - All software mentioned should be properly \cite-d.
-%% - All abbreviations should be introduced.
-%% - Unless the expansions of abbreviations are proper names (like "Journal
-%%   of Statistical Software" above) they should be in sentence case (like
-%%   "generalized linear models" below).
-
-\section[Introduction: Count data regression in R]{Introduction: Count data regression in \proglang{R}} \label{sec:intro}
-
-\begin{leftbar}
-
-The introduction is in principle ``as usual''. However, it should usually embed
-both the implemented \emph{methods} and the \emph{software} into the respective
-relevant literature. For the latter both competing and complementary software
-should be discussed (within the same software environment and beyond), bringing
-out relative (dis)advantages. All software mentioned should be properly
-\verb|\cite{}|d. (See also Appendix~\ref{app:bibtex} for more details on
-\textsc{Bib}{\TeX}.)
-
-For writing about software JSS requires authors to use the markup
-\verb|\proglang{}| (programming languages and large programmable systems),
-\verb|\pkg{}| (software packages), \verb|\code{}| (functions, commands,
-arguments, etc.). If there is such markup in (sub)section titles (as above), a
-plain text version has to be provided in the {\LaTeX} command as well. Below we
-also illustrate how abbrevations should be introduced and citation commands can
-be employed. See the {\LaTeX} code for more details.
-\end{leftbar}
-
-TODO (Standard Template Library containers, virtual classes, static
-variables, function pointers, templates, macros).
+\section{Introduction: previous software for change-point detection} \label{sec:intro}
 
@@
 
-%% -- Manuscript ---------------------------------------------------------------
-
-%% - In principle "as usual" again.
-%% - When using equations (e.g., {equation}, {eqnarray}, {align}, etc.
-%%   avoid empty lines before and after the equation (which would signal a new
-%%   paragraph.
-%% - When describing longer chunks of code that are _not_ meant for execution
-%%   (e.g., a function synopsis or list of arguments), the environment {Code}
-%%   is recommended. Alternatively, a plain {verbatim} can also be used.
-%%   (For executed code see the next section.)
+TODO \code{fastcpd} package \citep{fastcpd2024} (nice related work table).
 
@@
 
-The basic Poisson regression model for count data is a special case of the GLM
-framework \cite{McCullagh+Nelder:1989}. It describes the dependence of a count
-response variable $y_i$ ($i = 1, \dots, n$) by assuming a Poisson distribution
-$y_i \sim \mathrm{Pois}(\mu_i)$. The dependence of the conditional mean
-$\E[y_i \, | \, x_i] = \mu_i$ on the regressors $x_i$ is then specified via a
-log link and a linear predictor
+\subsection{C++ features}
+
+We used several features of C++ to implement \pkg{binsegRcpp} efficiently.
+In binary segmentation, a segment is split in each iteration, which results in up to two new segments that must be considered to split (see Section~\ref{sec:algo} for details).
+Computing the cost of splitting a segment of size $N$ is $O(N\log N)$ time for L1/Laplace losses, and $O(N)$ time for the normal/Poisson losses.
+After computing the cost for two new segments, the cost of all splittable segments must be considered, in order to identify the min cost segment to split next.
+We therefore propose using two different kinds of C++ Standard Template Library containers to store and retreive the cost of segments which have previously been considered, but not yet split (Table~\ref{tab:containers}).
+First, we propose using the \code{multiset} container (red-black tree), which can be used to achieve best case time complexity.
+The \code{multiset} keeps segments sorted by their cost values, so offers constant $O(1)$ time retreival of the min cost segment to split next; insertion takes logarithmic $O(\log n)$ time in the number of segments $n$ currently stored in the container.
+Second, we propose using the \code{list} container (doubly-linked list), which is used as a baseline, to demonstrate the benefits of the multiset container.
+The \code{list} keeps segments stored in the order in which they were inserted, so offers constant $O(1)$ time insertion of new elements, and linear $O(n)$ time retreival of the min cost segment to split, for a container with $n$ segments currently stored.
+
+A C++ virtual class \code{Container} is used to represent the methods that must be provided by either type of container: \code{insert}, \code{get_best}.
+We used a C macro, \code{CMAKER}, to declare the two sub-classes:
+\begin{verbatim}
+#define CMAKER(CONTAINER, INSERT, GET_BEST)...
+CMAKER(multiset, insert, segment_container.begin())
+CMAKER(list, push_back, std::min_element(
+ segment_container.begin(),segment_container.end()))
+\end{verbatim}
+The code above is expanded by the C pre-processor, to two declarations of \code{Container} sub-classes.
+
+Another C macro is used to declare the loss functions which require the cumulative median:
+\begin{verbatim}
+#define ABS_DIST(NAME, VARIANCE)...
+ABS_DIST(l1, false)
+ABS_DIST(laplace, true)
+\end{verbatim}
+The loss function which requires a variance estimate is the Laplace negative log likelihood; the simple L1 loss is obtained without a variance estimate.
+A \code{class absDistribution} implements the common operations for the two loss functions, including the cumulative median in $O(N\log N)$ time using the algorithm of \citet{Drouin2017mmit}.
+
+Another C macro is used to declare the loss functions which require the cumulative sum:
+\begin{verbatim}
+#define CUM_DIST(NAME, COMPUTE, ERROR, VARIANCE)...
+CUM_DIST(mean_norm, RSS, , false)
+CUM_DIST(poisson,
+ (mean>0) ? (mean*N - log(mean)*sum) : ( (sum==0) ? 0 : INFINITY ),
+ if(round(value)!=value)return ERROR_DATA_MUST_BE_INTEGER_FOR_POISSON_LOSS;
+ if(value < 0)return ERROR_DATA_MUST_BE_NON_NEGATIVE_FOR_POISSON_LOSS;,
+ false)
+CUM_DIST(meanvar_norm,
+ (var>max_zero_var) ? (RSS/var+N*log(2*M_PI*var))/2 : INFINITY,
+ , true)
+\end{verbatim}
+The code below declares three loss functions based on the cumulative sum: \code{mean_norm} is the residual sum of squares (RSS), \code{poisson} loss is for count data, and \code{meanvar_norm} is the normal model with change in mean and variance.
+The \code{COMPUTE} parmater specifies the loss, in terms of variables such as \code{mean} and \code{sum} which are computed in a \code{class CumDistribution} which contains the logic common to these loss functions.
+
+Additionally, the \code{CUM_DIST} and \code{ABS_DIST} macros generate code that populates a static \code{unordered_map} which can be queried to obtain a list of distributions which are supported by the C++ code:
 %
-\begin{equation} \label{eq:mean}
-\log(\mu_i) \quad = \quad x_i^\top \beta,
-\end{equation}
+<<>>=
+binsegRcpp::get_distribution_info()$dist
+@ 
 %
-where the regression coefficients $\beta$ are estimated by maximum likelihood
-(ML) using the iterative weighted least squares (IWLS) algorithm.
-
-\begin{leftbar}
-Note that around the \verb|{equation}| above there should be no spaces (avoided
-in the {\LaTeX} code by \verb|%| lines) so that ``normal'' spacing is used and
-not a new paragraph started.
-\end{leftbar}
-
-\proglang{R} provides a very flexible implementation of the general GLM
-framework in the function \fct{glm} \citep{Chambers+Hastie:1992} in the
-\pkg{stats} package. Its most important arguments are
-\begin{Code}
-glm(formula, data, subset, na.action, weights, offset,
-  family = gaussian, start = NULL, control = glm.control(...),
-  model = TRUE, y = TRUE, x = FALSE, ...)
-\end{Code}
-where \code{formula} plus \code{data} is the now standard way of specifying
-regression relationships in \proglang{R}/\proglang{S} introduced in
-\cite{Chambers+Hastie:1992}. The remaining arguments in the first line
-(\code{subset}, \code{na.action}, \code{weights}, and \code{offset}) are also
-standard  for setting up formula-based regression models in
-\proglang{R}/\proglang{S}. The arguments in the second line control aspects
-specific to GLMs while the arguments in the last line specify which components
-are returned in the fitted model object (of class \class{glm} which inherits
-from \class{lm}). For further arguments to \fct{glm} (including alternative
-specifications of starting values) see \code{?glm}. For estimating a Poisson
-model \code{family = poisson} has to be specified.
-
-\begin{leftbar}
-As the synopsis above is a code listing that is not meant to be executed,
-one can use either the dedicated \verb|{Code}| environment or a simple
-\verb|{verbatim}| environment for this. Again, spaces before and after should be
-avoided.
-
-Finally, there might be a reference to a \verb|{table}| such as
-Table~\ref{tab:overview}. Usually, these are placed at the top of the page
-(\verb|[t!]|), centered (\verb|\centering|), with a caption below the table,
-column headers and captions in sentence style, and if possible avoiding vertical
-lines.
-\end{leftbar}
+To illustrate the performence benefits of using log-time  containers  were used to efficiently , virtual classes, static
+variables, function pointers, templates, macros).
 
@@
 and \code{CUSUM} and \code{min.th} columns
-TODO.\code{fpop::multiBinSeg} result is a list with element
+TODO.
+\code{fpop::multiBinSeg} result is a list with element
 \code{J.est}, which is a vector of \code{max.changes} square loss
@@
     heap/multiset & $O(\log n)$ &
-    % $O(\log n)$ &
+    % $O(\log N)$ &
                     $O(1)$ &
@@
 
-%% -- Illustrations ------------------------------------------------------------
-
-%% - Virtually all JSS manuscripts list source code along with the generated
-%%   output. The style files provide dedicated environments for this.
-%% - In R, the environments {Sinput} and {Soutput} - as produced by Sweave() or
-%%   or knitr using the render_sweave() hook - are used (without the need to
-%%   load Sweave.sty).
-%% - Equivalently, {CodeInput} and {CodeOutput} can be used.
-%% - The code input should use "the usual" command prompt in the respective
-%%   software system.
-%% - For R code, the prompt "R> " should be used with "+  " as the
-%%   continuation prompt.
-%% - Comments within the code chunks should be avoided - these should be made
-%%   within the regular LaTeX text.
-
 \section{Illustrations} \label{sec:illustrations}
 
-For a simple illustration of basic Poisson and NB count regression the
-\code{quine} data from the \pkg{MASS} package is used. This provides the number
-of \code{Days} that children were absent from school in Australia in a
-particular year, along with several covariates that can be employed as regressors.
-The data can be loaded by
+TODO
 %
@@
 %
-and a basic frequency distribution of the response variable is displayed in
-Figure~\ref{fig:quine}.
-
-\begin{leftbar}
-For code input and output, the style files provide dedicated environments.
-Either the ``agnostic'' \verb|{CodeInput}| and \verb|{CodeOutput}| can be used
-or, equivalently, the environments \verb|{Sinput}| and \verb|{Soutput}| as
-produced by \fct{Sweave} or \pkg{knitr} when using the \code{render_sweave()}
-hook. Please make sure that all code is properly spaced, e.g., using
-\code{y = a + b * x} and \emph{not} \code{y=a+b*x}. Moreover, code input should
-use ``the usual'' command prompt in the respective software system. For
-\proglang{R} code, the prompt \code{"R> "} should be used with \code{"+  "} as
-the continuation prompt. Generally, comments within the code chunks should be
-avoided -- and made in the regular {\LaTeX} text instead. Finally, empty lines
-before and after code input/output should be avoided (see above).
-\end{leftbar}
+TODO
 
@@
 
-As a first model for the \code{quine} data, we fit the basic Poisson regression
-model. (Note that JSS prefers when the second line of code is indented by two
-spaces.)
-%
-<<poisson>>=
-m_pois <- glm(Days ~ (Eth + Sex + Age + Lrn)^2, data = quine,
-  family = poisson)
-@
-%
-To account for potential overdispersion we also consider a negative binomial
-GLM.
-%
-<<negbin>>=
-library("MASS")
-m_nbin <- glm.nb(Days ~ (Eth + Sex + Age + Lrn)^2, data = quine)
-@
-%
-In a comparison with the BIC the latter model is clearly preferred.
-%
-<<comparison>>=
-BIC(m_pois, m_nbin)
-@
-%
-Hence, the full summary of that model is shown below.
-%
-<<summary>>=
-summary(m_nbin)
-@
-
-%% -- Summary/conclusions/discussion -------------------------------------------
-
 \section{Summary and discussion} \label{sec:summary}
@@
 
-%% -- Optional special unnumbered sections -------------------------------------
-
-\section*{Computational details}
-
-\begin{leftbar}
-If necessary or useful, information about certain computational details
-such as version numbers, operating systems, or compilers could be included
-in an unnumbered section. Also, auxiliary packages (say, for visualizations,
-maps, tables, \dots) that are not cited in the main text can be credited here.
-\end{leftbar}
-
-The results in this paper were obtained using
-\proglang{R}~\Sexpr{paste(R.Version()[6:7], collapse = ".")} with the
-\pkg{MASS}~\Sexpr{packageVersion("MASS")} package. \proglang{R} itself
-and all packages used are available from the Comprehensive
-\proglang{R} Archive Network (CRAN) at
-\url{https://CRAN.R-project.org/}.
-
-\section*{Acknowledgments}
-
-\begin{leftbar}
-All acknowledgments (note the AE spelling) should be collected in this
-unnumbered section before the references. It may contain the usual information
-about funding and feedback from colleagues/reviewers/etc. Furthermore,
-information such as relative contributions of the authors may be added here
-(if any).
-\end{leftbar}
-
-%% -- Bibliography -------------------------------------------------------------
-%% - References need to be provided in a .bib BibTeX database.
-%% - All references should be made with \cite, \citet, \citep, \citealp etc.
-%%   (and never hard-coded). See the FAQ for details.
-%% - JSS-specific markup (\proglang, \pkg, \code) should be used in the .bib.
-%% - Titles in the .bib should be in title case.
-%% - DOIs should be included where available.
-
 \bibliography{refs}
-
-%% -- Appendix (if any) --------------------------------------------------------
-%% - After the bibliography with page break.
-%% - With proper section titles and _not_ just "Appendix".
 
@@
   \section{More technical details} \label{app:technical}
-
-\begin{leftbar}
-Appendices can be included after the bibliography (with a page break). Each
-section within the appendix should have a proper section title (rather than
-just \emph{Appendix}).
-
-For more technical style details, please check out JSS's style FAQ at
-\url{https://www.jstatsoft.org/pages/view/style#frequently-asked-questions}
-which includes the following topics:
-\begin{itemize}
-  \item Title vs.\ sentence case.
-  \item Graphics formatting.
-  \item Naming conventions.
-  \item Turning JSS manuscripts into \proglang{R} package vignettes.
-  \item Trouble shooting.
-  \item Many other potentially helpful details\dots
-\end{itemize}
-\end{leftbar}
-
-\section[Using BibTeX]{Using \textsc{Bib}{\TeX}} \label{app:bibtex}
-
-\begin{leftbar}
-References need to be provided in a \textsc{Bib}{\TeX} file (\code{.bib}). All
-references should be made with \verb|\cite|, \verb|\citet|, \verb|\citep|,
-\verb|\citealp| etc.\ (and never hard-coded). This commands yield different
-formats of author-year citations and allow to include additional details (e.g.,
-pages, chapters, \dots) in brackets. In case you are not familiar with these
-commands see the JSS style FAQ for details.
-
-Cleaning up \textsc{Bib}{\TeX} files is a somewhat tedious task -- especially
-when acquiring the entries automatically from mixed online sources. However,
-it is important that informations are complete and presented in a consistent
-style to avoid confusions. JSS requires the following format.
-\begin{itemize}
-  \item JSS-specific markup (\verb|\proglang|, \verb|\pkg|, \verb|\code|) should
-    be used in the references.
-  \item Titles should be in title case.
-  \item Journal titles should not be abbreviated and in title case.
-  \item DOIs should be included where available.
-  \item Software should be properly cited as well. For \proglang{R} packages
-    \code{citation("pkgname")} typically provides a good starting point.
-\end{itemize}
-\end{leftbar}
+  TODO
 
@@
 
-%% -----------------------------------------------------------------------------
-
 \end{document}

### 2025-04-29 article.Rnw
begin tables

@@
 The code below declares three loss functions based on the cumulative sum: \code{mean_norm} is the residual sum of squares (RSS), \code{poisson} loss is for count data, and \code{meanvar_norm} is the normal model with change in mean and variance.
-The \code{COMPUTE} parmater specifies the loss, in terms of variables such as \code{mean} and \code{sum} which are computed in a \code{class CumDistribution} which contains the logic common to these loss functions.
+The \code{COMPUTE} parameter specifies the loss, in terms of six variables: \code{mean}, \code{var}, \code{N}, \code{sum}, \code{squares}, \code{max_zero_var} (Table~\ref{tab:cum-params}).
+These variables are computed in a \code{class CumDistribution} which contains the logic common to these loss functions (cumulative sum, etc).
+Note that \code{mean} and \code{var} parameters  code can compute the loss of held-out validation data (not used to compute model parameters), with respect to a subset TODO.
+The \code{ERROR} parameter is an optional block of code, that is used to return an error status code, if there are unusual input data.
+
+We assume there is a sequence of $N$ data $x_1,\dots,x_N\in\mathbb R$,
+with corresponding weights $w_1,\dots,w_N>0$. Binary segmentation can
+be interpreted as attempting to
+
+To define the loss functions based on the cumulative sum
+
+\begin{table}
+  \begin{tabular}{ccc}
+    Parameter & Description & Definition \\
+    \hline
+    $M=$\code{mean} & model parameter & $T/W$ \\
+    \code{var} & model parameter &  $S/W + M(M-2T/W)$
+
+  \end{tabular}
+\end{table}
 

### 2025-05-02 article.Rnw
binsegRcpp storage vs wbs and changepoint

@@
 %% -- LaTeX packages and custom commands ---------------------------------------
-
+\usepackage{amsmath}
+\DeclareMathOperator*{\argmin}{arg\, min}
 %% recommended packages
@@
 <<preliminaries, echo=FALSE, results=hide>>=
-options(prompt = "R> ", continue = "+  ", width = 70, useFancyQuotes = FALSE)
+options(prompt = "R> ", continue = "+  ", width = 80, useFancyQuotes = FALSE)
 library("MASS")
@@
 
+We assume there is a sequence of $N$ data $x_1,\dots,x_N\in\mathbb R$, in which we want to find abrupt changes.
+We assume that each data point $i\in\{1,\dots,N\}$ has a corresponding weight $w_i>0$.
+Binary
+segmentation can be interpreted as attempting to find model parameters
+$\theta$ which minimize a given weighted loss function $w_i
+\ell(\theta, x_i)$. For example, minimizing the square loss $w_i
+\ell(\theta, x_i) = w_i (\theta-x_i)^2$ corresponds to maximizing the
+normal log-likelihood, when the variance is uniform.
+
+\paragraph{Choice of observation-specific weights.}
+Weights can be used to efficiently handle two particular kinds of data.
+First, in the case of data which are sampled uniformly across
+space/time, typically the weights are constant ($w_i=1$ for all $i$),
+but other weights can be used for data which are sampled non-uniformly
+(for example daily time series with some missing days).
+Second, in the case of data with runs of identical values, a run-length encoding can be used to convert the raw/unweighted data sequence into a compressed/weighted data sequence that can be more efficiently processed.
+For example, in genomics, DNA sequence reads are aligned to a reference genome, and a coverage profile is computed by counting the number of aligned reads at each reference position.
+Since each DNA sequence read is about 100 bases long, the coverage profile has runs of identical values, which can be converted to a run-length encoding for $\approx 100\times$ speedups.
+
+\paragraph{Min loss.}
+The iterative algorithm involves computing the min loss over
+candidate segments, each of which is defined by a start $j$ and end
+$e$ position ($1<j<e<N$).
+
+\begin{equation}
+L_{j,e} = \min_{\theta} \sum_{i=j}^e w_i \ell(\theta, x_i).
+\end{equation}
+The minimization problem is over the model parameter $\theta$ (segment-specific center and possibly scale).
+
+\paragraph{Best loss difference.}
+Since different segments have different loss values before and after splitting, the algorithm must choose the next segment to split by minimizing a loss difference value.
+Let $D_{j,e}(t)=L_{j,t}+L_{t+1,n}-L_{j,e}<0$ be the loss difference after splitting segment $(j,e)$ after $t$.
+The $L_{j,e}$ term is the loss before the split, and $L_{j,t}+L_{t+1,n}$ is the loss after the split, which is necessarily smaller.
+Let $\delta_{j,e},\tau_{j,e}=\min,\argmin_{t\in \{j,\dots,e-1\}} D_{j,e}(t)$ be the best loss difference/split on segment $(j,e)$.
+%% We can re-write this minimization as
+%% \begin{equation}
+%%   f_{j,e}=\min,\argmin -L_{j,e}+ \begin{cases}
+%%     D_{j,e}(j) = L_{j,j}+L_{j+1,e} \\
+%%     %D_{j,e}(j+1) = L_{j,j+1}+L_{j+2,e} \\
+%%     \vdots\\
+%%     D_{j,e}(e-1) = L_{j,e-1}+L_{e,e} \\
+%%   \end{cases}
+%%   \label{eq:fje}
+%% \end{equation}
+\paragraph{Efficient computation.}
+To compute the loss differences $D_{j,e}(j),\dots,D_{j,e}(e-1)$, we need to compute the sequence of values $L_{j,j}, \dots, L_{j,e-1}$ (loss values for possible new segments before candidate split points), and the sequence of values $L_{j+1,e}, \dots, L_{e,e}$ (loss values for possible new segments after candidate split points).
+These sequences can be computed efficiently for the five loss functions that we implemented in \code{binsegRcpp}.
+The time complexity of computing these sequences depends on the loss function, and the size of the segment to split, $s=e-j+1$.
+For the case of negative log likelihood of Normal and Poisson distributions, the cumulative sum trick can be used to compute these sequences in linear $O(s)$ time.
+For the case of the L1 loss and negative log likelihood of Laplace distribution, these sequences can be computed efficiently in log-linear $O(s\log s)$ time using the cumulative median algorithm of \citet{Drouin2017mmit}.
+
+\subsection{Algorithm pseudo-code}
+
+The binary segmentation algorithm needs to keep track of a set of segments that can be split, $\mathcal S_k$ for iteration $k$.
+Each segment $(j,e,d,t)\in\mathcal S_k$ is represented by the start/end positions $(j,e)$ along with the corresponding min loss difference $d<0$, as well as the best split point $t\in\{j,\dots,e-1\}$.
+
+\paragraph{Initialization.}
+The initialization of the algorithm is as follows.
+Let $\mathcal L_1=L_{1,n}$ be the loss with one segment,
+and let $\delta_1,t_1= f_{1,n}$ be the best loss difference and split point for the entire data.
+We initialize the set of segments to split as $\mathcal S_1=\{(1,N,d_1,t_1)\}$.
+
+\paragraph{Recursive update rules.}
+Let $K\in\{ 2, \dots, N\}$ be the user-specified max number of segments to consider.
+The binary segmentation algorithm recursively computes the following quantities for each $k\in\{2,\dots,K\}$.
+First, the best segment to split is computed by greedy minimization of the loss decrease over all splittable segments.
+\begin{equation}
+  j^*_k, e^*_k, d^*_k, t^*_k = \argmin_{(j,e,d,t)\in\mathcal S_{k-1}} d.
+\end{equation}
+Ideally, the minimization above takes $O(1)$ constant time, if the set of splittable segments is a red-black tree (C++ STL multiset).
+As a baseline, we also consider a linked list, which results in a sub-optimal $O(k)$ linear time search (see Section~\ref{sec:data-structures} for details).
+
+Second, the total loss is computed by adding the best loss decrease $d^*_k$ to the best loss $L_{k-1}$ at the previous iteration, 
+\begin{equation}
+  \mathcal L_k=\mathcal L_{k-1} +d^*_k.
+\end{equation}
+
+Third, we construct a set of two new segments to consider:
+\begin{equation}
+  \mathcal N_k=\{
+  (j^*_k, t^*_k),
+  (t^*_k+1,e^*_k)
+  \}.
+\end{equation}
+
+Fourth, we subset the two new segments, to consider only the segments that could possibly be split:
+\begin{equation}
+  \label{eq:Vk}
+  \mathcal V_k=\{(j,e,f_{j,e})\mid (j,e)\in\mathcal N_k,\, j<e\}.
+\end{equation}
+
+Finally, we update the set of splittable segments, by removing the currently split segment, and inserting the new segments that could be split:
+\begin{equation}
+  \mathcal S_k=
+[\mathcal S_{k-1}\setminus (j^*_k, e^*_k, d^*_k, t^*_k)]
+\cup\mathcal V_k.
+\end{equation}
+TODO time.
+
+\subsection{Modification for min segment length parameter}
+
+Sometimes there is prior knowledge that there should be no segments
+with fewer data points than $m\in\{1,2,\dots\}$. In that case, $m$ is
+referred to as a min segment length parameter, and there are two simple
+modifications to the equations above.
+First, the best split (\ref{eq:fje}) is re-defined as
+\begin{equation}
+f^m_{j,e}=\min,\argmin_{t\in\{
+j+m-1,\dots,e-m
+\}} D_{j,e}(t).
+\end{equation}
+Second, the set of splittable segments (\ref{eq:Vk}) is re-defined as
+\begin{equation}
+  \mathcal V^m_k=\{(j,e,f_{j,e})\mid (j,e)\in\mathcal N_k,\, 
+  e-j+1\geq 2m
+  \}.
+\end{equation}
+
 \subsection{C++ features}
@@
 \end{verbatim}
-The code below declares three loss functions based on the cumulative sum: \code{mean_norm} is the residual sum of squares (RSS), \code{poisson} loss is for count data, and \code{meanvar_norm} is the normal model with change in mean and variance.
+The code above declares three loss functions based on the cumulative sum: \code{mean_norm} is the residual sum of squares (RSS), \code{poisson} loss is for count data, and \code{meanvar_norm} is the normal model with change in mean and variance.
 The \code{COMPUTE} parameter specifies the loss, in terms of six variables: \code{mean}, \code{var}, \code{N}, \code{sum}, \code{squares}, \code{max_zero_var} (Table~\ref{tab:cum-params}).
 These variables are computed in a \code{class CumDistribution} which contains the logic common to these loss functions (cumulative sum, etc).
-Note that \code{mean} and \code{var} parameters  code can compute the loss of held-out validation data (not used to compute model parameters), with respect to a subset TODO.
+Note that the code supports cross-validation, in the sense that the input train set can be divided into a subtrain set (used to compute model parameters) and a validation set (used to evaluate the computed model parameters).
+The \code{mean} and \code{var} variables are model parameters, computed using the subtrain set, whereas the other variables depend on the data (subtrain or validation set).
 The \code{ERROR} parameter is an optional block of code, that is used to return an error status code, if there are unusual input data.
 
-We assume there is a sequence of $N$ data $x_1,\dots,x_N\in\mathbb R$,
-with corresponding weights $w_1,\dots,w_N>0$. Binary segmentation can
-be interpreted as attempting to
-
-To define the loss functions based on the cumulative sum
-
+To define the loss functions based on the cumulative sum TODO
 \begin{table}
@@
 % "2022.3.29"     "2.2.3"       "1.4" "2019.8.26"
-  version & 2022.3.29 & 2.2.3 & 1.4 &2019.8.26 & 1.1.6\\
+% https://github.com/HaotianXu/changepoints does not have binary segmentation.
+  version & 2022.3.29 & 2.2.3 & 1.4 & 2019.8.26 & 1.1.6\\
 \hline
@@
   correct & yes & no & yes & yes & no \\
-  losses & C++ & C & L2 & L2 & Python \\
+  losses & 5 & 6 & 1 & 1 & 10 \\
+  language & C++ & C & C & C++ & Python \\
   % all propose some loss computation.
@@
 \end{tabular}
-\caption{\label{tab:overview} TODO}
+\caption{\label{tab:overview}
+  Properties (rows) of five binary segmentation libraries (columns).
+Properties are weights (can specify observation-specific positive weights for the loss), max segs (can specify maximum number of segments parameter), dim (dimensions of input data), correct (does code return correct model parameters?), losses (number of loss functions implemented), language (programming language to implement loss), storage (technique used to store cost of splitting a segment), space (space complexity in number of segments $S$), cumsum (cumulative sum used), best (time complexity in best case data), worst (time complexity in worst case data), CV (support for cross-validation), params (parameters are returned).
+}
 \end{table}
@@
 \end{table}
+
+\section{Storage and retreival of segment-specific model parameters}
+
+To simplify the discussion in this section, we use ``segment mean'' to represent arbitrary segment-specific model parameters.
+For example, segment-specific model parameters can be the mean/variance of the normal distribution, or median/scale of the Laplace distribution.
+A unique feature that is provided by \pkg{binsegRcpp}, but not provided by other implementations of binary segmentation, is efficient storage and retrieval of arbitrary segment mean parameters.
+In contrast, \code{changepoint::cpt.mean} returns only the segment means for the selected model, and other implementations do not return any segment means (but the returned change-point variables could be used to compute estimates).
+
+%data_vec <- c(5, 0, 20, 15, -10, 0)
+For example, we consider a simple example with six data points.
+<<>>=
+ex6_df <- data.frame(value = c(1, -7, 8, 10, 2, 4))
+ex6_binsegRcpp <- binsegRcpp::binseg("mean_norm", ex6_df$value, max.segments = 4)
+ex6_binsegRcpp$splits[, .(segments, end, before.mean, after.mean,
+  invalidates.index, invalidates.after)]
+@
+The output above is a table with four rows, one for each model size.
+For a max number of segments $M$, this representation of model parameters uses $M$ rows, so is asymptotically linear, $O(M)$.
+The first row represents the trivial model: one segment across all data points, with \code{end=6} equal to the index of the last data point.
+The model parameter of the trivial model is \code{before.mean=3}, which is the mean of all of the data.
+The other rows represent the segments that result from the recursive splitting algorithm.
+For example, in the second row, \code{end=2} means that the first change-point results in a split between data points 2 and 3.
+The \code{before.mean = -3} and \code{after.mean=6} values are the means of the resulting segments, before and after that split.
+The other columns, \code{invalidates.index} and \code{invalidates.after}, indicate the previous means which are no longer valid, after the given split.
+For example, in the second row, \code{invalidates.index=1} identifies the first row, and \code{invalidates.after=0} identifies the mean before the split (not after), which implies that \code{before.mean=3} is no longer valid, after performing the split in the second row.
+In the third row, we see a new split at \code{end=4}, resulting in new parameters \code{before.mean=9} and \code{after.mean=3}, which invalidate the previous mean in row 2 after the split (\code{after.mean=6}).
+
+In general, for any number of segments $K$, we can compute a table of
+estimated means (one row for each segment), using the following
+algorithm.  First, consider only the first $K$ rows of the
+\code{splits} table.  Next, set values in \code{before.mean} and
+\code{after.mean} columns to missing (\code{NA}), based on values in
+\code{invalidates.index} and \code{invalidates.after} columns.
+Finally, sort the \code{end} values, and use the corresponding
+non-missing mean values.
+For $K$ segments, this algorithm is $O(K log K)$ time due to the sort.
+It is implemented in the \code{coef} method, as shown in the example code below:
+
+<<>>=
+(ex6_segments <- coef(ex6_binsegRcpp, 2:4))
+@
+
+The table above contains one row per segment, for three model sizes (two to four segments).
+It can be used to plot the segment means and change-point variables, using the code below.
+
+<< ex6, fig=TRUE, height=3 >>=
+library(ggplot2)
+ex6_df$position <- 1:nrow(ex6_df)
+ggplot() + geom_point(aes(position, value), data = ex6_df) +
+  geom_segment(aes(start.pos, y = mean, xend = end.pos, yend = mean),
+    color = "green", data = ex6_segments) +
+  geom_vline(aes(xintercept = start.pos), color="green",
+    linetype = "dashed", data = ex6_segments[1 < start]) +
+  facet_grid(segments ~ ., labeller = label_both)
+@
+
+The figure above shows the data points in black, along with the segmentation model in green (dashed vertical change-points, and solid horizontal segment means).
+
+\subsection{Comparison with changepoint}
+
+To do the analogous computation with \pkg{changepoint}, we can use the code below.
+
+<<>>=
+ex6_changepoint <- changepoint::cpt.mean(ex6_df$value,
+  method = "BinSeg", Q = 3, penalty = "Manual", pen.value = 0)
+changepoint::cpts.full(ex6_changepoint)
+@
+
+The output above is the matrix of change-points, one row for each iteration of binary segmentation.
+One issue with the output above is that the change-points are incorrect in rows 2 and 3.
+For example, row 2 represents the model with two change-points, which appear as 0 and 2 in the matrix above.
+However, for the model with two change-points, the expected result is segments that end at data points 2, 4, and 6.
+These results indicate that the implementation of binary segmentation in \pkg{changepoint} can yield incorrect change-point detection results.
+
+Another issue with \pkg{changepoint} is that the storage of change-points is relatively inefficient.
+For a maximum number of change-points $Q$, it uses a $Q\times Q$ matrix of integers to represent the change-points, so the storage space is quadratic $O(Q^2)$.
+In contrast, \pkg{binsegRcpp} uses only linear space, $O(Q)$, which is asymptotically more efficient.
+Therefore, \pkg{changepoint} can only handle models with relatively small numbers of change-points, whereas \pkg{binsegRcpp} can efficiently store arbitrarily large change-point models.
+
+A final issue is illustrated via the code below,
+%rbind(changepoint = changepoint::param.est(ex6_changepoint)$mean,
+%      expected = ex6_segments[segments==3, mean])
+<<>>=
+changepoint::param.est(ex6_changepoint)
+@
+
+The result above shows the estimated segment means, for which there are two issues.
+First, there are only three values, so these should be the segment means for the model with two change-points.
+However, these three values are incorrect (the correct segment means are 0, 9, and 3).
+Another issue is that the segment means are only reported for one model size,
+and it is not possible to retreive means for other model sizes.
+
+\subsection{Comparison with wbs}
+
+Like \pkg{binsegRcpp}, the \pkg{wbs} package provides a correct implementation of binary segmentation.
+However, it is missing some key features, such as the ability to specify a max number of segments.
+The \pkg{wbs} code below computes the full path of binary segmentation models.
+
+<<>>=
+ex6_wbs <- wbs::sbs(ex6_df$value)
+data.table::data.table(ex6_wbs$res)[order(-min.th)]
+@
+
+The \code{res} table above is similar to the \code{splits} table from \pkg{binsegRcpp}.
+It has one row per split in binary segmentation, so there are five rows in this example (because there are six data points).
+The \code{cpt} column is the index of the last data point before the detected change-point, which is consistent with the \code{end} column from \pkg{binsegRcpp}.
+The \pkg{wbs} package implements binary segmentation using recursive function calls in C (depth-first search, which means it is impossible to specify a max number of segments).
+To compute the change-points for some number of change-points $Q$, the \code{res} table must be sorted, and only the top $Q$ rows be kept.
+For example, the model with $Q=2$ change-points can be represented using the first two rows of the table above (change-points at 2 and 4).
+However, the \pkg{wbs} package does not support retreival of segment-specific model parameters such as segment means.
+If the user wants segment mean estimates, they must be computed using the returned change-points.
 

### 2025-05-02 article.Rnw
figure 1 update

@@
 % https://github.com/HaotianXu/changepoints does not have binary segmentation.
-  version & 2022.3.29 & 2.2.3 & 1.4 & 2019.8.26 & 1.1.6\\
+  version & 2022.3.29 & 2.2.3 & 1.4 & 2019.8.26 & 1.1.9\\
 \hline
@@
   dim & one & one & one & multi & multi \\
-  correct & yes & no & yes & yes & no \\
+  correct & yes & no & yes & yes & yes \\
   losses & 5 & 6 & 1 & 1 & 10 \\
@@
   \centering
-  \includegraphics[width=\textwidth]{figure-neuroblastoma.png}
+  \includegraphics[width=\textwidth]{changepoint.bug.png}
   \caption{\label{fig:neuroblastoma} For a real cancer DNA copy number
     data set with 273 observations, we show the first four
-    changepoints detected by several different implementations of
-    binary segmentation.}
+    change-points models for two different R packages that provide code for
+    binary segmentation: \pkg{binsegRcpp} and \pkg{changepoint}.}
 \end{figure}

### 2025-05-02 changepoint.bug.R

@@ -5,4 +5,90 @@ changepoint::cpt.mean(1:5, method="BinSeg", Q=3)@cpts.full
 changepoint::cpt.mean(1:6, method="BinSeg", Q=4)@cpts.full
 changepoint::cpt.mean(1:8, method="BinSeg", Q=3)@cpts.full
 
-
+library(data.table)
+data(neuroblastoma,package="neuroblastoma")
+nb.profiles <- data.table(neuroblastoma[["profiles"]])
+one.pid.chr <- nb.profiles[profile.id==2 & chromosome==2]
+one.pid.chr[, data.i := .I]
+data.vec <- one.pid.chr$logratio
+N.data <- length(data.vec)
+cum.data.vec <- cumsum(c(0,data.vec))
+max.segs <- 5
+max.changes <- max.segs-1
+binseg.fit <- binsegRcpp::binseg_normal(data.vec, max.segs)
+cpt.fit <- changepoint::cpt.mean(data.vec, method="BinSeg", Q=max.changes)
+loss.dt.list <- list()
+seg.dt.list <- list()
+for(n.changes in 1:max.changes){
+  n.segs <- n.changes+1L
+  end.list <- list(
+    changepoint=c(N.data, cpt.fit@cpts.full[n.changes,]),
+    binsegRcpp=coef(binseg.fit, n.segs)$end
+  )
+  for(package in names(end.list)){
+    end <- sort(end.list[[package]])
+    seg.size <- diff(c(0,end))
+    seg.mean <- (cum.data.vec[end+1]-cum.data.vec[end-seg.size+1])/seg.size
+    data.mean <- rep(seg.mean, seg.size)
+    seg.dt.list[[paste(n.changes, package)]] <- data.table(
+      n.changes,
+      package,
+      start=end-seg.size+1L,
+      end,
+      mean=seg.mean)
+    loss.dt.list[[paste(n.changes, package)]] <- data.table(
+      n.changes, package, total.square.loss=sum((data.mean-data.vec)^2))
+  }
+}
+loss.dt <- do.call(rbind, loss.dt.list)
+dcast(loss.dt, package ~ n.changes, value.var="total.square.loss")
+seg.dt <- do.call(rbind, seg.dt.list)
+loss.y <- c(
+  binsegRcpp=6,
+  changepoint=4)
+library(ggplot2)
+gg <- ggplot()+
+  theme_bw()+
+  scale_color_manual(values=c(binsegRcpp="red", changepoint="black"))+
[…]

### 2025-05-02 figure-neuroblastoma.R

@@ -1,4 +1,3 @@
-source("packages.R")
 if(require(JuliaCall)){
   julia_setup()
   julia_install_package_if_needed("Changepoints")
@@ -16,10 +15,8 @@ if(require(JuliaCall)){
   ## as C++ vector, push_back is constant and
 }
 ## https://github.com/STOR-i/Changepoints.jl#segmentation-with-bs
-Sys.setenv(RETICULATE_PYTHON=if(.Platform$OS.type=="unix")
-  "/home/tdhock/.local/share/r-miniconda/envs/cs570s22/bin/python"
-  else "~/Miniconda3/envs/cs570s22/python.exe")
-reticulate::use_condaenv("cs570s22", required=TRUE)
+Sys.setenv(RETICULATE_PYTHON="~/miniconda3/envs/ruptures/bin/python")
+reticulate::use_condaenv("ruptures", required=TRUE)
 ruptures <- reticulate::import("ruptures")
 data(neuroblastoma,package="neuroblastoma")
 nb.profiles <- data.table(neuroblastoma[["profiles"]])
@@ -56,11 +53,11 @@ ruptures.ord <- list()
 for(n.changes in 1:max.changes){
   n.segs <- n.changes+1L
   end.list <- list(
-    ##"fpop::multiBinSeg"=c(N.data,fpop.fit$t.est[1:n.changes]),
+    "fpop::multiBinSeg"=c(N.data,fpop.fit$t.est[1:n.changes]),
     ruptures=binseg_instance$fit(data.mat)$predict(n_bkps=n.changes),
     changepoint=c(N.data,cpt.fit@cpts.full[n.changes,]),
-    ##wbs=wbs.dt[, c(N.data, cpt)][1:n.segs],
-    "binsegRcpp=fpop=wbs"=coef(binseg.fit, n.segs)$end
+    wbs=wbs.dt[, c(N.data, cpt)][1:n.segs],
+    "binsegRcpp=fpop=wbs=ruptures"=coef(binseg.fit, n.segs)$end
   )
   ruptures.ord[[n.changes]] <- with(
     end.list, ruptures[!ruptures %in% last.ruptures])
@@ -81,24 +78,8 @@ for(n.changes in 1:max.changes){
   }
 }
 loss.dt <- do.call(rbind, loss.dt.list)
-
+dcast(loss.dt, package ~ n.changes, value.var="total.square.loss")
 seg.dt <- do.call(rbind, seg.dt.list)
-split.dt.list <- list()
-change.list <- list(
[…]

### 2025-05-02 ruptures_bug.py

@@ -1,3 +1,4 @@
+#https://github.com/deepcharles/ruptures/issues/242
 import ruptures as rpt
 import numpy as np
 rpt.version.version
@@ -18,4 +19,38 @@ expected_break_dict = {
     1:[2,4],
     2:[2,3,4],
     3:[1,2,3,4]
+    }
+
+
+
+
+#https://github.com/deepcharles/ruptures/issues/244
+import ruptures as rpt
+import numpy as np
+rpt.version.version
+data_list = [0,0.3,0.2,0.1, 10,11,12,13]
+N_data = len(data_list)
+data_mat = np.array(data_list).reshape(N_data,1)
+algo = rpt.Binseg(model="normal",jump=1).fit(data_mat)
+computed_break_dict={n_bkps:algo.predict(n_bkps=n_bkps) for n_bkps in range(4)}
+computed_break_dict
+expected_break_dict = {
+    0:[8],
+    1:[4,8],
+    2:[4,6,8],
+    3:[2,4,6,8]
+    }
+
+#from article.Rnw
+data_list=[1, -7, 8, 10, 2, 4]
+N_data = len(data_list)
+data_mat = np.array(data_list).reshape(N_data,1)
+algo = rpt.Binseg(model="l2",min_size=1,jump=1).fit(data_mat)
+computed_break_dict={n_bkps:algo.predict(n_bkps=n_bkps) for n_bkps in range(4)}
+computed_break_dict
+expected_break_dict = {
+    0:[8],
+    1:[4,8],
+    2:[4,6,8],
+    3:[2,4,6,8]
     }

### 2025-05-02 article.Rnw
rpt

@@
   extremely fast for $N$ data and $K$ segments, $O(N K)$ in the worst
-  case, and $O(N \log K)$ in the best case. In practice existing
-  implementations such as ruptures (\proglang{Python} module) and
-  changepoint (\proglang{R} package) are much slower, and in fact
-  sometimes do not return a correct result. We present the
-  \proglang{R} package \pkg{binsegRcpp}, which provides an efficient
-  C++ implementation of binary segmentation, and always returns
-  correct results. We discuss some C++ coding techniques
+  case, and $O(N \log K)$ in the best case. In practice, existing
+  implementations such as \pkg{ruptures} (\proglang{Python} module) and
+  \pkg{changepoint} (\proglang{R} package) are substantially slower.
+  Additionally, \pkg{changepoint} sometimes returns incorrect results.
+  We present the \proglang{R} package \pkg{binsegRcpp}, which provides an efficient
+  and correct C++ implementation of binary segmentation.
+  We discuss some C++ coding techniques
   which were used to avoid repetition, for efficiency, and for
@@
   empirical comparisons with other implementations of binary
-  segmentation in \proglang{R} packages \pkg{fpop}, \pkg{blockcpd} and
+  segmentation in other packages: \pkg{ruptures}, \pkg{fpop}, \pkg{changepoint} and
   \pkg{wbs}. }

### 2025-05-05 article.Rnw
loss table
blockcpd column

@@
 %% - \Abstract{} almost as usual
-\Abstract{ Binary segmentation is a classic algorithm for detecting
+\Abstract{
+  Binary segmentation is a classic algorithm for detecting
   changepoints in sequential data. In theory, using a simple loss
@@
   segmentation in other packages: \pkg{ruptures}, \pkg{fpop}, \pkg{changepoint} and
-  \pkg{wbs}. }
+  \pkg{wbs}.
+}
 
@@
     \code{var} & model parameter &  $S/W + M(M-2T/W)$
+  \end{tabular}
+  \caption{\label{tab:params}TODO}
+\end{table}
 
+\begin{table*}
+  \centering
+  \begin{tabular}{ccccccc}
+    Name & Loss & Distrib. & Center & Scale & Params & Complexity \\
+    \hline
+    \texttt{mean\_norm} & Square & Normal & yes & no & 1 & $O(S)$ \\
+    \texttt{meanvar\_norm} & Neg. Log. Lik. & Normal & yes & yes & 2 & $O(S)$ \\
+    \texttt{poisson} & Neg. Log. Lik. & Poisson & yes & yes & 1 & $O(S)$ \\
+    \texttt{l1} & Absolute & Laplace & yes & no & 1 & $O(S\log S)$ \\
+    \texttt{laplace} & Absolute & Laplace & yes & yes & 2 & $O(S\log S)$ \\
   \end{tabular}
-\end{table}
+  \caption{\label{tab:losses}
+    TODO
+  }
+\end{table*}
 
@@
   \centering
-\begin{tabular}{l|ccccc}
-  package & binsegRcpp & changepoint & wbs & fpop & ruptures \\
-  function & binseg & cpt.mean & sbs & multiBinSeg & Binseg \\
+  \scriptsize
+\begin{tabular}{l|cccccc}
+  package & binsegRcpp & changepoint & wbs & fpop & ruptures & blockcpd\\
+  function & binseg & cpt.mean & sbs & multiBinSeg & Binseg & fit\_blockcpd\\
 %   > ruptures$version$version
@@
 % https://github.com/HaotianXu/changepoints does not have binary segmentation.
-  version & 2022.3.29 & 2.2.3 & 1.4 & 2019.8.26 & 1.1.9\\
+  version & 2022.3.29 & 2.2.3 & 1.4 & 2019.8.26 & 1.1.9 & 1.0.0\\
 \hline
-  weights & yes & no & no & no & no \\
-  max segs & yes & yes & no & yes & yes \\
-  dim & one & one & one & multi & multi \\
+  weights & yes & no & no & no & no & no\\
+  max segs & yes & yes & no & yes & yes & yes\\
+  dim & one & one & one & multi & multi & multi \\
   correct & yes & no & yes & yes & yes \\
-  losses & 5 & 6 & 1 & 1 & 10 \\
-  language & C++ & C & C & C++ & Python \\
+  losses & 5 & 6 & 1 & 1 & 10 & 5\\
+  language & C++ & C & C & C++ & Python & C++\\
   % all propose some loss computation.
-  storage & STL multiset & arrays & recursion & heap & LRU cache \\
-  space & $O(S)$ & $O(S^2)$ & $O(S)$ & $O(S)$ & $O(S)$ \\
-  cumsum & yes & yes & yes & yes & no \\
-  best & $O(N\log N)$ & $O(N^3)$ & $O(N\log N)$ & $O(N\log N)$ & $>O(N\log N)$ \\
-  worst & $O(N^2)$ & $O(N^3)$ & $O(N^2)$ & $O(N^2)$ & $O(N^2)$ \\
-  CV & yes & no & no & no & no\\
-  params & yes & largest & no & no & no \\
+  storage & STL multiset & arrays & recursion & heap & LRU cache & heap \\
+  space & $O(S)$ & $O(S^2)$ & $O(S)$ & $O(S)$ & $O(S)$ & $O(S)$ \\
+  cumsum & yes & yes & yes & yes & no & yes\\
+  best & $O(N\log N)$ & $O(N^3)$ & $O(N\log N)$ & $O(N\log N)$ & $>O(N\log N)$ & $O(N\log N)$\\
+  worst & $O(N^2)$ & $O(N^3)$ & $O(N^2)$ & $O(N^2)$ & $O(N^2)$ & $O(N^2)$\\
+  CV & yes & no & no & no & no & no\\
+  params & all & one & no & no & no &one\\
 \end{tabular}

### 2025-05-06 article.Rnw
figure heap

@@
     \texttt{mean\_norm} & Square & Normal & yes & no & 1 & $O(S)$ \\
-    \texttt{meanvar\_norm} & Neg. Log. Lik. & Normal & yes & yes & 2 & $O(S)$ \\
-    \texttt{poisson} & Neg. Log. Lik. & Poisson & yes & yes & 1 & $O(S)$ \\
+    \texttt{meanvar\_norm} & Neg. Log Lik. & Normal & yes & yes & 2 & $O(S)$ \\
+    \texttt{poisson} & Neg. Log Lik. & Poisson & yes & yes & 1 & $O(S)$ \\
     \texttt{l1} & Absolute & Laplace & yes & no & 1 & $O(S\log S)$ \\
-    \texttt{laplace} & Absolute & Laplace & yes & yes & 2 & $O(S\log S)$ \\
+    \texttt{laplace} & Neg. Log Lik & Laplace & yes & yes & 2 & $O(S\log S)$ \\
   \end{tabular}
@@
 ex6_df <- data.frame(value = c(1, -7, 8, 10, 2, 4))
-ex6_binsegRcpp <- binsegRcpp::binseg("mean_norm", ex6_df$value, max.segments = 4)
+ex6_binsegRcpp <- binsegRcpp::binseg(
+  "mean_norm", ex6_df$value, max.segments = 4)
 ex6_binsegRcpp$splits[, .(segments, end, before.mean, after.mean,

### 2025-05-06 figure-heap-data.R

@@ -1,0 +1,51 @@
+Sys.setenv(RETICULATE_PYTHON="~/miniconda3/envs/ruptures/bin/python")
+reticulate::use_condaenv("ruptures", required=TRUE)
+
+ruptures <- reticulate::import("ruptures")
+binseg_l2 <- ruptures$Binseg(min_size=1L, jump=1L)
+
+gen_data <- function(N.data)list(
+  flat=rnorm(N.data),
+  best=1:N.data,
+  worst=rep(c(0,1),l=N.data))
+expr.list <- atime::atime_grid(list(
+  container=c("priority_queue","list","multiset")),
+  ##case=names(gen_data(2))),
+  binsegRcpp={
+    binseg.fit <- binsegRcpp::binseg(
+      "mean_norm", data_vec, max.segs, container.str = container)
+    sort(binseg.fit$splits$end)
+  },
+  expr.param.sep="\n")
+atime_list <- atime::atime(
+  N=2^seq(2, 30),
+  ##N=2^seq(2, 4),
+  setup={
+    max.segs <- as.integer(N/2)
+    max.changes <- max.segs-1L
+    set.seed(1)
+    ## data_list <- gen_data(N)
+    ## mat_list <- lapply(data_list, cbind)
+    data_vec <- 1:N
+    data_mat <- cbind(data_vec)
+  },
+  ruptures={
+    binseg_l2$fit(data_mat)$predict(n_bkps=max.changes)
+  },
+  changepoint={
+    cpt.fit <- changepoint::cpt.mean(data_vec, method="BinSeg", Q=max.changes)
+    sort(c(N, cpt.fit@cpts.full[max.changes,]))
+  },
+  verbose=TRUE,
+  result=TRUE,
+  seconds.limit=5,
+  expr.list=expr.list)
+atime_list$measurements[N<20, .(N, expr.name, paste(result))]
+library(data.table)
+atime_wide <- dcast(
+  atime_list$measurements[!is.na(container)],
+  N ~ container,
+  value.var="result")
+stopifnot(atime_wide[, identical(multiset, priority_queue)])
+atime_list$measurements[, result := NULL][]
+save(atime_list, file="figure-heap-data.RData")

### 2025-05-06 figure-heap.R

@@ -1,0 +1,44 @@
+(objs <- load("figure-heap-data.RData"))
+requireNamespace("atime")
+plot(atime_list)
+
+fun_list <- atime:::references_funs[c("N","N \\log N", "N^2", "N^3")]
+fun_list[["N^1.5"]] <- function(N)1.5 * log10(N)
+ref_list <- atime::references_best(atime_list, fun_list)
+ref_list$plot.references <- ref_list$references[(
+  fun.name %in% c("N","N^2") & unit=="kilobytes"
+)|(
+  fun.name != "N" & unit=="seconds"
+)]
+png("figure-heap-refs.png",width=14,height=6,units="in",res=200)
+plot(ref_list)
+dev.off()
+
+pred_list <- predict(ref_list)
+png("figure-heap-pred.png",width=12,height=4.5,units="in",res=200)
+pkg.colors <- c(
+  changepoint="#E41A1C",
+  ruptures="#377EB8",
+  "binsegRcpp\ncontainer=multiset"="black",
+  "binsegRcpp\ncontainer=priority_queue"="grey30",
+  "binsegRcpp\ncontainer=list"="grey50")
+plot(pred_list)+
+  facet_null()+
+  scale_y_log10(
+    "Computation time (seconds, log scale)")+
+  scale_x_log10(
+    "N = number of data to segment (max segments = N/2)",
+    breaks=10^seq(1,7),
+    limits=c(NA, 1e9))+
+  scale_color_manual(values=pkg.colors)+
+  scale_fill_manual(values=pkg.colors)
+dev.off()
+
+norm_dt <- pred_list$prediction[, ratio := max(N)/N][, .(expr.name, ratio)][pred_list$meas, on="expr.name"][unit=="seconds"][, norm_N := N*ratio][]
+library(ggplot2)
+ggplot()+
+  geom_line(aes(
+    norm_N, empirical, color=expr.name),
+    data=norm_dt)+
+  scale_x_log10()+
+  scale_y_log10()

### 2025-05-06 article.Rnw
figure-heap-refs-other

@@
 %% - \AND starts a new line, \And does not.
-\author{Toby Dylan Hocking~\orcid{0000-0002-3146-0865}\\Northern Arizona University}
+\author{Toby Dylan Hocking~\orcid{0000-0002-3146-0865}\\Université de Sherbrooke}
 \Plainauthor{Toby Dylan Hocking}
@@
 %% - \Shorttitle{} with LaTeX markup (if any), used as running title
-\title{Comparing binsegRcpp with other implementations of binary segmentation for changepoint detection}
-\Plaintitle{Comparing binsegRcpp with other implementations of binary segmentation for changepoint detection}
-\Shorttitle{binsegRcpp for changepoint detection}
+\title{Comparing \pkg{binsegRcpp} with other implementations of binary segmentation for change-point detection}
+\Plaintitle{Comparing binsegRcpp with other implementations of binary segmentation for change-point detection}
+\Shorttitle{binsegRcpp for efficient change-point detection}
 
@@
   Binary segmentation is a classic algorithm for detecting
-  changepoints in sequential data. In theory, using a simple loss
-  function like Gaussian or Poisson, binary segmentation should be
-  extremely fast for $N$ data and $K$ segments, $O(N K)$ in the worst
-  case, and $O(N \log K)$ in the best case. In practice, existing
-  implementations such as \pkg{ruptures} (\proglang{Python} module) and
-  \pkg{changepoint} (\proglang{R} package) are substantially slower.
-  Additionally, \pkg{changepoint} sometimes returns incorrect results.
-  We present the \proglang{R} package \pkg{binsegRcpp}, which provides an efficient
-  and correct C++ implementation of binary segmentation.
-  We discuss some C++ coding techniques
-  which were used to avoid repetition, for efficiency, and for
-  readability.  We additionally include detailed theoretical and
-  empirical comparisons with other implementations of binary
-  segmentation in other packages: \pkg{ruptures}, \pkg{fpop}, \pkg{changepoint} and
+  change-points in sequential data. In theory, using a simple loss
+  function like Normal or Poisson, binary segmentation should be
+  extremely fast for $N$ data and $K$ segments: asymptotically $O(N K)$ time in the worst
+  case, and $O(N \log K)$ time in the best case.
+  In practice,
+  other implementations can be asymptotically slower, and can 
+  return incorrect results. We propose \pkg{binsegRcpp}, an \proglang{R} package which
+  provides a correct \proglang{C++} implementation, with the expected asymptotic time complexity.
+  %% In practice, existing
+  %% implementations such as \pkg{ruptures} (\proglang{Python} module) and
+  %% \pkg{changepoint} (\proglang{R} package) are asymptotically slower.
+  %% Additionally, \pkg{changepoint} sometimes returns incorrect results.
+  %% We present the \proglang{R} package \pkg{binsegRcpp}, which provides an efficient
+  %% and correct C++ implementation of binary segmentation.
+  We discuss several important \proglang{C++} coding techniques,
+  and include detailed comparisons with other implementations of binary
+  segmentation: \pkg{ruptures}, \pkg{fpop}, \pkg{changepoint}, \pkg{blockcpd}, and
   \pkg{wbs}.
@@
 %% - Should be comma-separated and in sentence case.
-\Keywords{JSS, style guide, comma-separated, not capitalized, \proglang{R}}
-\Plainkeywords{JSS, style guide, comma-separated, not capitalized, R}
+\Keywords{\proglang{C++}, \proglang{R}, binary segmentation, change-point}
+\Plainkeywords{C++, R, binary segmentation, change-point}
 
@@
 TODO \code{fastcpd} package \citep{fastcpd2024} (nice related work table).
+
+TODO \pkg{blockcpd} package TODO \citep{blockcpd2021,Prates2021} .
 
@@
   \section{More technical details} \label{app:technical}
-  TODO
+
+\begin{figure}[t!]
+  \centering
+  \includegraphics[width=\textwidth]{figure-heap-pred}
+  \caption{\label{fig:heap-pred} Comparing computation time of \pkg{binsegRcpp} with three containers, and two baselines: \pkg{changepoint} and \pkg{ruptures}.}
+\end{figure}
+
+\begin{figure}[t!]
+  \centering
+  \includegraphics[width=\textwidth]{figure-heap-refs-binsegRcpp}
+  \caption{\label{fig:heap-refs} Estimation of asymptotic time (seconds) and memory (kilobytes) usage for \pkg{binsegRcpp}.
+    Black curves are empirical measurements, and violet curves are references.
+    These data suggest that both list and multiset containers are $O(N)$;
+    list is $O(N^2)$ time and multiset is $O(N\log N)$ time.  }
+\end{figure}
+
+\begin{figure}[t!]
+  \centering
+  \includegraphics[width=\textwidth]{figure-heap-refs-other}
+  \caption{\label{fig:heap-refs} Estimation of asymptotic time (seconds) and memory (kilobytes) usage for \pkg{changepoint} and \pkg{ruptures}.
+    Black curves are empirical measurements, and violet curves are references.
+    These data suggest that \pkg{changepoint} is $O(N^2)$ space and $O(N^3)$ time;
+    \pkg{ruptures} is $O(N)$ space and $O(N^{1.5})$ time.
+  }
+\end{figure}
 

### 2025-05-06 figure-heap.R

@@ -1,5 +1,6 @@
 (objs <- load("figure-heap-data.RData"))
 requireNamespace("atime")
+atime_list$measurements[, expr.name := sub("container=", "", expr.name)]
 plot(atime_list)
 
 fun_list <- atime:::references_funs[c("N","N \\log N", "N^2", "N^3")]
@@ -10,26 +11,46 @@ ref_list$plot.references <- ref_list$references[(
 )|(
   fun.name != "N" & unit=="seconds"
 )]
-png("figure-heap-refs.png",width=14,height=6,units="in",res=200)
-plot(ref_list)
+##ref_list$measurements <- ref_list
+png("figure-heap-refs-binsegRcpp.png",width=6,height=3.5,units="in",res=200)
+plot_algos <- function(regex){
+  regex_list <- ref_list
+  regex_list$measurements <- ref_list$measurements[grepl(regex,expr.name)]
+  regex_list$plot.references <- ref_list$plot.references[grepl(regex,expr.name)]
+  plot(regex_list)+
+    geom_blank(aes(
+      N, empirical),
+      data=ref_list$measurements[,.(N=1000,empirical,unit)])+
+    scale_x_log10(
+      breaks=10^seq(1,7,by=1))+
+    theme(
+      axis.text.x=element_text(angle=30,hjust=1))
+}
+plot_algos("list|multiset")
+dev.off()
+
+png("figure-heap-refs-other.png",width=6,height=3.5,units="in",res=200)
+plot_algos("change|rupt")
 dev.off()
 
 pred_list <- predict(ref_list)
-png("figure-heap-pred.png",width=12,height=4.5,units="in",res=200)
+png("figure-heap-pred.png",width=8,height=3,units="in",res=200)
 pkg.colors <- c(
   changepoint="#E41A1C",
   ruptures="#377EB8",
-  "binsegRcpp\ncontainer=multiset"="black",
-  "binsegRcpp\ncontainer=priority_queue"="grey30",
-  "binsegRcpp\ncontainer=list"="grey50")
+  "binsegRcpp\nmultiset"="black",
+  "binsegRcpp\npriority_queue"="grey30",
+  "binsegRcpp\nlist"="grey50")
 plot(pred_list)+
   facet_null()+
   scale_y_log10(
-    "Computation time (seconds, log scale)")+
+    "Computation time (seconds, log scale)",
+    breaks=10^seq(-3,1),
+    limits=c(NA,1e2))+
[…]

### 2025-05-06 figure-heap-refs-binsegRcpp.tex
tikz figures

% Created by tikzDevice version 0.12.6 on 2025-05-06 20:55:50
% !TEX encoding = UTF-8 Unicode
\begin{tikzpicture}[x=1pt,y=1pt]
\definecolor{fillColor}{RGB}{255,255,255}
\path[use as bounding box,fill=fillColor,fill opacity=0.00] (0,0) rectangle (433.62,216.81);
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (433.62,216.81);
\definecolor{drawColor}{RGB}{255,255,255}
\definecolor{fillColor}{RGB}{255,255,255}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] ( -0.00,  0.00) rectangle (433.62,216.81);
\end{scope}
\begin{scope}
\path[clip] ( 46.86,116.18) rectangle (226.46,185.23);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] ( 46.86,116.18) rectangle (226.46,185.23);
\definecolor{drawColor}{gray}{0.92}

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 46.86,124.04) --
	(226.46,124.04);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 46.86,141.98) --
	(226.46,141.98);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 46.86,159.91) --
	(226.46,159.91);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 46.86,177.84) --
	(226.46,177.84);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 51.34,116.18) --
	( 51.34,185.23);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 87.49,116.18) --
	( 87.49,185.23);

\path[draw=drawColor,line width= 0.3pt,line join=round] (123.65,116.18) --
	(123.65,185.23);

\path[draw=drawColor,line width= 0.3pt,line join=round] (159.81,116.18) --
	(159.81,185.23);

\path[draw=drawColor,line width= 0.3pt,line join=round] (195.97,116.18) --
	(195.97,185.23);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 46.86,133.01) --
	(226.46,133.01);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 46.86,150.94) --
	(226.46,150.94);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 46.86,168.88) --
	(226.46,168.88);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 69.42,116.18) --
	( 69.42,185.23);

\path[draw=drawColor,line width= 0.6pt,line join=round] (105.57,116.18) --
	(105.57,185.23);

\path[draw=drawColor,line width= 0.6pt,line join=round] (141.73,116.18) --
	(141.73,185.23);

\path[draw=drawColor,line width= 0.6pt,line join=round] (177.89,116.18) --
	(177.89,185.23);

\path[draw=drawColor,line width= 0.6pt,line join=round] (214.04,116.18) --
	(214.04,185.23);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 2.3pt,line join=round] ( 55.03,147.52) --
	( 65.91,140.30) --
	( 76.80,140.31) --
	( 87.68,140.47) --
	( 98.56,141.29) --
	(109.45,142.56) --
	(120.33,143.85) --
	(131.22,145.59) --
	(142.10,147.70) --
	(152.99,150.07) --
	(163.87,152.59) --
	(174.76,155.20) --
	(185.64,157.85) --
	(196.52,160.53) --
	(207.41,163.21) --
	(218.29,165.91);
\definecolor{drawColor}{RGB}{238,130,238}

\path[draw=drawColor,line width= 1.1pt,line join=round] (142.10,128.12) --
	(152.99,133.52) --
	(163.87,138.92) --
	(174.76,144.31) --
	(185.64,149.71) --
	(196.52,155.11) --
	(207.41,160.51) --
	(218.29,165.91);

\path[draw=drawColor,line width= 1.1pt,line join=round] ( 65.91,128.12) --
	( 76.80,130.82) --
	( 87.68,133.52) --
	( 98.56,136.22) --
	(109.45,138.92) --
	(120.33,141.62) --
	(131.22,144.31) --
	(142.10,147.01) --
	(152.99,149.71) --
	(163.87,152.41) --
	(174.76,155.11) --
	(185.64,157.81) --
	(196.52,160.51) --
	(207.41,163.21) --
	(218.29,165.91);
\end{scope}
\begin{scope}
\path[clip] ( 46.86,116.18) rectangle (226.46,185.23);
\definecolor{drawColor}{RGB}{0,0,0}
\definecolor{fillColor}{RGB}{238,130,238}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (133.87,125.28) --
	(142.10,128.12) --
	(150.33,125.28) --
	(150.33,116.41) --
	(133.87,116.41) --
	cycle;

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] ( 59.93,125.28) --
	( 65.91,128.12) --
	( 71.90,125.28) --
	( 71.90,116.41) --
	( 59.93,116.41) --
	cycle;

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at (142.10,118.39) {$N^2$};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at ( 65.91,118.39) {$N$};
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 46.86,116.18) rectangle (226.46,185.23);
\end{scope}
\begin{scope}
\path[clip] ( 46.86, 41.62) rectangle (226.46,110.68);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] ( 46.86, 41.62) rectangle (226.46,110.68);
\definecolor{drawColor}{gray}{0.92}

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 46.86, 42.48) --
	(226.46, 42.48);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 46.86, 55.84) --
	(226.46, 55.84);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 46.86, 69.20) --
	(226.46, 69.20);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 46.86, 82.55) --
	(226.46, 82.55);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 46.86, 95.91) --
	(226.46, 95.91);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 46.86,109.27) --
	(226.46,109.27);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 51.34, 41.62) --
	( 51.34,110.68);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 87.49, 41.62) --
	( 87.49,110.68);

\path[draw=drawColor,line width= 0.3pt,line join=round] (123.65, 41.62) --
	(123.65,110.68);

\path[draw=drawColor,line width= 0.3pt,line join=round] (159.81, 41.62) --
	(159.81,110.68);

\path[draw=drawColor,line width= 0.3pt,line join=round] (195.97, 41.62) --
	(195.97,110.68);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 46.86, 49.16) --
	(226.46, 49.16);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 46.86, 62.52) --
	(226.46, 62.52);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 46.86, 75.88) --
	(226.46, 75.88);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 46.86, 89.23) --
	(226.46, 89.23);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 46.86,102.59) --
	(226.46,102.59);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 69.42, 41.62) --
	( 69.42,110.68);

\path[draw=drawColor,line width= 0.6pt,line join=round] (105.57, 41.62) --
	(105.57,110.68);

\path[draw=drawColor,line width= 0.6pt,line join=round] (141.73, 41.62) --
	(141.73,110.68);

\path[draw=drawColor,line width= 0.6pt,line join=round] (177.89, 41.62) --
	(177.89,110.68);

\path[draw=drawColor,line width= 0.6pt,line join=round] (214.04, 41.62) --
	(214.04,110.68);
\definecolor{drawColor}{RGB}{190,190,190}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 46.86, 98.57) -- (226.46, 98.57);
\definecolor{fillColor}{RGB}{0,0,0}

\path[fill=fillColor,fill opacity=0.50] ( 55.03, 53.39) --
	( 65.91, 53.70) --
	( 76.80, 53.42) --
	( 87.68, 53.62) --
	( 98.56, 53.81) --
	(109.45, 52.36) --
	(120.33, 60.85) --
	(131.22, 50.50) --
	(142.10, 53.19) --
	(152.99, 58.46) --
	(163.87, 65.47) --
	(174.76, 72.35) --
	(185.64, 79.70) --
	(196.52, 88.57) --
	(207.41, 97.32) --
	(218.29,107.54) --
	(218.29,106.40) --
	(207.41, 97.20) --
	(196.52, 88.37) --
	(185.64, 79.22) --
	(174.76, 72.10) --
	(163.87, 65.19) --
	(152.99, 58.08) --
	(142.10, 51.77) --
	(131.22, 48.59) --
	(120.33, 47.68) --
	(109.45, 48.68) --
	( 98.56, 51.61) --
	( 87.68, 51.71) --
	( 76.80, 51.86) --
	( 65.91, 51.48) --
	( 55.03, 51.51) --
	cycle;

\path[] ( 55.03, 53.39) --
	( 65.91, 53.70) --
	( 76.80, 53.42) --
	( 87.68, 53.62) --
	( 98.56, 53.81) --
	(109.45, 52.36) --
	(120.33, 60.85) --
	(131.22, 50.50) --
	(142.10, 53.19) --
	(152.99, 58.46) --
	(163.87, 65.47) --
	(174.76, 72.35) --
	(185.64, 79.70) --
	(196.52, 88.57) --
	(207.41, 97.32) --
	(218.29,107.54);

\path[] (218.29,106.40) --
	(207.41, 97.20) --
	(196.52, 88.37) --
	(185.64, 79.22) --
	(174.76, 72.10) --
	(163.87, 65.19) --
	(152.99, 58.08) --
	(142.10, 51.77) --
	(131.22, 48.59) --
	(120.33, 47.68) --
	(109.45, 48.68) --
	( 98.56, 51.61) --
	( 87.68, 51.71) --
	( 76.80, 51.86) --
	( 65.91, 51.48) --
	( 55.03, 51.51);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 2.3pt,line join=round] ( 55.03, 51.80) --
	( 65.91, 51.96) --
	( 76.80, 52.23) --
	( 87.68, 51.97) --
	( 98.56, 51.99) --
	(109.45, 49.61) --
	(120.33, 48.49) --
	(131.22, 49.12) --
	(142.10, 52.46) --
	(152.99, 58.30) --
	(163.87, 65.29) --
	(174.76, 72.20) --
	(185.64, 79.51) --
	(196.52, 88.44) --
	(207.41, 97.24) --
	(218.29,106.51);
\definecolor{drawColor}{RGB}{238,130,238}

\path[draw=drawColor,line width= 1.1pt,line join=round] ( 87.68, 51.15) --
	( 98.56, 56.23) --
	(109.45, 61.15) --
	(120.33, 65.94) --
	(131.22, 70.65) --
	(142.10, 75.28) --
	(152.99, 79.85) --
	(163.87, 84.38) --
	(174.76, 88.87) --
	(185.64, 93.32) --
	(196.52, 97.74) --
	(207.41,102.13) --
	(218.29,106.51);

\path[draw=drawColor,line width= 1.1pt,line join=round] (120.33, 52.22) --
	(131.22, 58.25) --
	(142.10, 64.28) --
	(152.99, 70.31) --
	(163.87, 76.35) --
	(174.76, 82.38) --
	(185.64, 88.41) --
	(196.52, 94.44) --
	(207.41,100.48) --
	(218.29,106.51);

\path[draw=drawColor,line width= 1.1pt,line join=round] (152.99, 58.25) --
	(163.87, 66.29) --
	(174.76, 74.34) --
	(185.64, 82.38) --
	(196.52, 90.42) --
	(207.41, 98.46) --
	(218.29,106.51);

\path[draw=drawColor,line width= 1.1pt,line join=round] (174.76, 58.25) --
	(185.64, 70.31) --
	(196.52, 82.38) --
	(207.41, 94.44) --
	(218.29,106.51);
\end{scope}
\begin{scope}
\path[clip] ( 46.86, 41.62) rectangle (226.46,110.68);
\definecolor{drawColor}{RGB}{0,0,0}
\definecolor{fillColor}{RGB}{238,130,238}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] ( 73.97, 48.31) --
	( 87.68, 51.15) --
	(101.39, 48.31) --
	(101.39, 40.20) --
	( 73.97, 40.20) --
	cycle;

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (111.80, 48.31) --
	(120.33, 52.22) --
	(128.87, 48.31) --
	(128.87, 40.20) --
	(111.80, 40.20) --
	cycle;

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (145.45, 48.31) --
	(152.99, 58.25) --
	(160.52, 48.31) --
	(160.52, 40.20) --
	(145.45, 40.20) --
	cycle;

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (167.22, 48.31) --
	(174.76, 58.25) --
	(182.29, 48.31) --
	(182.29, 40.20) --
	(167.22, 40.20) --
	cycle;

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.71] at ( 87.68, 43.40) {$N \log N$};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.71] at (120.33, 43.40) {$N^{1.5}$};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.90] at (152.99, 42.12) {$N^2$};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.90] at (174.76, 42.12) {$N^3$};
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 46.86, 41.62) rectangle (226.46,110.68);
\end{scope}
\begin{scope}
\path[clip] (231.96,116.18) rectangle (411.55,185.23);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (231.96,116.18) rectangle (411.55,185.23);
\definecolor{drawColor}{gray}{0.92}

\path[draw=drawColor,line width= 0.3pt,line join=round] (231.96,124.04) --
	(411.55,124.04);

\path[draw=drawColor,line width= 0.3pt,line join=round] (231.96,141.98) --
	(411.55,141.98);

\path[draw=drawColor,line width= 0.3pt,line join=round] (231.96,159.91) --
	(411.55,159.91);

\path[draw=drawColor,line width= 0.3pt,line join=round] (231.96,177.84) --
	(411.55,177.84);

\path[draw=drawColor,line width= 0.3pt,line join=round] (237.48,116.18) --
	(237.48,185.23);

\path[draw=drawColor,line width= 0.3pt,line join=round] (263.31,116.18) --
	(263.31,185.23);

\path[draw=drawColor,line width= 0.3pt,line join=round] (289.14,116.18) --
	(289.14,185.23);

\path[draw=drawColor,line width= 0.3pt,line join=round] (314.96,116.18) --
	(314.96,185.23);

\path[draw=drawColor,line width= 0.3pt,line join=round] (340.79,116.18) --
	(340.79,185.23);

\path[draw=drawColor,line width= 0.3pt,line join=round] (366.62,116.18) --
	(366.62,185.23);

\path[draw=drawColor,line width= 0.3pt,line join=round] (392.44,116.18) --
	(392.44,185.23);

\path[draw=drawColor,line width= 0.6pt,line join=round] (231.96,133.01) --
	(411.55,133.01);

\path[draw=drawColor,line width= 0.6pt,line join=round] (231.96,150.94) --
	(411.55,150.94);

\path[draw=drawColor,line width= 0.6pt,line join=round] (231.96,168.88) --
	(411.55,168.88);

\path[draw=drawColor,line width= 0.6pt,line join=round] (250.40,116.18) --
	(250.40,185.23);

\path[draw=drawColor,line width= 0.6pt,line join=round] (276.22,116.18) --
	(276.22,185.23);

\path[draw=drawColor,line width= 0.6pt,line join=round] (302.05,116.18) --
	(302.05,185.23);

\path[draw=drawColor,line width= 0.6pt,line join=round] (327.88,116.18) --
	(327.88,185.23);

\path[draw=drawColor,line width= 0.6pt,line join=round] (353.70,116.18) --
	(353.70,185.23);

\path[draw=drawColor,line width= 0.6pt,line join=round] (379.53,116.18) --
	(379.53,185.23);

\path[draw=drawColor,line width= 0.6pt,line join=round] (405.36,116.18) --
	(405.36,185.23);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 2.3pt,line join=round] (240.12,140.30) --
	(247.89,140.30) --
	(255.67,140.31) --
	(263.44,140.47) --
	(271.22,141.29) --
	(278.99,142.56) --
	(286.77,143.85) --
	(294.54,145.59) --
	(302.32,147.70) --
	(310.09,150.07) --
	(317.87,152.59) --
	(325.64,155.20) --
	(333.41,157.85) --
	(341.19,160.53) --
	(348.96,163.22) --
	(356.74,165.91) --
	(364.51,168.60) --
	(372.29,171.30) --
	(380.06,174.00) --
	(387.84,176.70) --
	(395.61,179.40) --
	(403.39,182.10);
\definecolor{drawColor}{RGB}{238,130,238}

\path[draw=drawColor,line width= 1.1pt,line join=round] (325.64,128.11) --
	(333.41,133.51) --
	(341.19,138.91) --
	(348.96,144.31) --
	(356.74,149.71) --
	(364.51,155.11) --
	(372.29,160.50) --
	(380.06,165.90) --
	(387.84,171.30) --
	(395.61,176.70) --
	(403.39,182.10);

\path[draw=drawColor,line width= 1.1pt,line join=round] (247.89,128.11) --
	(255.67,130.81) --
	(263.44,133.51) --
	(271.22,136.21) --
	(278.99,138.91) --
	(286.77,141.61) --
	(294.54,144.31) --
	(302.32,147.01) --
	(310.09,149.71) --
	(317.87,152.41) --
	(325.64,155.11) --
	(333.41,157.80) --
	(341.19,160.50) --
	(348.96,163.20) --
	(356.74,165.90) --
	(364.51,168.60) --
	(372.29,171.30) --
	(380.06,174.00) --
	(387.84,176.70) --
	(395.61,179.40) --
	(403.39,182.10);
\end{scope}
\begin{scope}
\path[clip] (231.96,116.18) rectangle (411.55,185.23);
\definecolor{drawColor}{RGB}{0,0,0}
\definecolor{fillColor}{RGB}{238,130,238}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (317.41,125.27) --
	(325.64,128.11) --
	(333.87,125.27) --
	(333.87,116.40) --
	(317.41,116.40) --
	cycle;

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (241.91,125.27) --
	(247.89,128.11) --
	(253.88,125.27) --
	(253.88,116.40) --
	(241.91,116.40) --
	cycle;

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at (325.64,118.38) {$N^2$};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at (247.89,118.38) {$N$};
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (231.96,116.18) rectangle (411.55,185.23);
\end{scope}
\begin{scope}
\path[clip] (231.96, 41.62) rectangle (411.55,110.68);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (231.96, 41.62) rectangle (411.55,110.68);
\definecolor{drawColor}{gray}{0.92}

\path[draw=drawColor,line width= 0.3pt,line join=round] (231.96, 42.48) --
	(411.55, 42.48);

\path[draw=drawColor,line width= 0.3pt,line join=round] (231.96, 55.84) --
	(411.55, 55.84);

\path[draw=drawColor,line width= 0.3pt,line join=round] (231.96, 69.20) --
	(411.55, 69.20);

\path[draw=drawColor,line width= 0.3pt,line join=round] (231.96, 82.55) --
	(411.55, 82.55);

\path[draw=drawColor,line width= 0.3pt,line join=round] (231.96, 95.91) --
	(411.55, 95.91);

\path[draw=drawColor,line width= 0.3pt,line join=round] (231.96,109.27) --
	(411.55,109.27);

\path[draw=drawColor,line width= 0.3pt,line join=round] (237.48, 41.62) --
	(237.48,110.68);

\path[draw=drawColor,line width= 0.3pt,line join=round] (263.31, 41.62) --
	(263.31,110.68);

\path[draw=drawColor,line width= 0.3pt,line join=round] (289.14, 41.62) --
	(289.14,110.68);

\path[draw=drawColor,line width= 0.3pt,line join=round] (314.96, 41.62) --
	(314.96,110.68);

\path[draw=drawColor,line width= 0.3pt,line join=round] (340.79, 41.62) --
	(340.79,110.68);

\path[draw=drawColor,line width= 0.3pt,line join=round] (366.62, 41.62) --
	(366.62,110.68);

\path[draw=drawColor,line width= 0.3pt,line join=round] (392.44, 41.62) --
	(392.44,110.68);

\path[draw=drawColor,line width= 0.6pt,line join=round] (231.96, 49.16) --
	(411.55, 49.16);

\path[draw=drawColor,line width= 0.6pt,line join=round] (231.96, 62.52) --
	(411.55, 62.52);

\path[draw=drawColor,line width= 0.6pt,line join=round] (231.96, 75.88) --
	(411.55, 75.88);

\path[draw=drawColor,line width= 0.6pt,line join=round] (231.96, 89.23) --
	(411.55, 89.23);

\path[draw=drawColor,line width= 0.6pt,line join=round] (231.96,102.59) --
	(411.55,102.59);

\path[draw=drawColor,line width= 0.6pt,line join=round] (250.40, 41.62) --
	(250.40,110.68);

\path[draw=drawColor,line width= 0.6pt,line join=round] (276.22, 41.62) --
	(276.22,110.68);

\path[draw=drawColor,line width= 0.6pt,line join=round] (302.05, 41.62) --
	(302.05,110.68);

\path[draw=drawColor,line width= 0.6pt,line join=round] (327.88, 41.62) --
	(327.88,110.68);

\path[draw=drawColor,line width= 0.6pt,line join=round] (353.70, 41.62) --
	(353.70,110.68);

\path[draw=drawColor,line width= 0.6pt,line join=round] (379.53, 41.62) --
	(379.53,110.68);

\path[draw=drawColor,line width= 0.6pt,line join=round] (405.36, 41.62) --
	(405.36,110.68);
\definecolor{drawColor}{RGB}{190,190,190}

\path[draw=drawColor,line width= 0.6pt,line join=round] (231.96, 98.57) -- (411.55, 98.57);
\definecolor{fillColor}{RGB}{0,0,0}

\path[fill=fillColor,fill opacity=0.50] (240.12, 57.71) --
	(247.89, 53.91) --
	(255.67, 52.68) --
	(263.44, 63.57) --
	(271.22, 52.51) --
	(278.99, 49.84) --
	(286.77, 48.48) --
	(294.54, 49.49) --
	(302.32, 49.66) --
	(310.09, 52.74) --
	(317.87, 62.26) --
	(325.64, 63.29) --
	(333.41, 61.67) --
	(341.19, 66.65) --
	(348.96, 70.51) --
	(356.74, 73.89) --
	(364.51, 81.12) --
	(372.29, 82.11) --
	(380.06, 86.16) --
	(387.84, 90.86) --
	(395.61, 95.14) --
	(403.39, 99.71) --
	(403.39, 99.32) --
	(395.61, 94.61) --
	(387.84, 90.49) --
	(380.06, 85.91) --
	(372.29, 81.88) --
	(364.51, 77.39) --
	(356.74, 73.46) --
	(348.96, 69.25) --
	(341.19, 66.07) --
	(333.41, 61.08) --
	(325.64, 57.16) --
	(317.87, 54.48) --
	(310.09, 51.11) --
	(302.32, 48.90) --
	(294.54, 47.85) --
	(286.77, 47.05) --
	(278.99, 48.88) --
	(271.22, 51.55) --
	(263.44, 51.67) --
	(255.67, 51.49) --
	(247.89, 51.36) --
	(240.12, 52.74) --
	cycle;

\path[] (240.12, 57.71) --
	(247.89, 53.91) --
	(255.67, 52.68) --
	(263.44, 63.57) --
	(271.22, 52.51) --
	(278.99, 49.84) --
	(286.77, 48.48) --
	(294.54, 49.49) --
	(302.32, 49.66) --
	(310.09, 52.74) --
	(317.87, 62.26) --
	(325.64, 63.29) --
	(333.41, 61.67) --
	(341.19, 66.65) --
	(348.96, 70.51) --
	(356.74, 73.89) --
	(364.51, 81.12) --
	(372.29, 82.11) --
	(380.06, 86.16) --
	(387.84, 90.86) --
	(395.61, 95.14) --
	(403.39, 99.71);

\path[] (403.39, 99.32) --
	(395.61, 94.61) --
	(387.84, 90.49) --
	(380.06, 85.91) --
	(372.29, 81.88) --
	(364.51, 77.39) --
	(356.74, 73.46) --
	(348.96, 69.25) --
	(341.19, 66.07) --
	(333.41, 61.08) --
	(325.64, 57.16) --
	(317.87, 54.48) --
	(310.09, 51.11) --
	(302.32, 48.90) --
	(294.54, 47.85) --
	(286.77, 47.05) --
	(278.99, 48.88) --
	(271.22, 51.55) --
	(263.44, 51.67) --
	(255.67, 51.49) --
	(247.89, 51.36) --
	(240.12, 52.74);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 2.3pt,line join=round] (240.12, 55.09) --
	(247.89, 51.58) --
	(255.67, 51.81) --
	(263.44, 51.98) --
	(271.22, 51.79) --
	(278.99, 49.27) --
	(286.77, 47.55) --
	(294.54, 48.12) --
	(302.32, 49.04) --
	(310.09, 51.52) --
	(317.87, 54.89) --
	(325.64, 57.43) --
	(333.41, 61.36) --
	(341.19, 66.31) --
	(348.96, 69.59) --
	(356.74, 73.56) --
	(364.51, 77.55) --
	(372.29, 81.95) --
	(380.06, 86.08) --
	(387.84, 90.61) --
	(395.61, 94.71) --
	(403.39, 99.49);
\definecolor{drawColor}{RGB}{238,130,238}

\path[draw=drawColor,line width= 1.1pt,line join=round] (317.87, 51.47) --
	(325.64, 55.96) --
	(333.41, 60.41) --
	(341.19, 64.83) --
	(348.96, 69.23) --
	(356.74, 73.60) --
	(364.51, 77.96) --
	(372.29, 82.29) --
	(380.06, 86.61) --
	(387.84, 90.91) --
	(395.61, 95.21) --
	(403.39, 99.49);

\path[draw=drawColor,line width= 1.1pt,line join=round] (341.19, 51.23) --
	(348.96, 57.26) --
	(356.74, 63.29) --
	(364.51, 69.32) --
	(372.29, 75.36) --
	(380.06, 81.39) --
	(387.84, 87.42) --
	(395.61, 93.45) --
	(403.39, 99.49);

\path[draw=drawColor,line width= 1.1pt,line join=round] (356.74, 51.23) --
	(364.51, 59.27) --
	(372.29, 67.31) --
	(380.06, 75.36) --
	(387.84, 83.40) --
	(395.61, 91.44) --
	(403.39, 99.49);

\path[draw=drawColor,line width= 1.1pt,line join=round] (372.29, 51.23) --
	(380.06, 63.29) --
	(387.84, 75.36) --
	(395.61, 87.42) --
	(403.39, 99.49);
\end{scope}
\begin{scope}
\path[clip] (231.96, 41.62) rectangle (411.55,110.68);
\definecolor{drawColor}{RGB}{0,0,0}
\definecolor{fillColor}{RGB}{238,130,238}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (304.01, 48.38) --
	(317.87, 51.47) --
	(331.72, 48.38) --
	(331.72, 40.20) --
	(304.01, 40.20) --
	cycle;

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (332.23, 48.38) --
	(341.19, 51.23) --
	(349.47, 48.38) --
	(349.47, 40.20) --
	(332.23, 40.20) --
	cycle;

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (349.47, 48.38) --
	(356.74, 51.23) --
	(364.68, 48.38) --
	(364.68, 40.20) --
	(349.47, 40.20) --
	cycle;

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (364.68, 48.38) --
	(372.29, 51.23) --
	(379.90, 48.38) --
	(379.90, 40.20) --
	(364.68, 40.20) --
	cycle;

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.72] at (317.87, 43.42) {$N \log N$};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.72] at (340.85, 43.42) {$N^{1.5}$};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.91] at (357.07, 42.12) {$N^2$};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.91] at (372.29, 42.12) {$N^3$};
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (231.96, 41.62) rectangle (411.55,110.68);
\end{scope}
\begin{scope}
\path[clip] ( 46.86,185.23) rectangle (226.46,211.31);
\definecolor{drawColor}{gray}{0.20}
\definecolor{fillColor}{gray}{0.85}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] ( 46.86,185.23) rectangle (226.46,211.31);
\definecolor{drawColor}{gray}{0.10}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (136.66,199.99) {binsegRcpp};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (136.66,190.49) {list};
\end{scope}
\begin{scope}
\path[clip] (231.96,185.23) rectangle (411.55,211.31);
\definecolor{drawColor}{gray}{0.20}
\definecolor{fillColor}{gray}{0.85}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (231.96,185.23) rectangle (411.55,211.31);
\definecolor{drawColor}{gray}{0.10}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (321.75,199.99) {binsegRcpp};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (321.75,190.49) {multiset};
\end{scope}
\begin{scope}
\path[clip] (411.55,116.18) rectangle (428.12,185.23);
\definecolor{drawColor}{gray}{0.20}
\definecolor{fillColor}{gray}{0.85}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (411.55,116.18) rectangle (428.12,185.23);
\definecolor{drawColor}{gray}{0.10}

\node[text=drawColor,rotate=-90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (416.80,150.71) {kilobytes};
\end{scope}
\begin{scope}
\path[clip] (411.55, 41.62) rectangle (428.12,110.68);
\definecolor{drawColor}{gray}{0.20}
\definecolor{fillColor}{gray}{0.85}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (411.55, 41.62) rectangle (428.12,110.68);
\definecolor{drawColor}{gray}{0.10}

\node[text=drawColor,rotate=-90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (416.80, 76.15) {seconds};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (433.62,216.81);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 69.42, 38.87) --
	( 69.42, 41.62);

\path[draw=drawColor,line width= 0.6pt,line join=round] (105.57, 38.87) --
	(105.57, 41.62);

\path[draw=drawColor,line width= 0.6pt,line join=round] (141.73, 38.87) --
	(141.73, 41.62);

\path[draw=drawColor,line width= 0.6pt,line join=round] (177.89, 38.87) --
	(177.89, 41.62);

\path[draw=drawColor,line width= 0.6pt,line join=round] (214.04, 38.87) --
	(214.04, 41.62);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (433.62,216.81);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,rotate= 30.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 72.45, 31.42) {1e+01};

\node[text=drawColor,rotate= 30.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (108.60, 31.42) {1e+02};

\node[text=drawColor,rotate= 30.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (144.76, 31.42) {1e+03};

\node[text=drawColor,rotate= 30.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (180.92, 31.42) {1e+04};

\node[text=drawColor,rotate= 30.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (217.07, 31.42) {1e+05};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (433.62,216.81);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] (250.40, 38.87) --
	(250.40, 41.62);

\path[draw=drawColor,line width= 0.6pt,line join=round] (276.22, 38.87) --
	(276.22, 41.62);

\path[draw=drawColor,line width= 0.6pt,line join=round] (302.05, 38.87) --
	(302.05, 41.62);

\path[draw=drawColor,line width= 0.6pt,line join=round] (327.88, 38.87) --
	(327.88, 41.62);

\path[draw=drawColor,line width= 0.6pt,line join=round] (353.70, 38.87) --
	(353.70, 41.62);

\path[draw=drawColor,line width= 0.6pt,line join=round] (379.53, 38.87) --
	(379.53, 41.62);

\path[draw=drawColor,line width= 0.6pt,line join=round] (405.36, 38.87) --
	(405.36, 41.62);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (433.62,216.81);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,rotate= 30.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (253.43, 31.42) {1e+01};

\node[text=drawColor,rotate= 30.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (279.25, 31.42) {1e+02};

\node[text=drawColor,rotate= 30.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (305.08, 31.42) {1e+03};

\node[text=drawColor,rotate= 30.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (330.91, 31.42) {1e+04};

\node[text=drawColor,rotate= 30.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (356.73, 31.42) {1e+05};

\node[text=drawColor,rotate= 30.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (382.56, 31.42) {1e+06};

\node[text=drawColor,rotate= 30.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (408.39, 31.42) {1e+07};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (433.62,216.81);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 41.91,129.98) {1e+01};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 41.91,147.91) {1e+03};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 41.91,165.84) {1e+05};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (433.62,216.81);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 44.11,133.01) --
	( 46.86,133.01);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 44.11,150.94) --
	( 46.86,150.94);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 44.11,168.88) --
	( 46.86,168.88);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (433.62,216.81);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 41.91, 46.13) {1e-03};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 41.91, 59.49) {1e-02};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 41.91, 72.85) {1e-01};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 41.91, 86.20) {1e+00};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 41.91, 99.56) {1e+01};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (433.62,216.81);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 44.11, 49.16) --
	( 46.86, 49.16);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 44.11, 62.52) --
	( 46.86, 62.52);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 44.11, 75.88) --
	( 46.86, 75.88);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 44.11, 89.23) --
	( 46.86, 89.23);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 44.11,102.59) --
	( 46.86,102.59);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (433.62,216.81);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at (229.21,  7.64) {N};
\end{scope}
\end{tikzpicture}

### 2025-05-06 figure-heap-refs-other.tex
tikz figures

% Created by tikzDevice version 0.12.6 on 2025-05-06 20:54:35
% !TEX encoding = UTF-8 Unicode
\begin{tikzpicture}[x=1pt,y=1pt]
\definecolor{fillColor}{RGB}{255,255,255}
\path[use as bounding box,fill=fillColor,fill opacity=0.00] (0,0) rectangle (505.89,505.89);
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (505.89,505.89);
\definecolor{drawColor}{RGB}{255,255,255}
\definecolor{fillColor}{RGB}{255,255,255}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (  0.00,  0.00) rectangle (505.89,505.89);
\end{scope}
\begin{scope}
\path[clip] ( 46.86,264.98) rectangle (262.59,483.82);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] ( 46.86,264.98) rectangle (262.59,483.82);
\definecolor{drawColor}{gray}{0.92}

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 46.86,289.91) --
	(262.59,289.91);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 46.86,346.74) --
	(262.59,346.74);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 46.86,403.56) --
	(262.59,403.56);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 46.86,460.39) --
	(262.59,460.39);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 50.63,264.98) --
	( 50.63,483.82);

\path[draw=drawColor,line width= 0.3pt,line join=round] (109.85,264.98) --
	(109.85,483.82);

\path[draw=drawColor,line width= 0.3pt,line join=round] (169.08,264.98) --
	(169.08,483.82);

\path[draw=drawColor,line width= 0.3pt,line join=round] (228.30,264.98) --
	(228.30,483.82);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 46.86,318.33) --
	(262.59,318.33);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 46.86,375.15) --
	(262.59,375.15);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 46.86,431.98) --
	(262.59,431.98);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 80.24,264.98) --
	( 80.24,483.82);

\path[draw=drawColor,line width= 0.6pt,line join=round] (139.46,264.98) --
	(139.46,483.82);

\path[draw=drawColor,line width= 0.6pt,line join=round] (198.69,264.98) --
	(198.69,483.82);

\path[draw=drawColor,line width= 0.6pt,line join=round] (257.92,264.98) --
	(257.92,483.82);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 2.3pt,line join=round] ( 56.67,398.60) --
	( 74.50,286.74) --
	( 92.33,305.45) --
	(110.16,319.22) --
	(127.98,338.53) --
	(145.81,354.75) --
	(163.64,370.70) --
	(181.47,387.04) --
	(199.30,403.70) --
	(217.13,420.58) --
	(234.96,437.56) --
	(252.79,454.61);
\definecolor{drawColor}{RGB}{238,130,238}

\path[draw=drawColor,line width= 1.1pt,line join=round] ( 92.33,300.65) --
	(110.16,317.76) --
	(127.98,334.86) --
	(145.81,351.97) --
	(163.64,369.07) --
	(181.47,386.18) --
	(199.30,403.29) --
	(217.13,420.39) --
	(234.96,437.50) --
	(252.79,454.61);

\path[draw=drawColor,line width= 1.1pt,line join=round] ( 56.67,360.52) --
	( 74.50,369.07) --
	( 92.33,377.63) --
	(110.16,386.18) --
	(127.98,394.73) --
	(145.81,403.29) --
	(163.64,411.84) --
	(181.47,420.39) --
	(199.30,428.95) --
	(217.13,437.50) --
	(234.96,446.05) --
	(252.79,454.61);
\end{scope}
\begin{scope}
\path[clip] ( 46.86,264.98) rectangle (262.59,483.82);
\definecolor{drawColor}{RGB}{0,0,0}
\definecolor{fillColor}{RGB}{238,130,238}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] ( 84.10,297.80) --
	( 92.33,300.65) --
	(100.56,297.80) --
	(100.56,288.94) --
	( 84.10,288.94) --
	cycle;

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] ( 50.68,297.80) --
	( 56.67,360.52) --
	( 62.66,297.80) --
	( 62.66,288.94) --
	( 50.68,288.94) --
	cycle;

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at ( 92.33,290.92) {$N^2$};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at ( 56.67,290.92) {$N$};
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 46.86,264.98) rectangle (262.59,483.82);
\end{scope}
\begin{scope}
\path[clip] ( 46.86, 40.64) rectangle (262.59,259.48);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] ( 46.86, 40.64) rectangle (262.59,259.48);
\definecolor{drawColor}{gray}{0.92}

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 46.86, 45.62) --
	(262.59, 45.62);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 46.86, 88.17) --
	(262.59, 88.17);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 46.86,130.71) --
	(262.59,130.71);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 46.86,173.25) --
	(262.59,173.25);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 46.86,215.80) --
	(262.59,215.80);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 46.86,258.34) --
	(262.59,258.34);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 50.63, 40.64) --
	( 50.63,259.48);

\path[draw=drawColor,line width= 0.3pt,line join=round] (109.85, 40.64) --
	(109.85,259.48);

\path[draw=drawColor,line width= 0.3pt,line join=round] (169.08, 40.64) --
	(169.08,259.48);

\path[draw=drawColor,line width= 0.3pt,line join=round] (228.30, 40.64) --
	(228.30,259.48);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 46.86, 66.89) --
	(262.59, 66.89);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 46.86,109.44) --
	(262.59,109.44);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 46.86,151.98) --
	(262.59,151.98);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 46.86,194.52) --
	(262.59,194.52);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 46.86,237.07) --
	(262.59,237.07);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 80.24, 40.64) --
	( 80.24,259.48);

\path[draw=drawColor,line width= 0.6pt,line join=round] (139.46, 40.64) --
	(139.46,259.48);

\path[draw=drawColor,line width= 0.6pt,line join=round] (198.69, 40.64) --
	(198.69,259.48);

\path[draw=drawColor,line width= 0.6pt,line join=round] (257.92, 40.64) --
	(257.92,259.48);
\definecolor{drawColor}{RGB}{190,190,190}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 46.86,224.26) -- (262.59,224.26);
\definecolor{fillColor}{RGB}{0,0,0}

\path[fill=fillColor,fill opacity=0.50] ( 56.67, 66.89) --
	( 74.50, 78.17) --
	( 92.33, 71.00) --
	(110.16, 72.26) --
	(127.98, 68.86) --
	(145.81, 70.52) --
	(163.64, 78.50) --
	(181.47,112.13) --
	(199.30,141.19) --
	(217.13,167.33) --
	(234.96,193.27) --
	(252.79,228.18) --
	(252.79,227.94) --
	(234.96,192.75) --
	(217.13,155.33) --
	(199.30,123.25) --
	(181.47, 94.27) --
	(163.64, 70.55) --
	(145.81, 55.19) --
	(127.98, 53.93) --
	(110.16, 59.50) --
	( 92.33, 55.94) --
	( 74.50, 63.02) --
	( 56.67, 50.59) --
	cycle;

\path[] ( 56.67, 66.89) --
	( 74.50, 78.17) --
	( 92.33, 71.00) --
	(110.16, 72.26) --
	(127.98, 68.86) --
	(145.81, 70.52) --
	(163.64, 78.50) --
	(181.47,112.13) --
	(199.30,141.19) --
	(217.13,167.33) --
	(234.96,193.27) --
	(252.79,228.18);

\path[] (252.79,227.94) --
	(234.96,192.75) --
	(217.13,155.33) --
	(199.30,123.25) --
	(181.47, 94.27) --
	(163.64, 70.55) --
	(145.81, 55.19) --
	(127.98, 53.93) --
	(110.16, 59.50) --
	( 92.33, 55.94) --
	( 74.50, 63.02) --
	( 56.67, 50.59);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 2.3pt,line join=round] ( 56.67, 52.89) --
	( 74.50, 69.63) --
	( 92.33, 57.53) --
	(110.16, 60.50) --
	(127.98, 57.01) --
	(145.81, 58.85) --
	(163.64, 71.74) --
	(181.47, 94.98) --
	(199.30,125.79) --
	(217.13,156.20) --
	(234.96,192.85) --
	(252.79,228.05);
\definecolor{drawColor}{RGB}{238,130,238}

\path[draw=drawColor,line width= 1.1pt,line join=round] ( 74.50, 72.89) --
	( 92.33, 91.02) --
	(110.16,107.95) --
	(127.98,124.12) --
	(145.81,139.78) --
	(163.64,155.05) --
	(181.47,170.03) --
	(199.30,184.79) --
	(217.13,199.35) --
	(234.96,213.77) --
	(252.79,228.05);

\path[draw=drawColor,line width= 1.1pt,line join=round] (110.16, 74.37) --
	(127.98, 93.58) --
	(145.81,112.79) --
	(163.64,132.00) --
	(181.47,151.21) --
	(199.30,170.42) --
	(217.13,189.63) --
	(234.96,208.84) --
	(252.79,228.05);

\path[draw=drawColor,line width= 1.1pt,line join=round] (145.81, 74.37) --
	(163.64, 99.99) --
	(181.47,125.60) --
	(199.30,151.21) --
	(217.13,176.83) --
	(234.96,202.44) --
	(252.79,228.05);

\path[draw=drawColor,line width= 1.1pt,line join=round] (181.47, 74.37) --
	(199.30,112.79) --
	(217.13,151.21) --
	(234.96,189.63) --
	(252.79,228.05);
\end{scope}
\begin{scope}
\path[clip] ( 46.86, 40.64) rectangle (262.59,259.48);
\definecolor{drawColor}{RGB}{0,0,0}
\definecolor{fillColor}{RGB}{238,130,238}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] ( 55.83, 70.05) --
	( 74.50, 72.89) --
	( 93.17, 70.05) --
	( 93.17, 59.24) --
	( 55.83, 59.24) --
	cycle;

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] ( 98.75, 70.05) --
	(110.16, 74.37) --
	(121.56, 70.05) --
	(121.56, 59.24) --
	( 98.75, 59.24) --
	cycle;

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (137.59, 70.05) --
	(145.81, 74.37) --
	(154.04, 70.05) --
	(154.04, 61.18) --
	(137.59, 61.18) --
	cycle;

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (173.24, 70.05) --
	(181.47, 74.37) --
	(189.70, 70.05) --
	(189.70, 61.18) --
	(173.24, 61.18) --
	cycle;

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at ( 74.50, 63.16) {$N \log N$};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at (110.16, 63.16) {$N^{1.5}$};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at (145.81, 63.16) {$N^2$};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at (181.47, 63.16) {$N^3$};
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 46.86, 40.64) rectangle (262.59,259.48);
\end{scope}
\begin{scope}
\path[clip] (268.09,264.98) rectangle (483.82,483.82);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (268.09,264.98) rectangle (483.82,483.82);
\definecolor{drawColor}{gray}{0.92}

\path[draw=drawColor,line width= 0.3pt,line join=round] (268.09,289.91) --
	(483.82,289.91);

\path[draw=drawColor,line width= 0.3pt,line join=round] (268.09,346.74) --
	(483.82,346.74);

\path[draw=drawColor,line width= 0.3pt,line join=round] (268.09,403.56) --
	(483.82,403.56);

\path[draw=drawColor,line width= 0.3pt,line join=round] (268.09,460.39) --
	(483.82,460.39);

\path[draw=drawColor,line width= 0.3pt,line join=round] (272.36,264.98) --
	(272.36,483.82);

\path[draw=drawColor,line width= 0.3pt,line join=round] (326.65,264.98) --
	(326.65,483.82);

\path[draw=drawColor,line width= 0.3pt,line join=round] (380.94,264.98) --
	(380.94,483.82);

\path[draw=drawColor,line width= 0.3pt,line join=round] (435.23,264.98) --
	(435.23,483.82);

\path[draw=drawColor,line width= 0.6pt,line join=round] (268.09,318.33) --
	(483.82,318.33);

\path[draw=drawColor,line width= 0.6pt,line join=round] (268.09,375.15) --
	(483.82,375.15);

\path[draw=drawColor,line width= 0.6pt,line join=round] (268.09,431.98) --
	(483.82,431.98);

\path[draw=drawColor,line width= 0.6pt,line join=round] (299.50,264.98) --
	(299.50,483.82);

\path[draw=drawColor,line width= 0.6pt,line join=round] (353.79,264.98) --
	(353.79,483.82);

\path[draw=drawColor,line width= 0.6pt,line join=round] (408.08,264.98) --
	(408.08,483.82);

\path[draw=drawColor,line width= 0.6pt,line join=round] (462.37,264.98) --
	(462.37,483.82);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 2.3pt,line join=round] (277.90,302.81) --
	(294.24,264.98) --
	(310.58,264.98) --
	(326.93,264.98) --
	(343.27,264.98) --
	(359.61,274.93) --
	(375.95,282.47) --
	(392.30,290.48) --
	(408.64,298.75) --
	(424.98,307.16) --
	(441.33,315.64) --
	(457.67,324.16) --
	(474.01,332.70);
\definecolor{drawColor}{RGB}{238,130,238}

\path[draw=drawColor,line width= 1.1pt,line join=round] (441.33,298.48) --
	(457.67,315.59) --
	(474.01,332.70);

\path[draw=drawColor,line width= 1.1pt,line join=round] (408.64,298.48) --
	(424.98,307.04) --
	(441.33,315.59) --
	(457.67,324.14) --
	(474.01,332.70);
\end{scope}
\begin{scope}
\path[clip] (268.09,264.98) rectangle (483.82,483.82);
\definecolor{drawColor}{RGB}{0,0,0}
\definecolor{fillColor}{RGB}{238,130,238}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (433.10,295.64) --
	(441.33,298.48) --
	(449.55,295.64) --
	(449.55,286.77) --
	(433.10,286.77) --
	cycle;

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (402.66,295.64) --
	(408.64,298.48) --
	(414.63,295.64) --
	(414.63,286.77) --
	(402.66,286.77) --
	cycle;

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at (441.33,288.75) {$N^2$};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at (408.64,288.75) {$N$};
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (268.09,264.98) rectangle (483.82,483.82);
\end{scope}
\begin{scope}
\path[clip] (268.09, 40.64) rectangle (483.82,259.48);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (268.09, 40.64) rectangle (483.82,259.48);
\definecolor{drawColor}{gray}{0.92}

\path[draw=drawColor,line width= 0.3pt,line join=round] (268.09, 45.62) --
	(483.82, 45.62);

\path[draw=drawColor,line width= 0.3pt,line join=round] (268.09, 88.17) --
	(483.82, 88.17);

\path[draw=drawColor,line width= 0.3pt,line join=round] (268.09,130.71) --
	(483.82,130.71);

\path[draw=drawColor,line width= 0.3pt,line join=round] (268.09,173.25) --
	(483.82,173.25);

\path[draw=drawColor,line width= 0.3pt,line join=round] (268.09,215.80) --
	(483.82,215.80);

\path[draw=drawColor,line width= 0.3pt,line join=round] (268.09,258.34) --
	(483.82,258.34);

\path[draw=drawColor,line width= 0.3pt,line join=round] (272.36, 40.64) --
	(272.36,259.48);

\path[draw=drawColor,line width= 0.3pt,line join=round] (326.65, 40.64) --
	(326.65,259.48);

\path[draw=drawColor,line width= 0.3pt,line join=round] (380.94, 40.64) --
	(380.94,259.48);

\path[draw=drawColor,line width= 0.3pt,line join=round] (435.23, 40.64) --
	(435.23,259.48);

\path[draw=drawColor,line width= 0.6pt,line join=round] (268.09, 66.89) --
	(483.82, 66.89);

\path[draw=drawColor,line width= 0.6pt,line join=round] (268.09,109.44) --
	(483.82,109.44);

\path[draw=drawColor,line width= 0.6pt,line join=round] (268.09,151.98) --
	(483.82,151.98);

\path[draw=drawColor,line width= 0.6pt,line join=round] (268.09,194.52) --
	(483.82,194.52);

\path[draw=drawColor,line width= 0.6pt,line join=round] (268.09,237.07) --
	(483.82,237.07);

\path[draw=drawColor,line width= 0.6pt,line join=round] (299.50, 40.64) --
	(299.50,259.48);

\path[draw=drawColor,line width= 0.6pt,line join=round] (353.79, 40.64) --
	(353.79,259.48);

\path[draw=drawColor,line width= 0.6pt,line join=round] (408.08, 40.64) --
	(408.08,259.48);

\path[draw=drawColor,line width= 0.6pt,line join=round] (462.37, 40.64) --
	(462.37,259.48);
\definecolor{drawColor}{RGB}{190,190,190}

\path[draw=drawColor,line width= 0.6pt,line join=round] (268.09,224.26) -- (483.82,224.26);
\definecolor{fillColor}{RGB}{0,0,0}

\path[fill=fillColor,fill opacity=0.50] (277.90, 74.10) --
	(294.24, 88.49) --
	(310.58, 89.28) --
	(326.93,102.21) --
	(343.27,116.54) --
	(359.61,120.50) --
	(375.95,130.05) --
	(392.30,145.66) --
	(408.64,161.32) --
	(424.98,177.20) --
	(441.33,203.86) --
	(457.67,213.34) --
	(474.01,234.51) --
	(474.01,233.63) --
	(457.67,212.74) --
	(441.33,193.95) --
	(424.98,176.96) --
	(408.64,160.86) --
	(392.30,145.25) --
	(375.95,129.73) --
	(359.61,114.31) --
	(343.27,107.68) --
	(326.93,100.18) --
	(310.58, 84.31) --
	(294.24, 73.17) --
	(277.90, 56.40) --
	cycle;

\path[] (277.90, 74.10) --
	(294.24, 88.49) --
	(310.58, 89.28) --
	(326.93,102.21) --
	(343.27,116.54) --
	(359.61,120.50) --
	(375.95,130.05) --
	(392.30,145.66) --
	(408.64,161.32) --
	(424.98,177.20) --
	(441.33,203.86) --
	(457.67,213.34) --
	(474.01,234.51);

\path[] (474.01,233.63) --
	(457.67,212.74) --
	(441.33,193.95) --
	(424.98,176.96) --
	(408.64,160.86) --
	(392.30,145.25) --
	(375.95,129.73) --
	(359.61,114.31) --
	(343.27,107.68) --
	(326.93,100.18) --
	(310.58, 84.31) --
	(294.24, 73.17) --
	(277.90, 56.40);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 2.3pt,line join=round] (277.90, 57.44) --
	(294.24, 74.95) --
	(310.58, 84.63) --
	(326.93,100.29) --
	(343.27,112.63) --
	(359.61,114.70) --
	(375.95,129.87) --
	(392.30,145.44) --
	(408.64,160.99) --
	(424.98,177.03) --
	(441.33,195.14) --
	(457.67,212.79) --
	(474.01,234.06);
\definecolor{drawColor}{RGB}{238,130,238}

\path[draw=drawColor,line width= 1.1pt,line join=round] (310.58, 82.84) --
	(326.93, 99.77) --
	(343.27,115.95) --
	(359.61,131.60) --
	(375.95,146.87) --
	(392.30,161.86) --
	(408.64,176.61) --
	(424.98,191.18) --
	(441.33,205.59) --
	(457.67,219.88) --
	(474.01,234.06);

\path[draw=drawColor,line width= 1.1pt,line join=round] (343.27, 80.37) --
	(359.61, 99.58) --
	(375.95,118.79) --
	(392.30,138.00) --
	(408.64,157.21) --
	(424.98,176.42) --
	(441.33,195.63) --
	(457.67,214.85) --
	(474.01,234.06);

\path[draw=drawColor,line width= 1.1pt,line join=round] (375.95, 80.37) --
	(392.30,105.99) --
	(408.64,131.60) --
	(424.98,157.21) --
	(441.33,182.83) --
	(457.67,208.44) --
	(474.01,234.06);

\path[draw=drawColor,line width= 1.1pt,line join=round] (408.64, 80.37) --
	(424.98,118.79) --
	(441.33,157.21) --
	(457.67,195.63) --
	(474.01,234.06);
\end{scope}
\begin{scope}
\path[clip] (268.09, 40.64) rectangle (483.82,259.48);
\definecolor{drawColor}{RGB}{0,0,0}
\definecolor{fillColor}{RGB}{238,130,238}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (291.91, 77.53) --
	(310.58, 82.84) --
	(329.26, 77.53) --
	(329.26, 66.72) --
	(291.91, 66.72) --
	cycle;

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (331.86, 77.53) --
	(343.27, 80.37) --
	(354.68, 77.53) --
	(354.68, 66.72) --
	(331.86, 66.72) --
	cycle;

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (367.73, 77.53) --
	(375.95, 80.37) --
	(384.18, 77.53) --
	(384.18, 68.66) --
	(367.73, 68.66) --
	cycle;

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (400.41, 77.53) --
	(408.64, 80.37) --
	(416.87, 77.53) --
	(416.87, 68.66) --
	(400.41, 68.66) --
	cycle;

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at (310.58, 70.64) {$N \log N$};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at (343.27, 70.64) {$N^{1.5}$};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at (375.95, 70.64) {$N^2$};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at (408.64, 70.64) {$N^3$};
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (268.09, 40.64) rectangle (483.82,259.48);
\end{scope}
\begin{scope}
\path[clip] ( 46.86,483.82) rectangle (262.59,500.39);
\definecolor{drawColor}{gray}{0.20}
\definecolor{fillColor}{gray}{0.85}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] ( 46.86,483.82) rectangle (262.59,500.39);
\definecolor{drawColor}{gray}{0.10}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (154.73,489.07) {changepoint};
\end{scope}
\begin{scope}
\path[clip] (268.09,483.82) rectangle (483.82,500.39);
\definecolor{drawColor}{gray}{0.20}
\definecolor{fillColor}{gray}{0.85}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (268.09,483.82) rectangle (483.82,500.39);
\definecolor{drawColor}{gray}{0.10}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (375.95,489.07) {ruptures};
\end{scope}
\begin{scope}
\path[clip] (483.82,264.98) rectangle (500.39,483.82);
\definecolor{drawColor}{gray}{0.20}
\definecolor{fillColor}{gray}{0.85}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (483.82,264.98) rectangle (500.39,483.82);
\definecolor{drawColor}{gray}{0.10}

\node[text=drawColor,rotate=-90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (489.07,374.40) {kilobytes};
\end{scope}
\begin{scope}
\path[clip] (483.82, 40.64) rectangle (500.39,259.48);
\definecolor{drawColor}{gray}{0.20}
\definecolor{fillColor}{gray}{0.85}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (483.82, 40.64) rectangle (500.39,259.48);
\definecolor{drawColor}{gray}{0.10}

\node[text=drawColor,rotate=-90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (489.07,150.06) {seconds};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (505.89,505.89);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 80.24, 37.89) --
	( 80.24, 40.64);

\path[draw=drawColor,line width= 0.6pt,line join=round] (139.46, 37.89) --
	(139.46, 40.64);

\path[draw=drawColor,line width= 0.6pt,line join=round] (198.69, 37.89) --
	(198.69, 40.64);

\path[draw=drawColor,line width= 0.6pt,line join=round] (257.92, 37.89) --
	(257.92, 40.64);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (505.89,505.89);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,rotate= 30.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 83.27, 30.44) {10};

\node[text=drawColor,rotate= 30.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (142.49, 30.44) {100};

\node[text=drawColor,rotate= 30.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (201.72, 30.44) {1000};

\node[text=drawColor,rotate= 30.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (260.95, 30.44) {10000};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (505.89,505.89);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] (299.50, 37.89) --
	(299.50, 40.64);

\path[draw=drawColor,line width= 0.6pt,line join=round] (353.79, 37.89) --
	(353.79, 40.64);

\path[draw=drawColor,line width= 0.6pt,line join=round] (408.08, 37.89) --
	(408.08, 40.64);

\path[draw=drawColor,line width= 0.6pt,line join=round] (462.37, 37.89) --
	(462.37, 40.64);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (505.89,505.89);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,rotate= 30.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (302.53, 30.44) {10};

\node[text=drawColor,rotate= 30.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (356.82, 30.44) {100};

\node[text=drawColor,rotate= 30.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (411.11, 30.44) {1000};

\node[text=drawColor,rotate= 30.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (465.40, 30.44) {10000};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (505.89,505.89);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 41.91,315.30) {1e+01};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 41.91,372.12) {1e+03};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 41.91,428.95) {1e+05};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (505.89,505.89);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 44.11,318.33) --
	( 46.86,318.33);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 44.11,375.15) --
	( 46.86,375.15);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 44.11,431.98) --
	( 46.86,431.98);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (505.89,505.89);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 41.91, 63.86) {1e-03};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 41.91,106.41) {1e-02};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 41.91,148.95) {1e-01};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 41.91,191.49) {1e+00};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 41.91,234.04) {1e+01};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (505.89,505.89);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 44.11, 66.89) --
	( 46.86, 66.89);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 44.11,109.44) --
	( 46.86,109.44);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 44.11,151.98) --
	( 46.86,151.98);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 44.11,194.52) --
	( 46.86,194.52);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 44.11,237.07) --
	( 46.86,237.07);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (505.89,505.89);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at (265.34,  7.64) {N};
\end{scope}
\end{tikzpicture}

### 2025-05-06 figure-heap.R

@@ -33,6 +33,29 @@ png("figure-heap-refs-other.png",width=6,height=3.5,units="in",res=200)
 plot_algos("change|rupt")
 dev.off()
 
+tikz_algos <- function(regex){
+  regex_list <- ref_list
+  regex_list$measurements <- ref_list$measurements[grepl(regex,expr.name)]
+  regex_list$plot.references <- ref_list$plot.references[grepl(regex,expr.name)][
+   ,fun.name := sprintf("$%s$", sub("1.5","{1.5}",fun.latex))][]
+  plot(regex_list)+
+    geom_blank(aes(
+      N, empirical),
+      data=ref_list$measurements[,.(N=1000,empirical,unit)])+
+    scale_x_log10(
+      breaks=10^seq(1,7,by=1))+
+    theme(
+      axis.text.x=element_text(angle=30,hjust=1))
+}
+library(tikzDevice)
+tikz("figure-heap-refs-binsegRcpp.tex",width=6,height=3)
+tikz_algos("list|multiset")
+dev.off()
+
+tikz("figure-heap-refs-other.tex")
+tikz_algos("change|rupt")
+dev.off()
+
 pred_list <- predict(ref_list)
 png("figure-heap-pred.png",width=8,height=3,units="in",res=200)
 pkg.colors <- c(

### 2025-05-06 figure-heap-refs-binsegRcpp.tex
direct label in tikz

@@
-% Created by tikzDevice version 0.12.6 on 2025-05-06 20:55:50
+% Created by tikzDevice version 0.12.6 on 2025-05-06 22:53:29
 % !TEX encoding = UTF-8 Unicode
@@
 \path[clip] ( 46.86,116.18) rectangle (226.46,185.23);
-\definecolor{fillColor}{RGB}{255,255,255}
+\definecolor{fillColor}{gray}{0.92}
 
 \path[fill=fillColor] ( 46.86,116.18) rectangle (226.46,185.23);
-\definecolor{drawColor}{gray}{0.92}
+\definecolor{drawColor}{RGB}{255,255,255}
 
-\path[draw=drawColor,line width= 0.3pt,line join=round] ( 46.86,124.04) --
-	(226.46,124.04);
+\path[draw=drawColor,line width= 0.3pt,line join=round] ( 46.86,121.16) --
+	(226.46,121.16);
 
-\path[draw=drawColor,line width= 0.3pt,line join=round] ( 46.86,141.98) --
-	(226.46,141.98);
+\path[draw=drawColor,line width= 0.3pt,line join=round] ( 46.86,137.47) --
+	(226.46,137.47);
 
-\path[draw=drawColor,line width= 0.3pt,line join=round] ( 46.86,159.91) --
-	(226.46,159.91);
+\path[draw=drawColor,line width= 0.3pt,line join=round] ( 46.86,153.77) --
+	(226.46,153.77);
 
-\path[draw=drawColor,line width= 0.3pt,line join=round] ( 46.86,177.84) --
-	(226.46,177.84);
+\path[draw=drawColor,line width= 0.3pt,line join=round] ( 46.86,170.08) --
+	(226.46,170.08);
 
@@
 
-\path[draw=drawColor,line width= 0.6pt,line join=round] ( 46.86,133.01) --
-	(226.46,133.01);
+\path[draw=drawColor,line width= 0.6pt,line join=round] ( 46.86,129.31) --
+	(226.46,129.31);
 
-\path[draw=drawColor,line width= 0.6pt,line join=round] ( 46.86,150.94) --
-	(226.46,150.94);
+\path[draw=drawColor,line width= 0.6pt,line join=round] ( 46.86,145.62) --
+	(226.46,145.62);
 
-\path[draw=drawColor,line width= 0.6pt,line join=round] ( 46.86,168.88) --
-	(226.46,168.88);
+\path[draw=drawColor,line width= 0.6pt,line join=round] ( 46.86,161.92) --
+	(226.46,161.92);
+
+\path[draw=drawColor,line width= 0.6pt,line join=round] ( 46.86,178.23) --
+	(226.46,178.23);
 
@@
 
-\path[draw=drawColor,line width= 2.3pt,line join=round] ( 55.03,147.52) --
-	( 65.91,140.30) --
-	( 76.80,140.31) --
-	( 87.68,140.47) --
-	( 98.56,141.29) --
-	(109.45,142.56) --
-	(120.33,143.85) --
-	(131.22,145.59) --
-	(142.10,147.70) --
-	(152.99,150.07) --
-	(163.87,152.59) --
-	(174.76,155.20) --
-	(185.64,157.85) --
-	(196.52,160.53) --
-	(207.41,163.21) --
-	(218.29,165.91);
+\path[draw=drawColor,line width= 2.3pt,line join=round] ( 55.03,150.65) --
+	( 65.91,144.09) --
+	( 76.80,144.10) --
+	( 87.68,144.24) --
+	( 98.56,144.99) --
+	(109.45,146.15) --
+	(120.33,147.32) --
+	(131.22,148.90) --
+	(142.10,150.82) --
+	(152.99,152.98) --
+	(163.87,155.27) --
+	(174.76,157.64) --
+	(185.64,160.05) --
+	(196.52,162.48) --
+	(207.41,164.93) --
+	(218.29,167.38);
 \definecolor{drawColor}{RGB}{238,130,238}
 
-\path[draw=drawColor,line width= 1.1pt,line join=round] (142.10,128.12) --
-	(152.99,133.52) --
-	(163.87,138.92) --
-	(174.76,144.31) --
-	(185.64,149.71) --
-	(196.52,155.11) --
-	(207.41,160.51) --
-	(218.29,165.91);
+\path[draw=drawColor,line width= 1.1pt,line join=round] (142.10,133.02) --
+	(152.99,137.93) --
+	(163.87,142.84) --
+	(174.76,147.74) --
+	(185.64,152.65) --
+	(196.52,157.56) --
+	(207.41,162.47) --
+	(218.29,167.38);
 
-\path[draw=drawColor,line width= 1.1pt,line join=round] ( 65.91,128.12) --
-	( 76.80,130.82) --
-	( 87.68,133.52) --
-	( 98.56,136.22) --
-	(109.45,138.92) --
-	(120.33,141.62) --
-	(131.22,144.31) --
-	(142.10,147.01) --
-	(152.99,149.71) --
-	(163.87,152.41) --
-	(174.76,155.11) --
-	(185.64,157.81) --
-	(196.52,160.51) --
-	(207.41,163.21) --
-	(218.29,165.91);
+\path[draw=drawColor,line width= 1.1pt,line join=round] ( 65.91,133.02) --
+	( 76.80,135.47) --
+	( 87.68,137.93) --
+	( 98.56,140.38) --
+	(109.45,142.84) --
+	(120.33,145.29) --
+	(131.22,147.74) --
+	(142.10,150.20) --
+	(152.99,152.65) --
+	(163.87,155.11) --
+	(174.76,157.56) --
+	(185.64,160.01) --
+	(196.52,162.47) --
+	(207.41,164.92) --
+	(218.29,167.38);
 \end{scope}
@@
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (133.87,125.28) --
-	(142.10,128.12) --
-	(150.33,125.28) --
-	(150.33,116.41) --
-	(133.87,116.41) --
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (133.87,130.17) --
+	(142.10,133.02) --
+	(150.33,130.17) --
+	(150.33,121.31) --
+	(133.87,121.31) --
 	cycle;
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] ( 59.93,125.28) --
-	( 65.91,128.12) --
-	( 71.90,125.28) --
-	( 71.90,116.41) --
-	( 59.93,116.41) --
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] ( 59.93,130.17) --
+	( 65.91,133.02) --
+	( 71.90,130.17) --
+	( 71.90,121.31) --
+	( 59.93,121.31) --
 	cycle;
 
-\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at (142.10,118.39) {$N^2$};
+\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at (142.10,121.86) {$N^2$};
 
-\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at ( 65.91,118.39) {$N$};
-\definecolor{drawColor}{gray}{0.20}
-
-\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 46.86,116.18) rectangle (226.46,185.23);
+\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at ( 65.91,121.86) {$N$};
 \end{scope}
@@
 \path[clip] ( 46.86, 41.62) rectangle (226.46,110.68);
-\definecolor{fillColor}{RGB}{255,255,255}
+\definecolor{fillColor}{gray}{0.92}
 
 \path[fill=fillColor] ( 46.86, 41.62) rectangle (226.46,110.68);
-\definecolor{drawColor}{gray}{0.92}
+\definecolor{drawColor}{RGB}{255,255,255}
 
-\path[draw=drawColor,line width= 0.3pt,line join=round] ( 46.86, 42.48) --
-	(226.46, 42.48);
+\path[draw=drawColor,line width= 0.3pt,line join=round] ( 46.86, 45.09) --
+	(226.46, 45.09);
 
-\path[draw=drawColor,line width= 0.3pt,line join=round] ( 46.86, 55.84) --
-	(226.46, 55.84);
+\path[draw=drawColor,line width= 0.3pt,line join=round] ( 46.86, 68.35) --
+	(226.46, 68.35);
 
-\path[draw=drawColor,line width= 0.3pt,line join=round] ( 46.86, 69.20) --
-	(226.46, 69.20);
-
-\path[draw=drawColor,line width= 0.3pt,line join=round] ( 46.86, 82.55) --
-	(226.46, 82.55);
-
-\path[draw=drawColor,line width= 0.3pt,line join=round] ( 46.86, 95.91) --
-	(226.46, 95.91);
-
-\path[draw=drawColor,line width= 0.3pt,line join=round] ( 46.86,109.27) --
-	(226.46,109.27);
+\path[draw=drawColor,line width= 0.3pt,line join=round] ( 46.86, 91.60) --
+	(226.46, 91.60);
 
@@
 
-\path[draw=drawColor,line width= 0.6pt,line join=round] ( 46.86, 49.16) --
-	(226.46, 49.16);
+\path[draw=drawColor,line width= 0.6pt,line join=round] ( 46.86, 56.72) --
+	(226.46, 56.72);
 
-\path[draw=drawColor,line width= 0.6pt,line join=round] ( 46.86, 62.52) --
-	(226.46, 62.52);
+\path[draw=drawColor,line width= 0.6pt,line join=round] ( 46.86, 79.97) --
+	(226.46, 79.97);
 
-\path[draw=drawColor,line width= 0.6pt,line join=round] ( 46.86, 75.88) --
-	(226.46, 75.88);
-
-\path[draw=drawColor,line width= 0.6pt,line join=round] ( 46.86, 89.23) --
-	(226.46, 89.23);
-
-\path[draw=drawColor,line width= 0.6pt,line join=round] ( 46.86,102.59) --
-	(226.46,102.59);
+\path[draw=drawColor,line width= 0.6pt,line join=round] ( 46.86,103.23) --
+	(226.46,103.23);
 
@@
 	(214.04,110.68);
-\definecolor{drawColor}{RGB}{190,190,190}
-
-\path[draw=drawColor,line width= 0.6pt,line join=round] ( 46.86, 98.57) -- (226.46, 98.57);
 \definecolor{fillColor}{RGB}{0,0,0}
 
-\path[fill=fillColor,fill opacity=0.50] ( 55.03, 53.39) --
-	( 65.91, 53.70) --
-	( 76.80, 53.42) --
-	( 87.68, 53.62) --
-	( 98.56, 53.81) --
-	(109.45, 52.36) --
-	(120.33, 60.85) --
-	(131.22, 50.50) --
-	(142.10, 53.19) --
-	(152.99, 58.46) --
-	(163.87, 65.47) --
-	(174.76, 72.35) --
-	(185.64, 79.70) --
-	(196.52, 88.57) --
-	(207.41, 97.32) --
+\path[fill=fillColor,fill opacity=0.50] ( 55.03, 60.40) --
+	( 65.91, 60.67) --
+	( 76.80, 60.43) --
+	( 87.68, 60.60) --
+	( 98.56, 60.76) --
+	(109.45, 59.50) --
+	(120.33, 66.90) --
+	(131.22, 57.88) --
+	(142.10, 60.23) --
+	(152.99, 64.81) --
+	(163.87, 70.91) --
+	(174.76, 76.91) --
+	(185.64, 83.30) --
+	(196.52, 91.03) --
+	(207.41, 98.65) --
 	(218.29,107.54) --
-	(218.29,106.40) --
-	(207.41, 97.20) --
-	(196.52, 88.37) --
-	(185.64, 79.22) --
-	(174.76, 72.10) --
-	(163.87, 65.19) --
-	(152.99, 58.08) --
-	(142.10, 51.77) --
-	(131.22, 48.59) --
-	(120.33, 47.68) --
-	(109.45, 48.68) --
-	( 98.56, 51.61) --
-	( 87.68, 51.71) --
-	( 76.80, 51.86) --
-	( 65.91, 51.48) --
-	( 55.03, 51.51) --
+	(218.29,106.55) --
+	(207.41, 98.54) --
+	(196.52, 90.85) --
+	(185.64, 82.89) --
+	(174.76, 76.69) --
+	(163.87, 70.67) --
+	(152.99, 64.48) --
+	(142.10, 58.99) --
+	(131.22, 56.22) --
+	(120.33, 55.43) --
+	(109.45, 56.30) --
+	( 98.56, 58.85) --
+	( 87.68, 58.94) --
+	( 76.80, 59.07) --
+	( 65.91, 58.74) --
+	( 55.03, 58.77) --
 	cycle;
 
-\path[] ( 55.03, 53.39) --
-	( 65.91, 53.70) --
-	( 76.80, 53.42) --
-	( 87.68, 53.62) --
-	( 98.56, 53.81) --
-	(109.45, 52.36) --
-	(120.33, 60.85) --
-	(131.22, 50.50) --
-	(142.10, 53.19) --
-	(152.99, 58.46) --
-	(163.87, 65.47) --
-	(174.76, 72.35) --
-	(185.64, 79.70) --
-	(196.52, 88.57) --
-	(207.41, 97.32) --
+\path[] ( 55.03, 60.40) --
+	( 65.91, 60.67) --
+	( 76.80, 60.43) --
+	( 87.68, 60.60) --
+	( 98.56, 60.76) --
+	(109.45, 59.50) --
+	(120.33, 66.90) --
+	(131.22, 57.88) --
+	(142.10, 60.23) --
+	(152.99, 64.81) --
+	(163.87, 70.91) --
+	(174.76, 76.91) --
+	(185.64, 83.30) --
+	(196.52, 91.03) --
+	(207.41, 98.65) --
 	(218.29,107.54);
 
-\path[] (218.29,106.40) --
-	(207.41, 97.20) --
-	(196.52, 88.37) --
-	(185.64, 79.22) --
-	(174.76, 72.10) --
-	(163.87, 65.19) --
-	(152.99, 58.08) --
-	(142.10, 51.77) --
-	(131.22, 48.59) --
-	(120.33, 47.68) --
-	(109.45, 48.68) --
-	( 98.56, 51.61) --
-	( 87.68, 51.71) --
-	( 76.80, 51.86) --
-	( 65.91, 51.48) --
-	( 55.03, 51.51);
+\path[] (218.29,106.55) --
+	(207.41, 98.54) --
+	(196.52, 90.85) --
+	(185.64, 82.89) --
+	(174.76, 76.69) --
+	(163.87, 70.67) --
+	(152.99, 64.48) --
+	(142.10, 58.99) --
+	(131.22, 56.22) --
+	(120.33, 55.43) --
+	(109.45, 56.30) --
+	( 98.56, 58.85) --
+	( 87.68, 58.94) --
+	( 76.80, 59.07) --
+	( 65.91, 58.74) --
+	( 55.03, 58.77);
 \definecolor{drawColor}{RGB}{0,0,0}
 
-\path[draw=drawColor,line width= 2.3pt,line join=round] ( 55.03, 51.80) --
-	( 65.91, 51.96) --
-	( 76.80, 52.23) --
-	( 87.68, 51.97) --
-	( 98.56, 51.99) --
-	(109.45, 49.61) --
-	(120.33, 48.49) --
-	(131.22, 49.12) --
-	(142.10, 52.46) --
-	(152.99, 58.30) --
-	(163.87, 65.29) --
-	(174.76, 72.20) --
-	(185.64, 79.51) --
-	(196.52, 88.44) --
-	(207.41, 97.24) --
-	(218.29,106.51);
+\path[draw=drawColor,line width= 2.3pt,line join=round] ( 55.03, 59.02) --
+	( 65.91, 59.16) --
+	( 76.80, 59.39) --
+	( 87.68, 59.17) --
+	( 98.56, 59.18) --
+	(109.45, 57.11) --
+	(120.33, 56.13) --
+	(131.22, 56.68) --
+	(142.10, 59.59) --
+	(152.99, 64.67) --
+	(163.87, 70.76) --
+	(174.76, 76.78) --
+	(185.64, 83.14) --
+	(196.52, 90.91) --
+	(207.41, 98.58) --
+	(218.29,106.64);
 \definecolor{drawColor}{RGB}{238,130,238}
 
-\path[draw=drawColor,line width= 1.1pt,line join=round] ( 87.68, 51.15) --
-	( 98.56, 56.23) --
-	(109.45, 61.15) --
-	(120.33, 65.94) --
-	(131.22, 70.65) --
-	(142.10, 75.28) --
-	(152.99, 79.85) --
-	(163.87, 84.38) --
-	(174.76, 88.87) --
-	(185.64, 93.32) --
-	(196.52, 97.74) --
-	(207.41,102.13) --
-	(218.29,106.51);
+\path[draw=drawColor,line width= 1.1pt,line join=round] ( 87.68, 58.45) --
+	( 98.56, 62.87) --
+	(109.45, 67.15) --
+	(120.33, 71.33) --
+	(131.22, 75.42) --
+	(142.10, 79.46) --
+	(152.99, 83.44) --
+	(163.87, 87.38) --
+	(174.76, 91.28) --
+	(185.64, 95.16) --
+	(196.52, 99.01) --
+	(207.41,102.83) --
+	(218.29,106.64);
 
-\path[draw=drawColor,line width= 1.1pt,line join=round] (120.33, 52.22) --
-	(131.22, 58.25) --
-	(142.10, 64.28) --
-	(152.99, 70.31) --
-	(163.87, 76.35) --
-	(174.76, 82.38) --
-	(185.64, 88.41) --
-	(196.52, 94.44) --
-	(207.41,100.48) --
-	(218.29,106.51);
-
-\path[draw=drawColor,line width= 1.1pt,line join=round] (152.99, 58.25) --
-	(163.87, 66.29) --
-	(174.76, 74.34) --
-	(185.64, 82.38) --
-	(196.52, 90.42) --
-	(207.41, 98.46) --
-	(218.29,106.51);
-
-\path[draw=drawColor,line width= 1.1pt,line join=round] (174.76, 58.25) --
-	(185.64, 70.31) --
-	(196.52, 82.38) --
-	(207.41, 94.44) --
-	(218.29,106.51);
+\path[draw=drawColor,line width= 1.1pt,line join=round] (152.99, 64.63) --
+	(163.87, 71.63) --
+	(174.76, 78.64) --
+	(185.64, 85.64) --
+	(196.52, 92.64) --
+	(207.41, 99.64) --
+	(218.29,106.64);
 \end{scope}
@@
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] ( 73.97, 48.31) --
-	( 87.68, 51.15) --
-	(101.39, 48.31) --
-	(101.39, 40.20) --
-	( 73.97, 40.20) --
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] ( 69.01, 55.61) --
+	( 87.68, 58.45) --
+	(106.35, 55.61) --
+	(106.35, 44.80) --
+	( 69.01, 44.80) --
 	cycle;
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (111.80, 48.31) --
-	(120.33, 52.22) --
-	(128.87, 48.31) --
-	(128.87, 40.20) --
-	(111.80, 40.20) --
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (144.76, 55.61) --
+	(152.99, 64.63) --
+	(161.21, 55.61) --
+	(161.21, 46.74) --
+	(144.76, 46.74) --
 	cycle;
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (145.45, 48.31) --
-	(152.99, 58.25) --
-	(160.52, 48.31) --
-	(160.52, 40.20) --
-	(145.45, 40.20) --
-	cycle;
+\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at ( 87.68, 47.30) {$N \log N$};
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (167.22, 48.31) --
-	(174.76, 58.25) --
-	(182.29, 48.31) --
-	(182.29, 40.20) --
-	(167.22, 40.20) --
-	cycle;
-
-\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.71] at ( 87.68, 43.40) {$N \log N$};
-
-\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.71] at (120.33, 43.40) {$N^{1.5}$};
-
-\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.90] at (152.99, 42.12) {$N^2$};
-
-\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.90] at (174.76, 42.12) {$N^3$};
-\definecolor{drawColor}{gray}{0.20}
-
-\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 46.86, 41.62) rectangle (226.46,110.68);
+\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at (152.99, 47.30) {$N^2$};
 \end{scope}
@@
 \path[clip] (231.96,116.18) rectangle (411.55,185.23);
-\definecolor{fillColor}{RGB}{255,255,255}
+\definecolor{fillColor}{gray}{0.92}
 
 \path[fill=fillColor] (231.96,116.18) rectangle (411.55,185.23);
-\definecolor{drawColor}{gray}{0.92}
+\definecolor{drawColor}{RGB}{255,255,255}
 
-\path[draw=drawColor,line width= 0.3pt,line join=round] (231.96,124.04) --
-	(411.55,124.04);
+\path[draw=drawColor,line width= 0.3pt,line join=round] (231.96,121.16) --
+	(411.55,121.16);
 
-\path[draw=drawColor,line width= 0.3pt,line join=round] (231.96,141.98) --
-	(411.55,141.98);
+\path[draw=drawColor,line width= 0.3pt,line join=round] (231.96,137.47) --
+	(411.55,137.47);
 
-\path[draw=drawColor,line width= 0.3pt,line join=round] (231.96,159.91) --
-	(411.55,159.91);
+\path[draw=drawColor,line width= 0.3pt,line join=round] (231.96,153.77) --
+	(411.55,153.77);
 
-\path[draw=drawColor,line width= 0.3pt,line join=round] (231.96,177.84) --
-	(411.55,177.84);
+\path[draw=drawColor,line width= 0.3pt,line join=round] (231.96,170.08) --
+	(411.55,170.08);
 
@@
 
-\path[draw=drawColor,line width= 0.6pt,line join=round] (231.96,133.01) --
-	(411.55,133.01);
+\path[draw=drawColor,line width= 0.6pt,line join=round] (231.96,129.31) --
+	(411.55,129.31);
 
-\path[draw=drawColor,line width= 0.6pt,line join=round] (231.96,150.94) --
-	(411.55,150.94);
+\path[draw=drawColor,line width= 0.6pt,line join=round] (231.96,145.62) --
+	(411.55,145.62);
 
-\path[draw=drawColor,line width= 0.6pt,line join=round] (231.96,168.88) --
-	(411.55,168.88);
+\path[draw=drawColor,line width= 0.6pt,line join=round] (231.96,161.92) --
+	(411.55,161.92);
+
+\path[draw=drawColor,line width= 0.6pt,line join=round] (231.96,178.23) --
+	(411.55,178.23);
 
@@
 
-\path[draw=drawColor,line width= 2.3pt,line join=round] (240.12,140.30) --
-	(247.89,140.30) --
-	(255.67,140.31) --
-	(263.44,140.47) --
-	(271.22,141.29) --
-	(278.99,142.56) --
-	(286.77,143.85) --
-	(294.54,145.59) --
-	(302.32,147.70) --
-	(310.09,150.07) --
-	(317.87,152.59) --
-	(325.64,155.20) --
-	(333.41,157.85) --
-	(341.19,160.53) --
-	(348.96,163.22) --
-	(356.74,165.91) --
-	(364.51,168.60) --
-	(372.29,171.30) --
-	(380.06,174.00) --
-	(387.84,176.70) --
-	(395.61,179.40) --
+\path[draw=drawColor,line width= 2.3pt,line join=round] (240.12,144.09) --
+	(247.89,144.09) --
+	(255.67,144.10) --
+	(263.44,144.24) --
+	(271.22,144.99) --
+	(278.99,146.15) --
+	(286.77,147.32) --
+	(294.54,148.90) --
+	(302.32,150.82) --
+	(310.09,152.98) --
+	(317.87,155.27) --
+	(325.64,157.64) --
+	(333.41,160.05) --
+	(341.19,162.48) --
+	(348.96,164.93) --
+	(356.74,167.38) --
+	(364.51,169.83) --
+	(372.29,172.28) --
+	(380.06,174.73) --
+	(387.84,177.19) --
+	(395.61,179.64) --
 	(403.39,182.10);
@@
 
-\path[draw=drawColor,line width= 1.1pt,line join=round] (325.64,128.11) --
-	(333.41,133.51) --
-	(341.19,138.91) --
-	(348.96,144.31) --
-	(356.74,149.71) --
-	(364.51,155.11) --
-	(372.29,160.50) --
-	(380.06,165.90) --
-	(387.84,171.30) --
-	(395.61,176.70) --
+\path[draw=drawColor,line width= 1.1pt,line join=round] (325.64,133.01) --
+	(333.41,137.92) --
+	(341.19,142.83) --
+	(348.96,147.74) --
+	(356.74,152.65) --
+	(364.51,157.55) --
+	(372.29,162.46) --
+	(380.06,167.37) --
+	(387.84,172.28) --
+	(395.61,177.19) --
 	(403.39,182.10);
 
-\path[draw=drawColor,line width= 1.1pt,line join=round] (247.89,128.11) --
-	(255.67,130.81) --
-	(263.44,133.51) --
-	(271.22,136.21) --
-	(278.99,138.91) --
-	(286.77,141.61) --
-	(294.54,144.31) --
-	(302.32,147.01) --
-	(310.09,149.71) --
-	(317.87,152.41) --
-	(325.64,155.11) --
-	(333.41,157.80) --
-	(341.19,160.50) --
-	(348.96,163.20) --
-	(356.74,165.90) --
-	(364.51,168.60) --
-	(372.29,171.30) --
-	(380.06,174.00) --
-	(387.84,176.70) --
-	(395.61,179.40) --
+\path[draw=drawColor,line width= 1.1pt,line join=round] (247.89,133.01) --
+	(255.67,135.47) --
+	(263.44,137.92) --
+	(271.22,140.38) --
+	(278.99,142.83) --
+	(286.77,145.28) --
+	(294.54,147.74) --
+	(302.32,150.19) --
+	(310.09,152.65) --
+	(317.87,155.10) --
+	(325.64,157.55) --
+	(333.41,160.01) --
+	(341.19,162.46) --
+	(348.96,164.92) --
+	(356.74,167.37) --
+	(364.51,169.83) --
+	(372.29,172.28) --
+	(380.06,174.73) --
+	(387.84,177.19) --
+	(395.61,179.64) --
 	(403.39,182.10);
@@
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (317.41,125.27) --
-	(325.64,128.11) --
-	(333.87,125.27) --
-	(333.87,116.40) --
-	(317.41,116.40) --
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (317.41,130.17) --
+	(325.64,133.01) --
+	(333.87,130.17) --
+	(333.87,121.30) --
+	(317.41,121.30) --
 	cycle;
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (241.91,125.27) --
-	(247.89,128.11) --
-	(253.88,125.27) --
-	(253.88,116.40) --
-	(241.91,116.40) --
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (241.91,130.17) --
+	(247.89,133.01) --
+	(253.88,130.17) --
+	(253.88,121.30) --
+	(241.91,121.30) --
 	cycle;
 
-\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at (325.64,118.38) {$N^2$};
+\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at (325.64,121.86) {$N^2$};
 
-\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at (247.89,118.38) {$N$};
-\definecolor{drawColor}{gray}{0.20}
-
-\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (231.96,116.18) rectangle (411.55,185.23);
+\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at (247.89,121.86) {$N$};
 \end{scope}
@@
 \path[clip] (231.96, 41.62) rectangle (411.55,110.68);
-\definecolor{fillColor}{RGB}{255,255,255}
+\definecolor{fillColor}{gray}{0.92}
 
 \path[fill=fillColor] (231.96, 41.62) rectangle (411.55,110.68);
-\definecolor{drawColor}{gray}{0.92}
+\definecolor{drawColor}{RGB}{255,255,255}
 
-\path[draw=drawColor,line width= 0.3pt,line join=round] (231.96, 42.48) --
-	(411.55, 42.48);
+\path[draw=drawColor,line width= 0.3pt,line join=round] (231.96, 45.09) --
+	(411.55, 45.09);
 
-\path[draw=drawColor,line width= 0.3pt,line join=round] (231.96, 55.84) --
-	(411.55, 55.84);
+\path[draw=drawColor,line width= 0.3pt,line join=round] (231.96, 68.35) --
+	(411.55, 68.35);
 
-\path[draw=drawColor,line width= 0.3pt,line join=round] (231.96, 69.20) --
-	(411.55, 69.20);
-
-\path[draw=drawColor,line width= 0.3pt,line join=round] (231.96, 82.55) --
-	(411.55, 82.55);
-
-\path[draw=drawColor,line width= 0.3pt,line join=round] (231.96, 95.91) --
-	(411.55, 95.91);
-
-\path[draw=drawColor,line width= 0.3pt,line join=round] (231.96,109.27) --
-	(411.55,109.27);
+\path[draw=drawColor,line width= 0.3pt,line join=round] (231.96, 91.60) --
+	(411.55, 91.60);
 
@@
 
-\path[draw=drawColor,line width= 0.6pt,line join=round] (231.96, 49.16) --
-	(411.55, 49.16);
+\path[draw=drawColor,line width= 0.6pt,line join=round] (231.96, 56.72) --
+	(411.55, 56.72);
 
-\path[draw=drawColor,line width= 0.6pt,line join=round] (231.96, 62.52) --
-	(411.55, 62.52);
+\path[draw=drawColor,line width= 0.6pt,line join=round] (231.96, 79.97) --
+	(411.55, 79.97);
 
-\path[draw=drawColor,line width= 0.6pt,line join=round] (231.96, 75.88) --
-	(411.55, 75.88);
-
-\path[draw=drawColor,line width= 0.6pt,line join=round] (231.96, 89.23) --
-	(411.55, 89.23);
-
-\path[draw=drawColor,line width= 0.6pt,line join=round] (231.96,102.59) --
-	(411.55,102.59);
+\path[draw=drawColor,line width= 0.6pt,line join=round] (231.96,103.23) --
+	(411.55,103.23);
 
@@
 	(405.36,110.68);
-\definecolor{drawColor}{RGB}{190,190,190}
-
-\path[draw=drawColor,line width= 0.6pt,line join=round] (231.96, 98.57) -- (411.55, 98.57);
 \definecolor{fillColor}{RGB}{0,0,0}
 
-\path[fill=fillColor,fill opacity=0.50] (240.12, 57.71) --
-	(247.89, 53.91) --
-	(255.67, 52.68) --
-	(263.44, 63.57) --
-	(271.22, 52.51) --
-	(278.99, 49.84) --
-	(286.77, 48.48) --
-	(294.54, 49.49) --
-	(302.32, 49.66) --
-	(310.09, 52.74) --
-	(317.87, 62.26) --
-	(325.64, 63.29) --
-	(333.41, 61.67) --
-	(341.19, 66.65) --
-	(348.96, 70.51) --
-	(356.74, 73.89) --
-	(364.51, 81.12) --
-	(372.29, 82.11) --
-	(380.06, 86.16) --
-	(387.84, 90.86) --
-	(395.61, 95.14) --
-	(403.39, 99.71) --
-	(403.39, 99.32) --
-	(395.61, 94.61) --
-	(387.84, 90.49) --
-	(380.06, 85.91) --
-	(372.29, 81.88) --
-	(364.51, 77.39) --
-	(356.74, 73.46) --
-	(348.96, 69.25) --
-	(341.19, 66.07) --
-	(333.41, 61.08) --
-	(325.64, 57.16) --
-	(317.87, 54.48) --
-	(310.09, 51.11) --
-	(302.32, 48.90) --
-	(294.54, 47.85) --
-	(286.77, 47.05) --
-	(278.99, 48.88) --
-	(271.22, 51.55) --
-	(263.44, 51.67) --
-	(255.67, 51.49) --
-	(247.89, 51.36) --
-	(240.12, 52.74) --
+\path[fill=fillColor,fill opacity=0.50] (240.12, 64.16) --
+	(247.89, 60.85) --
+	(255.67, 59.78) --
+	(263.44, 69.26) --
+	(271.22, 59.63) --
+	(278.99, 57.31) --
+	(286.77, 56.13) --
+	(294.54, 57.00) --
+	(302.32, 57.15) --
+	(310.09, 59.83) --
+	(317.87, 68.12) --
+	(325.64, 69.02) --
+	(333.41, 67.60) --
+	(341.19, 71.94) --
+	(348.96, 75.31) --
+	(356.74, 78.25) --
+	(364.51, 84.54) --
+	(372.29, 85.40) --
+	(380.06, 88.92) --
+	(387.84, 93.02) --
+	(395.61, 96.74) --
+	(403.39,100.73) --
+	(403.39,100.39) --
+	(395.61, 96.28) --
+	(387.84, 92.69) --
+	(380.06, 88.71) --
+	(372.29, 85.21) --
+	(364.51, 81.29) --
+	(356.74, 77.87) --
+	(348.96, 74.21) --
+	(341.19, 71.44) --
+	(333.41, 67.10) --
+	(325.64, 63.68) --
+	(317.87, 61.35) --
+	(310.09, 58.41) --
+	(302.32, 56.50) --
+	(294.54, 55.57) --
+	(286.77, 54.88) --
+	(278.99, 56.47) --
+	(271.22, 58.80) --
+	(263.44, 58.90) --
+	(255.67, 58.75) --
+	(247.89, 58.63) --
+	(240.12, 59.84) --
 	cycle;
 
-\path[] (240.12, 57.71) --
-	(247.89, 53.91) --
-	(255.67, 52.68) --
-	(263.44, 63.57) --
-	(271.22, 52.51) --
-	(278.99, 49.84) --
-	(286.77, 48.48) --
-	(294.54, 49.49) --
-	(302.32, 49.66) --
-	(310.09, 52.74) --
-	(317.87, 62.26) --
-	(325.64, 63.29) --
-	(333.41, 61.67) --
-	(341.19, 66.65) --
-	(348.96, 70.51) --
-	(356.74, 73.89) --
-	(364.51, 81.12) --
-	(372.29, 82.11) --
-	(380.06, 86.16) --
-	(387.84, 90.86) --
-	(395.61, 95.14) --
-	(403.39, 99.71);
+\path[] (240.12, 64.16) --
+	(247.89, 60.85) --
+	(255.67, 59.78) --
+	(263.44, 69.26) --
+	(271.22, 59.63) --
+	(278.99, 57.31) --
+	(286.77, 56.13) --
+	(294.54, 57.00) --
+	(302.32, 57.15) --
+	(310.09, 59.83) --
+	(317.87, 68.12) --
+	(325.64, 69.02) --
+	(333.41, 67.60) --
+	(341.19, 71.94) --
+	(348.96, 75.31) --
+	(356.74, 78.25) --
+	(364.51, 84.54) --
+	(372.29, 85.40) --
+	(380.06, 88.92) --
+	(387.84, 93.02) --
+	(395.61, 96.74) --
+	(403.39,100.73);
 
-\path[] (403.39, 99.32) --
-	(395.61, 94.61) --
-	(387.84, 90.49) --
-	(380.06, 85.91) --
-	(372.29, 81.88) --
-	(364.51, 77.39) --
-	(356.74, 73.46) --
-	(348.96, 69.25) --
-	(341.19, 66.07) --
-	(333.41, 61.08) --
-	(325.64, 57.16) --
-	(317.87, 54.48) --
-	(310.09, 51.11) --
-	(302.32, 48.90) --
-	(294.54, 47.85) --
-	(286.77, 47.05) --
-	(278.99, 48.88) --
-	(271.22, 51.55) --
-	(263.44, 51.67) --
-	(255.67, 51.49) --
-	(247.89, 51.36) --
-	(240.12, 52.74);
+\path[] (403.39,100.39) --
+	(395.61, 96.28) --
+	(387.84, 92.69) --
+	(380.06, 88.71) --
+	(372.29, 85.21) --
+	(364.51, 81.29) --
+	(356.74, 77.87) --
+	(348.96, 74.21) --
+	(341.19, 71.44) --
+	(333.41, 67.10) --
+	(325.64, 63.68) --
+	(317.87, 61.35) --
+	(310.09, 58.41) --
+	(302.32, 56.50) --
+	(294.54, 55.57) --
+	(286.77, 54.88) --
+	(278.99, 56.47) --
+	(271.22, 58.80) --
+	(263.44, 58.90) --
+	(255.67, 58.75) --
+	(247.89, 58.63) --
+	(240.12, 59.84);
 \definecolor{drawColor}{RGB}{0,0,0}
 
-\path[draw=drawColor,line width= 2.3pt,line join=round] (240.12, 55.09) --
-	(247.89, 51.58) --
-	(255.67, 51.81) --
-	(263.44, 51.98) --
-	(271.22, 51.79) --
-	(278.99, 49.27) --
-	(286.77, 47.55) --
-	(294.54, 48.12) --
-	(302.32, 49.04) --
-	(310.09, 51.52) --
-	(317.87, 54.89) --
-	(325.64, 57.43) --
-	(333.41, 61.36) --
-	(341.19, 66.31) --
-	(348.96, 69.59) --
-	(356.74, 73.56) --
-	(364.51, 77.55) --
-	(372.29, 81.95) --
-	(380.06, 86.08) --
-	(387.84, 90.61) --
-	(395.61, 94.71) --
-	(403.39, 99.49);
+\path[draw=drawColor,line width= 2.3pt,line join=round] (240.12, 61.88) --
+	(247.89, 58.83) --
+	(255.67, 59.02) --
+	(263.44, 59.17) --
+	(271.22, 59.01) --
+	(278.99, 56.82) --
+	(286.77, 55.32) --
+	(294.54, 55.81) --
+	(302.32, 56.61) --
+	(310.09, 58.78) --
+	(317.87, 61.71) --
+	(325.64, 63.92) --
+	(333.41, 67.34) --
+	(341.19, 71.65) --
+	(348.96, 74.51) --
+	(356.74, 77.96) --
+	(364.51, 81.44) --
+	(372.29, 85.27) --
+	(380.06, 88.86) --
+	(387.84, 92.81) --
+	(395.61, 96.37) --
+	(403.39,100.53);
 \definecolor{drawColor}{RGB}{238,130,238}
 
-\path[draw=drawColor,line width= 1.1pt,line join=round] (317.87, 51.47) --
-	(325.64, 55.96) --
-	(333.41, 60.41) --
-	(341.19, 64.83) --
-	(348.96, 69.23) --
-	(356.74, 73.60) --
-	(364.51, 77.96) --
-	(372.29, 82.29) --
-	(380.06, 86.61) --
-	(387.84, 90.91) --
-	(395.61, 95.21) --
-	(403.39, 99.49);
+\path[draw=drawColor,line width= 1.1pt,line join=round] (317.87, 58.73) --
+	(325.64, 62.64) --
+	(333.41, 66.51) --
+	(341.19, 70.36) --
+	(348.96, 74.19) --
+	(356.74, 78.00) --
+	(364.51, 81.79) --
+	(372.29, 85.56) --
+	(380.06, 89.32) --
+	(387.84, 93.07) --
+	(395.61, 96.80) --
+	(403.39,100.53);
 
-\path[draw=drawColor,line width= 1.1pt,line join=round] (341.19, 51.23) --
-	(348.96, 57.26) --
-	(356.74, 63.29) --
-	(364.51, 69.32) --
-	(372.29, 75.36) --
-	(380.06, 81.39) --
-	(387.84, 87.42) --
-	(395.61, 93.45) --
-	(403.39, 99.49);
-
-\path[draw=drawColor,line width= 1.1pt,line join=round] (356.74, 51.23) --
-	(364.51, 59.27) --
-	(372.29, 67.31) --
-	(380.06, 75.36) --
-	(387.84, 83.40) --
-	(395.61, 91.44) --
-	(403.39, 99.49);
-
-\path[draw=drawColor,line width= 1.1pt,line join=round] (372.29, 51.23) --
-	(380.06, 63.29) --
-	(387.84, 75.36) --
-	(395.61, 87.42) --
-	(403.39, 99.49);
+\path[draw=drawColor,line width= 1.1pt,line join=round] (356.74, 58.52) --
+	(364.51, 65.52) --
+	(372.29, 72.52) --
+	(380.06, 79.52) --
+	(387.84, 86.52) --
+	(395.61, 93.53) --
+	(403.39,100.53);
 \end{scope}
@@
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (304.01, 48.38) --
-	(317.87, 51.47) --
-	(331.72, 48.38) --
-	(331.72, 40.20) --
-	(304.01, 40.20) --
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (299.19, 55.67) --
+	(317.87, 58.73) --
+	(336.54, 55.67) --
+	(336.54, 44.86) --
+	(299.19, 44.86) --
 	cycle;
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (332.23, 48.38) --
-	(341.19, 51.23) --
-	(349.47, 48.38) --
-	(349.47, 40.20) --
-	(332.23, 40.20) --
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (348.51, 55.67) --
+	(356.74, 58.52) --
+	(364.97, 55.67) --
+	(364.97, 46.81) --
+	(348.51, 46.81) --
 	cycle;
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (349.47, 48.38) --
-	(356.74, 51.23) --
-	(364.68, 48.38) --
-	(364.68, 40.20) --
-	(349.47, 40.20) --
-	cycle;
+\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at (317.87, 47.36) {$N \log N$};
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (364.68, 48.38) --
-	(372.29, 51.23) --
-	(379.90, 48.38) --
-	(379.90, 40.20) --
-	(364.68, 40.20) --
-	cycle;
-
-\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.72] at (317.87, 43.42) {$N \log N$};
-
-\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.72] at (340.85, 43.42) {$N^{1.5}$};
-
-\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.91] at (357.07, 42.12) {$N^2$};
-
-\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.91] at (372.29, 42.12) {$N^3$};
-\definecolor{drawColor}{gray}{0.20}
-
-\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (231.96, 41.62) rectangle (411.55,110.68);
+\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at (356.74, 47.36) {$N^2$};
 \end{scope}
@@
 \path[clip] ( 46.86,185.23) rectangle (226.46,211.31);
-\definecolor{drawColor}{gray}{0.20}
 \definecolor{fillColor}{gray}{0.85}
 
-\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] ( 46.86,185.23) rectangle (226.46,211.31);
+\path[fill=fillColor] ( 46.86,185.23) rectangle (226.46,211.31);
 \definecolor{drawColor}{gray}{0.10}
@@
 \path[clip] (231.96,185.23) rectangle (411.55,211.31);
-\definecolor{drawColor}{gray}{0.20}
 \definecolor{fillColor}{gray}{0.85}
 
-\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (231.96,185.23) rectangle (411.55,211.31);
+\path[fill=fillColor] (231.96,185.23) rectangle (411.55,211.31);
 \definecolor{drawColor}{gray}{0.10}
@@
 \path[clip] (411.55,116.18) rectangle (428.12,185.23);
-\definecolor{drawColor}{gray}{0.20}
 \definecolor{fillColor}{gray}{0.85}
 
-\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (411.55,116.18) rectangle (428.12,185.23);
+\path[fill=fillColor] (411.55,116.18) rectangle (428.12,185.23);
 \definecolor{drawColor}{gray}{0.10}
@@
 \path[clip] (411.55, 41.62) rectangle (428.12,110.68);
-\definecolor{drawColor}{gray}{0.20}
 \definecolor{fillColor}{gray}{0.85}
 
-\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (411.55, 41.62) rectangle (428.12,110.68);
+\path[fill=fillColor] (411.55, 41.62) rectangle (428.12,110.68);
 \definecolor{drawColor}{gray}{0.10}
@@
 
-\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 41.91,129.98) {1e+01};
+\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 41.91,126.28) {1e+00};
 
-\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 41.91,147.91) {1e+03};
+\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 41.91,142.59) {1e+02};
 
-\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 41.91,165.84) {1e+05};
+\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 41.91,158.89) {1e+04};
+
+\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 41.91,175.20) {1e+06};
 \end{scope}
@@
 
-\path[draw=drawColor,line width= 0.6pt,line join=round] ( 44.11,133.01) --
-	( 46.86,133.01);
+\path[draw=drawColor,line width= 0.6pt,line join=round] ( 44.11,129.31) --
+	( 46.86,129.31);
 
-\path[draw=drawColor,line width= 0.6pt,line join=round] ( 44.11,150.94) --
-	( 46.86,150.94);
+\path[draw=drawColor,line width= 0.6pt,line join=round] ( 44.11,145.62) --
+	( 46.86,145.62);
 
-\path[draw=drawColor,line width= 0.6pt,line join=round] ( 44.11,168.88) --
-	( 46.86,168.88);
+\path[draw=drawColor,line width= 0.6pt,line join=round] ( 44.11,161.92) --
+	( 46.86,161.92);
+
+\path[draw=drawColor,line width= 0.6pt,line join=round] ( 44.11,178.23) --
+	( 46.86,178.23);
 \end{scope}
@@
 
-\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 41.91, 46.13) {1e-03};
+\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 41.91, 53.69) {1e-03};
 
-\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 41.91, 59.49) {1e-02};
+\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 41.91, 76.94) {1e-01};
 
-\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 41.91, 72.85) {1e-01};
-
-\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 41.91, 86.20) {1e+00};
-
-\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 41.91, 99.56) {1e+01};
+\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 41.91,100.20) {1e+01};
 \end{scope}
@@
 
-\path[draw=drawColor,line width= 0.6pt,line join=round] ( 44.11, 49.16) --
-	( 46.86, 49.16);
+\path[draw=drawColor,line width= 0.6pt,line join=round] ( 44.11, 56.72) --
+	( 46.86, 56.72);
 
-\path[draw=drawColor,line width= 0.6pt,line join=round] ( 44.11, 62.52) --
-	( 46.86, 62.52);
+\path[draw=drawColor,line width= 0.6pt,line join=round] ( 44.11, 79.97) --
+	( 46.86, 79.97);
 
-\path[draw=drawColor,line width= 0.6pt,line join=round] ( 44.11, 75.88) --
-	( 46.86, 75.88);
-
-\path[draw=drawColor,line width= 0.6pt,line join=round] ( 44.11, 89.23) --
-	( 46.86, 89.23);
-
-\path[draw=drawColor,line width= 0.6pt,line join=round] ( 44.11,102.59) --
-	( 46.86,102.59);
+\path[draw=drawColor,line width= 0.6pt,line join=round] ( 44.11,103.23) --
+	( 46.86,103.23);
 \end{scope}
@@
 
-\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at (229.21,  7.64) {N};
+\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at (229.21,  7.64) {$N$ = number of data to segment};
 \end{scope}

### 2025-05-06 figure-heap-refs-other.tex
direct label in tikz

@@
-% Created by tikzDevice version 0.12.6 on 2025-05-06 20:54:35
+% Created by tikzDevice version 0.12.6 on 2025-05-06 22:53:33
 % !TEX encoding = UTF-8 Unicode
@@
 \definecolor{fillColor}{RGB}{255,255,255}
-\path[use as bounding box,fill=fillColor,fill opacity=0.00] (0,0) rectangle (505.89,505.89);
+\path[use as bounding box,fill=fillColor,fill opacity=0.00] (0,0) rectangle (433.62,216.81);
 \begin{scope}
-\path[clip] (  0.00,  0.00) rectangle (505.89,505.89);
+\path[clip] (  0.00,  0.00) rectangle (433.62,216.81);
 \definecolor{drawColor}{RGB}{255,255,255}
@@
 
-\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (  0.00,  0.00) rectangle (505.89,505.89);
+\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] ( -0.00,  0.00) rectangle (433.62,216.81);
 \end{scope}
 \begin{scope}
-\path[clip] ( 46.86,264.98) rectangle (262.59,483.82);
-\definecolor{fillColor}{RGB}{255,255,255}
+\path[clip] ( 46.86,120.44) rectangle (226.46,194.74);
+\definecolor{fillColor}{gray}{0.92}
 
-\path[fill=fillColor] ( 46.86,264.98) rectangle (262.59,483.82);
-\definecolor{drawColor}{gray}{0.92}
+\path[fill=fillColor] ( 46.86,120.44) rectangle (226.46,194.74);
+\definecolor{drawColor}{RGB}{255,255,255}
 
-\path[draw=drawColor,line width= 0.3pt,line join=round] ( 46.86,289.91) --
-	(262.59,289.91);
+\path[draw=drawColor,line width= 0.3pt,line join=round] ( 46.86,125.99) --
+	(226.46,125.99);
 
-\path[draw=drawColor,line width= 0.3pt,line join=round] ( 46.86,346.74) --
-	(262.59,346.74);
+\path[draw=drawColor,line width= 0.3pt,line join=round] ( 46.86,145.23) --
+	(226.46,145.23);
 
-\path[draw=drawColor,line width= 0.3pt,line join=round] ( 46.86,403.56) --
-	(262.59,403.56);
+\path[draw=drawColor,line width= 0.3pt,line join=round] ( 46.86,164.47) --
+	(226.46,164.47);
 
-\path[draw=drawColor,line width= 0.3pt,line join=round] ( 46.86,460.39) --
-	(262.59,460.39);
+\path[draw=drawColor,line width= 0.3pt,line join=round] ( 46.86,183.70) --
+	(226.46,183.70);
 
-\path[draw=drawColor,line width= 0.3pt,line join=round] ( 50.63,264.98) --
-	( 50.63,483.82);
+\path[draw=drawColor,line width= 0.3pt,line join=round] ( 50.00,120.44) --
+	( 50.00,194.74);
 
-\path[draw=drawColor,line width= 0.3pt,line join=round] (109.85,264.98) --
-	(109.85,483.82);
+\path[draw=drawColor,line width= 0.3pt,line join=round] ( 99.30,120.44) --
+	( 99.30,194.74);
 
-\path[draw=drawColor,line width= 0.3pt,line join=round] (169.08,264.98) --
-	(169.08,483.82);
+\path[draw=drawColor,line width= 0.3pt,line join=round] (148.61,120.44) --
+	(148.61,194.74);
 
-\path[draw=drawColor,line width= 0.3pt,line join=round] (228.30,264.98) --
-	(228.30,483.82);
+\path[draw=drawColor,line width= 0.3pt,line join=round] (197.91,120.44) --
+	(197.91,194.74);
 
-\path[draw=drawColor,line width= 0.6pt,line join=round] ( 46.86,318.33) --
-	(262.59,318.33);
+\path[draw=drawColor,line width= 0.6pt,line join=round] ( 46.86,135.61) --
+	(226.46,135.61);
 
-\path[draw=drawColor,line width= 0.6pt,line join=round] ( 46.86,375.15) --
-	(262.59,375.15);
+\path[draw=drawColor,line width= 0.6pt,line join=round] ( 46.86,154.85) --
+	(226.46,154.85);
 
-\path[draw=drawColor,line width= 0.6pt,line join=round] ( 46.86,431.98) --
-	(262.59,431.98);
+\path[draw=drawColor,line width= 0.6pt,line join=round] ( 46.86,174.08) --
+	(226.46,174.08);
 
-\path[draw=drawColor,line width= 0.6pt,line join=round] ( 80.24,264.98) --
-	( 80.24,483.82);
+\path[draw=drawColor,line width= 0.6pt,line join=round] ( 46.86,193.32) --
+	(226.46,193.32);
 
-\path[draw=drawColor,line width= 0.6pt,line join=round] (139.46,264.98) --
-	(139.46,483.82);
+\path[draw=drawColor,line width= 0.6pt,line join=round] ( 74.65,120.44) --
+	( 74.65,194.74);
 
-\path[draw=drawColor,line width= 0.6pt,line join=round] (198.69,264.98) --
-	(198.69,483.82);
+\path[draw=drawColor,line width= 0.6pt,line join=round] (123.95,120.44) --
+	(123.95,194.74);
 
-\path[draw=drawColor,line width= 0.6pt,line join=round] (257.92,264.98) --
-	(257.92,483.82);
+\path[draw=drawColor,line width= 0.6pt,line join=round] (173.26,120.44) --
+	(173.26,194.74);
+
+\path[draw=drawColor,line width= 0.6pt,line join=round] (222.56,120.44) --
+	(222.56,194.74);
 \definecolor{drawColor}{RGB}{0,0,0}
 
-\path[draw=drawColor,line width= 2.3pt,line join=round] ( 56.67,398.60) --
-	( 74.50,286.74) --
-	( 92.33,305.45) --
-	(110.16,319.22) --
-	(127.98,338.53) --
-	(145.81,354.75) --
-	(163.64,370.70) --
-	(181.47,387.04) --
-	(199.30,403.70) --
-	(217.13,420.58) --
-	(234.96,437.56) --
-	(252.79,454.61);
+\path[draw=drawColor,line width= 2.3pt,line join=round] ( 55.03,172.40) --
+	( 69.87,134.54) --
+	( 84.71,140.87) --
+	( 99.55,145.53) --
+	(114.40,152.07) --
+	(129.24,157.56) --
+	(144.08,162.96) --
+	(158.92,168.49) --
+	(173.77,174.13) --
+	(188.61,179.84) --
+	(203.45,185.59) --
+	(218.29,191.36);
 \definecolor{drawColor}{RGB}{238,130,238}
 
-\path[draw=drawColor,line width= 1.1pt,line join=round] ( 92.33,300.65) --
-	(110.16,317.76) --
-	(127.98,334.86) --
-	(145.81,351.97) --
-	(163.64,369.07) --
-	(181.47,386.18) --
-	(199.30,403.29) --
-	(217.13,420.39) --
-	(234.96,437.50) --
-	(252.79,454.61);
+\path[draw=drawColor,line width= 1.1pt,line join=round] ( 84.71,139.25) --
+	( 99.55,145.04) --
+	(114.40,150.83) --
+	(129.24,156.62) --
+	(144.08,162.41) --
+	(158.92,168.20) --
+	(173.77,173.99) --
+	(188.61,179.78) --
+	(203.45,185.57) --
+	(218.29,191.36);
 
-\path[draw=drawColor,line width= 1.1pt,line join=round] ( 56.67,360.52) --
-	( 74.50,369.07) --
-	( 92.33,377.63) --
-	(110.16,386.18) --
-	(127.98,394.73) --
-	(145.81,403.29) --
-	(163.64,411.84) --
-	(181.47,420.39) --
-	(199.30,428.95) --
-	(217.13,437.50) --
-	(234.96,446.05) --
-	(252.79,454.61);
+\path[draw=drawColor,line width= 1.1pt,line join=round] ( 55.03,159.51) --
+	( 69.87,162.41) --
+	( 84.71,165.30) --
+	( 99.55,168.20) --
+	(114.40,171.09) --
+	(129.24,173.99) --
+	(144.08,176.89) --
+	(158.92,179.78) --
+	(173.77,182.68) --
+	(188.61,185.57) --
+	(203.45,188.47) --
+	(218.29,191.36);
 \end{scope}
 \begin{scope}
-\path[clip] ( 46.86,264.98) rectangle (262.59,483.82);
+\path[clip] ( 46.86,120.44) rectangle (226.46,194.74);
 \definecolor{drawColor}{RGB}{0,0,0}
@@
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] ( 84.10,297.80) --
-	( 92.33,300.65) --
-	(100.56,297.80) --
-	(100.56,288.94) --
-	( 84.10,288.94) --
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] ( 76.48,136.40) --
+	( 84.71,139.25) --
+	( 92.94,136.40) --
+	( 92.94,127.54) --
+	( 76.48,127.54) --
 	cycle;
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] ( 50.68,297.80) --
-	( 56.67,360.52) --
-	( 62.66,297.80) --
-	( 62.66,288.94) --
-	( 50.68,288.94) --
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] ( 49.04,136.40) --
+	( 55.03,159.51) --
+	( 61.01,136.40) --
+	( 61.01,127.54) --
+	( 49.04,127.54) --
 	cycle;
 
-\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at ( 92.33,290.92) {$N^2$};
+\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at ( 84.71,128.09) {$N^2$};
 
-\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at ( 56.67,290.92) {$N$};
-\definecolor{drawColor}{gray}{0.20}
-
-\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 46.86,264.98) rectangle (262.59,483.82);
+\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at ( 55.03,128.09) {$N$};
 \end{scope}
 \begin{scope}
-\path[clip] ( 46.86, 40.64) rectangle (262.59,259.48);
-\definecolor{fillColor}{RGB}{255,255,255}
+\path[clip] ( 46.86, 40.64) rectangle (226.46,114.94);
+\definecolor{fillColor}{gray}{0.92}
 
-\path[fill=fillColor] ( 46.86, 40.64) rectangle (262.59,259.48);
-\definecolor{drawColor}{gray}{0.92}
+\path[fill=fillColor] ( 46.86, 40.64) rectangle (226.46,114.94);
+\definecolor{drawColor}{RGB}{255,255,255}
 
-\path[draw=drawColor,line width= 0.3pt,line join=round] ( 46.86, 45.62) --
-	(262.59, 45.62);
+\path[draw=drawColor,line width= 0.3pt,line join=round] ( 46.86, 44.40) --
+	(226.46, 44.40);
 
-\path[draw=drawColor,line width= 0.3pt,line join=round] ( 46.86, 88.17) --
-	(262.59, 88.17);
+\path[draw=drawColor,line width= 0.3pt,line join=round] ( 46.86, 71.59) --
+	(226.46, 71.59);
 
-\path[draw=drawColor,line width= 0.3pt,line join=round] ( 46.86,130.71) --
-	(262.59,130.71);
+\path[draw=drawColor,line width= 0.3pt,line join=round] ( 46.86, 98.78) --
+	(226.46, 98.78);
 
-\path[draw=drawColor,line width= 0.3pt,line join=round] ( 46.86,173.25) --
-	(262.59,173.25);
+\path[draw=drawColor,line width= 0.3pt,line join=round] ( 50.00, 40.64) --
+	( 50.00,114.94);
 
-\path[draw=drawColor,line width= 0.3pt,line join=round] ( 46.86,215.80) --
-	(262.59,215.80);
+\path[draw=drawColor,line width= 0.3pt,line join=round] ( 99.30, 40.64) --
+	( 99.30,114.94);
 
-\path[draw=drawColor,line width= 0.3pt,line join=round] ( 46.86,258.34) --
-	(262.59,258.34);
+\path[draw=drawColor,line width= 0.3pt,line join=round] (148.61, 40.64) --
+	(148.61,114.94);
 
-\path[draw=drawColor,line width= 0.3pt,line join=round] ( 50.63, 40.64) --
-	( 50.63,259.48);
+\path[draw=drawColor,line width= 0.3pt,line join=round] (197.91, 40.64) --
+	(197.91,114.94);
 
-\path[draw=drawColor,line width= 0.3pt,line join=round] (109.85, 40.64) --
-	(109.85,259.48);
+\path[draw=drawColor,line width= 0.6pt,line join=round] ( 46.86, 58.00) --
+	(226.46, 58.00);
 
-\path[draw=drawColor,line width= 0.3pt,line join=round] (169.08, 40.64) --
-	(169.08,259.48);
+\path[draw=drawColor,line width= 0.6pt,line join=round] ( 46.86, 85.19) --
+	(226.46, 85.19);
 
-\path[draw=drawColor,line width= 0.3pt,line join=round] (228.30, 40.64) --
-	(228.30,259.48);
+\path[draw=drawColor,line width= 0.6pt,line join=round] ( 46.86,112.38) --
+	(226.46,112.38);
 
-\path[draw=drawColor,line width= 0.6pt,line join=round] ( 46.86, 66.89) --
-	(262.59, 66.89);
+\path[draw=drawColor,line width= 0.6pt,line join=round] ( 74.65, 40.64) --
+	( 74.65,114.94);
 
-\path[draw=drawColor,line width= 0.6pt,line join=round] ( 46.86,109.44) --
-	(262.59,109.44);
+\path[draw=drawColor,line width= 0.6pt,line join=round] (123.95, 40.64) --
+	(123.95,114.94);
 
-\path[draw=drawColor,line width= 0.6pt,line join=round] ( 46.86,151.98) --
-	(262.59,151.98);
+\path[draw=drawColor,line width= 0.6pt,line join=round] (173.26, 40.64) --
+	(173.26,114.94);
 
-\path[draw=drawColor,line width= 0.6pt,line join=round] ( 46.86,194.52) --
-	(262.59,194.52);
-
-\path[draw=drawColor,line width= 0.6pt,line join=round] ( 46.86,237.07) --
-	(262.59,237.07);
-
-\path[draw=drawColor,line width= 0.6pt,line join=round] ( 80.24, 40.64) --
-	( 80.24,259.48);
-
-\path[draw=drawColor,line width= 0.6pt,line join=round] (139.46, 40.64) --
-	(139.46,259.48);
-
-\path[draw=drawColor,line width= 0.6pt,line join=round] (198.69, 40.64) --
-	(198.69,259.48);
-
-\path[draw=drawColor,line width= 0.6pt,line join=round] (257.92, 40.64) --
-	(257.92,259.48);
-\definecolor{drawColor}{RGB}{190,190,190}
-
-\path[draw=drawColor,line width= 0.6pt,line join=round] ( 46.86,224.26) -- (262.59,224.26);
+\path[draw=drawColor,line width= 0.6pt,line join=round] (222.56, 40.64) --
+	(222.56,114.94);
 \definecolor{fillColor}{RGB}{0,0,0}
 
-\path[fill=fillColor,fill opacity=0.50] ( 56.67, 66.89) --
-	( 74.50, 78.17) --
-	( 92.33, 71.00) --
-	(110.16, 72.26) --
-	(127.98, 68.86) --
-	(145.81, 70.52) --
-	(163.64, 78.50) --
-	(181.47,112.13) --
-	(199.30,141.19) --
-	(217.13,167.33) --
-	(234.96,193.27) --
-	(252.79,228.18) --
-	(252.79,227.94) --
-	(234.96,192.75) --
-	(217.13,155.33) --
-	(199.30,123.25) --
-	(181.47, 94.27) --
-	(163.64, 70.55) --
-	(145.81, 55.19) --
-	(127.98, 53.93) --
-	(110.16, 59.50) --
-	( 92.33, 55.94) --
-	( 74.50, 63.02) --
-	( 56.67, 50.59) --
+\path[fill=fillColor,fill opacity=0.50] ( 55.03, 58.00) --
+	( 69.87, 61.60) --
+	( 84.71, 59.31) --
+	( 99.55, 59.71) --
+	(114.40, 58.63) --
+	(129.24, 59.16) --
+	(144.08, 61.71) --
+	(158.92, 72.45) --
+	(173.77, 81.74) --
+	(188.61, 90.10) --
+	(203.45, 98.38) --
+	(218.29,109.54) --
+	(218.29,109.46) --
+	(203.45, 98.22) --
+	(188.61, 86.26) --
+	(173.77, 76.01) --
+	(158.92, 66.75) --
+	(144.08, 59.17) --
+	(129.24, 54.26) --
+	(114.40, 53.86) --
+	( 99.55, 55.64) --
+	( 84.71, 54.50) --
+	( 69.87, 56.76) --
+	( 55.03, 52.79) --
 	cycle;
 
-\path[] ( 56.67, 66.89) --
-	( 74.50, 78.17) --
-	( 92.33, 71.00) --
-	(110.16, 72.26) --
-	(127.98, 68.86) --
-	(145.81, 70.52) --
-	(163.64, 78.50) --
-	(181.47,112.13) --
-	(199.30,141.19) --
-	(217.13,167.33) --
-	(234.96,193.27) --
-	(252.79,228.18);
+\path[] ( 55.03, 58.00) --
+	( 69.87, 61.60) --
+	( 84.71, 59.31) --
+	( 99.55, 59.71) --
+	(114.40, 58.63) --
+	(129.24, 59.16) --
+	(144.08, 61.71) --
+	(158.92, 72.45) --
+	(173.77, 81.74) --
+	(188.61, 90.10) --
+	(203.45, 98.38) --
+	(218.29,109.54);
 
-\path[] (252.79,227.94) --
-	(234.96,192.75) --
-	(217.13,155.33) --
-	(199.30,123.25) --
-	(181.47, 94.27) --
-	(163.64, 70.55) --
-	(145.81, 55.19) --
-	(127.98, 53.93) --
-	(110.16, 59.50) --
-	( 92.33, 55.94) --
-	( 74.50, 63.02) --
-	( 56.67, 50.59);
+\path[] (218.29,109.46) --
+	(203.45, 98.22) --
+	(188.61, 86.26) --
+	(173.77, 76.01) --
+	(158.92, 66.75) --
+	(144.08, 59.17) --
+	(129.24, 54.26) --
+	(114.40, 53.86) --
+	( 99.55, 55.64) --
+	( 84.71, 54.50) --
+	( 69.87, 56.76) --
+	( 55.03, 52.79);
 \definecolor{drawColor}{RGB}{0,0,0}
 
-\path[draw=drawColor,line width= 2.3pt,line join=round] ( 56.67, 52.89) --
-	( 74.50, 69.63) --
-	( 92.33, 57.53) --
-	(110.16, 60.50) --
-	(127.98, 57.01) --
-	(145.81, 58.85) --
-	(163.64, 71.74) --
-	(181.47, 94.98) --
-	(199.30,125.79) --
-	(217.13,156.20) --
-	(234.96,192.85) --
-	(252.79,228.05);
+\path[draw=drawColor,line width= 2.3pt,line join=round] ( 55.03, 53.52) --
+	( 69.87, 58.87) --
+	( 84.71, 55.00) --
+	( 99.55, 55.96) --
+	(114.40, 54.84) --
+	(129.24, 55.43) --
+	(144.08, 59.55) --
+	(158.92, 66.97) --
+	(173.77, 76.82) --
+	(188.61, 86.54) --
+	(203.45, 98.25) --
+	(218.29,109.50);
 \definecolor{drawColor}{RGB}{238,130,238}
 
-\path[draw=drawColor,line width= 1.1pt,line join=round] ( 74.50, 72.89) --
-	( 92.33, 91.02) --
-	(110.16,107.95) --
-	(127.98,124.12) --
-	(145.81,139.78) --
-	(163.64,155.05) --
-	(181.47,170.03) --
-	(199.30,184.79) --
-	(217.13,199.35) --
-	(234.96,213.77) --
-	(252.79,228.05);
+\path[draw=drawColor,line width= 1.1pt,line join=round] (129.24, 60.39) --
+	(144.08, 68.57) --
+	(158.92, 76.76) --
+	(173.77, 84.94) --
+	(188.61, 93.13) --
+	(203.45,101.31) --
+	(218.29,109.50);
 
-\path[draw=drawColor,line width= 1.1pt,line join=round] (110.16, 74.37) --
-	(127.98, 93.58) --
-	(145.81,112.79) --
-	(163.64,132.00) --
-	(181.47,151.21) --
-	(199.30,170.42) --
-	(217.13,189.63) --
-	(234.96,208.84) --
-	(252.79,228.05);
-
-\path[draw=drawColor,line width= 1.1pt,line join=round] (145.81, 74.37) --
-	(163.64, 99.99) --
-	(181.47,125.60) --
-	(199.30,151.21) --
-	(217.13,176.83) --
-	(234.96,202.44) --
-	(252.79,228.05);
-
-\path[draw=drawColor,line width= 1.1pt,line join=round] (181.47, 74.37) --
-	(199.30,112.79) --
-	(217.13,151.21) --
-	(234.96,189.63) --
-	(252.79,228.05);
+\path[draw=drawColor,line width= 1.1pt,line join=round] (158.92, 60.39) --
+	(173.77, 72.67) --
+	(188.61, 84.94) --
+	(203.45, 97.22) --
+	(218.29,109.50);
 \end{scope}
 \begin{scope}
-\path[clip] ( 46.86, 40.64) rectangle (262.59,259.48);
+\path[clip] ( 46.86, 40.64) rectangle (226.46,114.94);
 \definecolor{drawColor}{RGB}{0,0,0}
@@
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] ( 55.83, 70.05) --
-	( 74.50, 72.89) --
-	( 93.17, 70.05) --
-	( 93.17, 59.24) --
-	( 55.83, 59.24) --
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (121.01, 57.54) --
+	(129.24, 60.39) --
+	(137.47, 57.54) --
+	(137.47, 48.68) --
+	(121.01, 48.68) --
 	cycle;
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] ( 98.75, 70.05) --
-	(110.16, 74.37) --
-	(121.56, 70.05) --
-	(121.56, 59.24) --
-	( 98.75, 59.24) --
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (150.70, 57.54) --
+	(158.92, 60.39) --
+	(167.15, 57.54) --
+	(167.15, 48.68) --
+	(150.70, 48.68) --
 	cycle;
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (137.59, 70.05) --
-	(145.81, 74.37) --
-	(154.04, 70.05) --
-	(154.04, 61.18) --
-	(137.59, 61.18) --
-	cycle;
+\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at (129.24, 49.23) {$N^2$};
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (173.24, 70.05) --
-	(181.47, 74.37) --
-	(189.70, 70.05) --
-	(189.70, 61.18) --
-	(173.24, 61.18) --
-	cycle;
-
-\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at ( 74.50, 63.16) {$N \log N$};
-
-\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at (110.16, 63.16) {$N^{1.5}$};
-
-\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at (145.81, 63.16) {$N^2$};
-
-\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at (181.47, 63.16) {$N^3$};
-\definecolor{drawColor}{gray}{0.20}
-
-\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 46.86, 40.64) rectangle (262.59,259.48);
+\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at (158.92, 49.23) {$N^3$};
 \end{scope}
 \begin{scope}
-\path[clip] (268.09,264.98) rectangle (483.82,483.82);
-\definecolor{fillColor}{RGB}{255,255,255}
+\path[clip] (231.96,120.44) rectangle (411.55,194.74);
+\definecolor{fillColor}{gray}{0.92}
 
-\path[fill=fillColor] (268.09,264.98) rectangle (483.82,483.82);
-\definecolor{drawColor}{gray}{0.92}
+\path[fill=fillColor] (231.96,120.44) rectangle (411.55,194.74);
+\definecolor{drawColor}{RGB}{255,255,255}
 
-\path[draw=drawColor,line width= 0.3pt,line join=round] (268.09,289.91) --
-	(483.82,289.91);
+\path[draw=drawColor,line width= 0.3pt,line join=round] (231.96,125.99) --
+	(411.55,125.99);
 
-\path[draw=drawColor,line width= 0.3pt,line join=round] (268.09,346.74) --
-	(483.82,346.74);
+\path[draw=drawColor,line width= 0.3pt,line join=round] (231.96,145.23) --
+	(411.55,145.23);
 
-\path[draw=drawColor,line width= 0.3pt,line join=round] (268.09,403.56) --
-	(483.82,403.56);
+\path[draw=drawColor,line width= 0.3pt,line join=round] (231.96,164.47) --
+	(411.55,164.47);
 
-\path[draw=drawColor,line width= 0.3pt,line join=round] (268.09,460.39) --
-	(483.82,460.39);
+\path[draw=drawColor,line width= 0.3pt,line join=round] (231.96,183.70) --
+	(411.55,183.70);
 
-\path[draw=drawColor,line width= 0.3pt,line join=round] (272.36,264.98) --
-	(272.36,483.82);
+\path[draw=drawColor,line width= 0.3pt,line join=round] (235.51,120.44) --
+	(235.51,194.74);
 
-\path[draw=drawColor,line width= 0.3pt,line join=round] (326.65,264.98) --
-	(326.65,483.82);
+\path[draw=drawColor,line width= 0.3pt,line join=round] (280.70,120.44) --
+	(280.70,194.74);
 
-\path[draw=drawColor,line width= 0.3pt,line join=round] (380.94,264.98) --
-	(380.94,483.82);
+\path[draw=drawColor,line width= 0.3pt,line join=round] (325.90,120.44) --
+	(325.90,194.74);
 
-\path[draw=drawColor,line width= 0.3pt,line join=round] (435.23,264.98) --
-	(435.23,483.82);
+\path[draw=drawColor,line width= 0.3pt,line join=round] (371.10,120.44) --
+	(371.10,194.74);
 
-\path[draw=drawColor,line width= 0.6pt,line join=round] (268.09,318.33) --
-	(483.82,318.33);
+\path[draw=drawColor,line width= 0.6pt,line join=round] (231.96,135.61) --
+	(411.55,135.61);
 
-\path[draw=drawColor,line width= 0.6pt,line join=round] (268.09,375.15) --
-	(483.82,375.15);
+\path[draw=drawColor,line width= 0.6pt,line join=round] (231.96,154.85) --
+	(411.55,154.85);
 
-\path[draw=drawColor,line width= 0.6pt,line join=round] (268.09,431.98) --
-	(483.82,431.98);
+\path[draw=drawColor,line width= 0.6pt,line join=round] (231.96,174.08) --
+	(411.55,174.08);
 
-\path[draw=drawColor,line width= 0.6pt,line join=round] (299.50,264.98) --
-	(299.50,483.82);
+\path[draw=drawColor,line width= 0.6pt,line join=round] (231.96,193.32) --
+	(411.55,193.32);
 
-\path[draw=drawColor,line width= 0.6pt,line join=round] (353.79,264.98) --
-	(353.79,483.82);
+\path[draw=drawColor,line width= 0.6pt,line join=round] (258.11,120.44) --
+	(258.11,194.74);
 
-\path[draw=drawColor,line width= 0.6pt,line join=round] (408.08,264.98) --
-	(408.08,483.82);
+\path[draw=drawColor,line width= 0.6pt,line join=round] (303.30,120.44) --
+	(303.30,194.74);
 
-\path[draw=drawColor,line width= 0.6pt,line join=round] (462.37,264.98) --
-	(462.37,483.82);
+\path[draw=drawColor,line width= 0.6pt,line join=round] (348.50,120.44) --
+	(348.50,194.74);
+
+\path[draw=drawColor,line width= 0.6pt,line join=round] (393.69,120.44) --
+	(393.69,194.74);
 \definecolor{drawColor}{RGB}{0,0,0}
 
-\path[draw=drawColor,line width= 2.3pt,line join=round] (277.90,302.81) --
-	(294.24,264.98) --
-	(310.58,264.98) --
-	(326.93,264.98) --
-	(343.27,264.98) --
-	(359.61,274.93) --
-	(375.95,282.47) --
-	(392.30,290.48) --
-	(408.64,298.75) --
-	(424.98,307.16) --
-	(441.33,315.64) --
-	(457.67,324.16) --
-	(474.01,332.70);
+\path[draw=drawColor,line width= 2.3pt,line join=round] (240.12,139.98) --
+	(253.73,120.44) --
+	(267.33,120.44) --
+	(280.94,120.44) --
+	(294.54,120.44) --
+	(308.15,130.54) --
+	(321.75,133.09) --
+	(335.36,135.80) --
+	(348.96,138.60) --
+	(362.57,141.45) --
+	(376.17,144.32) --
+	(389.78,147.21) --
+	(403.39,150.09);
 \definecolor{drawColor}{RGB}{238,130,238}
 
-\path[draw=drawColor,line width= 1.1pt,line join=round] (441.33,298.48) --
-	(457.67,315.59) --
-	(474.01,332.70);
+\path[draw=drawColor,line width= 1.1pt,line join=round] (376.17,138.51) --
+	(389.78,144.30) --
+	(403.39,150.09);
 
-\path[draw=drawColor,line width= 1.1pt,line join=round] (408.64,298.48) --
-	(424.98,307.04) --
-	(441.33,315.59) --
-	(457.67,324.14) --
-	(474.01,332.70);
+\path[draw=drawColor,line width= 1.1pt,line join=round] (348.96,138.51) --
+	(362.57,141.41) --
+	(376.17,144.30) --
+	(389.78,147.20) --
+	(403.39,150.09);
 \end{scope}
 \begin{scope}
-\path[clip] (268.09,264.98) rectangle (483.82,483.82);
+\path[clip] (231.96,120.44) rectangle (411.55,194.74);
 \definecolor{drawColor}{RGB}{0,0,0}
@@
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (433.10,295.64) --
-	(441.33,298.48) --
-	(449.55,295.64) --
-	(449.55,286.77) --
-	(433.10,286.77) --
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (367.95,135.67) --
+	(376.17,138.51) --
+	(384.40,135.67) --
+	(384.40,126.80) --
+	(367.95,126.80) --
 	cycle;
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (402.66,295.64) --
-	(408.64,298.48) --
-	(414.63,295.64) --
-	(414.63,286.77) --
-	(402.66,286.77) --
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (342.98,135.67) --
+	(348.96,138.51) --
+	(354.95,135.67) --
+	(354.95,126.80) --
+	(342.98,126.80) --
 	cycle;
 
-\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at (441.33,288.75) {$N^2$};
+\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at (376.17,127.36) {$N^2$};
 
-\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at (408.64,288.75) {$N$};
-\definecolor{drawColor}{gray}{0.20}
-
-\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (268.09,264.98) rectangle (483.82,483.82);
+\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at (348.96,127.36) {$N$};
 \end{scope}
 \begin{scope}
-\path[clip] (268.09, 40.64) rectangle (483.82,259.48);
-\definecolor{fillColor}{RGB}{255,255,255}
+\path[clip] (231.96, 40.64) rectangle (411.55,114.94);
+\definecolor{fillColor}{gray}{0.92}
 
-\path[fill=fillColor] (268.09, 40.64) rectangle (483.82,259.48);
-\definecolor{drawColor}{gray}{0.92}
+\path[fill=fillColor] (231.96, 40.64) rectangle (411.55,114.94);
+\definecolor{drawColor}{RGB}{255,255,255}
 
-\path[draw=drawColor,line width= 0.3pt,line join=round] (268.09, 45.62) --
-	(483.82, 45.62);
+\path[draw=drawColor,line width= 0.3pt,line join=round] (231.96, 44.40) --
+	(411.55, 44.40);
 
-\path[draw=drawColor,line width= 0.3pt,line join=round] (268.09, 88.17) --
-	(483.82, 88.17);
+\path[draw=drawColor,line width= 0.3pt,line join=round] (231.96, 71.59) --
+	(411.55, 71.59);
 
-\path[draw=drawColor,line width= 0.3pt,line join=round] (268.09,130.71) --
-	(483.82,130.71);
+\path[draw=drawColor,line width= 0.3pt,line join=round] (231.96, 98.78) --
+	(411.55, 98.78);
 
-\path[draw=drawColor,line width= 0.3pt,line join=round] (268.09,173.25) --
-	(483.82,173.25);
+\path[draw=drawColor,line width= 0.3pt,line join=round] (235.51, 40.64) --
+	(235.51,114.94);
 
-\path[draw=drawColor,line width= 0.3pt,line join=round] (268.09,215.80) --
-	(483.82,215.80);
+\path[draw=drawColor,line width= 0.3pt,line join=round] (280.70, 40.64) --
+	(280.70,114.94);
 
-\path[draw=drawColor,line width= 0.3pt,line join=round] (268.09,258.34) --
-	(483.82,258.34);
+\path[draw=drawColor,line width= 0.3pt,line join=round] (325.90, 40.64) --
+	(325.90,114.94);
 
-\path[draw=drawColor,line width= 0.3pt,line join=round] (272.36, 40.64) --
-	(272.36,259.48);
+\path[draw=drawColor,line width= 0.3pt,line join=round] (371.10, 40.64) --
+	(371.10,114.94);
 
-\path[draw=drawColor,line width= 0.3pt,line join=round] (326.65, 40.64) --
-	(326.65,259.48);
+\path[draw=drawColor,line width= 0.6pt,line join=round] (231.96, 58.00) --
+	(411.55, 58.00);
 
-\path[draw=drawColor,line width= 0.3pt,line join=round] (380.94, 40.64) --
-	(380.94,259.48);
+\path[draw=drawColor,line width= 0.6pt,line join=round] (231.96, 85.19) --
+	(411.55, 85.19);
 
-\path[draw=drawColor,line width= 0.3pt,line join=round] (435.23, 40.64) --
-	(435.23,259.48);
+\path[draw=drawColor,line width= 0.6pt,line join=round] (231.96,112.38) --
+	(411.55,112.38);
 
-\path[draw=drawColor,line width= 0.6pt,line join=round] (268.09, 66.89) --
-	(483.82, 66.89);
+\path[draw=drawColor,line width= 0.6pt,line join=round] (258.11, 40.64) --
+	(258.11,114.94);
 
-\path[draw=drawColor,line width= 0.6pt,line join=round] (268.09,109.44) --
-	(483.82,109.44);
+\path[draw=drawColor,line width= 0.6pt,line join=round] (303.30, 40.64) --
+	(303.30,114.94);
 
-\path[draw=drawColor,line width= 0.6pt,line join=round] (268.09,151.98) --
-	(483.82,151.98);
+\path[draw=drawColor,line width= 0.6pt,line join=round] (348.50, 40.64) --
+	(348.50,114.94);
 
-\path[draw=drawColor,line width= 0.6pt,line join=round] (268.09,194.52) --
-	(483.82,194.52);
-
-\path[draw=drawColor,line width= 0.6pt,line join=round] (268.09,237.07) --
-	(483.82,237.07);
-
-\path[draw=drawColor,line width= 0.6pt,line join=round] (299.50, 40.64) --
-	(299.50,259.48);
-
-\path[draw=drawColor,line width= 0.6pt,line join=round] (353.79, 40.64) --
-	(353.79,259.48);
-
-\path[draw=drawColor,line width= 0.6pt,line join=round] (408.08, 40.64) --
-	(408.08,259.48);
-
-\path[draw=drawColor,line width= 0.6pt,line join=round] (462.37, 40.64) --
-	(462.37,259.48);
-\definecolor{drawColor}{RGB}{190,190,190}
-
-\path[draw=drawColor,line width= 0.6pt,line join=round] (268.09,224.26) -- (483.82,224.26);
+\path[draw=drawColor,line width= 0.6pt,line join=round] (393.69, 40.64) --
+	(393.69,114.94);
 \definecolor{fillColor}{RGB}{0,0,0}
 
-\path[fill=fillColor,fill opacity=0.50] (277.90, 74.10) --
-	(294.24, 88.49) --
-	(310.58, 89.28) --
-	(326.93,102.21) --
-	(343.27,116.54) --
-	(359.61,120.50) --
-	(375.95,130.05) --
-	(392.30,145.66) --
-	(408.64,161.32) --
-	(424.98,177.20) --
-	(441.33,203.86) --
-	(457.67,213.34) --
-	(474.01,234.51) --
-	(474.01,233.63) --
-	(457.67,212.74) --
-	(441.33,193.95) --
-	(424.98,176.96) --
-	(408.64,160.86) --
-	(392.30,145.25) --
-	(375.95,129.73) --
-	(359.61,114.31) --
-	(343.27,107.68) --
-	(326.93,100.18) --
-	(310.58, 84.31) --
-	(294.24, 73.17) --
-	(277.90, 56.40) --
+\path[fill=fillColor,fill opacity=0.50] (240.12, 60.30) --
+	(253.73, 64.90) --
+	(267.33, 65.15) --
+	(280.94, 69.28) --
+	(294.54, 73.86) --
+	(308.15, 75.13) --
+	(321.75, 78.18) --
+	(335.36, 83.17) --
+	(348.96, 88.17) --
+	(362.57, 93.25) --
+	(376.17,101.77) --
+	(389.78,104.80) --
+	(403.39,111.56) --
+	(403.39,111.28) --
+	(389.78,104.61) --
+	(376.17, 98.60) --
+	(362.57, 93.17) --
+	(348.96, 88.03) --
+	(335.36, 83.04) --
+	(321.75, 78.08) --
+	(308.15, 73.15) --
+	(294.54, 71.03) --
+	(280.94, 68.63) --
+	(267.33, 63.57) --
+	(253.73, 60.00) --
+	(240.12, 54.65) --
 	cycle;
 
-\path[] (277.90, 74.10) --
-	(294.24, 88.49) --
-	(310.58, 89.28) --
-	(326.93,102.21) --
-	(343.27,116.54) --
-	(359.61,120.50) --
-	(375.95,130.05) --
-	(392.30,145.66) --
-	(408.64,161.32) --
-	(424.98,177.20) --
-	(441.33,203.86) --
-	(457.67,213.34) --
-	(474.01,234.51);
+\path[] (240.12, 60.30) --
+	(253.73, 64.90) --
+	(267.33, 65.15) --
+	(280.94, 69.28) --
+	(294.54, 73.86) --
+	(308.15, 75.13) --
+	(321.75, 78.18) --
+	(335.36, 83.17) --
+	(348.96, 88.17) --
+	(362.57, 93.25) --
+	(376.17,101.77) --
+	(389.78,104.80) --
+	(403.39,111.56);
 
-\path[] (474.01,233.63) --
-	(457.67,212.74) --
-	(441.33,193.95) --
-	(424.98,176.96) --
-	(408.64,160.86) --
-	(392.30,145.25) --
-	(375.95,129.73) --
-	(359.61,114.31) --
-	(343.27,107.68) --
-	(326.93,100.18) --
-	(310.58, 84.31) --
-	(294.24, 73.17) --
-	(277.90, 56.40);
+\path[] (403.39,111.28) --
+	(389.78,104.61) --
+	(376.17, 98.60) --
+	(362.57, 93.17) --
+	(348.96, 88.03) --
+	(335.36, 83.04) --
+	(321.75, 78.08) --
+	(308.15, 73.15) --
+	(294.54, 71.03) --
+	(280.94, 68.63) --
+	(267.33, 63.57) --
+	(253.73, 60.00) --
+	(240.12, 54.65);
 \definecolor{drawColor}{RGB}{0,0,0}
 
-\path[draw=drawColor,line width= 2.3pt,line join=round] (277.90, 57.44) --
-	(294.24, 74.95) --
-	(310.58, 84.63) --
-	(326.93,100.29) --
-	(343.27,112.63) --
-	(359.61,114.70) --
-	(375.95,129.87) --
-	(392.30,145.44) --
-	(408.64,160.99) --
-	(424.98,177.03) --
-	(441.33,195.14) --
-	(457.67,212.79) --
-	(474.01,234.06);
+\path[draw=drawColor,line width= 2.3pt,line join=round] (240.12, 54.98) --
+	(253.73, 60.57) --
+	(267.33, 63.67) --
+	(280.94, 68.67) --
+	(294.54, 72.61) --
+	(308.15, 73.27) --
+	(321.75, 78.12) --
+	(335.36, 83.10) --
+	(348.96, 88.07) --
+	(362.57, 93.19) --
+	(376.17, 98.98) --
+	(389.78,104.62) --
+	(403.39,111.42);
 \definecolor{drawColor}{RGB}{238,130,238}
 
-\path[draw=drawColor,line width= 1.1pt,line join=round] (310.58, 82.84) --
-	(326.93, 99.77) --
-	(343.27,115.95) --
-	(359.61,131.60) --
-	(375.95,146.87) --
-	(392.30,161.86) --
-	(408.64,176.61) --
-	(424.98,191.18) --
-	(441.33,205.59) --
-	(457.67,219.88) --
-	(474.01,234.06);
+\path[draw=drawColor,line width= 1.1pt,line join=round] (267.33, 63.09) --
+	(280.94, 68.50) --
+	(294.54, 73.67) --
+	(308.15, 78.68) --
+	(321.75, 83.56) --
+	(335.36, 88.35) --
+	(348.96, 93.06) --
+	(362.57, 97.72) --
+	(376.17,102.32) --
+	(389.78,106.89) --
+	(403.39,111.42);
 
-\path[draw=drawColor,line width= 1.1pt,line join=round] (343.27, 80.37) --
-	(359.61, 99.58) --
-	(375.95,118.79) --
-	(392.30,138.00) --
-	(408.64,157.21) --
-	(424.98,176.42) --
-	(441.33,195.63) --
-	(457.67,214.85) --
-	(474.01,234.06);
+\path[draw=drawColor,line width= 1.1pt,line join=round] (294.54, 62.31) --
+	(308.15, 68.44) --
+	(321.75, 74.58) --
+	(335.36, 80.72) --
+	(348.96, 86.86) --
+	(362.57, 93.00) --
+	(376.17, 99.14) --
+	(389.78,105.28) --
+	(403.39,111.42);
 
-\path[draw=drawColor,line width= 1.1pt,line join=round] (375.95, 80.37) --
-	(392.30,105.99) --
-	(408.64,131.60) --
-	(424.98,157.21) --
-	(441.33,182.83) --
-	(457.67,208.44) --
-	(474.01,234.06);
-
-\path[draw=drawColor,line width= 1.1pt,line join=round] (408.64, 80.37) --
-	(424.98,118.79) --
-	(441.33,157.21) --
-	(457.67,195.63) --
-	(474.01,234.06);
+\path[draw=drawColor,line width= 1.1pt,line join=round] (321.75, 62.31) --
+	(335.36, 70.49) --
+	(348.96, 78.68) --
+	(362.57, 86.86) --
+	(376.17, 95.05) --
+	(389.78,103.23) --
+	(403.39,111.42);
 \end{scope}
 \begin{scope}
-\path[clip] (268.09, 40.64) rectangle (483.82,259.48);
+\path[clip] (231.96, 40.64) rectangle (411.55,114.94);
 \definecolor{drawColor}{RGB}{0,0,0}
@@
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (291.91, 77.53) --
-	(310.58, 82.84) --
-	(329.26, 77.53) --
-	(329.26, 66.72) --
-	(291.91, 66.72) --
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (247.22, 59.46) --
+	(267.33, 63.09) --
+	(284.57, 59.46) --
+	(284.57, 48.65) --
+	(247.22, 48.65) --
 	cycle;
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (331.86, 77.53) --
-	(343.27, 80.37) --
-	(354.68, 77.53) --
-	(354.68, 66.72) --
-	(331.86, 66.72) --
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (284.57, 59.46) --
+	(294.54, 62.31) --
+	(307.39, 59.46) --
+	(307.39, 48.65) --
+	(284.57, 48.65) --
 	cycle;
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (367.73, 77.53) --
-	(375.95, 80.37) --
-	(384.18, 77.53) --
-	(384.18, 68.66) --
-	(367.73, 68.66) --
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (313.52, 59.46) --
+	(321.75, 62.31) --
+	(329.98, 59.46) --
+	(329.98, 50.60) --
+	(313.52, 50.60) --
 	cycle;
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (400.41, 77.53) --
-	(408.64, 80.37) --
-	(416.87, 77.53) --
-	(416.87, 68.66) --
-	(400.41, 68.66) --
-	cycle;
+\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at (265.90, 51.15) {$N \log N$};
 
-\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at (310.58, 70.64) {$N \log N$};
+\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at (295.98, 51.15) {$N^{1.5}$};
 
-\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at (343.27, 70.64) {$N^{1.5}$};
+\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at (321.75, 51.15) {$N^2$};
+\end{scope}
+\begin{scope}
+\path[clip] ( 46.86,194.74) rectangle (226.46,211.31);
+\definecolor{fillColor}{gray}{0.85}
 
-\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at (375.95, 70.64) {$N^2$};
+\path[fill=fillColor] ( 46.86,194.74) rectangle (226.46,211.31);
+\definecolor{drawColor}{gray}{0.10}
 
-\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at (408.64, 70.64) {$N^3$};
+\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (136.66,199.99) {changepoint};
+\end{scope}
+\begin{scope}
+\path[clip] (231.96,194.74) rectangle (411.55,211.31);
+\definecolor{fillColor}{gray}{0.85}
+
+\path[fill=fillColor] (231.96,194.74) rectangle (411.55,211.31);
+\definecolor{drawColor}{gray}{0.10}
+
+\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (321.75,199.99) {ruptures};
+\end{scope}
+\begin{scope}
+\path[clip] (411.55,120.44) rectangle (428.12,194.74);
+\definecolor{fillColor}{gray}{0.85}
+
+\path[fill=fillColor] (411.55,120.44) rectangle (428.12,194.74);
+\definecolor{drawColor}{gray}{0.10}
+
+\node[text=drawColor,rotate=-90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (416.80,157.59) {kilobytes};
+\end{scope}
+\begin{scope}
+\path[clip] (411.55, 40.64) rectangle (428.12,114.94);
+\definecolor{fillColor}{gray}{0.85}
+
+\path[fill=fillColor] (411.55, 40.64) rectangle (428.12,114.94);
+\definecolor{drawColor}{gray}{0.10}
+
+\node[text=drawColor,rotate=-90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (416.80, 77.79) {seconds};
+\end{scope}
+\begin{scope}
+\path[clip] (  0.00,  0.00) rectangle (433.62,216.81);
 \definecolor{drawColor}{gray}{0.20}
 
-\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (268.09, 40.64) rectangle (483.82,259.48);
+\path[draw=drawColor,line width= 0.6pt,line join=round] ( 74.65, 37.89) --
+	( 74.65, 40.64);
+
+\path[draw=drawColor,line width= 0.6pt,line join=round] (123.95, 37.89) --
+	(123.95, 40.64);
+
+\path[draw=drawColor,line width= 0.6pt,line join=round] (173.26, 37.89) --
+	(173.26, 40.64);
+
+\path[draw=drawColor,line width= 0.6pt,line join=round] (222.56, 37.89) --
+	(222.56, 40.64);
 \end{scope}
 \begin{scope}
-\path[clip] ( 46.86,483.82) rectangle (262.59,500.39);
-\definecolor{drawColor}{gray}{0.20}
-\definecolor{fillColor}{gray}{0.85}
+\path[clip] (  0.00,  0.00) rectangle (433.62,216.81);
+\definecolor{drawColor}{gray}{0.30}
 
-\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] ( 46.86,483.82) rectangle (262.59,500.39);
-\definecolor{drawColor}{gray}{0.10}
+\node[text=drawColor,rotate= 30.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 77.68, 30.44) {10};
 
-\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (154.73,489.07) {changepoint};
+\node[text=drawColor,rotate= 30.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (126.98, 30.44) {100};
+
+\node[text=drawColor,rotate= 30.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (176.29, 30.44) {1000};
+
+\node[text=drawColor,rotate= 30.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (225.59, 30.44) {10000};
 \end{scope}
 \begin{scope}
-\path[clip] (268.09,483.82) rectangle (483.82,500.39);
+\path[clip] (  0.00,  0.00) rectangle (433.62,216.81);
 \definecolor{drawColor}{gray}{0.20}
-\definecolor{fillColor}{gray}{0.85}
 
-\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (268.09,483.82) rectangle (483.82,500.39);
-\definecolor{drawColor}{gray}{0.10}
+\path[draw=drawColor,line width= 0.6pt,line join=round] (258.11, 37.89) --
+	(258.11, 40.64);
 
-\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (375.95,489.07) {ruptures};
+\path[draw=drawColor,line width= 0.6pt,line join=round] (303.30, 37.89) --
+	(303.30, 40.64);
+
+\path[draw=drawColor,line width= 0.6pt,line join=round] (348.50, 37.89) --
+	(348.50, 40.64);
+
+\path[draw=drawColor,line width= 0.6pt,line join=round] (393.69, 37.89) --
+	(393.69, 40.64);
 \end{scope}
 \begin{scope}
-\path[clip] (483.82,264.98) rectangle (500.39,483.82);
-\definecolor{drawColor}{gray}{0.20}
-\definecolor{fillColor}{gray}{0.85}
+\path[clip] (  0.00,  0.00) rectangle (433.62,216.81);
+\definecolor{drawColor}{gray}{0.30}
 
-\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (483.82,264.98) rectangle (500.39,483.82);
-\definecolor{drawColor}{gray}{0.10}
+\node[text=drawColor,rotate= 30.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (261.14, 30.44) {10};
 
-\node[text=drawColor,rotate=-90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (489.07,374.40) {kilobytes};
+\node[text=drawColor,rotate= 30.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (306.33, 30.44) {100};
+
+\node[text=drawColor,rotate= 30.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (351.53, 30.44) {1000};
+
+\node[text=drawColor,rotate= 30.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (396.72, 30.44) {10000};
 \end{scope}
 \begin{scope}
-\path[clip] (483.82, 40.64) rectangle (500.39,259.48);
-\definecolor{drawColor}{gray}{0.20}
-\definecolor{fillColor}{gray}{0.85}
+\path[clip] (  0.00,  0.00) rectangle (433.62,216.81);
+\definecolor{drawColor}{gray}{0.30}
 
-\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (483.82, 40.64) rectangle (500.39,259.48);
-\definecolor{drawColor}{gray}{0.10}
+\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 41.91,132.58) {1e+00};
 
-\node[text=drawColor,rotate=-90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (489.07,150.06) {seconds};
+\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 41.91,151.82) {1e+02};
+
+\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 41.91,171.05) {1e+04};
+
+\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 41.91,190.29) {1e+06};
 \end{scope}
 \begin{scope}
-\path[clip] (  0.00,  0.00) rectangle (505.89,505.89);
+\path[clip] (  0.00,  0.00) rectangle (433.62,216.81);
 \definecolor{drawColor}{gray}{0.20}
 
-\path[draw=drawColor,line width= 0.6pt,line join=round] ( 80.24, 37.89) --
-	( 80.24, 40.64);
+\path[draw=drawColor,line width= 0.6pt,line join=round] ( 44.11,135.61) --
+	( 46.86,135.61);
 
-\path[draw=drawColor,line width= 0.6pt,line join=round] (139.46, 37.89) --
-	(139.46, 40.64);
+\path[draw=drawColor,line width= 0.6pt,line join=round] ( 44.11,154.85) --
+	( 46.86,154.85);
 
-\path[draw=drawColor,line width= 0.6pt,line join=round] (198.69, 37.89) --
-	(198.69, 40.64);
+\path[draw=drawColor,line width= 0.6pt,line join=round] ( 44.11,174.08) --
+	( 46.86,174.08);
 
-\path[draw=drawColor,line width= 0.6pt,line join=round] (257.92, 37.89) --
-	(257.92, 40.64);
+\path[draw=drawColor,line width= 0.6pt,line join=round] ( 44.11,193.32) --
+	( 46.86,193.32);
 \end{scope}
 \begin{scope}
-\path[clip] (  0.00,  0.00) rectangle (505.89,505.89);
+\path[clip] (  0.00,  0.00) rectangle (433.62,216.81);
 \definecolor{drawColor}{gray}{0.30}
 
-\node[text=drawColor,rotate= 30.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 83.27, 30.44) {10};
+\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 41.91, 54.97) {1e-03};
 
-\node[text=drawColor,rotate= 30.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (142.49, 30.44) {100};
+\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 41.91, 82.16) {1e-01};
 
-\node[text=drawColor,rotate= 30.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (201.72, 30.44) {1000};
-
-\node[text=drawColor,rotate= 30.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (260.95, 30.44) {10000};
+\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 41.91,109.35) {1e+01};
 \end{scope}
 \begin{scope}
-\path[clip] (  0.00,  0.00) rectangle (505.89,505.89);
+\path[clip] (  0.00,  0.00) rectangle (433.62,216.81);
 \definecolor{drawColor}{gray}{0.20}
 
-\path[draw=drawColor,line width= 0.6pt,line join=round] (299.50, 37.89) --
-	(299.50, 40.64);
+\path[draw=drawColor,line width= 0.6pt,line join=round] ( 44.11, 58.00) --
+	( 46.86, 58.00);
 
-\path[draw=drawColor,line width= 0.6pt,line join=round] (353.79, 37.89) --
-	(353.79, 40.64);
+\path[draw=drawColor,line width= 0.6pt,line join=round] ( 44.11, 85.19) --
+	( 46.86, 85.19);
 
-\path[draw=drawColor,line width= 0.6pt,line join=round] (408.08, 37.89) --
-	(408.08, 40.64);
-
-\path[draw=drawColor,line width= 0.6pt,line join=round] (462.37, 37.89) --
-	(462.37, 40.64);
+\path[draw=drawColor,line width= 0.6pt,line join=round] ( 44.11,112.38) --
+	( 46.86,112.38);
 \end{scope}
 \begin{scope}
-\path[clip] (  0.00,  0.00) rectangle (505.89,505.89);
-\definecolor{drawColor}{gray}{0.30}
-
-\node[text=drawColor,rotate= 30.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (302.53, 30.44) {10};
-
-\node[text=drawColor,rotate= 30.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (356.82, 30.44) {100};
-
-\node[text=drawColor,rotate= 30.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (411.11, 30.44) {1000};
-
-\node[text=drawColor,rotate= 30.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (465.40, 30.44) {10000};
-\end{scope}
-\begin{scope}
-\path[clip] (  0.00,  0.00) rectangle (505.89,505.89);
-\definecolor{drawColor}{gray}{0.30}
-
-\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 41.91,315.30) {1e+01};
-
-\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 41.91,372.12) {1e+03};
-
-\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 41.91,428.95) {1e+05};
-\end{scope}
-\begin{scope}
-\path[clip] (  0.00,  0.00) rectangle (505.89,505.89);
-\definecolor{drawColor}{gray}{0.20}
-
-\path[draw=drawColor,line width= 0.6pt,line join=round] ( 44.11,318.33) --
-	( 46.86,318.33);
-
-\path[draw=drawColor,line width= 0.6pt,line join=round] ( 44.11,375.15) --
-	( 46.86,375.15);
-
-\path[draw=drawColor,line width= 0.6pt,line join=round] ( 44.11,431.98) --
-	( 46.86,431.98);
-\end{scope}
-\begin{scope}
-\path[clip] (  0.00,  0.00) rectangle (505.89,505.89);
-\definecolor{drawColor}{gray}{0.30}
-
-\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 41.91, 63.86) {1e-03};
-
-\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 41.91,106.41) {1e-02};
-
-\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 41.91,148.95) {1e-01};
-
-\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 41.91,191.49) {1e+00};
-
-\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 41.91,234.04) {1e+01};
-\end{scope}
-\begin{scope}
-\path[clip] (  0.00,  0.00) rectangle (505.89,505.89);
-\definecolor{drawColor}{gray}{0.20}
-
-\path[draw=drawColor,line width= 0.6pt,line join=round] ( 44.11, 66.89) --
-	( 46.86, 66.89);
-
-\path[draw=drawColor,line width= 0.6pt,line join=round] ( 44.11,109.44) --
-	( 46.86,109.44);
-
-\path[draw=drawColor,line width= 0.6pt,line join=round] ( 44.11,151.98) --
-	( 46.86,151.98);
-
-\path[draw=drawColor,line width= 0.6pt,line join=round] ( 44.11,194.52) --
-	( 46.86,194.52);
-
-\path[draw=drawColor,line width= 0.6pt,line join=round] ( 44.11,237.07) --
-	( 46.86,237.07);
-\end{scope}
-\begin{scope}
-\path[clip] (  0.00,  0.00) rectangle (505.89,505.89);
+\path[clip] (  0.00,  0.00) rectangle (433.62,216.81);
 \definecolor{drawColor}{RGB}{0,0,0}
 
-\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at (265.34,  7.64) {N};
+\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at (229.21,  7.64) {$N$ = number of data to segment};
 \end{scope}

### 2025-05-06 figure-heap.R

@@ -8,9 +8,13 @@ fun_list[["N^1.5"]] <- function(N)1.5 * log10(N)
 ref_list <- atime::references_best(atime_list, fun_list)
 ref_list$plot.references <- ref_list$references[(
   fun.name %in% c("N","N^2") & unit=="kilobytes"
+)|(unit=="seconds" & ( (
+  grepl("binsegRcpp", expr.name) & fun.name %in% c("N log N", "N^2")
 )|(
-  fun.name != "N" & unit=="seconds"
-)]
+  expr.name=="ruptures" & !fun.name%in%c("N^3","N")
+)|(
+  expr.name=="changepoint" & fun.name%in%c("N^2","N^3")
+)))]
 ##ref_list$measurements <- ref_list
 png("figure-heap-refs-binsegRcpp.png",width=6,height=3.5,units="in",res=200)
 plot_algos <- function(regex){
@@ -33,28 +37,54 @@ png("figure-heap-refs-other.png",width=6,height=3.5,units="in",res=200)
 plot_algos("change|rupt")
 dev.off()
 
-tikz_algos <- function(regex){
-  regex_list <- ref_list
-  regex_list$measurements <- ref_list$measurements[grepl(regex,expr.name)]
-  regex_list$plot.references <- ref_list$plot.references[grepl(regex,expr.name)][
-   ,fun.name := sprintf("$%s$", sub("1.5","{1.5}",fun.latex))][]
-  plot(regex_list)+
+tikz_algos <- function(f.tex,regex){
+  trans <- function(DT)DT[,expr.name:=sub("_","\\_",expr.name,fixed=TRUE)]
+  meas <- trans(ref_list$measurements[grepl(regex,expr.name)])
+  ref.dt <- trans(ref_list$plot.references[grepl(regex,expr.name)])[,fun.name := sprintf("$%s$", sub("1.5","{1.5}",fun.latex))][]
+  ##plot(regex_list)+
+  emp.color <- "black"
+  ref.color <- "violet"
+  gg <- ggplot()+
+    facet_grid(unit ~ expr.name, scales="free")+
+    ggplot2::geom_ribbon(ggplot2::aes(
+      N, ymin = min, 
+      ymax = max, group = expr.name),
+      data = meas[unit == "seconds"],
+      fill = emp.color,
+      alpha = 0.5) +
+    ggplot2::geom_line(ggplot2::aes(
+      N, empirical, group = expr.name),
+      size = 2,
+      color = emp.color, 
+      data = meas) +
[…]

### 2025-05-07 article.Rnw
tikz figures
direct label in tikz
N^1.4

@@
 
-\subsection{C++ features}
+\subsection{Choice of container}
 
@@
 After computing the cost for two new segments, the cost of all splittable segments must be considered, in order to identify the min cost segment to split next.
-We therefore propose using two different kinds of C++ Standard Template Library containers to store and retreive the cost of segments which have previously been considered, but not yet split (Table~\ref{tab:containers}).
-First, we propose using the \code{multiset} container (red-black tree), which can be used to achieve best case time complexity.
-The \code{multiset} keeps segments sorted by their cost values, so offers constant $O(1)$ time retreival of the min cost segment to split next; insertion takes logarithmic $O(\log n)$ time in the number of segments $n$ currently stored in the container.
-Second, we propose using the \code{list} container (doubly-linked list), which is used as a baseline, to demonstrate the benefits of the multiset container.
+We propose using three different kinds of C++ Standard Template Library containers to store and retreive the cost of segments which have previously been considered, but not yet split (Table~\ref{tab:containers}).
+First, we propose using the \code{multiset} container (red-black tree) and \code{priority\_queue} container (heap), which can be used to achieve best case time complexity.
+These containers keep segments sorted by their cost values, so offer constant $O(1)$ time retreival of the min cost segment to split next; insertion takes logarithmic $O(\log n)$ time in the number of segments $n$ currently stored in the container.
+Second, we propose using the \code{list} container (doubly-linked list), which is used as a baseline, to demonstrate the benefits of the other containers.
 The \code{list} keeps segments stored in the order in which they were inserted, so offers constant $O(1)$ time insertion of new elements, and linear $O(n)$ time retreival of the min cost segment to split, for a container with $n$ segments currently stored.
 
-A C++ virtual class \code{Container} is used to represent the methods that must be provided by either type of container: \code{insert}, \code{get_best}.
-We used a C macro, \code{CMAKER}, to declare the two sub-classes:
-\begin{verbatim}
-#define CMAKER(CONTAINER, INSERT, GET_BEST)...
-CMAKER(multiset, insert, segment_container.begin())
-CMAKER(list, push_back, std::min_element(
- segment_container.begin(),segment_container.end()))
-\end{verbatim}
-The code above is expanded by the C pre-processor, to two declarations of \code{Container} sub-classes.
+%% A C++ virtual class \code{Container} is used to represent the methods that must be provided by either type of container: \code{insert}, \code{get_best}.
+%% We used a C macro, \code{CMAKER}, to declare the two sub-classes:
+%% \begin{verbatim}
+%% #define CMAKER(CONTAINER, INSERT, GET_BEST)...
+%% CMAKER(multiset, insert, segment_container.begin())
+%% CMAKER(list, push_back, std::min_element(
+%%  segment_container.begin(),segment_container.end()))
+%% \end{verbatim}
+%% The code above is expanded by the C pre-processor, to two declarations of \code{Container} sub-classes.
 
-Another C macro is used to declare the loss functions which require the cumulative median:
+\subsection{Choice of loss function}
+
+C macros are used to declare the different loss functions.
+For loss functions which require the cumulative median, we use the C++ code below
 \begin{verbatim}
 #define ABS_DIST(NAME, VARIANCE)...
+ABS_DIST(laplace, true)
 ABS_DIST(l1, false)
-ABS_DIST(laplace, true)
 \end{verbatim}
@@
 CUM_DIST(mean_norm, RSS, , false)
+CUM_DIST(meanvar_norm,
+ (var>max_zero_var) ? (RSS/var+N*log(2*M_PI*var))/2 : INFINITY,
+ , true)
 CUM_DIST(poisson,
@@
  false)
-CUM_DIST(meanvar_norm,
- (var>max_zero_var) ? (RSS/var+N*log(2*M_PI*var))/2 : INFINITY,
- , true)
 \end{verbatim}
@@
   \centering
-  \scriptsize
+  \small
 \begin{tabular}{l|cccccc}
@@
   dim & one & one & one & multi & multi & multi \\
-  correct & yes & no & yes & yes & yes \\
+  correct & yes & no & yes & yes & yes & yes\\
   losses & 5 & 6 & 1 & 1 & 10 & 5\\
@@
   % all propose some loss computation.
-  storage & STL multiset & arrays & recursion & heap & LRU cache & heap \\
+  storage & heap & arrays & recursion & heap & LRU cache & heap \\
   space & $O(S)$ & $O(S^2)$ & $O(S)$ & $O(S)$ & $O(S)$ & $O(S)$ \\
   cumsum & yes & yes & yes & yes & no & yes\\
-  best & $O(N\log N)$ & $O(N^3)$ & $O(N\log N)$ & $O(N\log N)$ & $>O(N\log N)$ & $O(N\log N)$\\
+  best & $O(N\log N)$ & $O(N^3)$ & $O(N\log N)$ & $O(N\log N)$ & $O(N^{1.4})$ & $O(N\log N)$\\
   worst & $O(N^2)$ & $O(N^3)$ & $O(N^2)$ & $O(N^2)$ & $O(N^2)$ & $O(N^2)$\\
@@
 \caption{\label{tab:overview}
-  Properties (rows) of five binary segmentation libraries (columns).
+  Properties (rows) of six binary segmentation libraries (columns).
 Properties are weights (can specify observation-specific positive weights for the loss), max segs (can specify maximum number of segments parameter), dim (dimensions of input data), correct (does code return correct model parameters?), losses (number of loss functions implemented), language (programming language to implement loss), storage (technique used to store cost of splitting a segment), space (space complexity in number of segments $S$), cumsum (cumulative sum used), best (time complexity in best case data), worst (time complexity in worst case data), CV (support for cross-validation), params (parameters are returned).
@@
 For example, in the second row, \code{end=2} means that the first change-point results in a split between data points 2 and 3.
-The \code{before.mean = -3} and \code{after.mean=6} values are the means of the resulting segments, before and after that split.
-The other columns, \code{invalidates.index} and \code{invalidates.after}, indicate the previous means which are no longer valid, after the given split.
+The \code{before.mean = -3} is the mean up to data point 2, and \code{after.mean=6} is the mean from data point 3 to the end.
+The other columns, \code{invalidates.index} and \code{invalidates.after}, indicate the previous means which are no longer valid, for models with the given split.
 For example, in the second row, \code{invalidates.index=1} identifies the first row, and \code{invalidates.after=0} identifies the mean before the split (not after), which implies that \code{before.mean=3} is no longer valid, after performing the split in the second row.
-In the third row, we see a new split at \code{end=4}, resulting in new parameters \code{before.mean=9} and \code{after.mean=3}, which invalidate the previous mean in row 2 after the split (\code{after.mean=6}).
+In the third row, we see a new split at \code{end=4}, resulting in new parameters \code{before.mean=9} and \code{after.mean=3}, which invalidate the previous mean in row 2 after the split (\code{after.mean=6}), while the other mean in row 2 remains valid (\code{before.mean = -3}).
 
@@
 non-missing mean values.
-For $K$ segments, this algorithm is $O(K log K)$ time due to the sort.
+For $K$ segments, this algorithm is $O(K \log K)$ time due to the sort.
 It is implemented in the \code{coef} method, as shown in the example code below:
@@
 First, there are only three values, so these should be the segment means for the model with two change-points.
-However, these three values are incorrect (the correct segment means are 0, 9, and 3).
+However, these three values are incorrect (the correct segment means are -3, 9, and 3).
 Another issue is that the segment means are only reported for one model size,
@@
   \includegraphics[width=\textwidth]{figure-heap-pred}
-  \caption{\label{fig:heap-pred} Comparing computation time of \pkg{binsegRcpp} with three containers, and two baselines: \pkg{changepoint} and \pkg{ruptures}.}
+  \caption{\label{fig:heap-pred} Comparing best-case computation time of \pkg{binsegRcpp} with three containers, and two baselines: \pkg{changepoint} and \pkg{ruptures}.
+    Square loss and best case data were used (integers from 1 to $N$, result in equal splits).
+  }
 \end{figure}
@@
   \centering
-  \includegraphics[width=\textwidth]{figure-heap-refs-binsegRcpp}
-  \caption{\label{fig:heap-refs} Estimation of asymptotic time (seconds) and memory (kilobytes) usage for \pkg{binsegRcpp}.
-    Black curves are empirical measurements, and violet curves are references.
-    These data suggest that both list and multiset containers are $O(N)$;
-    list is $O(N^2)$ time and multiset is $O(N\log N)$ time.  }
+  %\includegraphics[width=\textwidth]{figure-heap-refs-binsegRcpp}
+  \input{figure-heap-refs-binsegRcpp}
+  \caption{\label{fig:heap-refs}
+    Estimating best-case asymptotic complexity for time (seconds) and memory (kilobytes).
+    Black curves are empirical measurements for \pkg{binsegRcpp} with two containers, and violet curves are references.
+    These data suggest that both list and multiset containers are $O(N)$ space
+    list is $O(N^2)$ time and multiset is $O(N\log N)$ time.
+    Square loss and best case data were used (integers from 1 to $N$, result in equal splits).
+  }
 \end{figure}
@@
   \centering
-  \includegraphics[width=\textwidth]{figure-heap-refs-other}
-  \caption{\label{fig:heap-refs} Estimation of asymptotic time (seconds) and memory (kilobytes) usage for \pkg{changepoint} and \pkg{ruptures}.
-    Black curves are empirical measurements, and violet curves are references.
+  %\includegraphics[width=\textwidth]{figure-heap-refs-other}
+  \input{figure-heap-refs-other}
+  \caption{\label{fig:heap-refs}
+    Estimating best-case asymptotic complexity for time (seconds) and memory (kilobytes).
+    Black curves are empirical measurements for \pkg{changepoint} and \pkg{ruptures}, and violet curves are references.
     These data suggest that \pkg{changepoint} is $O(N^2)$ space and $O(N^3)$ time;
-    \pkg{ruptures} is $O(N)$ space and $O(N^{1.5})$ time.
+    \pkg{ruptures} is $O(N)$ space and $\approx O(N^{1.4})$ time.
+    Square loss and best case data were used (integers from 1 to $N$, result in equal splits).
   }

### 2025-05-07 figure-heap-refs-binsegRcpp.tex
N^1.4

@@
-% Created by tikzDevice version 0.12.6 on 2025-05-06 22:53:29
+% Created by tikzDevice version 0.12.6 on 2025-05-07 07:42:32
 % !TEX encoding = UTF-8 Unicode
@@
 
-\path[draw=drawColor,line width= 2.3pt,line join=round] ( 55.03,150.65) --
+\path[draw=drawColor,line width= 2.3pt,line join=round] ( 55.03,144.09) --
 	( 65.91,144.09) --
@@
 
-\path[draw=drawColor,line width= 0.3pt,line join=round] ( 46.86, 45.09) --
-	(226.46, 45.09);
+\path[draw=drawColor,line width= 0.3pt,line join=round] ( 46.86, 48.78) --
+	(226.46, 48.78);
 
-\path[draw=drawColor,line width= 0.3pt,line join=round] ( 46.86, 68.35) --
-	(226.46, 68.35);
+\path[draw=drawColor,line width= 0.3pt,line join=round] ( 46.86, 70.29) --
+	(226.46, 70.29);
 
-\path[draw=drawColor,line width= 0.3pt,line join=round] ( 46.86, 91.60) --
-	(226.46, 91.60);
+\path[draw=drawColor,line width= 0.3pt,line join=round] ( 46.86, 91.80) --
+	(226.46, 91.80);
 
@@
 
-\path[draw=drawColor,line width= 0.6pt,line join=round] ( 46.86, 56.72) --
-	(226.46, 56.72);
+\path[draw=drawColor,line width= 0.6pt,line join=round] ( 46.86, 59.54) --
+	(226.46, 59.54);
 
-\path[draw=drawColor,line width= 0.6pt,line join=round] ( 46.86, 79.97) --
-	(226.46, 79.97);
+\path[draw=drawColor,line width= 0.6pt,line join=round] ( 46.86, 81.05) --
+	(226.46, 81.05);
 
-\path[draw=drawColor,line width= 0.6pt,line join=round] ( 46.86,103.23) --
-	(226.46,103.23);
+\path[draw=drawColor,line width= 0.6pt,line join=round] ( 46.86,102.56) --
+	(226.46,102.56);
 
@@
 
-\path[fill=fillColor,fill opacity=0.50] ( 55.03, 60.40) --
-	( 65.91, 60.67) --
-	( 76.80, 60.43) --
-	( 87.68, 60.60) --
-	( 98.56, 60.76) --
-	(109.45, 59.50) --
-	(120.33, 66.90) --
-	(131.22, 57.88) --
-	(142.10, 60.23) --
-	(152.99, 64.81) --
-	(163.87, 70.91) --
-	(174.76, 76.91) --
-	(185.64, 83.30) --
-	(196.52, 91.03) --
-	(207.41, 98.65) --
+\path[fill=fillColor,fill opacity=0.50] ( 55.03, 60.60) --
+	( 65.91, 61.69) --
+	( 76.80, 58.47) --
+	( 87.68, 58.55) --
+	( 98.56, 59.43) --
+	(109.45, 59.33) --
+	(120.33, 60.10) --
+	(131.22, 61.82) --
+	(142.10, 64.29) --
+	(152.99, 67.15) --
+	(163.87, 73.58) --
+	(174.76, 79.78) --
+	(185.64, 84.15) --
+	(196.52, 93.41) --
+	(207.41, 99.34) --
 	(218.29,107.54) --
-	(218.29,106.55) --
-	(207.41, 98.54) --
-	(196.52, 90.85) --
-	(185.64, 82.89) --
-	(174.76, 76.69) --
-	(163.87, 70.67) --
-	(152.99, 64.48) --
-	(142.10, 58.99) --
-	(131.22, 56.22) --
-	(120.33, 55.43) --
-	(109.45, 56.30) --
-	( 98.56, 58.85) --
-	( 87.68, 58.94) --
-	( 76.80, 59.07) --
-	( 65.91, 58.74) --
-	( 55.03, 58.77) --
+	(218.29,106.75) --
+	(207.41, 99.27) --
+	(196.52, 93.30) --
+	(185.64, 84.11) --
+	(174.76, 79.60) --
+	(163.87, 72.14) --
+	(152.99, 66.65) --
+	(142.10, 62.14) --
+	(131.22, 61.06) --
+	(120.33, 58.46) --
+	(109.45, 57.76) --
+	( 98.56, 57.98) --
+	( 87.68, 57.21) --
+	( 76.80, 57.18) --
+	( 65.91, 57.79) --
+	( 55.03, 57.45) --
 	cycle;
 
-\path[] ( 55.03, 60.40) --
-	( 65.91, 60.67) --
-	( 76.80, 60.43) --
-	( 87.68, 60.60) --
-	( 98.56, 60.76) --
-	(109.45, 59.50) --
-	(120.33, 66.90) --
-	(131.22, 57.88) --
-	(142.10, 60.23) --
-	(152.99, 64.81) --
-	(163.87, 70.91) --
-	(174.76, 76.91) --
-	(185.64, 83.30) --
-	(196.52, 91.03) --
-	(207.41, 98.65) --
+\path[] ( 55.03, 60.60) --
+	( 65.91, 61.69) --
+	( 76.80, 58.47) --
+	( 87.68, 58.55) --
+	( 98.56, 59.43) --
+	(109.45, 59.33) --
+	(120.33, 60.10) --
+	(131.22, 61.82) --
+	(142.10, 64.29) --
+	(152.99, 67.15) --
+	(163.87, 73.58) --
+	(174.76, 79.78) --
+	(185.64, 84.15) --
+	(196.52, 93.41) --
+	(207.41, 99.34) --
 	(218.29,107.54);
 
-\path[] (218.29,106.55) --
-	(207.41, 98.54) --
-	(196.52, 90.85) --
-	(185.64, 82.89) --
-	(174.76, 76.69) --
-	(163.87, 70.67) --
-	(152.99, 64.48) --
-	(142.10, 58.99) --
-	(131.22, 56.22) --
-	(120.33, 55.43) --
-	(109.45, 56.30) --
-	( 98.56, 58.85) --
-	( 87.68, 58.94) --
-	( 76.80, 59.07) --
-	( 65.91, 58.74) --
-	( 55.03, 58.77);
+\path[] (218.29,106.75) --
+	(207.41, 99.27) --
+	(196.52, 93.30) --
+	(185.64, 84.11) --
+	(174.76, 79.60) --
+	(163.87, 72.14) --
+	(152.99, 66.65) --
+	(142.10, 62.14) --
+	(131.22, 61.06) --
+	(120.33, 58.46) --
+	(109.45, 57.76) --
+	( 98.56, 57.98) --
+	( 87.68, 57.21) --
+	( 76.80, 57.18) --
+	( 65.91, 57.79) --
+	( 55.03, 57.45);
 \definecolor{drawColor}{RGB}{0,0,0}
 
-\path[draw=drawColor,line width= 2.3pt,line join=round] ( 55.03, 59.02) --
-	( 65.91, 59.16) --
-	( 76.80, 59.39) --
-	( 87.68, 59.17) --
-	( 98.56, 59.18) --
-	(109.45, 57.11) --
-	(120.33, 56.13) --
-	(131.22, 56.68) --
-	(142.10, 59.59) --
-	(152.99, 64.67) --
-	(163.87, 70.76) --
-	(174.76, 76.78) --
-	(185.64, 83.14) --
-	(196.52, 90.91) --
-	(207.41, 98.58) --
-	(218.29,106.64);
+\path[draw=drawColor,line width= 2.3pt,line join=round] ( 55.03, 60.14) --
+	( 65.91, 58.16) --
+	( 76.80, 57.45) --
+	( 87.68, 57.47) --
+	( 98.56, 58.51) --
+	(109.45, 58.09) --
+	(120.33, 59.09) --
+	(131.22, 61.38) --
+	(142.10, 63.34) --
+	(152.99, 66.94) --
+	(163.87, 72.43) --
+	(174.76, 79.66) --
+	(185.64, 84.13) --
+	(196.52, 93.39) --
+	(207.41, 99.29) --
+	(218.29,107.49);
 \definecolor{drawColor}{RGB}{238,130,238}
 
-\path[draw=drawColor,line width= 1.1pt,line join=round] ( 87.68, 58.45) --
-	( 98.56, 62.87) --
-	(109.45, 67.15) --
-	(120.33, 71.33) --
-	(131.22, 75.42) --
-	(142.10, 79.46) --
-	(152.99, 83.44) --
-	(163.87, 87.38) --
-	(174.76, 91.28) --
-	(185.64, 95.16) --
-	(196.52, 99.01) --
-	(207.41,102.83) --
-	(218.29,106.64);
+\path[draw=drawColor,line width= 1.1pt,line join=round] ( 76.80, 58.64) --
+	( 87.68, 62.92) --
+	( 98.56, 67.01) --
+	(109.45, 70.97) --
+	(120.33, 74.83) --
+	(131.22, 78.62) --
+	(142.10, 82.35) --
+	(152.99, 86.03) --
+	(163.87, 89.68) --
+	(174.76, 93.29) --
+	(185.64, 96.87) --
+	(196.52,100.43) --
+	(207.41,103.97) --
+	(218.29,107.49);
 
-\path[draw=drawColor,line width= 1.1pt,line join=round] (152.99, 64.63) --
-	(163.87, 71.63) --
-	(174.76, 78.64) --
-	(185.64, 85.64) --
-	(196.52, 92.64) --
-	(207.41, 99.64) --
-	(218.29,106.64);
+\path[draw=drawColor,line width= 1.1pt,line join=round] (142.10, 62.16) --
+	(152.99, 68.64) --
+	(163.87, 75.11) --
+	(174.76, 81.59) --
+	(185.64, 88.06) --
+	(196.52, 94.54) --
+	(207.41,101.02) --
+	(218.29,107.49);
 \end{scope}
@@
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] ( 69.01, 55.61) --
-	( 87.68, 58.45) --
-	(106.35, 55.61) --
-	(106.35, 44.80) --
-	( 69.01, 44.80) --
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] ( 58.12, 55.80) --
+	( 76.80, 58.64) --
+	( 95.47, 55.80) --
+	( 95.47, 44.99) --
+	( 58.12, 44.99) --
 	cycle;
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (144.76, 55.61) --
-	(152.99, 64.63) --
-	(161.21, 55.61) --
-	(161.21, 46.74) --
-	(144.76, 46.74) --
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (133.87, 55.80) --
+	(142.10, 62.16) --
+	(150.33, 55.80) --
+	(150.33, 46.93) --
+	(133.87, 46.93) --
 	cycle;
 
-\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at ( 87.68, 47.30) {$N \log N$};
+\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at ( 76.80, 47.49) {$N \log N$};
 
-\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at (152.99, 47.30) {$N^2$};
+\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at (142.10, 47.49) {$N^2$};
 \end{scope}
@@
 
-\path[draw=drawColor,line width= 0.3pt,line join=round] (231.96, 45.09) --
-	(411.55, 45.09);
+\path[draw=drawColor,line width= 0.3pt,line join=round] (231.96, 48.78) --
+	(411.55, 48.78);
 
-\path[draw=drawColor,line width= 0.3pt,line join=round] (231.96, 68.35) --
-	(411.55, 68.35);
+\path[draw=drawColor,line width= 0.3pt,line join=round] (231.96, 70.29) --
+	(411.55, 70.29);
 
-\path[draw=drawColor,line width= 0.3pt,line join=round] (231.96, 91.60) --
-	(411.55, 91.60);
+\path[draw=drawColor,line width= 0.3pt,line join=round] (231.96, 91.80) --
+	(411.55, 91.80);
 
@@
 
-\path[draw=drawColor,line width= 0.6pt,line join=round] (231.96, 56.72) --
-	(411.55, 56.72);
+\path[draw=drawColor,line width= 0.6pt,line join=round] (231.96, 59.54) --
+	(411.55, 59.54);
 
-\path[draw=drawColor,line width= 0.6pt,line join=round] (231.96, 79.97) --
-	(411.55, 79.97);
+\path[draw=drawColor,line width= 0.6pt,line join=round] (231.96, 81.05) --
+	(411.55, 81.05);
 
-\path[draw=drawColor,line width= 0.6pt,line join=round] (231.96,103.23) --
-	(411.55,103.23);
+\path[draw=drawColor,line width= 0.6pt,line join=round] (231.96,102.56) --
+	(411.55,102.56);
 
@@
 
-\path[fill=fillColor,fill opacity=0.50] (240.12, 64.16) --
-	(247.89, 60.85) --
-	(255.67, 59.78) --
-	(263.44, 69.26) --
-	(271.22, 59.63) --
-	(278.99, 57.31) --
-	(286.77, 56.13) --
-	(294.54, 57.00) --
-	(302.32, 57.15) --
-	(310.09, 59.83) --
-	(317.87, 68.12) --
-	(325.64, 69.02) --
-	(333.41, 67.60) --
-	(341.19, 71.94) --
-	(348.96, 75.31) --
-	(356.74, 78.25) --
-	(364.51, 84.54) --
-	(372.29, 85.40) --
-	(380.06, 88.92) --
-	(387.84, 93.02) --
-	(395.61, 96.74) --
-	(403.39,100.73) --
-	(403.39,100.39) --
+\path[fill=fillColor,fill opacity=0.50] (240.12, 58.27) --
+	(247.89, 58.20) --
+	(255.67, 57.82) --
+	(263.44, 73.31) --
+	(271.22, 58.78) --
+	(278.99, 58.24) --
+	(286.77, 59.39) --
+	(294.54, 61.03) --
+	(302.32, 60.46) --
+	(310.09, 62.27) --
+	(317.87, 64.16) --
+	(325.64, 74.97) --
+	(333.41, 69.26) --
+	(341.19, 78.78) --
+	(348.96, 76.56) --
+	(356.74, 79.86) --
+	(364.51, 82.84) --
+	(372.29, 85.90) --
+	(380.06, 89.36) --
+	(387.84, 93.13) --
 	(395.61, 96.28) --
-	(387.84, 92.69) --
-	(380.06, 88.71) --
-	(372.29, 85.21) --
-	(364.51, 81.29) --
-	(356.74, 77.87) --
-	(348.96, 74.21) --
-	(341.19, 71.44) --
-	(333.41, 67.10) --
-	(325.64, 63.68) --
-	(317.87, 61.35) --
-	(310.09, 58.41) --
-	(302.32, 56.50) --
-	(294.54, 55.57) --
-	(286.77, 54.88) --
-	(278.99, 56.47) --
-	(271.22, 58.80) --
-	(263.44, 58.90) --
-	(255.67, 58.75) --
-	(247.89, 58.63) --
-	(240.12, 59.84) --
+	(403.39,100.05) --
+	(403.39, 99.70) --
+	(395.61, 96.10) --
+	(387.84, 92.95) --
+	(380.06, 89.10) --
+	(372.29, 85.75) --
+	(364.51, 82.37) --
+	(356.74, 79.13) --
+	(348.96, 75.93) --
+	(341.19, 73.00) --
+	(333.41, 69.00) --
+	(325.64, 66.39) --
+	(317.87, 63.54) --
+	(310.09, 61.16) --
+	(302.32, 59.96) --
+	(294.54, 58.76) --
+	(286.77, 58.17) --
+	(278.99, 57.49) --
+	(271.22, 57.71) --
+	(263.44, 57.10) --
+	(255.67, 57.13) --
+	(247.89, 57.13) --
+	(240.12, 57.43) --
 	cycle;
 
-\path[] (240.12, 64.16) --
-	(247.89, 60.85) --
-	(255.67, 59.78) --
-	(263.44, 69.26) --
-	(271.22, 59.63) --
-	(278.99, 57.31) --
-	(286.77, 56.13) --
-	(294.54, 57.00) --
-	(302.32, 57.15) --
-	(310.09, 59.83) --
-	(317.87, 68.12) --
-	(325.64, 69.02) --
-	(333.41, 67.60) --
-	(341.19, 71.94) --
-	(348.96, 75.31) --
-	(356.74, 78.25) --
-	(364.51, 84.54) --
-	(372.29, 85.40) --
-	(380.06, 88.92) --
-	(387.84, 93.02) --
-	(395.61, 96.74) --
-	(403.39,100.73);
+\path[] (240.12, 58.27) --
+	(247.89, 58.20) --
+	(255.67, 57.82) --
+	(263.44, 73.31) --
+	(271.22, 58.78) --
+	(278.99, 58.24) --
+	(286.77, 59.39) --
+	(294.54, 61.03) --
+	(302.32, 60.46) --
+	(310.09, 62.27) --
+	(317.87, 64.16) --
+	(325.64, 74.97) --
+	(333.41, 69.26) --
+	(341.19, 78.78) --
+	(348.96, 76.56) --
+	(356.74, 79.86) --
+	(364.51, 82.84) --
+	(372.29, 85.90) --
+	(380.06, 89.36) --
+	(387.84, 93.13) --
+	(395.61, 96.28) --
+	(403.39,100.05);
 
-\path[] (403.39,100.39) --
-	(395.61, 96.28) --
-	(387.84, 92.69) --
-	(380.06, 88.71) --
-	(372.29, 85.21) --
-	(364.51, 81.29) --
-	(356.74, 77.87) --
-	(348.96, 74.21) --
-	(341.19, 71.44) --
-	(333.41, 67.10) --
-	(325.64, 63.68) --
-	(317.87, 61.35) --
-	(310.09, 58.41) --
-	(302.32, 56.50) --
-	(294.54, 55.57) --
-	(286.77, 54.88) --
-	(278.99, 56.47) --
-	(271.22, 58.80) --
-	(263.44, 58.90) --
-	(255.67, 58.75) --
-	(247.89, 58.63) --
-	(240.12, 59.84);
+\path[] (403.39, 99.70) --
+	(395.61, 96.10) --
+	(387.84, 92.95) --
+	(380.06, 89.10) --
+	(372.29, 85.75) --
+	(364.51, 82.37) --
+	(356.74, 79.13) --
+	(348.96, 75.93) --
+	(341.19, 73.00) --
+	(333.41, 69.00) --
+	(325.64, 66.39) --
+	(317.87, 63.54) --
+	(310.09, 61.16) --
+	(302.32, 59.96) --
+	(294.54, 58.76) --
+	(286.77, 58.17) --
+	(278.99, 57.49) --
+	(271.22, 57.71) --
+	(263.44, 57.10) --
+	(255.67, 57.13) --
+	(247.89, 57.13) --
+	(240.12, 57.43);
 \definecolor{drawColor}{RGB}{0,0,0}
 
-\path[draw=drawColor,line width= 2.3pt,line join=round] (240.12, 61.88) --
-	(247.89, 58.83) --
-	(255.67, 59.02) --
-	(263.44, 59.17) --
-	(271.22, 59.01) --
-	(278.99, 56.82) --
-	(286.77, 55.32) --
-	(294.54, 55.81) --
-	(302.32, 56.61) --
-	(310.09, 58.78) --
-	(317.87, 61.71) --
-	(325.64, 63.92) --
-	(333.41, 67.34) --
-	(341.19, 71.65) --
-	(348.96, 74.51) --
-	(356.74, 77.96) --
-	(364.51, 81.44) --
-	(372.29, 85.27) --
-	(380.06, 88.86) --
-	(387.84, 92.81) --
-	(395.61, 96.37) --
-	(403.39,100.53);
+\path[draw=drawColor,line width= 2.3pt,line join=round] (240.12, 57.78) --
+	(247.89, 57.37) --
+	(255.67, 57.22) --
+	(263.44, 57.27) --
+	(271.22, 57.94) --
+	(278.99, 57.68) --
+	(286.77, 58.62) --
+	(294.54, 59.17) --
+	(302.32, 60.06) --
+	(310.09, 61.74) --
+	(317.87, 63.77) --
+	(325.64, 66.87) --
+	(333.41, 69.04) --
+	(341.19, 73.56) --
+	(348.96, 76.03) --
+	(356.74, 79.22) --
+	(364.51, 82.56) --
+	(372.29, 85.84) --
+	(380.06, 89.21) --
+	(387.84, 93.10) --
+	(395.61, 96.17) --
+	(403.39, 99.93);
 \definecolor{drawColor}{RGB}{238,130,238}
 
-\path[draw=drawColor,line width= 1.1pt,line join=round] (317.87, 58.73) --
-	(325.64, 62.64) --
-	(333.41, 66.51) --
-	(341.19, 70.36) --
-	(348.96, 74.19) --
-	(356.74, 78.00) --
-	(364.51, 81.79) --
-	(372.29, 85.56) --
-	(380.06, 89.32) --
-	(387.84, 93.07) --
-	(395.61, 96.80) --
-	(403.39,100.53);
+\path[draw=drawColor,line width= 1.1pt,line join=round] (317.87, 61.28) --
+	(325.64, 64.89) --
+	(333.41, 68.47) --
+	(341.19, 72.03) --
+	(348.96, 75.57) --
+	(356.74, 79.09) --
+	(364.51, 82.60) --
+	(372.29, 86.09) --
+	(380.06, 89.56) --
+	(387.84, 93.03) --
+	(395.61, 96.48) --
+	(403.39, 99.93);
 
-\path[draw=drawColor,line width= 1.1pt,line join=round] (356.74, 58.52) --
-	(364.51, 65.52) --
-	(372.29, 72.52) --
-	(380.06, 79.52) --
-	(387.84, 86.52) --
-	(395.61, 93.53) --
-	(403.39,100.53);
+\path[draw=drawColor,line width= 1.1pt,line join=round] (356.74, 61.08) --
+	(364.51, 67.55) --
+	(372.29, 74.03) --
+	(380.06, 80.50) --
+	(387.84, 86.98) --
+	(395.61, 93.45) --
+	(403.39, 99.93);
 \end{scope}
@@
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (299.19, 55.67) --
-	(317.87, 58.73) --
-	(336.54, 55.67) --
-	(336.54, 44.86) --
-	(299.19, 44.86) --
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (299.19, 58.23) --
+	(317.87, 61.28) --
+	(336.54, 58.23) --
+	(336.54, 47.42) --
+	(299.19, 47.42) --
 	cycle;
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (348.51, 55.67) --
-	(356.74, 58.52) --
-	(364.97, 55.67) --
-	(364.97, 46.81) --
-	(348.51, 46.81) --
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (348.51, 58.23) --
+	(356.74, 61.08) --
+	(364.97, 58.23) --
+	(364.97, 49.37) --
+	(348.51, 49.37) --
 	cycle;
 
-\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at (317.87, 47.36) {$N \log N$};
+\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at (317.87, 49.92) {$N \log N$};
 
-\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at (356.74, 47.36) {$N^2$};
+\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at (356.74, 49.92) {$N^2$};
 \end{scope}
@@
 
-\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 41.91, 53.69) {1e-03};
+\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 41.91, 56.51) {1e-03};
 
-\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 41.91, 76.94) {1e-01};
+\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 41.91, 78.02) {1e-01};
 
-\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 41.91,100.20) {1e+01};
+\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 41.91, 99.53) {1e+01};
 \end{scope}
@@
 
-\path[draw=drawColor,line width= 0.6pt,line join=round] ( 44.11, 56.72) --
-	( 46.86, 56.72);
+\path[draw=drawColor,line width= 0.6pt,line join=round] ( 44.11, 59.54) --
+	( 46.86, 59.54);
 
-\path[draw=drawColor,line width= 0.6pt,line join=round] ( 44.11, 79.97) --
-	( 46.86, 79.97);
+\path[draw=drawColor,line width= 0.6pt,line join=round] ( 44.11, 81.05) --
+	( 46.86, 81.05);
 
-\path[draw=drawColor,line width= 0.6pt,line join=round] ( 44.11,103.23) --
-	( 46.86,103.23);
+\path[draw=drawColor,line width= 0.6pt,line join=round] ( 44.11,102.56) --
+	( 46.86,102.56);
 \end{scope}

### 2025-05-07 figure-heap-refs-other.tex
N^1.4

@@
-% Created by tikzDevice version 0.12.6 on 2025-05-06 22:53:33
+% Created by tikzDevice version 0.12.6 on 2025-05-07 07:42:34
 % !TEX encoding = UTF-8 Unicode
@@
 
-\path[draw=drawColor,line width= 2.3pt,line join=round] ( 55.03,172.40) --
+\path[draw=drawColor,line width= 2.3pt,line join=round] ( 55.03,120.44) --
 	( 69.87,134.54) --
@@
 
-\path[draw=drawColor,line width= 0.3pt,line join=round] ( 46.86, 44.40) --
-	(226.46, 44.40);
+\path[draw=drawColor,line width= 0.3pt,line join=round] ( 46.86, 48.79) --
+	(226.46, 48.79);
 
-\path[draw=drawColor,line width= 0.3pt,line join=round] ( 46.86, 71.59) --
-	(226.46, 71.59);
+\path[draw=drawColor,line width= 0.3pt,line join=round] ( 46.86, 74.31) --
+	(226.46, 74.31);
 
-\path[draw=drawColor,line width= 0.3pt,line join=round] ( 46.86, 98.78) --
-	(226.46, 98.78);
+\path[draw=drawColor,line width= 0.3pt,line join=round] ( 46.86, 99.82) --
+	(226.46, 99.82);
 
@@
 
-\path[draw=drawColor,line width= 0.6pt,line join=round] ( 46.86, 58.00) --
-	(226.46, 58.00);
+\path[draw=drawColor,line width= 0.6pt,line join=round] ( 46.86, 61.55) --
+	(226.46, 61.55);
 
-\path[draw=drawColor,line width= 0.6pt,line join=round] ( 46.86, 85.19) --
-	(226.46, 85.19);
+\path[draw=drawColor,line width= 0.6pt,line join=round] ( 46.86, 87.07) --
+	(226.46, 87.07);
 
-\path[draw=drawColor,line width= 0.6pt,line join=round] ( 46.86,112.38) --
-	(226.46,112.38);
+\path[draw=drawColor,line width= 0.6pt,line join=round] ( 46.86,112.58) --
+	(226.46,112.58);
 
@@
 
-\path[fill=fillColor,fill opacity=0.50] ( 55.03, 58.00) --
-	( 69.87, 61.60) --
-	( 84.71, 59.31) --
-	( 99.55, 59.71) --
-	(114.40, 58.63) --
-	(129.24, 59.16) --
-	(144.08, 61.71) --
-	(158.92, 72.45) --
-	(173.77, 81.74) --
-	(188.61, 90.10) --
-	(203.45, 98.38) --
-	(218.29,109.54) --
-	(218.29,109.46) --
-	(203.45, 98.22) --
-	(188.61, 86.26) --
-	(173.77, 76.01) --
-	(158.92, 66.75) --
-	(144.08, 59.17) --
-	(129.24, 54.26) --
-	(114.40, 53.86) --
-	( 99.55, 55.64) --
-	( 84.71, 54.50) --
-	( 69.87, 56.76) --
-	( 55.03, 52.79) --
+\path[fill=fillColor,fill opacity=0.50] ( 55.03, 56.99) --
+	( 69.87, 57.03) --
+	( 84.71, 57.32) --
+	( 99.55, 60.60) --
+	(114.40, 59.12) --
+	(129.24, 60.73) --
+	(144.08, 64.48) --
+	(158.92, 69.92) --
+	(173.77, 78.10) --
+	(188.61, 89.04) --
+	(203.45, 99.45) --
+	(218.29,110.08) --
+	(218.29,109.75) --
+	(203.45, 98.66) --
+	(188.61, 87.82) --
+	(173.77, 77.86) --
+	(158.92, 69.30) --
+	(144.08, 62.34) --
+	(129.24, 57.82) --
+	(114.40, 55.88) --
+	( 99.55, 58.18) --
+	( 84.71, 53.31) --
+	( 69.87, 52.48) --
+	( 55.03, 52.46) --
 	cycle;
 
-\path[] ( 55.03, 58.00) --
-	( 69.87, 61.60) --
-	( 84.71, 59.31) --
-	( 99.55, 59.71) --
-	(114.40, 58.63) --
-	(129.24, 59.16) --
-	(144.08, 61.71) --
-	(158.92, 72.45) --
-	(173.77, 81.74) --
-	(188.61, 90.10) --
-	(203.45, 98.38) --
-	(218.29,109.54);
+\path[] ( 55.03, 56.99) --
+	( 69.87, 57.03) --
+	( 84.71, 57.32) --
+	( 99.55, 60.60) --
+	(114.40, 59.12) --
+	(129.24, 60.73) --
+	(144.08, 64.48) --
+	(158.92, 69.92) --
+	(173.77, 78.10) --
+	(188.61, 89.04) --
+	(203.45, 99.45) --
+	(218.29,110.08);
 
-\path[] (218.29,109.46) --
-	(203.45, 98.22) --
-	(188.61, 86.26) --
-	(173.77, 76.01) --
-	(158.92, 66.75) --
-	(144.08, 59.17) --
-	(129.24, 54.26) --
-	(114.40, 53.86) --
-	( 99.55, 55.64) --
-	( 84.71, 54.50) --
-	( 69.87, 56.76) --
-	( 55.03, 52.79);
+\path[] (218.29,109.75) --
+	(203.45, 98.66) --
+	(188.61, 87.82) --
+	(173.77, 77.86) --
+	(158.92, 69.30) --
+	(144.08, 62.34) --
+	(129.24, 57.82) --
+	(114.40, 55.88) --
+	( 99.55, 58.18) --
+	( 84.71, 53.31) --
+	( 69.87, 52.48) --
+	( 55.03, 52.46);
 \definecolor{drawColor}{RGB}{0,0,0}
 
-\path[draw=drawColor,line width= 2.3pt,line join=round] ( 55.03, 53.52) --
-	( 69.87, 58.87) --
-	( 84.71, 55.00) --
-	( 99.55, 55.96) --
-	(114.40, 54.84) --
-	(129.24, 55.43) --
-	(144.08, 59.55) --
-	(158.92, 66.97) --
-	(173.77, 76.82) --
-	(188.61, 86.54) --
-	(203.45, 98.25) --
-	(218.29,109.50);
+\path[draw=drawColor,line width= 2.3pt,line join=round] ( 55.03, 53.27) --
+	( 69.87, 52.94) --
+	( 84.71, 53.71) --
+	( 99.55, 59.11) --
+	(114.40, 56.54) --
+	(129.24, 58.38) --
+	(144.08, 62.55) --
+	(158.92, 69.56) --
+	(173.77, 77.91) --
+	(188.61, 87.87) --
+	(203.45, 98.84) --
+	(218.29,109.81);
 \definecolor{drawColor}{RGB}{238,130,238}
 
-\path[draw=drawColor,line width= 1.1pt,line join=round] (129.24, 60.39) --
-	(144.08, 68.57) --
-	(158.92, 76.76) --
-	(173.77, 84.94) --
-	(188.61, 93.13) --
-	(203.45,101.31) --
-	(218.29,109.50);
+\path[draw=drawColor,line width= 1.1pt,line join=round] (129.24, 63.73) --
+	(144.08, 71.41) --
+	(158.92, 79.09) --
+	(173.77, 86.77) --
+	(188.61, 94.45) --
+	(203.45,102.13) --
+	(218.29,109.81);
 
-\path[draw=drawColor,line width= 1.1pt,line join=round] (158.92, 60.39) --
-	(173.77, 72.67) --
-	(188.61, 84.94) --
-	(203.45, 97.22) --
-	(218.29,109.50);
+\path[draw=drawColor,line width= 1.1pt,line join=round] (158.92, 63.73) --
+	(173.77, 75.25) --
+	(188.61, 86.77) --
+	(203.45, 98.29) --
+	(218.29,109.81);
 \end{scope}
@@
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (121.01, 57.54) --
-	(129.24, 60.39) --
-	(137.47, 57.54) --
-	(137.47, 48.68) --
-	(121.01, 48.68) --
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (121.01, 60.88) --
+	(129.24, 63.73) --
+	(137.47, 60.88) --
+	(137.47, 52.02) --
+	(121.01, 52.02) --
 	cycle;
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (150.70, 57.54) --
-	(158.92, 60.39) --
-	(167.15, 57.54) --
-	(167.15, 48.68) --
-	(150.70, 48.68) --
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (150.70, 60.88) --
+	(158.92, 63.73) --
+	(167.15, 60.88) --
+	(167.15, 52.02) --
+	(150.70, 52.02) --
 	cycle;
 
-\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at (129.24, 49.23) {$N^2$};
+\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at (129.24, 52.57) {$N^2$};
 
-\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at (158.92, 49.23) {$N^3$};
+\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at (158.92, 52.57) {$N^3$};
 \end{scope}
@@
 
-\path[draw=drawColor,line width= 2.3pt,line join=round] (240.12,139.98) --
+\path[draw=drawColor,line width= 2.3pt,line join=round] (240.12,120.44) --
 	(253.73,120.44) --
@@
 
-\path[draw=drawColor,line width= 0.3pt,line join=round] (231.96, 44.40) --
-	(411.55, 44.40);
+\path[draw=drawColor,line width= 0.3pt,line join=round] (231.96, 48.79) --
+	(411.55, 48.79);
 
-\path[draw=drawColor,line width= 0.3pt,line join=round] (231.96, 71.59) --
-	(411.55, 71.59);
+\path[draw=drawColor,line width= 0.3pt,line join=round] (231.96, 74.31) --
+	(411.55, 74.31);
 
-\path[draw=drawColor,line width= 0.3pt,line join=round] (231.96, 98.78) --
-	(411.55, 98.78);
+\path[draw=drawColor,line width= 0.3pt,line join=round] (231.96, 99.82) --
+	(411.55, 99.82);
 
@@
 
-\path[draw=drawColor,line width= 0.6pt,line join=round] (231.96, 58.00) --
-	(411.55, 58.00);
+\path[draw=drawColor,line width= 0.6pt,line join=round] (231.96, 61.55) --
+	(411.55, 61.55);
 
-\path[draw=drawColor,line width= 0.6pt,line join=round] (231.96, 85.19) --
-	(411.55, 85.19);
+\path[draw=drawColor,line width= 0.6pt,line join=round] (231.96, 87.07) --
+	(411.55, 87.07);
 
-\path[draw=drawColor,line width= 0.6pt,line join=round] (231.96,112.38) --
-	(411.55,112.38);
+\path[draw=drawColor,line width= 0.6pt,line join=round] (231.96,112.58) --
+	(411.55,112.58);
 
@@
 
-\path[fill=fillColor,fill opacity=0.50] (240.12, 60.30) --
-	(253.73, 64.90) --
-	(267.33, 65.15) --
-	(280.94, 69.28) --
-	(294.54, 73.86) --
-	(308.15, 75.13) --
-	(321.75, 78.18) --
-	(335.36, 83.17) --
-	(348.96, 88.17) --
-	(362.57, 93.25) --
-	(376.17,101.77) --
-	(389.78,104.80) --
+\path[fill=fillColor,fill opacity=0.50] (240.12, 59.86) --
+	(253.73, 60.11) --
+	(267.33, 63.24) --
+	(280.94, 67.07) --
+	(294.54, 71.40) --
+	(308.15, 76.33) --
+	(321.75, 81.19) --
+	(335.36, 85.66) --
+	(348.96, 90.13) --
+	(362.57, 97.75) --
+	(376.17,100.03) --
+	(389.78,105.46) --
 	(403.39,111.56) --
-	(403.39,111.28) --
-	(389.78,104.61) --
-	(376.17, 98.60) --
-	(362.57, 93.17) --
-	(348.96, 88.03) --
-	(335.36, 83.04) --
-	(321.75, 78.08) --
-	(308.15, 73.15) --
-	(294.54, 71.03) --
-	(280.94, 68.63) --
-	(267.33, 63.57) --
-	(253.73, 60.00) --
-	(240.12, 54.65) --
+	(403.39,111.44) --
+	(389.78,105.19) --
+	(376.17, 99.67) --
+	(362.57, 94.81) --
+	(348.96, 89.66) --
+	(335.36, 84.90) --
+	(321.75, 80.54) --
+	(308.15, 75.94) --
+	(294.54, 71.18) --
+	(280.94, 66.41) --
+	(267.33, 61.49) --
+	(253.73, 57.03) --
+	(240.12, 53.14) --
 	cycle;
 
-\path[] (240.12, 60.30) --
-	(253.73, 64.90) --
-	(267.33, 65.15) --
-	(280.94, 69.28) --
-	(294.54, 73.86) --
-	(308.15, 75.13) --
-	(321.75, 78.18) --
-	(335.36, 83.17) --
-	(348.96, 88.17) --
-	(362.57, 93.25) --
-	(376.17,101.77) --
-	(389.78,104.80) --
+\path[] (240.12, 59.86) --
+	(253.73, 60.11) --
+	(267.33, 63.24) --
+	(280.94, 67.07) --
+	(294.54, 71.40) --
+	(308.15, 76.33) --
+	(321.75, 81.19) --
+	(335.36, 85.66) --
+	(348.96, 90.13) --
+	(362.57, 97.75) --
+	(376.17,100.03) --
+	(389.78,105.46) --
 	(403.39,111.56);
 
-\path[] (403.39,111.28) --
-	(389.78,104.61) --
-	(376.17, 98.60) --
-	(362.57, 93.17) --
-	(348.96, 88.03) --
-	(335.36, 83.04) --
-	(321.75, 78.08) --
-	(308.15, 73.15) --
-	(294.54, 71.03) --
-	(280.94, 68.63) --
-	(267.33, 63.57) --
-	(253.73, 60.00) --
-	(240.12, 54.65);
+\path[] (403.39,111.44) --
+	(389.78,105.19) --
+	(376.17, 99.67) --
+	(362.57, 94.81) --
+	(348.96, 89.66) --
+	(335.36, 84.90) --
+	(321.75, 80.54) --
+	(308.15, 75.94) --
+	(294.54, 71.18) --
+	(280.94, 66.41) --
+	(267.33, 61.49) --
+	(253.73, 57.03) --
+	(240.12, 53.14);
 \definecolor{drawColor}{RGB}{0,0,0}
 
-\path[draw=drawColor,line width= 2.3pt,line join=round] (240.12, 54.98) --
-	(253.73, 60.57) --
-	(267.33, 63.67) --
-	(280.94, 68.67) --
-	(294.54, 72.61) --
-	(308.15, 73.27) --
-	(321.75, 78.12) --
-	(335.36, 83.10) --
-	(348.96, 88.07) --
-	(362.57, 93.19) --
-	(376.17, 98.98) --
-	(389.78,104.62) --
-	(403.39,111.42);
+\path[draw=drawColor,line width= 2.3pt,line join=round] (240.12, 53.54) --
+	(253.73, 57.15) --
+	(267.33, 61.65) --
+	(280.94, 66.59) --
+	(294.54, 71.33) --
+	(308.15, 75.99) --
+	(321.75, 80.60) --
+	(335.36, 85.23) --
+	(348.96, 89.69) --
+	(362.57, 94.88) --
+	(376.17, 99.75) --
+	(389.78,105.24) --
+	(403.39,111.49);
 \definecolor{drawColor}{RGB}{238,130,238}
 
-\path[draw=drawColor,line width= 1.1pt,line join=round] (267.33, 63.09) --
-	(280.94, 68.50) --
-	(294.54, 73.67) --
-	(308.15, 78.68) --
-	(321.75, 83.56) --
-	(335.36, 88.35) --
-	(348.96, 93.06) --
-	(362.57, 97.72) --
-	(376.17,102.32) --
-	(389.78,106.89) --
-	(403.39,111.42);
+\path[draw=drawColor,line width= 1.1pt,line join=round] (253.73, 60.71) --
+	(267.33, 66.14) --
+	(280.94, 71.22) --
+	(294.54, 76.07) --
+	(308.15, 80.76) --
+	(321.75, 85.34) --
+	(335.36, 89.84) --
+	(348.96, 94.26) --
+	(362.57, 98.63) --
+	(376.17,102.95) --
+	(389.78,107.24) --
+	(403.39,111.49);
 
-\path[draw=drawColor,line width= 1.1pt,line join=round] (294.54, 62.31) --
-	(308.15, 68.44) --
-	(321.75, 74.58) --
-	(335.36, 80.72) --
-	(348.96, 86.86) --
-	(362.57, 93.00) --
-	(376.17, 99.14) --
-	(389.78,105.28) --
-	(403.39,111.42);
+\path[draw=drawColor,line width= 1.1pt,line join=round] (280.94, 63.10) --
+	(294.54, 68.47) --
+	(308.15, 73.85) --
+	(321.75, 79.23) --
+	(335.36, 84.60) --
+	(348.96, 89.98) --
+	(362.57, 95.36) --
+	(376.17,100.73) --
+	(389.78,106.11) --
+	(403.39,111.49);
 
-\path[draw=drawColor,line width= 1.1pt,line join=round] (321.75, 62.31) --
-	(335.36, 70.49) --
-	(348.96, 78.68) --
-	(362.57, 86.86) --
-	(376.17, 95.05) --
-	(389.78,103.23) --
-	(403.39,111.42);
+\path[draw=drawColor,line width= 1.1pt,line join=round] (321.75, 65.40) --
+	(335.36, 73.08) --
+	(348.96, 80.76) --
+	(362.57, 88.44) --
+	(376.17, 96.12) --
+	(389.78,103.81) --
+	(403.39,111.49);
 \end{scope}
@@
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (247.22, 59.46) --
-	(267.33, 63.09) --
-	(284.57, 59.46) --
-	(284.57, 48.65) --
-	(247.22, 48.65) --
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (233.62, 57.86) --
+	(253.73, 60.71) --
+	(270.96, 57.86) --
+	(270.96, 47.05) --
+	(233.62, 47.05) --
 	cycle;
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (284.57, 59.46) --
-	(294.54, 62.31) --
-	(307.39, 59.46) --
-	(307.39, 48.65) --
-	(284.57, 48.65) --
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (270.96, 57.86) --
+	(280.94, 63.10) --
+	(293.78, 57.86) --
+	(293.78, 47.05) --
+	(270.96, 47.05) --
 	cycle;
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (313.52, 59.46) --
-	(321.75, 62.31) --
-	(329.98, 59.46) --
-	(329.98, 50.60) --
-	(313.52, 50.60) --
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (313.52, 57.86) --
+	(321.75, 65.40) --
+	(329.98, 57.86) --
+	(329.98, 49.00) --
+	(313.52, 49.00) --
 	cycle;
 
-\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at (265.90, 51.15) {$N \log N$};
+\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at (252.29, 49.55) {$N \log N$};
 
-\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at (295.98, 51.15) {$N^{1.5}$};
+\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at (282.37, 49.55) {$N^{1.4}$};
 
-\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at (321.75, 51.15) {$N^2$};
+\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at (321.75, 49.55) {$N^2$};
 \end{scope}
@@
 
-\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 41.91, 54.97) {1e-03};
+\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 41.91, 58.52) {1e-03};
 
-\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 41.91, 82.16) {1e-01};
+\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 41.91, 84.04) {1e-01};
 
-\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 41.91,109.35) {1e+01};
+\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 41.91,109.55) {1e+01};
 \end{scope}
@@
 
-\path[draw=drawColor,line width= 0.6pt,line join=round] ( 44.11, 58.00) --
-	( 46.86, 58.00);
+\path[draw=drawColor,line width= 0.6pt,line join=round] ( 44.11, 61.55) --
+	( 46.86, 61.55);
 
-\path[draw=drawColor,line width= 0.6pt,line join=round] ( 44.11, 85.19) --
-	( 46.86, 85.19);
+\path[draw=drawColor,line width= 0.6pt,line join=round] ( 44.11, 87.07) --
+	( 46.86, 87.07);
 
-\path[draw=drawColor,line width= 0.6pt,line join=round] ( 44.11,112.38) --
-	( 46.86,112.38);
+\path[draw=drawColor,line width= 0.6pt,line join=round] ( 44.11,112.58) --
+	( 46.86,112.58);
 \end{scope}

### 2025-05-07 figure-heap.R

@@ -4,7 +4,8 @@ atime_list$measurements[, expr.name := sub("container=", "", expr.name)]
 plot(atime_list)
 
 fun_list <- atime:::references_funs[c("N","N \\log N", "N^2", "N^3")]
-fun_list[["N^1.5"]] <- function(N)1.5 * log10(N)
+myexp <- 1.4
+fun_list[[paste0("N^",myexp)]] <- function(N)myexp * log10(N)
 ref_list <- atime::references_best(atime_list, fun_list)
 ref_list$plot.references <- ref_list$references[(
   fun.name %in% c("N","N^2") & unit=="kilobytes"
@@ -37,10 +38,41 @@ png("figure-heap-refs-other.png",width=6,height=3.5,units="in",res=200)
 plot_algos("change|rupt")
 dev.off()
 
+pred_list <- predict(ref_list)
+png("figure-heap-pred.png",width=8,height=3,units="in",res=200)
+pkg.colors <- c(
+  changepoint="#E41A1C",
+  ruptures="#377EB8",
+  "binsegRcpp\nmultiset"="black",
+  "binsegRcpp\npriority_queue"="grey30",
+  "binsegRcpp\nlist"="grey50")
+plot(pred_list)+
+  facet_null()+
+  scale_y_log10(
+    "Computation time (seconds, log scale)",
+    breaks=10^seq(-3,1),
+    limits=c(NA,200))+
+  scale_x_log10(
+    "N = number of data to segment (max segments = N/2)",
+    breaks=10^seq(1,7),
+    limits=c(NA, 1e8))+
+  scale_color_manual(values=pkg.colors)+
+  scale_fill_manual(values=pkg.colors)
+dev.off()
+
+norm_dt <- pred_list$prediction[, ratio := max(N)/N][, .(expr.name, ratio)][pred_list$meas, on="expr.name"][unit=="seconds"][, norm_N := N*ratio][]
+library(ggplot2)
+ggplot()+
+  geom_line(aes(
+    norm_N, empirical, color=expr.name),
+    data=norm_dt)+
+  scale_x_log10()+
+  scale_y_log10()
+
 tikz_algos <- function(f.tex,regex){
   trans <- function(DT)DT[,expr.name:=sub("_","\\_",expr.name,fixed=TRUE)]
   meas <- trans(ref_list$measurements[grepl(regex,expr.name)])
[…]

### 2025-05-08 article.Rnw
intro

@@
   In practice,
-  other implementations can be asymptotically slower, and can 
-  return incorrect results. We propose \pkg{binsegRcpp}, an \proglang{R} package which
-  provides a correct \proglang{C++} implementation, with the expected asymptotic time complexity.
+  existing implementations can be asymptotically slower, and can 
+  return incorrect results.
+  We propose \pkg{binsegRcpp}, an \proglang{R} package which
+  provides a correct \proglang{C++} implementation,
+  with the expected asymptotic time complexity.
   %% In practice, existing
@@
   and include detailed comparisons with other implementations of binary
-  segmentation: \pkg{ruptures}, \pkg{fpop}, \pkg{changepoint}, \pkg{blockcpd}, and
+  segmentation:
+  \pkg{ruptures}, \pkg{fpop}, \pkg{changepoint}, \pkg{blockcpd}, and
   \pkg{wbs}.
@@
 
-TODO \code{fpop::multiBinSeg} \citep{Maidstone2017}.
+Change-point detection is an important problem in sequential data
+analysis, that occurs in a variety of different applications, such as
+genomics \citep{Hocking2013bioinformatics}, neuroscience
+\citep{Jewell2019biostatistics}, and medical monitoring
+\citep{Fotoohinasab2020segmentationEMBC}, among others.
+The focus of
+this article is change-point detection in the off-line setting, in
+which we assume there is a sequence of $N$ observed data, and we want
+an algorithm to identify the precise locations of any abrupt changes
+that may be present \citep{Truong2020}. We assume there is
+user-specified maximum number of change-points $K$, and we would like
+to compute a sequence of models from 1 to $K$ change-points.
 
-TODO \code{wbs::sbs} \citep{Baranowski2019}.
+Using
+the classical dynamic programming algorithm of
+\citet{segment-neighborhood}, optimal change-points for a given cost
+function, and a sequence of model sizes $k\in\{1,\dots,K\}$, can be
+computed in $O(N^2 K)$ time.
+%\citep{Jackson2005}
+More recently, the pruned dynamic programming algorithms of
+\citet{Maidstone2017} can be used to compute the same sequence of $K$
+optimal models, with the same $O(N^2 K)$ worst case time complexity,
+but faster $O(KN)$ time best case time complexity.  Other dynamic
+programming algorithms such as PELT \citep{pelt} and DuST
+\citep{truong2024efficient} compute a single model in best case linear
+time $O(N)$, worst case quadratic time $O(N^2)$.  They may be used as
+a sub-routine to efficiently compute a range of $K$ change-point
+models, using the CROCS algorithm \citep{Liehrmann2021chipseq}, which
+again results in best case $O(N K)$ time, worst case $O(N^2 K)$ time.
+
+Binary segmentation is a heuristic algorithm for computing a sequence of models from 1 to $K$ change-points \citep{binary-segmentation}.
+Using a simple loss function like the square loss, the best case of binary segmentation is $O(N \log K)$ time \citep{Hocking2024finite}, which is much faster than dynamic programming.
+The worst case of binary segmentation is $O(N K)$ time, which is the same speed as the best case of dynamic programming.
+
+\begin{table}[t!]
+  \centering
+  \small
+\begin{tabular}{l|cccccc}
+%   > ruptures$version$version
+% [1] "1.1.6"
+% > sapply(c("binsegRcpp", "changepoint", "wbs", "fpop"), function(p)paste(packageVersion(p)))
+%  binsegRcpp changepoint         wbs        fpop 
+% "2022.3.29"     "2.2.3"       "1.4" "2019.8.26"
+% https://github.com/HaotianXu/changepoints does not have binary segmentation.
+  package & binsegRcpp & changepoint & wbs & fpop & ruptures & blockcpd\\
+  function & binseg & cpt.mean & sbs & multiBinSeg & Binseg & fit\_blockcpd\\
+  version & 2022.3.29 & 2.2.3 & 1.4 & 2019.8.26 & 1.1.9 & 1.0.0\\
+\hline
+weights & yes & no & no & no & no & no\\
+min len & yes & yes & no & no & yes & yes\\
+max segs & yes & yes & no & yes & yes & yes\\
+dim & one & one & one & multi & multi & multi \\
+correct & yes & no & yes & yes & yes & yes\\
+losses & 5 & 6 & 1 & 1 & 10 & 5\\
+language & C++ & C & C & C++ & Python & C++\\
+% all propose some loss computation.
+storage & heap & arrays & recursion & heap & LRU cache & heap \\
+space & $O(S)$ & $O(S^2)$ & $O(S)$ & $O(S)$ & $O(S)$ & $O(S)$ \\
+cumsum & yes & yes & yes & yes & no & yes\\
+best & $O(N\log N)$ & $O(N^3)$ & $O(N\log N)$ & $O(N\log N)$ & $O(N^{1.4})$ & $O(N\log N)$\\
+worst & $O(N^2)$ & $O(N^3)$ & $O(N^2)$ & $O(N^2)$ & $O(N^2)$ & $O(N^2)$\\
+CV & yes & no & no & no & no & no\\
+params & all & one & no & no & no &one\\
+\end{tabular}
+\caption{\label{tab:packages}
+  Properties (rows) of six binary segmentation libraries (columns).
+Properties are weights (can specify observation-specific positive weights for the loss), min len (can specify minimum segment length parameter), max segs (can specify maximum number of segments parameter), dim (dimensions of input data), correct (does code return correct model parameters?), losses (number of loss functions implemented), language (programming language to implement loss), storage (technique used to store cost of splitting a segment), space (space complexity in number of segments $S$), cumsum (cumulative sum used), best (time complexity in best case data), worst (time complexity in worst case data), CV (support for cross-validation), params (parameters returned, one or all model sizes).
+}
+\end{table}
+
+There are several free/open-source implementations of binary
+segmentation, most of which use C/C++ code for efficiency
+(Table~\ref{tab:packages}).  The \pkg{fpop} package in R contains the
+\code{fpop::multiBinSeg} function, which implements the square loss
+for multi-dimensional data \citep{Maidstone2017}.  The \pkg{wbs}
+package in R contains the \code{wbs::sbs} function, which implements
+the square loss for univariate data \citep{Baranowski2019}.  Since it
+uses recursive function calls, it does not support the max segments
+parameter $K$, and instead always computes the full path of models
+(from 1 to $N-1$ change-points).
 
@@
 variables, function pointers, templates, macros).
-
-\begin{table}[t!]
-  \centering
-  \small
-\begin{tabular}{l|cccccc}
-  package & binsegRcpp & changepoint & wbs & fpop & ruptures & blockcpd\\
-  function & binseg & cpt.mean & sbs & multiBinSeg & Binseg & fit\_blockcpd\\
-%   > ruptures$version$version
-% [1] "1.1.6"
-% > sapply(c("binsegRcpp", "changepoint", "wbs", "fpop"), function(p)paste(packageVersion(p)))
-%  binsegRcpp changepoint         wbs        fpop 
-% "2022.3.29"     "2.2.3"       "1.4" "2019.8.26"
-% https://github.com/HaotianXu/changepoints does not have binary segmentation.
-  version & 2022.3.29 & 2.2.3 & 1.4 & 2019.8.26 & 1.1.9 & 1.0.0\\
-\hline
-  weights & yes & no & no & no & no & no\\
-  max segs & yes & yes & no & yes & yes & yes\\
-  dim & one & one & one & multi & multi & multi \\
-  correct & yes & no & yes & yes & yes & yes\\
-  losses & 5 & 6 & 1 & 1 & 10 & 5\\
-  language & C++ & C & C & C++ & Python & C++\\
-  % all propose some loss computation.
-  storage & heap & arrays & recursion & heap & LRU cache & heap \\
-  space & $O(S)$ & $O(S^2)$ & $O(S)$ & $O(S)$ & $O(S)$ & $O(S)$ \\
-  cumsum & yes & yes & yes & yes & no & yes\\
-  best & $O(N\log N)$ & $O(N^3)$ & $O(N\log N)$ & $O(N\log N)$ & $O(N^{1.4})$ & $O(N\log N)$\\
-  worst & $O(N^2)$ & $O(N^3)$ & $O(N^2)$ & $O(N^2)$ & $O(N^2)$ & $O(N^2)$\\
-  CV & yes & no & no & no & no & no\\
-  params & all & one & no & no & no &one\\
-\end{tabular}
-\caption{\label{tab:overview}
-  Properties (rows) of six binary segmentation libraries (columns).
-Properties are weights (can specify observation-specific positive weights for the loss), max segs (can specify maximum number of segments parameter), dim (dimensions of input data), correct (does code return correct model parameters?), losses (number of loss functions implemented), language (programming language to implement loss), storage (technique used to store cost of splitting a segment), space (space complexity in number of segments $S$), cumsum (cumulative sum used), best (time complexity in best case data), worst (time complexity in worst case data), CV (support for cross-validation), params (parameters are returned).
-}
-\end{table}
 

### 2025-05-08 article.Rnw
wbs

@@
 
+\subsection{Existing implementations of binary segmentation}
 There are several free/open-source implementations of binary
 segmentation, most of which use C/C++ code for efficiency
-(Table~\ref{tab:packages}).  The \pkg{fpop} package in R contains the
-\code{fpop::multiBinSeg} function, which implements the square loss
-for multi-dimensional data \citep{Maidstone2017}.  The \pkg{wbs}
-package in R contains the \code{wbs::sbs} function, which implements
-the square loss for univariate data \citep{Baranowski2019}.  Since it
-uses recursive function calls, it does not support the max segments
-parameter $K$, and instead always computes the full path of models
-(from 1 to $N-1$ change-points).
+(Table~\ref{tab:packages}).
+We compared our proposed implementation to these baselines in terms of functionality, correctness, and efficiency.
 
-TODO \code{changepoint::cpt.mean(method="BinSeg")} \citep{Killick2014,Killick2022}.
+The \pkg{fpop} package in R contains the \code{fpop::multiBinSeg}
+function, which implements the square loss for multi-dimensional data
+\citep{Maidstone2017}.
+
+The \pkg{wbs} package in R contains the \code{wbs::sbs} function,
+which implements the square loss for univariate data
+\citep{Baranowski2019}.  Since it uses recursive function calls, it
+does not support the max segments parameter $K$, and instead always
+computes the full path of models (from 1 to $N-1$ change-points).
+
+The \code{changepoint::cpt.mean(method="BinSeg")} \citep{Killick2014,Killick2022}.
 

### 2025-05-08 article.Rnw
done paragraphs based on table 1

@@
 %% - \Abstract{} almost as usual
-\Abstract{
-  Binary segmentation is a classic algorithm for detecting
+\Abstract{ Binary segmentation is a classic algorithm for detecting
   change-points in sequential data. In theory, using a simple loss
   function like Normal or Poisson, binary segmentation should be
-  extremely fast for $N$ data and $K$ segments: asymptotically $O(N K)$ time in the worst
-  case, and $O(N \log K)$ time in the best case.
-  In practice,
-  existing implementations can be asymptotically slower, and can 
-  return incorrect results.
-  We propose \pkg{binsegRcpp}, an \proglang{R} package which
-  provides a correct \proglang{C++} implementation,
-  with the expected asymptotic time complexity.
+  extremely fast for $N$ data and $S$ segments: asymptotically $O(N
+  S)$ time in the worst case, and $O(N \log S)$ time in the best case.
+  In practice, existing implementations can be asymptotically slower,
+  and can return incorrect results.  We propose \pkg{binsegRcpp}, an
+  \proglang{R} package which provides a correct \proglang{C++}
+  implementation, with the expected asymptotic time complexity.
   %% In practice, existing
@@
 that may be present \citep{Truong2020}. We assume there is
-user-specified maximum number of change-points $K$, and we would like
-to compute a sequence of models from 1 to $K$ change-points.
+user-specified maximum number of segments $S$, and we would like
+to compute a sequence of models from 1 to $S$ segments (0 to $S-1$ change-points).
 
@@
 \citet{segment-neighborhood}, optimal change-points for a given cost
-function, and a sequence of model sizes $k\in\{1,\dots,K\}$, can be
-computed in $O(N^2 K)$ time.
+function, and a sequence of model sizes $k\in\{1,\dots,S\}$, can be
+computed in $O(N^2 S)$ time.
 %\citep{Jackson2005}
 More recently, the pruned dynamic programming algorithms of
-\citet{Maidstone2017} can be used to compute the same sequence of $K$
-optimal models, with the same $O(N^2 K)$ worst case time complexity,
-but faster $O(KN)$ time best case time complexity.  Other dynamic
+\citet{Maidstone2017} can be used to compute the same sequence of $S$
+optimal models, with the same $O(N^2 S)$ worst case time complexity,
+but faster $O(NS)$ time best case time complexity.  Other dynamic
 programming algorithms such as PELT \citep{pelt} and DuST
@@
 time $O(N)$, worst case quadratic time $O(N^2)$.  They may be used as
-a sub-routine to efficiently compute a range of $K$ change-point
+a sub-routine to efficiently compute a range of $S$ change-point
 models, using the CROCS algorithm \citep{Liehrmann2021chipseq}, which
-again results in best case $O(N K)$ time, worst case $O(N^2 K)$ time.
+again results in best case $O(N S)$ time, worst case $O(N^2 S)$ time.
+\citet{fastcpd2024} propose a change-point algorithm based on dynamic programming, and provide comparisons with several previous algorithms.
 
-Binary segmentation is a heuristic algorithm for computing a sequence of models from 1 to $K$ change-points \citep{binary-segmentation}.
-Using a simple loss function like the square loss, the best case of binary segmentation is $O(N \log K)$ time \citep{Hocking2024finite}, which is much faster than dynamic programming.
-The worst case of binary segmentation is $O(N K)$ time, which is the same speed as the best case of dynamic programming.
+Binary segmentation is a heuristic algorithm for computing a sequence of models from 1 to $S$ change-points \citep{binary-segmentation}.
+Using a simple loss function like the square loss, the best case of binary segmentation is $O(N \log S)$ time \citep{Hocking2024finite}, which is much faster than dynamic programming.
+The worst case of binary segmentation is $O(N S)$ time, which is the same speed as the best case of dynamic programming.
 
@@
 \end{tabular}
-\caption{\label{tab:packages}
-  Properties (rows) of six binary segmentation libraries (columns).
-Properties are weights (can specify observation-specific positive weights for the loss), min len (can specify minimum segment length parameter), max segs (can specify maximum number of segments parameter), dim (dimensions of input data), correct (does code return correct model parameters?), losses (number of loss functions implemented), language (programming language to implement loss), storage (technique used to store cost of splitting a segment), space (space complexity in number of segments $S$), cumsum (cumulative sum used), best (time complexity in best case data), worst (time complexity in worst case data), CV (support for cross-validation), params (parameters returned, one or all model sizes).
-}
+\caption{\label{tab:packages} Properties (rows) of six binary
+  segmentation libraries (columns).  Properties are weights (can
+  specify observation-specific positive weights for the loss), min len
+  (can specify minimum segment length parameter), max segs (can
+  specify maximum number of segments parameter), dim (dimensions of
+  input data), correct (does code return correct model parameters?),
+  losses (number of loss functions implemented), language (programming
+  language to implement loss), storage (technique used to store cost
+  of splitting a segment), space (space complexity in number of
+  segments $S$), cumsum (cumulative sum used), best and worst case
+  time complexity (using $N$ data and $S=N$ segments), CV (support for
+  cross-validation), params (parameters returned, one or all model
+  sizes). }
 \end{table}
@@
 There are several free/open-source implementations of binary
-segmentation, most of which use C/C++ code for efficiency
-(Table~\ref{tab:packages}).
-We compared our proposed implementation to these baselines in terms of functionality, correctness, and efficiency.
+segmentation, most of which are \proglang{R} packages with
+\proglang{C}/\proglang{C++} code (Table~\ref{tab:packages}).  We
+compared our proposed implementation to these baselines in terms of
+functionality, correctness, and efficiency.  None of these existing
+implementations support cross-validation, nor observation-specific
+weights, which are unique features that we propose in \pkg{binsegRcpp}.
 
-The \pkg{fpop} package in R contains the \code{fpop::multiBinSeg}
-function, which implements the square loss for multi-dimensional data
-\citep{Maidstone2017}.
+The \pkg{changepoint} package in \proglang{R} contains the
+\code{cpt.mean} function which implements several
+different change-point detection algorithms, including binary
+segmentation via \code{method="BinSeg"}
+\citep{Killick2014,Killick2022}.  Six loss functions are supported,
+along with min segment length parameter, max segments parameter, and
+returning the segment-specific parameters for one model (not all).
+The
+implementation does not always return correct results, and requires
+more time/space than expected. For $S$ segments, an $S\times S$ matrix
+is used to represent the segment end positions, whereas only a vector
+of $S$ values is required. Therefore, the time complexity is $O(N
+S^2)$, much slower than the expected $O(N\log S)$. Likewise, the space
+complexity is $O(S^2)$ rather than the expected $O(S)$.
 
-The \pkg{wbs} package in R contains the \code{wbs::sbs} function,
+The \pkg{wbs} package in \proglang{R} contains the \code{sbs} function,
 which implements the square loss for univariate data
 \citep{Baranowski2019}. Since it uses recursive function calls, it
-does not support the max segments parameter $K$, and instead always
+does not support the max segments parameter $S$, and instead always
 computes the full path of models (from 1 to $N-1$ change-points).
+It does not implement min segment length, nor
+returning the segment mean parameters.
 
-The \code{changepoint::cpt.mean(method="BinSeg")} \citep{Killick2014,Killick2022}.
+The \pkg{fpop} package in \proglang{R} contains the \code{multiBinSeg}
+function, which implements the square loss for multi-dimensional data
+\citep{Maidstone2017}. It implements the max segments parameter, but
+not min segment length, nor
+returning the segment mean parameters.
 
-TODO \code{ruptures.Binseg} \citep{Truong2020}.
+The \pkg{ruptures} package in \proglang{python} contains the
+\code{Binseg} class, which implements binary segmentation for 10 loss
+functions \citep{Truong2020}. It supports min length parameter and max
+segments parameter. The cumulative sum is not used in the
+implementation of the square loss, which is normally required to
+achieve optimal best case time complexity. When running on $N$ best
+case data, with $S=N$ segments, we have observed approximately
+$O(N^{1.4})$ time complexity (see Results section), which is slower
+than the expected/optimal $O(N\log N)$ time complexity.
 
-TODO \code{fastcpd} package \citep{fastcpd2024} (nice related work table).
+The \pkg{blockcpd} package in \proglang{R} contains the
+\code{fit\_blockcpd} function, which implements several algorithms for
+change-point detection, including binary segmentation via
+\code{method="hierseg"} \citep{blockcpd2021,Prates2021}. It supports
+input of min segment length and maximum segments, and output of
+segment mean parameters for one model size (not all).
 
-TODO \pkg{blockcpd} package TODO \citep{blockcpd2021,Prates2021} .
-
-\section{Models and software} \label{sec:models}
+\section{Statistical model and algorithm} \label{sec:models}
 
@@
 \paragraph{Recursive update rules.}
-Let $K\in\{ 2, \dots, N\}$ be the user-specified max number of segments to consider.
-The binary segmentation algorithm recursively computes the following quantities for each $k\in\{2,\dots,K\}$.
+Let $S\in\{ 2, \dots, N\}$ be the user-specified max number of segments to consider.
+The binary segmentation algorithm recursively computes the following quantities for each $k\in\{2,\dots,S\}$.
 First, the best segment to split is computed by greedy minimization of the loss decrease over all splittable segments.
@@
 
-In general, for any number of segments $K$, we can compute a table of
-estimated means (one row for each segment), using the following
-algorithm.  First, consider only the first $K$ rows of the
-\code{splits} table.  Next, set values in \code{before.mean} and
-\code{after.mean} columns to missing (\code{NA}), based on values in
-\code{invalidates.index} and \code{invalidates.after} columns.
-Finally, sort the \code{end} values, and use the corresponding
-non-missing mean values.
-For $K$ segments, this algorithm is $O(K \log K)$ time due to the sort.
-It is implemented in the \code{coef} method, as shown in the example code below:
+In general, after running binary segmentation up to $S$ segments, we
+can compute a table of estimated means (one row for each segment), for
+any number of segments $k\in\{1,\dots,S\}$, using the following algorithm.
+First, consider only the first $k$ rows of the \code{splits} table.
+Next, set values in \code{before.mean} and \code{after.mean} columns
+to missing (\code{NA}), based on values in \code{invalidates.index}
+and \code{invalidates.after} columns.  Finally, sort the \code{end}
+values, and use the corresponding non-missing mean values.  For $k$
+segments, this algorithm is $O(k \log k)$ time due to the sort.  It is
+implemented in the \code{coef} method, as shown in the example code
+below:
 

### 2025-05-08 article.Rnw
distributions

@@
 
-We assume there is a sequence of $N$ data $x_1,\dots,x_N\in\mathbb R$, in which we want to find abrupt changes.
-We assume that each data point $i\in\{1,\dots,N\}$ has a corresponding weight $w_i>0$.
-Binary
-segmentation can be interpreted as attempting to find model parameters
-$\theta$ which minimize a given weighted loss function $w_i
-\ell(\theta, x_i)$. For example, minimizing the square loss $w_i
+We assume there is a sequence of $N$ data $x_1,\dots,x_N\in\mathbb R$,
+in which we want to find abrupt changes.  We
+further assume that on each segment there is a constant model
+parameter $\theta$, typically the mean (and perhaps
+variance/scale). We would like to find a sequence of change-points,
+such that each segment has a parameter $\theta$ which minimizes a loss
+function $\ell(\theta,x)$. The loss functions that we consider to
+minimize can be interpreted as maximizing the log likelihood of
+several common distributions (Normal, Poisson, Laplace).
+
+Furthermore, we assume that each data point $i\in\{1,\dots,N\}$ has a
+corresponding weight $w_i>0$, which is used in the loss function via
+$w_i \ell(\theta, x_i)$. For example, minimizing the square loss $w_i
 \ell(\theta, x_i) = w_i (\theta-x_i)^2$ corresponds to maximizing the
@@
 Since each DNA sequence read is about 100 bases long, the coverage profile has runs of identical values, which can be converted to a run-length encoding for $\approx 100\times$ speedups.
+
+\subsection{Distributions and loss functions}
+
+We consider five change-point models with loss functions $\ell$, each
+of which have a corresponding value for the \code{distribution.str}
+argument of the proposed function, \code{binsegRcpp::binseg}.
+
+\paragraph{Square loss, normal change in mean.}
+The total square loss is used for \verb|distribution.str="mean_norm"|.
+The segment-specific parameter $\theta\in\mathbb R$ is the mean.
+
+\paragraph{Normal change in mean and variance.}
+The total Gaussian negative log likelihood is used as the loss function for \verb|distribution.str="meanvar_norm"|.
+The segment-specific parameter $\theta\in\mathbb R^2$ is the mean and variance.
+The variance is estimated via
+\begin{equation}
+  Y/W -\mu(\mu-2X/W).
+\end{equation}
+TODO discuss max zero var, which is the largest value that should be
+considered numerically zero for a variance estimate. \code{INFINITY}
+is returned for the loss if the estimated variance is smaller than max
+zero var.
+
+\paragraph{Poisson change in rate parameter (mean and variance).}
+The total Poisson loss is used for \verb|distribution.str="poisson"|
+
+\paragraph{L1 loss, Laplace change in median.}
+The total absolute error is used for \verb|distribution.str="l1"|
+
+\paragraph{Laplace change in median and scale.}
+The total Laplace negative log likelihood is used for
+\verb|distribution.str="laplace"|
 

### 2025-05-09 article.Rnw
parameter abs

@@
 
-\subsection{Distributions and loss functions}
+\section{Distributions and loss functions}
 
@@
 argument of the proposed function, \code{binsegRcpp::binseg}.
+The five models can be divided in two categories.
 
-\paragraph{Square loss, normal change in mean.}
-The total square loss is used for \verb|distribution.str="mean_norm"|.
+\subsection{Loss functions based on absolute error}
+
+Two models have loss functions that involve the absolute error
+function, and the Laplace distribution. The probability density
+function of the Laplace distribution, for data $x\in\mathbb R$, with
+location parameter $\mu\in\mathbb R$, and scale parameter $b\in\mathbb
+R$, is
+\begin{equation}
+  \frac{1}{2b} \exp\left(
+  \frac{|x-\mu|}{-b}
+  \right).
+\end{equation}
+For \verb|distribution.str="l1"|, the segment-specific parameter
+$\theta=\mu\in\mathbb R$ is the median. In this case, maximizing the
+probability corresponds to minimizing the negative log likelihood,
+which simplifies to the absolute error loss function,
+\begin{equation}
+  \ell(\mu, x) = |\mu-x|.
+\end{equation}
+
+For \verb|distribution.str="laplace"|, the segment-specific parameter
+$\theta=[\mu,b]\in\mathbb R^2$ is a vector with two values, the median
+$\mu$ and scale $b$. In this case, the loss function we use is the
+negative log likelihood,
+\begin{equation}
+  \ell(\theta,x)=\log(2b) + |\mu-x|/b.
+\end{equation}
+
+%% using the cumulative median
+%% For loss functions which require the cumulative median, we use the C++ code below
+%% \begin{verbatim}
+%% #define ABS_DIST(NAME, SCALE)...
+%% ABS_DIST(laplace, true)
+%% ABS_DIST(l1, false)
+%% \end{verbatim}
+%% The loss function which requires a variance estimate is the Laplace negative log likelihood; the simple L1 loss is obtained without a variance estimate.
+A \code{class absDistribution} implements the common operations for the two loss functions, including parameter estimation via a cumulative median algorithm, which is required to efficiently compute the optimal split point on a segment.
+To implement the cumulative median, we used a modified version of the L1 loss minimization code proposed by \citet{Drouin2017mmit}.
+For a segment from $j$ to $e$ ($1\leq j\leq e\leq N$), the estimated parameters are
+\begin{eqnarray}
+  \mu^* &=& \argmin_\mu \sum_{i=j}^e w_i |\mu-x_i| \\
+  b^* &=& \frac{\sum_{i=j}^e w_i |\mu^*-x_i|}{\sum_{i=j}^e w_i}.
+\end{eqnarray}
+Note that for an even number of data, there is a range of $\mu^*$ values which minimize the absolute error.
+In that case, we define $\mu^*$ as the value in the middle of the range of values which minimize the absolute error (this is the traditional definition of the sample median).
+
+%% Another C macro is used to declare the loss functions which require the cumulative sum:
+%% \begin{verbatim}
+%% #define CUM_DIST(NAME, COMPUTE, ERROR, VARIANCE)...
+%% CUM_DIST(mean_norm, RSS, , false)
+%% CUM_DIST(meanvar_norm,
+%%  (var>max_zero_var) ? (RSS/var+N*log(2*M_PI*var))/2 : INFINITY,
+%%  , true)
+%% CUM_DIST(poisson,
+%%  (mean>0) ? (mean*N - log(mean)*sum) : ( (sum==0) ? 0 : INFINITY ),
+%%  if(round(value)!=value)return ERROR_DATA_MUST_BE_INTEGER_FOR_POISSON_LOSS;
+%%  if(value < 0)return ERROR_DATA_MUST_BE_NON_NEGATIVE_FOR_POISSON_LOSS;,
+%%  false)
+%% \end{verbatim}
+%% The code above declares three loss functions based on the cumulative sum: \code{mean_norm} is the residual sum of squares (RSS), \code{poisson} loss is for count data, and \code{meanvar_norm} is the normal model with change in mean and variance.
+%% The \code{COMPUTE} parameter specifies the loss, in terms of six variables: \code{mean}, \code{var}, \code{N}, \code{sum}, \code{squares}, \code{max_zero_var} (Table~\ref{tab:cum-params}).
+%% These variables are computed in a \code{class CumDistribution} which contains the logic common to these loss functions (cumulative sum, etc).
+
+Note that the code supports cross-validation, in the sense that the input train set can be divided into a subtrain set (used to compute model parameters) and a validation set (used to evaluate the computed model parameters).
+The \code{mean} and \code{var} variables are model parameters, computed using the subtrain set, whereas the other variables depend on the data (subtrain or validation set).
+The \code{ERROR} parameter is an optional block of code, that is used to return an error status code, if there are unusual input data.
+
+%% To define the loss functions based on the cumulative sum TODO
+%% \begin{table}
+%%   \begin{tabular}{ccc}
+%%     Parameter & Description & Definition \\
+%%     \hline
+%%     $M=$\code{mean} & model parameter & $T/W$ \\
+%%     \code{var} & model parameter &  $S/W + M(M-2T/W)$
+%%   \end{tabular}
+%%   \caption{\label{tab:params}TODO}
+%% \end{table}
+
+\begin{table*}
+  \centering
+  \begin{tabular}{ccccccc}
+    \code{distribution.str} & Loss & Distrib. & Center & Scale & Params & Complexity \\
+    \hline
+    \texttt{mean\_norm} & Square & Normal & yes & no & 1 & $O(N)$ \\
+    \texttt{meanvar\_norm} & Neg. Log Lik. & Normal & yes & yes & 2 & $O(N)$ \\
+    \texttt{poisson} & Neg. Log Lik. & Poisson & yes & yes & 1 & $O(N)$ \\
+    \texttt{l1} & Absolute & Laplace & yes & no & 1 & $O(N\log N)$ \\
+    \texttt{laplace} & Neg. Log Lik & Laplace & yes & yes & 2 & $O(N\log N)$ \\
+  \end{tabular}
+  \caption{\label{tab:losses} Different change-point models provided
+    by \code{binsegRcpp::binseg}.  All models support a change in
+    center parameter; two models have an additional scale parameter;
+    change in Poisson rate parameter affects both mean and variance.
+    Complexity is for computing best split on a segment with $N$ data.
+  }
+\end{table*}
+
+Additionally, the \code{CUM_DIST} and \code{ABS_DIST} macros generate code that populates a static \code{unordered_map} which can be queried to obtain a list of distributions which are supported by the C++ code:
+%
+<<>>=
+binsegRcpp::get_distribution_info()$dist
+@ 
+%
+
+\paragraph{Square loss, normal change in mean,} \verb|distribution.str="mean_norm"|.
 The segment-specific parameter $\theta\in\mathbb R$ is the mean.
 
-\paragraph{Normal change in mean and variance.}
-The total Gaussian negative log likelihood is used as the loss function for \verb|distribution.str="meanvar_norm"|.
+\paragraph{Normal change in mean and variance,} \verb|distribution.str="meanvar_norm"|.
+The total Gaussian negative log likelihood is used as the loss function for \verb
 The segment-specific parameter $\theta\in\mathbb R^2$ is the mean and variance.
@@
 
-\paragraph{Poisson change in rate parameter (mean and variance).}
-The total Poisson loss is used for \verb|distribution.str="poisson"|
-
-\paragraph{L1 loss, Laplace change in median.}
-The total absolute error is used for \verb|distribution.str="l1"|
-
-\paragraph{Laplace change in median and scale.}
-The total Laplace negative log likelihood is used for
-\verb|distribution.str="laplace"|
+\paragraph{Poisson change in rate parameter (mean and variance),} \verb|distribution.str="poisson"|.
+The total Poisson loss is used for
 
@@
 For the case of negative log likelihood of Normal and Poisson distributions, the cumulative sum trick can be used to compute these sequences in linear $O(s)$ time.
-For the case of the L1 loss and negative log likelihood of Laplace distribution, these sequences can be computed efficiently in log-linear $O(s\log s)$ time using the cumulative median algorithm of \citet{Drouin2017mmit}.
+For the case of the L1 loss and negative log likelihood of Laplace distribution, these sequences can be computed efficiently in log-linear $O(s\log s)$ time using a cumulative median algorithm.
+To implement the cumulative median, we used a modified version of the L1 loss minimization code proposed by \citet{Drouin2017mmit}.
 
@@
 
-\subsection{Choice of loss function}
-
-C macros are used to declare the different loss functions.
-For loss functions which require the cumulative median, we use the C++ code below
-\begin{verbatim}
-#define ABS_DIST(NAME, VARIANCE)...
-ABS_DIST(laplace, true)
-ABS_DIST(l1, false)
-\end{verbatim}
-The loss function which requires a variance estimate is the Laplace negative log likelihood; the simple L1 loss is obtained without a variance estimate.
-A \code{class absDistribution} implements the common operations for the two loss functions, including the cumulative median in $O(N\log N)$ time using the algorithm of \citet{Drouin2017mmit}.
-
-Another C macro is used to declare the loss functions which require the cumulative sum:
-\begin{verbatim}
-#define CUM_DIST(NAME, COMPUTE, ERROR, VARIANCE)...
-CUM_DIST(mean_norm, RSS, , false)
-CUM_DIST(meanvar_norm,
- (var>max_zero_var) ? (RSS/var+N*log(2*M_PI*var))/2 : INFINITY,
- , true)
-CUM_DIST(poisson,
- (mean>0) ? (mean*N - log(mean)*sum) : ( (sum==0) ? 0 : INFINITY ),
- if(round(value)!=value)return ERROR_DATA_MUST_BE_INTEGER_FOR_POISSON_LOSS;
- if(value < 0)return ERROR_DATA_MUST_BE_NON_NEGATIVE_FOR_POISSON_LOSS;,
- false)
-\end{verbatim}
-The code above declares three loss functions based on the cumulative sum: \code{mean_norm} is the residual sum of squares (RSS), \code{poisson} loss is for count data, and \code{meanvar_norm} is the normal model with change in mean and variance.
-The \code{COMPUTE} parameter specifies the loss, in terms of six variables: \code{mean}, \code{var}, \code{N}, \code{sum}, \code{squares}, \code{max_zero_var} (Table~\ref{tab:cum-params}).
-These variables are computed in a \code{class CumDistribution} which contains the logic common to these loss functions (cumulative sum, etc).
-Note that the code supports cross-validation, in the sense that the input train set can be divided into a subtrain set (used to compute model parameters) and a validation set (used to evaluate the computed model parameters).
-The \code{mean} and \code{var} variables are model parameters, computed using the subtrain set, whereas the other variables depend on the data (subtrain or validation set).
-The \code{ERROR} parameter is an optional block of code, that is used to return an error status code, if there are unusual input data.
-
-To define the loss functions based on the cumulative sum TODO
-\begin{table}
-  \begin{tabular}{ccc}
-    Parameter & Description & Definition \\
-    \hline
-    $M=$\code{mean} & model parameter & $T/W$ \\
-    \code{var} & model parameter &  $S/W + M(M-2T/W)$
-  \end{tabular}
-  \caption{\label{tab:params}TODO}
-\end{table}
-
-\begin{table*}
-  \centering
-  \begin{tabular}{ccccccc}
-    Name & Loss & Distrib. & Center & Scale & Params & Complexity \\
-    \hline
-    \texttt{mean\_norm} & Square & Normal & yes & no & 1 & $O(S)$ \\
-    \texttt{meanvar\_norm} & Neg. Log Lik. & Normal & yes & yes & 2 & $O(S)$ \\
-    \texttt{poisson} & Neg. Log Lik. & Poisson & yes & yes & 1 & $O(S)$ \\
-    \texttt{l1} & Absolute & Laplace & yes & no & 1 & $O(S\log S)$ \\
-    \texttt{laplace} & Neg. Log Lik & Laplace & yes & yes & 2 & $O(S\log S)$ \\
-  \end{tabular}
-  \caption{\label{tab:losses}
-    TODO
-  }
-\end{table*}
-
-Additionally, the \code{CUM_DIST} and \code{ABS_DIST} macros generate code that populates a static \code{unordered_map} which can be queried to obtain a list of distributions which are supported by the C++ code:
-%
-<<>>=
-binsegRcpp::get_distribution_info()$dist
-@ 
-%
 To illustrate the performence benefits of using log-time  containers  were used to efficiently , virtual classes, static

### 2025-05-09 article.Rnw
iterations

@@
 
-\section{Statistical model and algorithm} \label{sec:models}
+\section{Distributions and loss functions}
 
@@
 
-\section{Distributions and loss functions}
-
 We consider five change-point models with loss functions $\ell$, each
 of which have a corresponding value for the \code{distribution.str}
-argument of the proposed function, \code{binsegRcpp::binseg}.
-The five models can be divided in two categories.
+argument of the proposed function, \code{binsegRcpp::binseg}
+(Table~\ref{tab:losses}).  The five models can be divided in two
+categories: absolute error and cumulative sum.
 
@@
   \frac{1}{2b} \exp\left(
-  \frac{|x-\mu|}{-b}
+  \frac{|\mu-x|}{-b}
   \right).
@@
 \begin{equation}
-  \ell(\mu, x) = |\mu-x|.
+  \ell_{\texttt{l1}}(\mu, x) = |\mu-x|.
 \end{equation}
@@
 \begin{equation}
-  \ell(\theta,x)=\log(2b) + |\mu-x|/b.
+  \ell_{\texttt{laplace}}([\mu,b],x)=\log(2b) + |\mu-x|/b.
 \end{equation}
@@
 %% For loss functions which require the cumulative median, we use the C++ code below
-%% \begin{verbatim}
-%% #define ABS_DIST(NAME, SCALE)...
-%% ABS_DIST(laplace, true)
-%% ABS_DIST(l1, false)
-%% \end{verbatim}
 %% The loss function which requires a variance estimate is the Laplace negative log likelihood; the simple L1 loss is obtained without a variance estimate.
 A \code{class absDistribution} implements the common operations for the two loss functions, including parameter estimation via a cumulative median algorithm, which is required to efficiently compute the optimal split point on a segment.
-To implement the cumulative median, we used a modified version of the L1 loss minimization code proposed by \citet{Drouin2017mmit}.
+To implement the cumulative median for $N$ data in $O(N\log N)$ time, we used a modified version of the L1 loss minimization code proposed by \citet{Drouin2017mmit}.
 For a segment from $j$ to $e$ ($1\leq j\leq e\leq N$), the estimated parameters are
@@
 In that case, we define $\mu^*$ as the value in the middle of the range of values which minimize the absolute error (this is the traditional definition of the sample median).
+
+\subsection{Loss functions based on cumulative sums}
+
+There are three models which use the cumulative sum trick to find the
+best split for $N$ data in $O(N)$ time.
+In particular, we define the vectors $W,X,Y\in\mathbb R^{N+1}$ as follows:
+\begin{eqnarray}
+  \text{Cumulative weights: } W_0=0,\,&& \forall t\in\{1,\dots,N\},\, W_t=\sum_{i=1}^t w_i,\\
+  \text{Cumulative sum: }X_0=0,\,&& \forall t\in\{1,\dots,N\},\, X_t=\sum_{i=1}^t w_i x_i,\\
+  \text{Cumulative squares: }Y_0=0,\,&& \forall t\in\{1,\dots,N\},\, Y_t=\sum_{i=1}^t w_i x^2_i.
+\end{eqnarray}
+
+\begin{table*}
+  \centering
+  \begin{tabular}{ccccccc}
+    \code{distribution.str} & Loss & Distrib. & Center & Scale & Params & Complexity \\
+    \hline
+    \texttt{mean\_norm} & Square & Normal & yes & no & 1 & $O(N)$ \\
+    \texttt{meanvar\_norm} & Neg. Log Lik. & Normal & yes & yes & 2 & $O(N)$ \\
+    \texttt{poisson} & Neg. Log Lik. & Poisson & yes & yes & 1 & $O(N)$ \\
+    \texttt{l1} & Absolute & Laplace & yes & no & 1 & $O(N\log N)$ \\
+    \texttt{laplace} & Neg. Log Lik & Laplace & yes & yes & 2 & $O(N\log N)$ \\
+  \end{tabular}
+  \caption{\label{tab:losses} Different change-point models provided
+    by \code{binsegRcpp::binseg}.  All models support a change in
+    center parameter; two models have an additional scale parameter;
+    change in Poisson rate parameter affects both mean and variance.
+    Complexity is time for computing best split on a segment with $N$ data.
+  }
+\end{table*}
+
+For \verb|distribution.str="mean_norm"|, the normal change in mean model,
+the segment-specific parameter $\theta\in\mathbb R$ is the mean.
+For data $x\in\mathbb R$, mean $\mu\in\mathbb R$, and variance $\sigma^2>0$, the normal probability density function is
+\begin{equation}
+  \frac{1}{\sqrt{2\pi\sigma^2}}
+  \exp\left[
+    \frac{
+      (\mu-x)^2
+    }{
+      -2\sigma^2
+    }
+    \right].
+\end{equation}
+Maximizing the normal probability is equivalent to minimizing the negative log-likelihood, which simplifies to the square loss,
+\begin{equation}
+  \ell_{\texttt{mean\_norm}}(\mu, x) = (\mu-x)^2.
+\end{equation}
+For a segment from $j$ to $e$ ($1\leq j\leq e\leq N$), we use the notation $W_{j,e}=W_j-W_{e-1}$ for the sum of the weights over the segment, and similarly for $X_{j,e}$ and $Y_{j,e}$.
+Then the mean on the segment from $j$ to $e$ can be computed via
+\begin{equation}
+  \mu^* = X_{j,e}/W_{j,e}.
+\end{equation}
+
+For \verb|distribution.str="meanvar_norm"|, the normal change in mean
+and variance model, the segment-specific parameter $\theta=[\mu,\sigma^2]\in\mathbb R^2$ is a vector with two values, the mean $\mu$ and variance $\sigma^2$.
+The Gaussian negative log likelihood is used as the loss function,
+\begin{equation}
+\ell_{\texttt{meanvar\_norm}}([\mu,\sigma^2],x)=[\log(2\pi\sigma^2)+(\mu-x)^2/\sigma^2]/2.
+\end{equation}
+For a segment from $j$ to $e$ ($1\leq j\leq e\leq N$), the variance is estimated via
+\begin{equation}
+  \sigma^{2*}=Y_{j,e}/W_{j,e} -\mu^*(\mu^*-2X_{j,e}/W_{j,e}).
+\end{equation}
+%% TODO discuss max zero var, which is the largest value that should be
+%% considered numerically zero for a variance estimate. \code{INFINITY}
+%% is returned for the loss if the estimated variance is smaller than max
+%% zero var.
+
+For \verb|distribution.str="poisson"|, the Poisson model, the
+segment-specific parameter $\theta\geq 0$ is the rate.  The Poisson
+probability mass function for a data value $x\in\{0,1,2,\dots\}$ is
+\begin{equation}
+  \theta^x e^{-\theta} / (x!).
+\end{equation}
+Maximizing the probability is equivalent to minimizing the negative
+log likelihood, which simplifies to the Poisson loss,
+\begin{equation}
+  \ell_{\texttt{poisson}}(\theta,x)=\theta - x\log \theta.
+\end{equation}
 
@@
 
-Note that the code supports cross-validation, in the sense that the input train set can be divided into a subtrain set (used to compute model parameters) and a validation set (used to evaluate the computed model parameters).
-The \code{mean} and \code{var} variables are model parameters, computed using the subtrain set, whereas the other variables depend on the data (subtrain or validation set).
-The \code{ERROR} parameter is an optional block of code, that is used to return an error status code, if there are unusual input data.
-
 %% To define the loss functions based on the cumulative sum TODO
@@
 
-\begin{table*}
-  \centering
-  \begin{tabular}{ccccccc}
-    \code{distribution.str} & Loss & Distrib. & Center & Scale & Params & Complexity \\
-    \hline
-    \texttt{mean\_norm} & Square & Normal & yes & no & 1 & $O(N)$ \\
-    \texttt{meanvar\_norm} & Neg. Log Lik. & Normal & yes & yes & 2 & $O(N)$ \\
-    \texttt{poisson} & Neg. Log Lik. & Poisson & yes & yes & 1 & $O(N)$ \\
-    \texttt{l1} & Absolute & Laplace & yes & no & 1 & $O(N\log N)$ \\
-    \texttt{laplace} & Neg. Log Lik & Laplace & yes & yes & 2 & $O(N\log N)$ \\
-  \end{tabular}
-  \caption{\label{tab:losses} Different change-point models provided
-    by \code{binsegRcpp::binseg}.  All models support a change in
-    center parameter; two models have an additional scale parameter;
-    change in Poisson rate parameter affects both mean and variance.
-    Complexity is for computing best split on a segment with $N$ data.
-  }
-\end{table*}
+%% Additionally, the \code{CUM_DIST} and \code{ABS_DIST} macros generate code that populates a static \code{unordered_map} which can be queried to obtain a list of distributions which are supported by the C++ code:
+%% %
+%% <<>>=
+%% binsegRcpp::get_distribution_info()$dist
+%% @ 
+%% %
 
-Additionally, the \code{CUM_DIST} and \code{ABS_DIST} macros generate code that populates a static \code{unordered_map} which can be queried to obtain a list of distributions which are supported by the C++ code:
-%
-<<>>=
-binsegRcpp::get_distribution_info()$dist
-@ 
-%
+\section{Algorithm} \label{sec:algorithm}
+This section gives the equations for the binary segmentation algorithm.
 
-\paragraph{Square loss, normal change in mean,} \verb|distribution.str="mean_norm"|.
-The segment-specific parameter $\theta\in\mathbb R$ is the mean.
-
-\paragraph{Normal change in mean and variance,} \verb|distribution.str="meanvar_norm"|.
-The total Gaussian negative log likelihood is used as the loss function for \verb
-The segment-specific parameter $\theta\in\mathbb R^2$ is the mean and variance.
-The variance is estimated via
-\begin{equation}
-  Y/W -\mu(\mu-2X/W).
-\end{equation}
-TODO discuss max zero var, which is the largest value that should be
-considered numerically zero for a variance estimate. \code{INFINITY}
-is returned for the loss if the estimated variance is smaller than max
-zero var.
-
-\paragraph{Poisson change in rate parameter (mean and variance),} \verb|distribution.str="poisson"|.
-The total Poisson loss is used for
-
-\paragraph{Min loss.}
+\subsection{Min loss and difference}
 The iterative algorithm involves computing the min loss over
@@
 \end{equation}
-The minimization problem is over the model parameter $\theta$ (segment-specific center and possibly scale).
+The minimization problem is over the model parameter $\theta$ (which affect the center and possibly scale/variance).
 
@@
 The $L_{j,e}$ term is the loss before the split, and $L_{j,t}+L_{t+1,n}$ is the loss after the split, which is necessarily smaller.
-Let $\delta_{j,e},\tau_{j,e}=\min,\argmin_{t\in \{j,\dots,e-1\}} D_{j,e}(t)$ be the best loss difference/split on segment $(j,e)$.
+We define the best loss difference/split on segment $(j,e)$ as
+\begin{equation}
+  \label{eq:fje}
+  f_{j,e}=\min,\argmin_{t\in \{j,\dots,e-1\}} D_{j,e}(t).
+\end{equation}
 %% We can re-write this minimization as
@@
 Let $\mathcal L_1=L_{1,n}$ be the loss with one segment,
-and let $\delta_1,t_1= f_{1,n}$ be the best loss difference and split point for the entire data.
-We initialize the set of segments to split as $\mathcal S_1=\{(1,N,d_1,t_1)\}$.
+and let $ f_{1,n}$ be the best loss difference and split point for the entire data.
+We initialize the set of segments to split as $\mathcal S_1=\{(1,N,f_{1,n})\}$.
 
@@
 Let $S\in\{ 2, \dots, N\}$ be the user-specified max number of segments to consider.
-The binary segmentation algorithm recursively computes the following quantities for each $k\in\{2,\dots,S\}$.
+The binary segmentation algorithm recursively computes the following quantities for each iteration $k\in\{2,\dots,S\}$.
 First, the best segment to split is computed by greedy minimization of the loss decrease over all splittable segments.
@@
 \end{equation}
-Ideally, the minimization above takes $O(1)$ constant time, if the set of splittable segments is a red-black tree (C++ STL multiset).
-As a baseline, we also consider a linked list, which results in a sub-optimal $O(k)$ linear time search (see Section~\ref{sec:data-structures} for details).
+Ideally, the minimization above takes $O(1)$ constant time, if the set
+of splittable segments is stored in an efficient container (for example, C++ Standard Template Library multiset or priority\_queue). As a
+baseline, we also consider a linked list, which results in a
+sub-optimal $O(k)$ linear time search at iteration $k$ (see
+Section~\ref{sec:data-structures} for details).
 
-Second, the total loss is computed by adding the best loss decrease $d^*_k$ to the best loss $L_{k-1}$ at the previous iteration, 
+In the second update at iteration $k$, the total loss is computed by adding the best loss decrease $d^*_k$ to the best loss $L_{k-1}$ at the previous iteration, 
 \begin{equation}
@@
 \end{equation}
-TODO time.
+For an efficient implementation, the time complexity of these
+remove/insert operations should be sub-linear. The list container
+offers constant time insertion/removal; the multiset and
+priority\_queue containers offer constant time removal and log-time
+insertion.
 
@@
 
+\section{Other features}
+We used several features of C++ to implement \pkg{binsegRcpp} efficiently.
+\subsection{Cross-validation}
+Note that the code supports cross-validation, in the sense that the input train set can be divided into a subtrain set (used to compute model parameters) and a validation set (used to evaluate the computed model parameters).
+The \code{mean} and \code{var} variables are model parameters, computed using the subtrain set, whereas the other variables depend on the data (subtrain or validation set).
+The \code{ERROR} parameter is an optional block of code, that is used to return an error status code, if there are unusual input data.
 \subsection{Choice of container}
 
-We used several features of C++ to implement \pkg{binsegRcpp} efficiently.
 In binary segmentation, a segment is split in each iteration, which results in up to two new segments that must be considered to split (see Section~\ref{sec:algo} for details).

### 2025-05-12 article.Rnw
seq param est

@@
 \end{equation}
+For a segment from $j$ to $e$ ($1\leq j\leq e\leq N$), the estimated parameters are
+\begin{eqnarray}
+  \mu_{j,e}^* &=& \argmin_\mu \sum_{i=j}^e w_i |\mu-x_i| \\
+  b_{j,e}^* &=& \frac{\sum_{i=j}^e w_i |\mu^*-x_i|}{\sum_{i=j}^e w_i}.
+\end{eqnarray}
+Note that there could be a range of $\mu_{j,e}^*$ values which minimize the absolute error (for example with an even number of data and uniform weights).
+In that case, we define $\mu_{j,e}^*$ as the value in the middle of the range of values which minimize the absolute error (this is the traditional definition of the sample median).
 
@@
 %% The loss function which requires a variance estimate is the Laplace negative log likelihood; the simple L1 loss is obtained without a variance estimate.
-A \code{class absDistribution} implements the common operations for the two loss functions, including parameter estimation via a cumulative median algorithm, which is required to efficiently compute the optimal split point on a segment.
-To implement the cumulative median for $N$ data in $O(N\log N)$ time, we used a modified version of the L1 loss minimization code proposed by \citet{Drouin2017mmit}.
-For a segment from $j$ to $e$ ($1\leq j\leq e\leq N$), the estimated parameters are
-\begin{eqnarray}
-  \mu^* &=& \argmin_\mu \sum_{i=j}^e w_i |\mu-x_i| \\
-  b^* &=& \frac{\sum_{i=j}^e w_i |\mu^*-x_i|}{\sum_{i=j}^e w_i}.
-\end{eqnarray}
-Note that for an even number of data, there is a range of $\mu^*$ values which minimize the absolute error.
-In that case, we define $\mu^*$ as the value in the middle of the range of values which minimize the absolute error (this is the traditional definition of the sample median).
+A \code{class absDistribution} implements the common operations for
+the \texttt{l1} and \texttt{laplace} loss functions, including
+parameter estimation via a cumulative median algorithm, which is
+required to efficiently compute the optimal split point on a segment.
+In particular, when we search for the best split on a segment with
+$s=e-j+1$ data, we need to compute the sequence of parameter estimates
+$\mu^*_{j,j},\dots,\mu^*_{j,e-1}$ before the split, and
+$\mu^*_{j+1,e},\dots,\mu^*_{e,e}$ after the split.  Computing these
+parameter estimates and corresponding loss values can be done in
+$O(s\log s)$ time, using a modified version of the L1 loss
+minimization code proposed by \citet{Drouin2017mmit}.
 

### 2025-05-13 article.Rnw
two changes fig

@@
 
-\section{Distributions and loss functions}
+\section{Statistical model, distributions and loss functions}
 
 We assume there is a sequence of $N$ data $x_1,\dots,x_N\in\mathbb R$,
-in which we want to find abrupt changes.  We
-further assume that on each segment there is a constant model
-parameter $\theta$, typically the mean (and perhaps
-variance/scale). We would like to find a sequence of change-points,
-such that each segment has a parameter $\theta$ which minimizes a loss
-function $\ell(\theta,x)$. The loss functions that we consider to
-minimize can be interpreted as maximizing the log likelihood of
-several common distributions (Normal, Poisson, Laplace).
+in which we want to find abrupt changes.  For each index
+$i\in\{1,\dots,N\}$, we assume $x_i\sim\mathcal D(\theta_i)$, meaning
+the observed data value $x_i$ is drawn from some distribution
+$\mathcal D$ with un-observed parameter $\theta_i$, typically the mean
+(and perhaps variance/scale).  We would like to compute a sequence of
+parameter values, $\theta_1,\dots,\theta_N$, which is piecewise
+constant.  We would like to find a sequence of change-points, such
+that each segment has a parameter $\theta$ which minimizes a loss
+function $\ell(\theta,x)$ that depends on the distribution $\mathcal
+D$. The loss functions that we consider to minimize can be interpreted
+as maximizing the log likelihood of several common distributions
+(Normal, Poisson, Laplace).
 
@@
 function of the Laplace distribution, for data $x\in\mathbb R$, with
-location parameter $\mu\in\mathbb R$, and scale parameter $b\in\mathbb
-R$, is
+location parameter $\mu\in\mathbb R$, and scale parameter $b>0$, is
 \begin{equation}
@@
 %% The loss function which requires a variance estimate is the Laplace negative log likelihood; the simple L1 loss is obtained without a variance estimate.
-A \code{class absDistribution} implements the common operations for
-the \texttt{l1} and \texttt{laplace} loss functions, including
-parameter estimation via a cumulative median algorithm, which is
-required to efficiently compute the optimal split point on a segment.
-In particular, when we search for the best split on a segment with
-$s=e-j+1$ data, we need to compute the sequence of parameter estimates
-$\mu^*_{j,j},\dots,\mu^*_{j,e-1}$ before the split, and
-$\mu^*_{j+1,e},\dots,\mu^*_{e,e}$ after the split.  Computing these
-parameter estimates and corresponding loss values can be done in
-$O(s\log s)$ time, using a modified version of the L1 loss
+In the proposed C++ code, a \code{class absDistribution} implements
+the common operations for the \texttt{l1} and \texttt{laplace} loss
+functions, including parameter estimation via a cumulative median
+algorithm, which is required to efficiently compute the optimal split
+point on a segment.  In particular, when we search for the best split
+on a segment with $s=e-j+1$ data, we need to compute the sequence of
+parameter estimates $\mu^*_{j,j},\dots,\mu^*_{j,e-1}$ before the
+split, and $\mu^*_{j+1,e},\dots,\mu^*_{e,e}$ after the split.
+Computing these parameter estimates and corresponding loss values can
+be done in $O(s\log s)$ time, using a modified version of the L1 loss
 minimization code proposed by \citet{Drouin2017mmit}.
@@
 There are three models which use the cumulative sum trick to find the
-best split for $N$ data in $O(N)$ time.
-In particular, we define the vectors $W,X,Y\in\mathbb R^{N+1}$ as follows:
+best split on a segment with $s$ data in $O(s)$ time.
+In particular, for a sequence of $N$ data, we compute the vectors $W,X,Y\in\mathbb R^{N+1}$ in linear $O(N)$ time via:
 \begin{eqnarray}
-  \text{Cumulative weights: } W_0=0,\,&& \forall t\in\{1,\dots,N\},\, W_t=\sum_{i=1}^t w_i,\\
-  \text{Cumulative sum: }X_0=0,\,&& \forall t\in\{1,\dots,N\},\, X_t=\sum_{i=1}^t w_i x_i,\\
-  \text{Cumulative squares: }Y_0=0,\,&& \forall t\in\{1,\dots,N\},\, Y_t=\sum_{i=1}^t w_i x^2_i.
+  \text{Cumulative weights: } W_0=0,\,&& \forall t\in\{1,\dots,N\},\, W_t=\sum_{i=1}^t w_i = W_{t-1}+w_i,\\
+  \text{Cumulative sum: }X_0=0,\,&& \forall t\in\{1,\dots,N\},\, X_t=\sum_{i=1}^t w_i x_i=X_{t-1}+w_i x_i,\\
+  \text{Cumulative squares: }Y_0=0,\,&& \forall t\in\{1,\dots,N\},\, Y_t=\sum_{i=1}^t w_i x^2_i=Y_{t-1}+w_i x_i^2.
 \end{eqnarray}
@@
 For a segment from $j$ to $e$ ($1\leq j\leq e\leq N$), we use the notation $W_{j,e}=W_j-W_{e-1}$ for the sum of the weights over the segment, and similarly for $X_{j,e}$ and $Y_{j,e}$.
-Then the mean on the segment from $j$ to $e$ can be computed via
+Then the mean on the segment from $j$ to $e$ can be computed in constant $O(1)$ time via
 \begin{equation}
+  \label{eq:mu-star}
   \mu^* = X_{j,e}/W_{j,e}.
@@
 \end{equation}
-For a segment from $j$ to $e$ ($1\leq j\leq e\leq N$), the variance is estimated via
+For a segment from $j$ to $e$ ($1\leq j\leq e\leq N$), the variance estimate is computed in constant $O(1)$ time via
 \begin{equation}
@@
 For \verb|distribution.str="poisson"|, the Poisson model, the
-segment-specific parameter $\theta\geq 0$ is the rate.  The Poisson
+segment-specific parameter $\theta=\mu\geq 0$ is the rate.  The Poisson
 probability mass function for a data value $x\in\{0,1,2,\dots\}$ is
 \begin{equation}
-  \theta^x e^{-\theta} / (x!).
+  \mu^x e^{-\mu} / (x!).
 \end{equation}
@@
 \begin{equation}
-  \ell_{\texttt{poisson}}(\theta,x)=\theta - x\log \theta.
+  \ell_{\texttt{poisson}}(\mu,x)=\mu - x\log \mu.
 \end{equation}
+The Poisson rate parameter $\mu$ is estimated via the sample mean~(\ref{eq:mu-star}).
 
@@
 \section{Algorithm} \label{sec:algorithm}
-This section gives the equations for the binary segmentation algorithm.
+This section gives details of the binary segmentation algorithm.
 
-\subsection{Min loss and difference}
+\subsection{Efficient computation of optimal split point on a segment}
 The iterative algorithm involves computing the min loss over
@@
 \end{equation}
-The minimization problem is over the model parameter $\theta$ (which affect the center and possibly scale/variance).
+The minimization problem is over the model parameter $\theta$.
 
@@
 %% \end{equation}
-\paragraph{Efficient computation.}
+\paragraph{Efficient computation of optimal split point.}
 To compute the loss differences $D_{j,e}(j),\dots,D_{j,e}(e-1)$, we need to compute the sequence of values $L_{j,j}, \dots, L_{j,e-1}$ (loss values for possible new segments before candidate split points), and the sequence of values $L_{j+1,e}, \dots, L_{e,e}$ (loss values for possible new segments after candidate split points).
@@
 
-\subsection{Algorithm pseudo-code}
-
+\subsection{Containers for efficient retreival of next segment to split}
+\label{sec:data-structures}
 The binary segmentation algorithm needs to keep track of a set of segments that can be split, $\mathcal S_k$ for iteration $k$.
@@
 
+\begin{table}
+  \centering
+  \begin{tabular}{c|cc|cc|cc}
+    Splits: & \multicolumn{2}{c|}{one split, iteration $k$} & \multicolumn{2}{c|}{$S$ equal splits} & \multicolumn{2}{c}{$S$ unequal splits} \\
+    operation: & insert &
+    %remove &
+    argmin & insert & argmin & insert & argmin\\
+    \hline
+    list & $O(1)$ &
+    % $O(1)$ &
+    $O(k)$ &
+    $O(S)$ & $O(S^2)$ &
+    $O(S)$ & $O(S)$\\
+    \hline
+    heap/multiset & $O(\log k)$ &
+    % $O(\log N)$ &
+    $O(1)$ &
+    $O(S\log S)$ & $O(S)$ &
+    $O(S)$ & $O(S)$\\
+    \hline
+    Compute loss & \multicolumn{2}{c|}{---} & \multicolumn{2}{c|}{$O(N\log S)$} & \multicolumn{2}{c}{$O(NS)$}
+  \end{tabular}
+  \caption{
+    Asymptotic time complexity for container operations (insert new segment which could be split, argmin to find best segment to split next).
+    At iteration $k$, list takes $O(k)$ time, whereas heap/multiset takes $O(\log k)$ time.
+    Best case time for finding $S$ splits in $N$ data is $O(N\log S)$,
+    which can be achieved by heap/multiset, because insert $O(S\log S)$ time is faster.
+    However, best case time can not be achieved by list, because argmin $O(S^2)$ time is slower.
+  }
+  \label{tab:containers}
+\end{table}
+In binary segmentation, a segment is split in each iteration, which results in up to two new segments that must be considered to split. % (\ref{eq:Nk}).
+Computing the cost of splitting a segment of size $s$ is $O(s\log s)$ time for L1/Laplace losses, and $O(s)$ time for the normal/Poisson losses.
+After computing the cost for two new segments, the cost of all splittable segments must be considered, in order to identify the min cost segment to split next.
+We propose using three different kinds of C++ Standard Template Library containers to represent $\mathcal S_k$ (set of splittable segments at iteration $k$).
+Each container has a different implementation of storage and retreival of the cost of segments which have previously been considered, but not yet split (Table~\ref{tab:containers}).
+First, we propose using the \code{multiset} container (red-black tree) and \code{priority\_queue} container (heap), which can be used to achieve best case time complexity.
+These containers keep segments sorted by their cost values, so offer constant $O(1)$ time retreival of the min cost segment to split next; insertion takes logarithmic $O(\log n)$ time in the number of segments $n$ currently stored in the container.
+
+\paragraph{List container as baseline.}
+Second, we propose using the \code{list} container (doubly-linked list), which is used as a baseline, to demonstrate the benefits of the other containers.
+The \code{list} keeps segments stored in the order in which they were inserted, so offers constant $O(1)$ time insertion of new elements, and linear $O(n)$ time retreival of the min cost segment to split, for a container with $n$ segments currently stored.
+In the best case run-time of binary segmentation, there are equal sized splits at each iteration, with two new segments that could be split, but only one that is split (and the other one which stays in the set $\mathcal S_k$).
+Therefore the best case size of the set of splittable segments at iteration $k$ is linear, $|\mathcal S_k|=O(k)$, and the \code{list} container therefore takes $O(k)$ time to find the next segment to split.
+Going up to $S$ segments, the list therefore takes $O(S^2)$ time, which can dominate the overall time for large $S$, and prevent the code from achieving the best case $O(N \log S)$ time.
+
+\subsection{Algorithm pseudo-code}
+This section gives the update rules for the binary segmentation algorithm.
 \paragraph{Initialization.}
@@
 \end{equation}
-Ideally, the minimization above takes $O(1)$ constant time, if the set
+Ideally, the minimization above takes $O(1)$ constant time (independent of the number of segments that could be split), if the set
 of splittable segments is stored in an efficient container (for example, C++ Standard Template Library multiset or priority\_queue). As a
@@
 \begin{equation}
+  \label{eq:Nk}
   \mathcal N_k=\{
@@
 \end{equation}
-
-Fourth, we subset the two new segments, to consider only the segments that could possibly be split:
+New segments in $N_k$ could be too small to split, if there is only one data point on that segment.
+We therefore consider only the segments that are large enough to be split:
 \begin{equation}
@@
 For an efficient implementation, the time complexity of these
-remove/insert operations should be sub-linear. The list container
+remove/insert operations should be sub-linear in the container size,
+which is true of all three containers that we consider. The list container
 offers constant time insertion/removal; the multiset and
@@
 
-\section{Other features}
-We used several features of C++ to implement \pkg{binsegRcpp} efficiently.
-\subsection{Cross-validation}
-Note that the code supports cross-validation, in the sense that the input train set can be divided into a subtrain set (used to compute model parameters) and a validation set (used to evaluate the computed model parameters).
-The \code{mean} and \code{var} variables are model parameters, computed using the subtrain set, whereas the other variables depend on the data (subtrain or validation set).
-The \code{ERROR} parameter is an optional block of code, that is used to return an error status code, if there are unusual input data.
-\subsection{Choice of container}
-
-In binary segmentation, a segment is split in each iteration, which results in up to two new segments that must be considered to split (see Section~\ref{sec:algo} for details).
-Computing the cost of splitting a segment of size $N$ is $O(N\log N)$ time for L1/Laplace losses, and $O(N)$ time for the normal/Poisson losses.
-After computing the cost for two new segments, the cost of all splittable segments must be considered, in order to identify the min cost segment to split next.
-We propose using three different kinds of C++ Standard Template Library containers to store and retreive the cost of segments which have previously been considered, but not yet split (Table~\ref{tab:containers}).
-First, we propose using the \code{multiset} container (red-black tree) and \code{priority\_queue} container (heap), which can be used to achieve best case time complexity.
-These containers keep segments sorted by their cost values, so offer constant $O(1)$ time retreival of the min cost segment to split next; insertion takes logarithmic $O(\log n)$ time in the number of segments $n$ currently stored in the container.
-Second, we propose using the \code{list} container (doubly-linked list), which is used as a baseline, to demonstrate the benefits of the other containers.
-The \code{list} keeps segments stored in the order in which they were inserted, so offers constant $O(1)$ time insertion of new elements, and linear $O(n)$ time retreival of the min cost segment to split, for a container with $n$ segments currently stored.
-
-%% A C++ virtual class \code{Container} is used to represent the methods that must be provided by either type of container: \code{insert}, \code{get_best}.
-%% We used a C macro, \code{CMAKER}, to declare the two sub-classes:
-%% \begin{verbatim}
-%% #define CMAKER(CONTAINER, INSERT, GET_BEST)...
-%% CMAKER(multiset, insert, segment_container.begin())
-%% CMAKER(list, push_back, std::min_element(
-%%  segment_container.begin(),segment_container.end()))
-%% \end{verbatim}
-%% The code above is expanded by the C pre-processor, to two declarations of \code{Container} sub-classes.
-
-To illustrate the performence benefits of using log-time  containers  were used to efficiently , virtual classes, static
-variables, function pointers, templates, macros).
-
-\paragraph{Loss computation.} We run each algorithm on \code{N.data}
-points up to \code{max.segments =
-  max.changes+1}. \code{binsegRcpp::binseg} result is a list which
-contains a data table with \code{max.segments} rows and column
-\code{loss} that is the square loss.  \code{changepoint::cpt.mean}
-result is a list of class \code{cpt.range} with method \code{logLik}
-which returns the square loss of one of the models. \code{wbs::sbs}
-result is a list which contains a data frame with \code{N.data-1} rows
-and \code{CUSUM} and \code{min.th} columns
-TODO.
-\code{fpop::multiBinSeg} result is a list with element
-\code{J.est}, which is a vector of \code{max.changes} square loss
-decrease values.  \code{ruptures.Binseg.predict} result is a vector of
-segment ends for one model size, which can be passed to
-\code{sum_of_costs} method to compute the square loss.
-
-Three packages have implemented the normal change in mean and variance
-model. \code{binsegRcpp::binseg} loss values are the normal negative
-log likelihood (NLL). The \code{changepoint::logLik} function returns two
-times the NLL. The \code{ruptures.Binseg}
-loss is related via this equation,
-
-\begin{equation}
-  \text{NLL} = (\text{rupturesLoss} + N[1+\log(2\pi)])/2
-\end{equation}
-
-\begin{table}
-  \centering
-  \begin{tabular}{c|cc|cc|cc}
-    case/splits: & \multicolumn{2}{c}{one } & \multicolumn{2}{c}{best/equal} & \multicolumn{2}{c}{worst/unequal } \\
-    operation: & insert &
-                             %remove &
-                                      argmin & insert & argmin & insert & argmin\\
-    \hline
-    list & $O(1)$ &
-    % $O(1)$ &
-               $O(n)$ &
-    $O(S\log S)$ & $O(S)$ & $O(S)$ & $O(S)$\\
-    heap/multiset & $O(\log n)$ &
-    % $O(\log N)$ &
-                    $O(1)$ &
-                             $O(S)$ & $O(S^2)$ & $O(S)$ & $O(S)$\\
-    \hline
-    Find new split & \multicolumn{2}{c}{} & \multicolumn{2}{c}{$O(N\log S)$} & \multicolumn{2}{c}{$O(NS)$}
-  \end{tabular}
-  \caption{container properties}
-  \label{tab:containers}
-\end{table}
-
-\section{Storage and retreival of segment-specific model parameters}
-
+\section{Demonstration and comparison of code}
+In this section, we show example code for computing binary segmentation models using several software packages.
+We aim to emphasize how \pkg{binsegRcpp} makes it easy to retreive segment-specific model parameters.
 To simplify the discussion in this section, we use ``segment mean'' to represent arbitrary segment-specific model parameters.
 For example, segment-specific model parameters can be the mean/variance of the normal distribution, or median/scale of the Laplace distribution.
+
 A unique feature that is provided by \pkg{binsegRcpp}, but not provided by other implementations of binary segmentation, is efficient storage and retrieval of arbitrary segment mean parameters.
@@
 ex6_df <- data.frame(value = c(1, -7, 8, 10, 2, 4))
+@
+The code below can be used to compute models up to a max of 4 segments, using the square loss (normal change in mean, constant variance).
+<<>>=
 ex6_binsegRcpp <- binsegRcpp::binseg(
@@
 The output above is a table with four rows, one for each model size.
-For a max number of segments $M$, this representation of model parameters uses $M$ rows, so is asymptotically linear, $O(M)$.
+For a max number of segments $S$, this representation of model parameters uses $S$ rows, so is asymptotically linear space, $O(S)$.
 The first row represents the trivial model: one segment across all data points, with \code{end=6} equal to the index of the last data point.
@@
 
-In general, after running binary segmentation up to $S$ segments, we
-can compute a table of estimated means (one row for each segment), for
-any number of segments $k\in\{1,\dots,S\}$, using the following algorithm.
+In general, after running binary segmentation up to $S$ segments, the model parameters are stored in a table with $S$ rows.
+For any number of segments $k\in\{1,\dots,S\}$, we can compute a table of $k$ estimated means (one row for each segment), using the following algorithm.
 First, consider only the first $k$ rows of the \code{splits} table.
 Next, set values in \code{before.mean} and \code{after.mean} columns
-to missing (\code{NA}), based on values in \code{invalidates.index}
-and \code{invalidates.after} columns.  Finally, sort the \code{end}
+to missing (\code{NA}), based on values in columns \code{invalidates.index}
+and \code{invalidates.after}.  Finally, sort the \code{end}
 values, and use the corresponding non-missing mean values.  For $k$
@@
 
-<< ex6, fig=TRUE, height=3 >>=
+<< ex6, fig=FALSE, height=3 >>=
 library(ggplot2)
@@
 @
+The code above results in Figure~\ref{fig:ex6}.
 
-The figure above shows the data points in black, along with the segmentation model in green (dashed vertical change-points, and solid horizontal segment means).
+\begin{figure}[t!]
+  \centering
+  \includegraphics[width=\textwidth]{article-ex6}
+  \caption{\label{fig:ex6} Change-point models with 2 to 4 segments computed by \pkg{binsegRcpp}.
+    Data points are shown in black, along with the model in green (dashed vertical change-points, and solid horizontal segment means).
+    %and visualized using the mean parameters output via the \code{coef()} method.
+  }
+\end{figure}
 
@@
 
-To do the analogous computation with \pkg{changepoint}, we can use the code below.
+To attempt the analogous computation with \pkg{changepoint}, we can use the code below.
 
@@
 These results indicate that the implementation of binary segmentation in \pkg{changepoint} can yield incorrect change-point detection results.
+Incorrect results can also be observed in real data sets, such as in Figure~\ref{fig:neuroblastoma}, which shows that \pkg{changepoint} computes models with sub-optimal change-points and loss values.
+
+\begin{figure}[t!]
+  \centering
+  \includegraphics[width=\textwidth]{changepoint.bug.png}
+  \caption{\label{fig:neuroblastoma} For a real cancer DNA copy number
+    data set with 273 observations, we show the first four
+    change-points models for two different R packages that provide code for
+    binary segmentation: \pkg{binsegRcpp} and \pkg{changepoint}.
+  }
+\end{figure}
 
@@
 
-\section{Illustrations} \label{sec:illustrations}
+\subsection{Cross-validation}
+A unique feature of \pkg{binsegRcpp} is efficient cross-validation, which can be useful for selecting the number of change-points.
+The data input to \code{binseg()} is called the train set, which is divided into a subtrain set (used to compute model parameters) and a validation set (used to evaluate the computed model parameters).
+The assignment of data to each set is controlled by the \code{is.validation.vec} argument, which must be a logical vector (same length as the data to segment).
+For example, we can simulate a data set with two change-points:
+<<>>=
+set.seed(8)
+two_changes <- c(rnorm(7, 1), rnorm(10, 3), rnorm(5, 0))
+@
+Then we can create a vector to indicate that every other data point should be assigned to either subtrain or validation:
+<<>>=
+(is.validation.vec <- rep(c(TRUE, FALSE), l=length(two_changes)))
+@
+Then we can run binary segmentation with cross-validation via the code below:
+<<>>=
+valid_fit <- binsegRcpp::binseg("mean_norm", two_changes, 
+  is.validation.vec = is.validation.vec)
+valid_fit
+@
+The output above includes a table with one row per model size, and a column for the validation loss, which is minimized at 3 segments (as expected).
+The validation loss is computed as the total loss of the segment means with respect to the overlapping validation data points, as shown in Figure~\ref{fig:subtrain-validation}.
 
-TODO
-%
-<<data>>=
-data("quine", package = "MASS")
-@
-%
-TODO
+%% . To visualize the calculation, we can use the coef method to obtain a table of segment means, as in the code below.
+%% <<>>=
+%% (vsegs <- coef(valid_fit, 3L))
+%% @
+%% Then we create a table to represent the data sequence in the code below.
+%% << two-changes, fig=TRUE, height=2 >>=
+%% vdata <- data.table::data.table(value = two_changes,
+%%  position = seq_along(two_changes), set = ifelse(
+%%    is.validation.vec, "validation", "subtrain"))
+%% valid_join <- vsegs[vdata, .(position, mean, value, set, segments), on=.(start.pos<position, end.pos>position)]
+%% valid_join[, .(loss=sum((mean-value)^2)), by=.(set, segments)]
+%% valid_fit$splits[3, .(segments, loss, validation.loss)]
+%% gg <- ggplot() +
+%%   theme_bw()+
+%%   scale_fill_manual(values = c(subtrain = "black", validation = "white")) +
+%%   scale_color_manual(values = c("validation loss" = "violet", "subtrain mean" = "green")) +
+%%   geom_point(aes(position, value, fill = set), shape = 21, data = vdata) +
+%%   geom_segment(aes(
+%%     start.pos, mean, xend = end.pos, yend = mean, color=line),
+%%     data = data.frame(vsegs, line="subtrain mean")) + 
+%%     scale_x_continuous(breaks=1:22)+
+%%   geom_segment(aes(
+%%     position, mean,
+%%     color=line,
+%%     xend=position, yend=value),
+%%     size=1,
+%%     data=data.frame(valid_join[set=="validation"], line="validation loss"))+
+%%   theme(panel.grid.minor=element_blank())
+%% print(gg)
+%% @
+%% The data and model are visualized in Figure~\ref{fig:subtrain-validation}.
+\begin{figure}[t!]
+  \centering
+  \includegraphics[width=\textwidth]{article-two-changes}
+  \caption{\label{fig:subtrain-validation}
+    Change-point model parameters (segment means and change-point positions) are computed using the subtrain set; validation set is used for selecting the number of change-points.
+  }
+\end{figure}
+
+%% <<>>=
+%% @
+
+%% If the user wants to do model selection several subtrain/validation splits with the normal loss, the 
+%% <<>>=
+%% cv_fit <- binsegRcpp::binseg_normal_cv(two_changes)
+%% cv_fit$cv
+%% @
+
+%% A C++ virtual class \code{Container} is used to represent the methods that must be provided by either type of container: \code{insert}, \code{get_best}.
+%% We used a C macro, \code{CMAKER}, to declare the two sub-classes:
+%% \begin{verbatim}
+%% #define CMAKER(CONTAINER, INSERT, GET_BEST)...
+%% CMAKER(multiset, insert, segment_container.begin())
+%% CMAKER(list, push_back, std::min_element(
+%%  segment_container.begin(),segment_container.end()))
+%% \end{verbatim}
+%% The code above is expanded by the C pre-processor, to two declarations of \code{Container} sub-classes.
+
+%To illustrate the performence benefits of using log-time  containers  were used to efficiently , virtual classes, static variables, function pointers, templates, macros).
+
+%% \paragraph{Loss computation.} We run each algorithm on \code{N.data}
+%% points up to \code{max.segments =
+%%   max.changes+1}. The \code{binsegRcpp::binseg} result is a list which
+%% contains a data table with \code{max.segments} rows and column
+%% \code{loss} that is the square loss. The \code{changepoint::cpt.mean}
+%% result is a list of class \code{cpt.range} with method \code{logLik}
+%% which returns the square loss of one of the models. The \code{wbs::sbs}
+%% result is a list which contains a data frame with \code{N.data-1} rows
+%% (one for each possible split). The \code{fpop::multiBinSeg} result is a
+%% list with element \code{J.est}, which is a vector of
+%% \code{max.changes} square loss decrease values. The
+%% \code{ruptures.Binseg.predict} result is a vector of segment ends for
+%% one model size, which can be passed to \code{sum_of_costs} method to
+%% compute the square loss.
+
+%% Three packages have implemented the normal change in mean and variance
+%% model. \code{binsegRcpp::binseg} loss values are the normal negative
+%% log likelihood (NLL). The \code{changepoint::logLik} function returns two
+%% times the NLL. The \code{ruptures.Binseg}
+%% loss is related via this equation,
+
+%% \begin{equation}
+%%   \text{NLL} = (\text{rupturesLoss} + N[1+\log(2\pi)])/2
+%% \end{equation}
+
+\section{Empirical time complexity analysis} \label{sec:illustrations}
+This section presents results from empirical timings of the proposed
+\pkg{binsegRcpp} package, along with baselines.
+
+\subsection{Best case time using multiset but not list container}
+First, we wanted to empirically validate the expected time complexity
+of the proposed C++ code using the Standard Template Library
+containers (Table~\ref{tab:containers}).  We expected the multiset
+container to achieve the best case $O(N\log S)$ time for $N$ data and
+$S$ segments; we expected the list container to be slower, $O(S^2)$
+time, even for best case data.  To test this expectation empirically,
+we used \pkg{binsegRcpp} with synthetic data sequences of varying
+sizes $N\in\{2,4,8,\dots\}$, and a maximum number of segments $S=N/2$.
+For each $N$, the data sequence we used was $1,\dots,N$, which results
+in equal splits with the square loss: one segment of
+size $N$ split into two of size $N/2$, each of which is split into two
+of size $N/4$, etc.
+In this simulation, $S=O(N)$, so we expected time complexity of $O(N\log N)$ for the multiset, and $O(N^2)$ for the list.
+In Figure~\ref{fig:heap-refs-binsegRcpp}, it is clear that the empirical timings using the list container closely align with the $O(N^2)$ reference line, whereas the empirical timings using the multiset container closely align with the $O(N\log N)$ reference line.
+These data indicate that \pkg{binsegRcpp} with the multiset container indeed achieves the best case asymptotic time complexity, $O(N\log S)$.
+Futhermore, these data indicate that the best case time complexity can not be attained, if list container is used.
+In addition, Figure~\ref{fig:heap-refs-binsegRcpp} clearly shows the linear space complexity of both containers, $O(N)$.
 
@@
   \centering
-  \includegraphics[width=\textwidth]{changepoint.bug.png}
-  \caption{\label{fig:neuroblastoma} For a real cancer DNA copy number
-    data set with 273 observations, we show the first four
-    change-points models for two different R packages that provide code for
-    binary segmentation: \pkg{binsegRcpp} and \pkg{changepoint}.}
+  %\includegraphics[width=\textwidth]{figure-heap-refs-binsegRcpp}
+  \input{figure-heap-refs-binsegRcpp}
+  \caption{\label{fig:heap-refs-binsegRcpp}
+    Estimating best-case asymptotic complexity for time (seconds) and memory (kilobytes).
+    Black curves are empirical measurements for \pkg{binsegRcpp} with two containers, and violet curves are references.
+    These data suggest that both list and multiset containers are $O(N)$ space
+    list is $O(N^2)$ time and multiset is $O(N\log N)$ time.
+    Square loss and best case data were used (integers from 1 to $N$, result in equal splits).
+  }
 \end{figure}
+
+\subsection{Other packages which do not achieve best case time}
+Next, we wanted to empirically verify the best case asymptotic time
+complexity of two baselines, \pkg{changepoint} and \pkg{ruptures}. We
+expected that these baselines should have the asymptotically optimal
+time complexity, $O(N\log S)$ (only different from \pkg{binsegRcpp} by
+constant factors). To test this hypothesis, we ran these baselines in
+the same computational experiment described in the previous paragraph
+(square loss, best case data, $S=N/2$). We observed in Figure~\ref{fig:heap-refs-other} that
+\pkg{changepoint} takes $O(N^2)$ space, which is a larger asymptotic complexity class than expected, $O(N)$.
+Similarly, we observed in Figure~\ref{fig:heap-refs-other} that \pkg{changepoint} takes $O(N^3)$ time, which is again larger than the expected $O(N\log N)$ complexity.
+We observed in Figure~\ref{fig:heap-refs-other} that \pkg{ruptures} empirical timings align well with a $O(N^{1.4})$ asymptotic reference, which is clearly larger than $O(N\log N)$, but smaller than $O(N^2)$.
+Contrary to our expectations, these data indicate that the implementations of binary segmentation in both \pkg{changepoint} and \pkg{ruptures} do not achieve the best case asymptotic time complexity.
+
+\begin{figure}[t!]
+  \centering
+  %\includegraphics[width=\textwidth]{figure-heap-refs-other}
+  \input{figure-heap-refs-other}
+  \caption{\label{fig:heap-refs-other}
+    Estimating best-case asymptotic complexity for time (seconds) and memory (kilobytes).
+    Black curves are empirical measurements for \pkg{changepoint} and \pkg{ruptures}, and violet curves are references.
+    These data suggest that \pkg{changepoint} is $O(N^2)$ space and $O(N^3)$ time;
+    \pkg{ruptures} is $O(N)$ space and $\approx O(N^{1.4})$ time.
+    Square loss and best case data were used (integers from 1 to $N$, result in equal splits).
+  }
+\end{figure}
+
+\subsection{Comparing timings and throughput with baselines}
+In Figure~\ref{fig:heap-pred}, we show the previously described empirical timings (Figures~\ref{fig:heap-refs-binsegRcpp}--\ref{fig:heap-refs-other}) on the same axes, for comparison.
+Whereas there are not very large timing differences for small data sizes (all work in less than one second for up to 1000 data), there are large differences for data sizez of several thousand data or larger.
+We highlight the estimated throughput at 5 seconds, which is the data size $N$ that each package can handle in that time limit.
+It can be seen that \pkg{binsegRcpp} has about the same timings, using either \code{multiset} or \code{priority\_queue} containers (as expected because they both offer log-time retreival of the next segment to split).
+Also, \pkg{binsegRcpp} with \code{list} container is about 100$\times$ slower (as expected because of the linear time retreival of the next segment to split).
+It can be seen that \pkg{ruptures} and \pkg{changepoint} are significantly slower, with about $1000\times$ smaller throughput than \pkg{binsegRcpp} (thousands versus millions of data possible in a time limit of 5 seconds).
+These data indicate that for large data (several thousand or more), \pkg{binsegRcpp} is preferable for speed.
+
+\begin{figure}[t!]
+  \centering
+  \includegraphics[width=\textwidth]{figure-heap-pred}
+  \caption{\label{fig:heap-pred} Comparing best-case computation time of \pkg{binsegRcpp} with three containers, and two baselines: \pkg{changepoint} and \pkg{ruptures}.
+    Square loss and best case data were used (integers from 1 to $N$, result in equal splits).
+  }
+\end{figure}
+
+\subsection{Other packages which achieve best case time}
+We conducted empirical timings on other implementations of binary
+segmentation (\code{wbs::sbs}, \code{fpop::multiBinSeg}, 
+\code{blockcpd::heap}).
+In Figure~\ref{fig:timings}, we show timings for the square loss in three different data sets (best case, worst case, flat).
+It is clear that \code{fpop::multiBinSeg} and \code{wbs::sbs} have very similar asymptotic timings as \pkg{binsegRcpp}.
+In the worst case, all packages show the same slope (indicating quadratic time), except \pkg{changepoint} (which has a larger slope, indicating cubic time).
 
@@
 
-\begin{figure}[t!]
-  \centering
-  \includegraphics[width=\textwidth]{figure-timings-laplace.png}
-\caption{\label{fig:timings-laplace} Timings using L1 loss.}
-\end{figure}
+We also investigated the Poisson model (Figure~\ref{fig:timings-poisson}),
+and the normal change in mean and variance model
+(Figure~\ref{fig:timings-meanvar_norm}), which are implemented by
+\code{blockcpd::heap}.  It is clear that \code{blockcpd::heap} has the
+same asymptotic slope, but much larger constant factors, when compared
+with \pkg{binsegRcpp}.
 
@@
 \caption{\label{fig:timings-poisson} Timings using Poisson loss.}
-\end{figure}
-
-\begin{figure}[t!]
-  \centering
-  \includegraphics[width=\textwidth]{figure-compare-distributions.png}
-  \caption{\label{fig:compare-distributions} Timings using binsegRcpp multiset with
-    several different loss functions.}
 \end{figure}
@@
 
-  \section{More technical details} \label{app:technical}
+  \section{More empirical timings} \label{app:more-timings}
+
+  We also performed computational experiments involving the L1 loss
+  (Laplace change in median), which was implemented by \pkg{ruptures}.
+  Figure~\ref{fig:timings-laplace} shows that \pkg{ruptures} has slightly larger asymptotic slope, and much larger constant factors, than \pkg{binsegRcpp} (in the best case).
+
+  Figure~\ref{fig:compare-distributions} shows \pkg{binsegRcpp}
+  timings for four loss functions, and three cases (best, worst,
+  flat). It can be seen that the Laplace median model (L1 loss) has
+  slightly larger asymptotic slope in the best and flat cases (but
+  same quadratic slope in the worst case).
 
@@
   \centering
-  \includegraphics[width=\textwidth]{figure-heap-pred}
-  \caption{\label{fig:heap-pred} Comparing best-case computation time of \pkg{binsegRcpp} with three containers, and two baselines: \pkg{changepoint} and \pkg{ruptures}.
-    Square loss and best case data were used (integers from 1 to $N$, result in equal splits).
-  }
+  \includegraphics[width=\textwidth]{figure-timings-laplace.png}
+\caption{\label{fig:timings-laplace} Timings using L1 loss.}
 \end{figure}
@@
   \centering
-  %\includegraphics[width=\textwidth]{figure-heap-refs-binsegRcpp}
-  \input{figure-heap-refs-binsegRcpp}
-  \caption{\label{fig:heap-refs}
-    Estimating best-case asymptotic complexity for time (seconds) and memory (kilobytes).
-    Black curves are empirical measurements for \pkg{binsegRcpp} with two containers, and violet curves are references.
-    These data suggest that both list and multiset containers are $O(N)$ space
-    list is $O(N^2)$ time and multiset is $O(N\log N)$ time.
-    Square loss and best case data were used (integers from 1 to $N$, result in equal splits).
-  }
-\end{figure}
-
-\begin{figure}[t!]
-  \centering
-  %\includegraphics[width=\textwidth]{figure-heap-refs-other}
-  \input{figure-heap-refs-other}
-  \caption{\label{fig:heap-refs}
-    Estimating best-case asymptotic complexity for time (seconds) and memory (kilobytes).
-    Black curves are empirical measurements for \pkg{changepoint} and \pkg{ruptures}, and violet curves are references.
-    These data suggest that \pkg{changepoint} is $O(N^2)$ space and $O(N^3)$ time;
-    \pkg{ruptures} is $O(N)$ space and $\approx O(N^{1.4})$ time.
-    Square loss and best case data were used (integers from 1 to $N$, result in equal splits).
-  }
+  \includegraphics[width=\textwidth]{figure-compare-distributions.png}
+  \caption{\label{fig:compare-distributions} Timings using binsegRcpp multiset with
+    several different loss functions.}
 \end{figure}
